### 2022-12-10 ArticleDraft.Rmd
The first commit

---
title: "The First Draft"
authors:
  - name: "Mehrdad Samadishadlou"
    address: "Medical Nanotechnology gruop, Tabriz University of Medical Sciences, Tabriz, Iran"
date: "`r Sys.Date()`"
link-citation: yes
bibliography: For Article Refrence.bib
output:
  bookdown::pdf_book:
  bookdown::word_document2: default
#    base_format: rticles::bioinformatics_article
---

```{r setup, include=FALSE, cache.rebuild=TRUE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(knitr, captioner, bundesligR, stringr)
```

``` {r , echo=F}
# What to do next:

# Generating information related to the machine learning first layer using welcome file.

```

# Materials and Methods

## Microarray data collection

Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database ([https://www.ncbi.nlm.nih.gov/geo/](https://www.ncbi.nlm.nih.gov/geo/)). In order to come by sufficient classification power for MI samples and others, the sample size of the dataset should be relatively large. Therefore, GSE59867 and GSE56609 were selected as train set and GSE62646 and GSE54475 as validation set for both DEGs and pathway analyse as well as machine learning section. All datasets have been conducted using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform. Only healthy, stable CAD and early stage MI samples were selected from these datasets for further analysis. The basic information for the four GEO datasets evaluated in the current study is provided in Table 1.

|                | Platform | Healthy Control | CAD Control | MI  | Reference |
|----------------|----------|-----------------|-------------|-----|-----------|
| Training Set   |          |                 |             |     |           |
| GSE59867       | GPL6244  | \-              | 46          | 111 | [@67]     |
| GSE56609       | GPL6244  | 46              | \-          | \-  | [@09]     |
| Test Set |          |                 |             |     |           |
| GSE62646       | GPL6244  | \-              | 14          | 28  | [@46]     |
| GSE54475       | GPL6244  | 5               | \-          | \-  | [@75]     |

: Basic information of the 4 GEO microarray datasets.

## Pre-processing

Raw data (CEL files) of the four datasets were downloaded from the GEO and preprocessed using the fRMA package [@fRMA]. fRMA allows to preprocess individual microarray samples and combine them consistently for analysis. For each dataset, background correction is performed and then it is quantile normalized based on the reference distribution. During summarization, batch effects are removed and variances of the gene expressions are estimated by taking into account these probe-specific effects. For those multiple probe sets matched to the identical gene, the mean log fold change was retained. This way fRMA can be seen as a batch effect removal technique for different datasets produced using identical microarray platform. Thus, In order to ensure about batch effect removal, the principal component analysis and the relative log expression of all samples were plotted before and after fRMA [@BER].

## Differential expression analysis

The barcode algorithm proposed by McCall et al. [@Barcode] transforms the actual expression values into binary barcode values. Huge sets of samples were collected and normalized using fRMA for several platforms. The distribution of the expressed and non-expressed observed intensities for each gene is estimated using these normalized sets. Genes are deemed expressed (and their value coded to 1) or unexpressed (and their value coded to 0) according to the following equation:

$$
\hat{x}_{ij} = \left\{
  \begin{array}{ll}
    1 & \mbox{if } x_{ij} >= \mu^{ne} + C \times \sigma^{ne} \\
    0 & \mbox{otherwise}
  \end{array}
\right.
$$ where $x_{ij}$ is the normalized intensity of gene $i$ in sample $j$, $C$ is a user-defined parameter, $\sigma^{ne}$ is the standard deviation of the non-expressed distribution and $\mu^{ne}$ is the mean of the non-expressed distribution. The barcode representation of a sample is a vector of ones and zeros denoting which genes are estimated to be expressed (ones) and unexpressed (zeros). The barcode algorithm was implemented by the barcode function in the R fRMA package, and the default value of $C$ was used.

To determine if the expressed ratios differed in the MI group versus the healthy control group, Fisher's exact test for individual genes was carried out upon the barcode values. Genes with a false discovery rate (FDR) of $< 0.05$, which was calculated through the Benjamini-Hochberg (BH) procedure to adjust for multiple testing issue, were considered as differentially expressed genes.

## Functional and pathway enrichment analyses

Using the R clusterProfilter package, the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation were carried out on the differentially expressed genes. The GO analysis included biological process (BP), cellular component (CC) and molecular function (MF) categories. An adjusted P < 0.05 was considered to indicate a statistically significant difference. In these analyses, all default parameters were used.

## Identification of hub genes

The online Search Tool for the Retrieval of Interacting Genes (STRING) database [(http://string-db.org/)](http://string-db.org/) [50] was used to obtain the predicted interactions for the DEGs. The protein–protein interaction (PPI) network of the DEGs was visualized with Cytoscape software (Version 3.9.1, [http://www.cytoscape.org/](http://www.cytoscape.org/)). The CytoHubba plugin in Cytoscape features 12 different algorithms to analyse PPI network topology: ?? Maximal Clique Centrality (MCC), Density of Maximum Neighborhood Component (DMNC), Maximum Neighborhood Component (MNC), Degree, Component (EPC), BottleNeck, EcCentricity, Closeness, Radiality, Betweenness, Stress and ClusteringCoefficient [51]. The outputs of these algorithms can be integrated to identify hub genes.??

## Machine Learning

The machine learning analysis was performed using Python software, ver. 3.9, numpy [@numpy], pandas [@pandas] and Scikit-Learn packages [@scikit-learn]. In all ML analysis, the train datasets where divided into a train and a development set by 0.7:0.3 ratio.

A two layer architecture has been deployed to the data in order to maximize the prediction values. The first layer will predict whether a sample is healthy or not, and the second layer will separates MI from CAD in the samples which were predicted as not healthy in the first layer. To this end, a distinct ML model was trained for each layer.

### First layer: seperating healthy and not healthy samples

Different SVM models using linear, polynomial, and RBF kernels were trained using single miRNAs and different combinations of them. The receiver operating characteristic (ROC) curve were generated for them and area under curves (AUC) were calculated. The best combination were selected based on AUC, accuracy and ??.

### Second layer: seperating MI and CAD samples

In order to investigate, whether a single miRNAs could predict the presence of MI with good sensitivity and specificity or not, a simple SVM with linear kernel model has been trained using each differentially expressed miRNA and the ROC curve were generated. Moreover, we tested combination of miRNAs to evaluate their ability to improve the models predictive values. The miRNA combination with the highest AUC has been used to train different algorithms for the sake of finding the model with the best diagnosing capability. To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Multi-layer Perceptron (MLP), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) has been trained using the expression profile of the best combination miRNAs. All models were trained with their pre-set parameters with 10-fold cross-validation.

Criteria for choosing the best model was the highest accuracy on development set between the models with train accuracy of $>0.95$. The best model was hypertuned with scikit-opt package [@scikitopt] to get the best predictive performance. After finding hypertunning the best model based on the train and development sets, the model performance has been evaluated on the two unseen test datasets (GSE62646 and GSE54475).

# Results and Discussion

## Pre-processing

The PCA plot of the samples were shown in fig\@ref(fig:PCA). As it is clear, there is a complete separation between healthy samples and CAD and MI samples in primary data. Moreover,  there is a relative separation between samples with the same disease status in different datasets, which could be considered as a sign of batch effect [@BER]. After pre-processing data with fRMA algorithm, the separation between healthy samples and CAD and MI samples still remain, but the separation between samples with the same disease status has been removed, as a result of batch effect removal.

The RLE plot presented in fig\@ref(fig:PCA) also validates batch effect removal. For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in RLE plot, and inter-quantile distances would be grater than 0.1 [@BER]. The mentioned criteria is not met in primary data, but has been met after conducting fRMA algorithm.

```{r PCA, echo=F, out.width="100%", fig.cap="PCA and RLE plot for all samples before and after fRMA."}
knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/PCA and RLE.png")
```

## Differential expression analysis
```{r , echo=F}
DEGs <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Train/MI/DEGs MI-Healthy train.txt", check.names = F)
```

According to the cutoff criterion of $FDR < 0.05$, there are `r dim(DEGs)[1]-1` DEGs between the MI patients and the healthy controls. Among them, `r sum(DEGs$V4 == 'TRUE')` are over-expressed in MI, and `r sum(DEGs$V4 == 'FALSE')` are down-expressed in MI compared to the healthy controls.

## Gene ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) enrichment analyses of the DEGs.

To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses. Many biological functions enriched with the DEGs were associated with the immune cells, as expected. GO enrichment analysis in the cellular component (CC) category suggested that the robust DEGs were enriched in “secretory granule membrane” and “sazurophil granule” (fig\@ref(fig:enrich)A). In the biological process (BP) category, the robust DEGs were enriched in “mononuclear cell differentiation”, “lymphocyte differentiation”, "leukocyte activation involved in immune response", and "lymphocyte activation involved in immune response” (fig\@ref(fig:enrich)B). In the molecular function (MF) category, the robust DEGs were involved in “cadherin binding” and “MHC class I protein binding” (fig\@ref(fig:enrich)C). KEGG pathway analysis indicated that the robust DEGs were related to the following pathways: “Chemokine signaling pathway”, “Lipid and atherosclerosis”, “Hematopoietic cell lineage” and “Chronic myeloid leukemia” (fig\@ref(fig:enrich)D). The above results suggested that the abnormal expression of the DEGs may ??.

```{r enrich, echo=F, out.width="100%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the DEGs. (A) Biological process terms enriched with the DEGs. (B) Cellular component terms enriched with the DEGs. (C) Molecular function terms enriched with the DEGs. (D) KEGG analysis of the DEGs. The respective pathway involved in MI was identified by using the KEGG pathway database."}
knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Train/MI/enrichment.pdf")
```

## Machine Learning

Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 are differntially expressed miRNAs. The expression profile of these three miRNAs are presented in fig\@ref(fig:expression). The ROC curves of each miRNA is presented in

```{r expression, echo=F, out.width="100%", fig.cap="Barplot of differentially expressed miRNAs expression values."}
knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Expressionforall.pdf")
```

# Refrences

### 2022-12-11 ArticleDraft.Rmd
changing pca and rle plots

@@
 This way fRMA can be seen as a batch effect removal technique for different datasets produced using identical microarray platform.
-Thus, In order to ensure about batch effect removal, the principal component analysis and the relative log expression of all samples were plotted before and after fRMA [@BER].
+Thus, In order to ensure about batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
 
@@
 
-The PCA plot of the samples were shown in fig\@ref(fig:PCA).
+The PCA plot of the train samples were shown in fig\@ref(fig:PCA).
 As it is clear, there is a complete separation between healthy samples and CAD and MI samples in primary data.
@@
 ```{r PCA, echo=F, out.width="100%", fig.cap="PCA and RLE plot for all samples before and after fRMA."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/PCA and RLE.png")
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/PCA and RLE of trains.pdf")
 ```

### 2023-02-07 ArticleDraft.Rmd
After a long time

@@
 
-``` {r , echo=F}
+```{r , echo=F}
 # What to do next:
@@
 
-Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database ([https://www.ncbi.nlm.nih.gov/geo/](https://www.ncbi.nlm.nih.gov/geo/)).
-In order to come by sufficient classification power for MI samples and others, the sample size of the dataset should be relatively large.
-Therefore, GSE59867 and GSE56609 were selected as train set and GSE62646 and GSE54475 as validation set for both DEGs and pathway analyse as well as machine learning section.
+Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database (<https://www.ncbi.nlm.nih.gov/geo/>).
+In order to come by sufficient classification power between MI samples and others, the sample size of the dataset should be relatively large.
+Therefore, GSE59867 for MI and CAD samples, and GSE56609, GSE54475, and GSE23832 for healthy samples were selected.
 All datasets have been conducted using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
@@
 |          | Platform | Healthy Control | CAD Control | MI  | Reference |
-|----------------|----------|-----------------|-------------|-----|-----------|
-| Training Set   |          |                 |             |     |           |
+|----------|----------|-----------------|-------------|-----|-----------|
 | GSE59867 | GPL6244  | \-              | 46          | 111 | [@67]     |
 | GSE56609 | GPL6244  | 46              | \-          | \-  | [@09]     |
-| Test Set |          |                 |             |     |           |
-| GSE62646       | GPL6244  | \-              | 14          | 28  | [@46]     |
 | GSE54475 | GPL6244  | 5               | \-          | \-  | [@75]     |
 
-: Basic information of the 4 GEO microarray datasets.
+: Basic information of the 3 GEO microarray datasets.
 
@@
 fRMA allows to preprocess individual microarray samples and combine them consistently for analysis.
-For each dataset, background correction is performed and then it is quantile normalized based on the reference distribution.
+For each dataset, background correction is performed using RMA algorithm and then it is quantile normalized based on the reference distribution.
 During summarization, batch effects are removed and variances of the gene expressions are estimated by taking into account these probe-specific effects.
@@
 
+Also, the same procedure were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs in between them.
+
 ## Functional and pathway enrichment analyses
@@
 The GO analysis included biological process (BP), cellular component (CC) and molecular function (MF) categories.
-An adjusted P < 0.05 was considered to indicate a statistically significant difference.
+An adjusted P \< 0.05 was considered to indicate a statistically significant difference.
+Enrichment were counducted on the MI-healthy and CAD-healthy DEGs.
 In these analyses, all default parameters were used.
-
-## Identification of hub genes
-
-The online Search Tool for the Retrieval of Interacting Genes (STRING) database [(http://string-db.org/)](http://string-db.org/) [50] was used to obtain the predicted interactions for the DEGs.
-The protein–protein interaction (PPI) network of the DEGs was visualized with Cytoscape software (Version 3.9.1, [http://www.cytoscape.org/](http://www.cytoscape.org/)).
-The CytoHubba plugin in Cytoscape features 12 different algorithms to analyse PPI network topology: ??
-Maximal Clique Centrality (MCC), Density of Maximum Neighborhood Component (DMNC), Maximum Neighborhood Component (MNC), Degree, Component (EPC), BottleNeck, EcCentricity, Closeness, Radiality, Betweenness, Stress and ClusteringCoefficient [51].
-The outputs of these algorithms can be integrated to identify hub genes.??
 
@@
 3.9, numpy [@numpy], pandas [@pandas] and Scikit-Learn packages [@scikit-learn].
-In all ML analysis, the train datasets where divided into a train and a development set by 0.7:0.3 ratio.
+In all ML analysis, the datasets where divided into train and test set by 0.7:0.3 ratio and the reported results are the average of a 10-fold cross-validation.
 
-A two layer architecture has been deployed to the data in order to maximize the prediction values.
+In the first step, different models has been trained on all microRNAs in DEGs.
+Moreover, since the number of microRNAs in DEGs are limited, in order to reaching the highest predicting value, different models also trained on microRNAs selected by their individual AUC-ROC.
+These approach can provide an informative comparison between the predictive capabilities of set of microRNAs selected with two different above-mentioned approaches.
+
+### microRNAs in DEGs
+
+In this approach a two layer architecture has been deployed to the data in order to maximize the prediction values.
 The first layer will predict whether a sample is healthy or not, and the second layer will separates MI from CAD in the samples which were predicted as not healthy in the first layer.
 To this end, a distinct ML model was trained for each layer.
+Since there are limited numbers of microRNAs in DEGs both layers where trained with all microRNAs available.
+For both layers the receiver operating characteristic (ROC) curve were generated and area under curves (AUC) were calculated for each microRNAs in DEGs for further comparison with the models performance.
 
-### First layer: seperating healthy and not healthy samples
+#### First layer for seperating healthy and not-healthy samples:
 
-Different SVM models using linear, polynomial, and RBF kernels were trained using single miRNAs and different combinations of them.
-The receiver operating characteristic (ROC) curve were generated for them and area under curves (AUC) were calculated.
-The best combination were selected based on AUC, accuracy and ??.
+An SVM models using RBF kernels were hyper-tunned and trained using all microRNAs in DEGs.
+In order to handle the imbalance number of samples in groups (51 for healthy and 157 for not-healthy group), sample weight for not-healthy samples were set to 0.5.
 
-### Second layer: seperating MI and CAD samples
+#### Second layer for seperating MI and CAD samples:
 
-In order to investigate, whether a single miRNAs could predict the presence of MI with good sensitivity and specificity or not, a simple SVM with linear kernel model has been trained using each differentially expressed miRNA and the ROC curve were generated.
-Moreover, we tested combination of miRNAs to evaluate their ability to improve the models predictive values.
-The miRNA combination with the highest AUC has been used to train different algorithms for the sake of finding the model with the best diagnosing capability.
-To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Multi-layer Perceptron (MLP), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) has been trained using the expression profile of the best combination miRNAs.
+For the sake of reaching the highest classification performance using the set of three microRNAs different models were investigated.
+To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) has been trained using the expression profile of the miRNAs.
 All models were trained with their pre-set parameters with 10-fold cross-validation.
 
-Criteria for choosing the best model was the highest accuracy on development set between the models with train accuracy of $>0.95$.
-The best model was hypertuned with scikit-opt package [@scikitopt] to get the best predictive performance.
-After finding hypertunning the best model based on the train and development sets, the model performance has been evaluated on the two unseen test datasets (GSE62646 and GSE54475).
+Criteria for choosing the best model was the highest accuracy and AUC on the test set. The best model was hyper-tuned with scikit-opt package [@scikitopt] to get the best predictive performance.
+
+### microRNAs with the highest AUC-ROC
+
+Like the previous approach, two different models will be trained.
+The first layer for classifying samples to healthy and not-healthy and the second layer for separating MI and CAD samples.
+However, for keeping the number of microRNAs as low as possible microRNAs will be selected from the second layer (which are the microRNAs with the best performance in MI/CAD separation) and then thier performance will be evaluated in the first layer.
+
+#### Second layer for separating MI and CAD:
+
+AUC-ROC of all microRNAs for classifying MI and CAD samples has been calculated.
+For finding the number of microRNAs with the highest predictive values, the microRNAs with the highest individual AUC-ROC is adding to the set one-by-one and the AUC-ROC for the set is calculated.
+The set with the highest AUC-ROC has been selected as the set for the following steps.
+
+The selected number of microRNAs has been used to train different algorithms for the sake of finding the best model.
+As previous approach, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) has been trained using the expression profile of the selected miRNAs set.
+All models were trained with their pre-set parameters with 10-fold cross-validation.
+The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned with scikit-opt package [@scikitopt].
+
+#### First layer for finding healthy and not-healthy samples:
+
+The set of microRNAs selected for differentiating MI and CAD samples also used to diagnose not-healthy samples and the ROC curve were ploted for each microRNA.
+Moreover, expression values of the microRNAs in the set were used to train an SVM model with RBF kernel and its ROc curve were plotted also.
 
@@
 The PCA plot of the train samples were shown in fig\@ref(fig:PCA).
-As it is clear, there is a complete separation between healthy samples and CAD and MI samples in primary data.
-Moreover,  there is a relative separation between samples with the same disease status in different datasets, which could be considered as a sign of batch effect [@BER].
-After pre-processing data with fRMA algorithm, the separation between healthy samples and CAD and MI samples still remain, but the separation between samples with the same disease status has been removed, as a result of batch effect removal.
+As it is clear, there is a complete separation between healthy samples and CAD and MI samples in primary data which also remains after conducting fRMA on the data.
 
-The RLE plot presented in fig\@ref(fig:PCA) also validates batch effect removal.
+The RLE plot presented in fig\@ref(fig:PCA) validates batch effect removal.
 For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in RLE plot, and inter-quantile distances would be grater than 0.1 [@BER].
@@
 ```{r PCA, echo=F, out.width="100%", fig.cap="PCA and RLE plot for all samples before and after fRMA."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/PCA and RLE of trains.pdf")
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/PCA and RLE of trains.pdf")
 ```
@@
 ## Differential expression analysis
+
 ```{r , echo=F}
-DEGs <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Train/MI/DEGs MI-Healthy train.txt", check.names = F)
+M_H <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-Healthy.txt", check.names = F)
+
+M_C <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-CAD.txt", check.names = F)
+
+C_H <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs CAD-Healthy.txt", check.names = F)
 ```
 
-According to the cutoff criterion of $FDR < 0.05$, there are `r dim(DEGs)[1]-1` DEGs between the MI patients and the healthy controls.
-Among them, `r sum(DEGs$V4 == 'TRUE')` are over-expressed in MI, and `r sum(DEGs$V4 == 'FALSE')` are down-expressed in MI compared to the healthy controls.
+According to the cutoff criterion of $FDR < 0.05$, there are `r dim(M_H)[1]` DEGs between the MI patients and the healthy controls.
+Among them, `r sum(M_H$logFC > 0)` are up-regulated in MI, and `r sum(M_H$logFC < 0)` are down-regulated in MI compared to the healthy controls.
+For CAD and healthy groups there are `r dim(C_H)[1]` DEGs, `r sum(C_H$logFC > 0)` of them were up-regulated and `r sum(C_H$logFC < 0)` of them were down-regulated in CAD samples in comparison with healthy samples.
+For MI and CAD group the number of DEGs are `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples in comparison with CAD samples are `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively.
+All these data is summarized in table 2.
+
+|                 | No. of DEGs     | no. of up-regulated DEGs | no. of down-regulated DEGs | microRNAs                           |
+|---------------|---------------|---------------|---------------|---------------|
+| MI vs. Healthy  | `r dim(M_H)[1]` | `r sum(M_H$logFC > 0)`   | `r sum(M_H$logFC < 0)`     | hsa-miR-186, hsa-miR-21, hsa-miR-32 |
+| CAD vs. Healthy | `r dim(C_H)[1]` | `r sum(C_H$logFC > 0)`   | `r sum(C_H$logFC < 0)`     | hsa-miR-186, hsa-miR-21, hsa-miR-32 |
+| MI vs. CAD      | `r dim(M_C)[1]` | `r sum(M_C$logFC > 0)`   | `r sum(M_C$logFC < 0)`     | hsa-miR-186                         |
+
+: Total DEGs and name of microRNAs in DEGs.
 
@@
 
-To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses.
-Many biological functions enriched with the DEGs were associated with the immune cells, as expected.
-GO enrichment analysis in the cellular component (CC) category suggested that the robust DEGs were enriched in “secretory granule membrane” and “sazurophil granule” (fig\@ref(fig:enrich)A).
-In the biological process (BP) category, the robust DEGs were enriched in “mononuclear cell differentiation”, “lymphocyte differentiation”, "leukocyte activation involved in immune response", and "lymphocyte activation involved in immune response” (fig\@ref(fig:enrich)B).
-In the molecular function (MF) category, the robust DEGs were involved in “cadherin binding” and “MHC class I protein binding” (fig\@ref(fig:enrich)C).
-KEGG pathway analysis indicated that the robust DEGs were related to the following pathways: “Chemokine signaling pathway”, “Lipid and atherosclerosis”, “Hematopoietic cell lineage” and “Chronic myeloid leukemia” (fig\@ref(fig:enrich)D).
-The above results suggested that the abnormal expression of the DEGs may ??.
+To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
+For both set of DEGs, many biological functions enriched were associated with the immune cells, as expected.
 
-```{r enrich, echo=F, out.width="100%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the DEGs.
-(A) Biological process terms enriched with the DEGs.
-(B) Cellular component terms enriched with the DEGs.
-(C) Molecular function terms enriched with the DEGs.
-(D) KEGG analysis of the DEGs.
-The respective pathway involved in MI was identified by using the KEGG pathway database."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Train/MI/enrichment.pdf")
+For MI versus healthy samples, GO enrichment analysis in the biological process (BP) category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (fig\@ref(fig:MIHEnrich)A).
+In the cellular component (CC) category the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (fig\@ref(fig:MIHEnrich)B).
+In the molecular function (MF) category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (fig\@ref(fig:MIHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (fig\@ref(fig:MIHEnrich)D).
+
+```{r MIHEnrich, echo=F, out.width="95%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs.
+(A) Biological process terms.
+(B) Cellular component terms.
+(C) Molecular function terms.
+(D) KEGG analysis."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/enrich MI-Healthy.pdf")
+```
+
+The enrichment results for DEGs of CAD versus healthy samples are as follows.
+In the biological process (BP) category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (fig\@ref(fig:CADHEnrich)A).
+In the cellular component (CC) category the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (fig\@ref(fig:CADHEnrich)B).
+In the molecular function (MF) category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (fig\@ref(fig:CADHEnrich)C).
+KEGG pathway analysis indicated that the  DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (fig\@ref(fig:CADHEnrich)D).
+
+```{r CADHEnrich, echo=F, out.width="95%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs.
+(A) Biological process terms.
+(B) Cellular component terms.
+(C) Molecular function terms.
+(D) KEGG analysis."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/enrich CAD-Healthy.pdf")
 ```
@@
 
+### microRNAs in DEGs
+
 Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 are differntially expressed miRNAs.
-The expression profile of these three miRNAs are presented in fig\@ref(fig:expression).
-The ROC curves of each miRNA is presented in
+The expression profile of these three miRNAs are presented in fig\@ref(fig:DEMexp).
+The ROC curves of each miRNA for each layer is presented in fig\@ref(fig:miRROC).
+As it is clear from the fig\@ref(fig:miRROC)A, all these three microRNAs has an acceptable performance fro separating healthy and not-healthy samples.
+AUC for hsa-miR-21, hsa-miR-32, and hsa-miR-186 is 0.99, 1, and 0.91 respectively.
+Also the accuracy of each microRNA for classifying the samples to healthy and not-healthy groups on test set using a simple logistic regression model is 0.92, 0.98, and 0.89 for hsa-miR-21, hsa-miR-32, and hsa-miR-186 respectively.
+Moreover, In fig\@ref(fig:miRROC)B the ROC curve of each microRNA for classifying MI and CAD samples were presented.
+For  hsa-miR-21, hsa-miR-32, and hsa-miR-186 AUC on test set is 0.85; 0.7; and 0.82, and accuracy on test set is 0.78; 0.67; and 0.74.
 
-```{r expression, echo=F, out.width="100%", fig.cap="Barplot of differentially expressed miRNAs expression values."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Expressionforall.pdf")
+```{r DEMexp, echo=F, out.width="50%", fig.align='center', fig.cap="Expression profile of hsa-miR-186, hsa-miR-21, and hsa-miR32 in Healthy, CAD, and MI samples."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEGs healthy not-healthy/DEMs Expression.pdf")
+```
+
+```{r miRROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for single microRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/miRs ROCs.pdf")
+```
+
+#### First layer for healthy not-healthy seperation:
+
+Although single microRNAs have acceptable performance, but their prediction value can improve even more using them as a set.
+The ROC curve for the SVM model with rbf kernel trained with all three microRNAs is presented in fig\@ref(fig:DEMROC)A.
+The model has better performance in classification than single microRNAs.The AUC for the model is 1, and its accuracy on test set is 1 as well.
+
+```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for microRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs ROCs h not h cad mi.pdf")
+```
+
+#### Second layer for separating MI samples from CAD:
+
+Different models has been trained using expression values of three differentially microRNAs.
+The models 10-fold cross-validate AUC and accuracy on the test set is reported on fig\@ref(fig:DEMmodels).
+The best model from both AUC and accuracy point-of-view is an SVM model with linear kernel.
+The AUC and accuracy for this model with its preset values is 0.93 and 0.82 respectively.
+The model is hyper-tuned for C and gamma hyperparameters, and therefore the model showed a better performance.
+The ROC curve of the hyper-tuned model as the best model is presented in fig\@ref(fig:DEMROC)B.
+For this model the AUC reached to 0.95 and the accuracy improved to 0.85.
+Moreover, the sensitivity and specificity for the model is 0.91 and 0.71 respectively.
+
+```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) and accuracy of different models trained with three microRNAs in DEGs on the test set."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models DEGs mirs.pdf")
+```
+
+### AUC approach
+
+#### Second layer; MI form CAD:
+
+After calculating AUC for each microRNA, they are sorted and their performance as a set were investigated.
+The metric for finding the best set was AUC.
+The diagram for AUC against the number of microRNAs in the set is presented in fig\@ref(fig:AUCset).
+As it is clear, the AUC increase until  the number of microRNAs in the set reaches tho 6 and after that it drops.
+The AUC for separating MI samples from CAD using these microRNAs is 0.93.
+The microRNAs in the set are has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1.
+The expression values of these microRNAs in healthy, CAD, and MI samples is presented in  fig\@ref(fig:AUCexp).
+
+```{r AUCset, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) for sets containing increasing number of microRNAs with the highest individual AUC in MI/CAD separation."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC for mirs.pdf")
+```
+
+```{r AUCexp, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/ExpressionforAUC.pdf")
+```
+
+For finding the most model for training the best set, different models were trained using their pre-set values.
+Their AUC and accuracy results on the test set is presented in fig\@ref(fig:AUCmodels).
+The best model from AUC point of view is the LR and from accuracy point-of-view is the SVM model with polynomial kernel.
+For LR model the AUC and accuracy were 0.92 and 0.81; and for SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
+Both models were hyper-tuned and ROC curve for their best performance presented in fig\@ref(fig:AUCROC)A and B.
+The AUC and accuracy for LR model increased to 0.94 and 0.88 respectively; And for SVM model with polynomial kernel these values increased to 0.95 and 0.88 respectively.
+The sensitivity for LR and SVM models are 1 and 0.97; and the specificity for them is 0.57 and 0.64, respectively.
+
+```{r AUCmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) and accuracy of different models trained with three microRNAs in DEGs on the test set."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models AUC mirs.pdf")
+```
+
+```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of microRNAs selected by AUC on test set classification.
+(A) Logistic regression model for CAD and MI samples classification.
+(B) SVM with polynomial kernel for CAD and MI samples classification.
+(C) SVM with RBF kernel for healthy and not-healthy samples classification."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC best model ROCs.pdf")
 ```

### 2023-02-07 ArticleDraft.Rmd
material/methods and results complete without check

@@
 
-The PCA plot of the train samples were shown in fig\@ref(fig:PCA).
+The PCA plot of the train samples were shown in Fig\@ref(fig:PCA).
 As it is clear, there is a complete separation between healthy samples and CAD and MI samples in primary data which also remains after conducting fRMA on the data.
 
-The RLE plot presented in fig\@ref(fig:PCA) validates batch effect removal.
+The RLE plot presented in Fig\@ref(fig:PCA) validates batch effect removal.
 For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in RLE plot, and inter-quantile distances would be grater than 0.1 [@BER].
@@
 
-For MI versus healthy samples, GO enrichment analysis in the biological process (BP) category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (fig\@ref(fig:MIHEnrich)A).
-In the cellular component (CC) category the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (fig\@ref(fig:MIHEnrich)B).
-In the molecular function (MF) category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (fig\@ref(fig:MIHEnrich)C).
-KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (fig\@ref(fig:MIHEnrich)D).
+For MI versus healthy samples, GO enrichment analysis in the biological process (BP) category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Fig\@ref(fig:MIHEnrich)A).
+In the cellular component (CC) category the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:MIHEnrich)B).
+In the molecular function (MF) category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Fig\@ref(fig:MIHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:MIHEnrich)D).
 
@@
 The enrichment results for DEGs of CAD versus healthy samples are as follows.
-In the biological process (BP) category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (fig\@ref(fig:CADHEnrich)A).
-In the cellular component (CC) category the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (fig\@ref(fig:CADHEnrich)B).
-In the molecular function (MF) category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (fig\@ref(fig:CADHEnrich)C).
-KEGG pathway analysis indicated that the  DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (fig\@ref(fig:CADHEnrich)D).
+In the biological process (BP) category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Fig\@ref(fig:CADHEnrich)A).
+In the cellular component (CC) category the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:CADHEnrich)B).
+In the molecular function (MF) category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Fig\@ref(fig:CADHEnrich)C).
+KEGG pathway analysis indicated that the  DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:CADHEnrich)D).
 
@@
 Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 are differntially expressed miRNAs.
-The expression profile of these three miRNAs are presented in fig\@ref(fig:DEMexp).
-The ROC curves of each miRNA for each layer is presented in fig\@ref(fig:miRROC).
-As it is clear from the fig\@ref(fig:miRROC)A, all these three microRNAs has an acceptable performance fro separating healthy and not-healthy samples.
+The expression profile of these three miRNAs are presented in Fig\@ref(fig:DEMexp).
+The ROC curves of each miRNA for each layer is presented in Fig\@ref(fig:miRROC).
+As it is clear from the Fig\@ref(fig:miRROC)A, all these three microRNAs has an acceptable performance fro separating healthy and not-healthy samples.
 AUC for hsa-miR-21, hsa-miR-32, and hsa-miR-186 is 0.99, 1, and 0.91 respectively.
 Also the accuracy of each microRNA for classifying the samples to healthy and not-healthy groups on test set using a simple logistic regression model is 0.92, 0.98, and 0.89 for hsa-miR-21, hsa-miR-32, and hsa-miR-186 respectively.
-Moreover, In fig\@ref(fig:miRROC)B the ROC curve of each microRNA for classifying MI and CAD samples were presented.
+Moreover, In Fig\@ref(fig:miRROC)B the ROC curve of each microRNA for classifying MI and CAD samples were presented.
 For  hsa-miR-21, hsa-miR-32, and hsa-miR-186 AUC on test set is 0.85; 0.7; and 0.82, and accuracy on test set is 0.78; 0.67; and 0.74.
@@
 Although single microRNAs have acceptable performance, but their prediction value can improve even more using them as a set.
-The ROC curve for the SVM model with rbf kernel trained with all three microRNAs is presented in fig\@ref(fig:DEMROC)A.
+The ROC curve for the SVM model with rbf kernel trained with all three microRNAs is presented in Fig\@ref(fig:DEMROC)A.
 The model has better performance in classification than single microRNAs.The AUC for the model is 1, and its accuracy on test set is 1 as well.
+The confusion matrix for the model is presented in Fig\@ref(fig:DEMCF)A.
 
@@
 
+```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy samples classification.
+and (B) An SVM model with linear kernel for CAD and MI samples classification."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs Conf. matrix.pdf")
+```
+
 #### Second layer for separating MI samples from CAD:
@@
 Different models has been trained using expression values of three differentially microRNAs.
-The models 10-fold cross-validate AUC and accuracy on the test set is reported on fig\@ref(fig:DEMmodels).
+The models 10-fold cross-validate AUC and accuracy on the test set is reported on Fig\@ref(fig:DEMmodels).
 The best model from both AUC and accuracy point-of-view is an SVM model with linear kernel.
@@
 The model is hyper-tuned for C and gamma hyperparameters, and therefore the model showed a better performance.
-The ROC curve of the hyper-tuned model as the best model is presented in fig\@ref(fig:DEMROC)B.
+The ROC curve of the hyper-tuned model as the best model is presented in Fig\@ref(fig:DEMROC)B.
 For this model the AUC reached to 0.95 and the accuracy improved to 0.85.
-Moreover, the sensitivity and specificity for the model is 0.91 and 0.71 respectively.
+Moreover, the sensitivity and specificity for the model on the test set is 0.91 and 0.71 respectively.
+The confusion matrix for the hyper-tuned model is illustrated in Fig\@ref(fig:DEMCF)B.
 
@@
 The metric for finding the best set was AUC.
-The diagram for AUC against the number of microRNAs in the set is presented in fig\@ref(fig:AUCset).
+The diagram for AUC against the number of microRNAs in the set is presented in Fig\@ref(fig:AUCset).
 As it is clear, the AUC increase until  the number of microRNAs in the set reaches tho 6 and after that it drops.
@@
 The microRNAs in the set are has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1.
-The expression values of these microRNAs in healthy, CAD, and MI samples is presented in  fig\@ref(fig:AUCexp).
+The expression values of these microRNAs in healthy, CAD, and MI samples is presented in  Fig\@ref(fig:AUCexp).
+The ROC curve of the selected microRNAs for MI and CAD samples classification is illustrated in Fig\@ref(fig:miRROC)B.
 
 ```{r AUCset, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) for sets containing increasing number of microRNAs with the highest individual AUC in MI/CAD separation."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC for mirs.pdf")
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC for cad mi mirs.pdf")
 ```
@@
 For finding the most model for training the best set, different models were trained using their pre-set values.
-Their AUC and accuracy results on the test set is presented in fig\@ref(fig:AUCmodels).
+Their AUC and accuracy results on the test set is presented in Fig\@ref(fig:AUCmodels).
 The best model from AUC point of view is the LR and from accuracy point-of-view is the SVM model with polynomial kernel.
 For LR model the AUC and accuracy were 0.92 and 0.81; and for SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
-Both models were hyper-tuned and ROC curve for their best performance presented in fig\@ref(fig:AUCROC)A and B.
+Both models were hyper-tuned and ROC curve for their best performance presented in Fig\@ref(fig:AUCROC)A and B.
 The AUC and accuracy for LR model increased to 0.94 and 0.88 respectively; And for SVM model with polynomial kernel these values increased to 0.95 and 0.88 respectively.
 The sensitivity for LR and SVM models are 1 and 0.97; and the specificity for them is 0.57 and 0.64, respectively.
+The confusiom matrix for both models is illustrated in Fig\@ref(fig:AUCCF)A and B.
+
+```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) Logistic regression model for CAD and MI samples classification.
+(B) SVM with polynomial kernel for CAD and MI samples classification.
+(C) SVM with RBF kernel for healthy and not-healthy samples classification."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC Conf. matrix.pdf")
+```
 
@@
 
+#### First layer:
+
+Using the microRNAs set selected for classification of MI and CAD samples, an SVM model with RBF kernel has been trained for separating healthy from not-healthy samples.
+The ROC curve for the model is presented in Fig\@ref(fig:AUCROC)C and confusion matrix is illustrated in Fig\@ref(fig:AUCCF)C.
+Both AUC and accuracy for the model on the test set is equal to 1.
+
 # Refrences

### 2023-02-08 ArticleDraft.Rmd
Correcting MM and R verb tenses and story coherency

@@
 Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database (<https://www.ncbi.nlm.nih.gov/geo/>).
-In order to come by sufficient classification power between MI samples and others, the sample size of the dataset should be relatively large.
-Therefore, GSE59867 for MI and CAD samples, and GSE56609, GSE54475, and GSE23832 for healthy samples were selected.
-All datasets have been conducted using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
+In order to come by sufficient classification power between MI samples and others, the sample size should be relatively large.
+Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475 for healthy samples were selected.
+All samples were produced using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
 Only healthy, stable CAD and early stage MI samples were selected from these datasets for further analysis.
-The basic information for the four GEO datasets evaluated in the current study is provided in Table 1.
+The basic information for the three datasets evaluated in the current study is provided in Table 1.
 
@@
 
-: Basic information of the 3 GEO microarray datasets.
+: Basic information of the GEO microarray datasets.
 
@@
 
-Raw data (CEL files) of the four datasets were downloaded from the GEO and preprocessed using the fRMA package [@fRMA].
+Raw data (CEL files) of all datasets were downloaded from the GEO and pre-processed using the fRMA package [@fRMA].
 fRMA allows to preprocess individual microarray samples and combine them consistently for analysis.
-For each dataset, background correction is performed using RMA algorithm and then it is quantile normalized based on the reference distribution.
-During summarization, batch effects are removed and variances of the gene expressions are estimated by taking into account these probe-specific effects.
+For each dataset, background correction was performed using RMA algorithm and then it was quantile normalized based on the reference distribution.
+During summarization, batch effects were removed and variances of the gene expressions were estimated by taking into account these probe-specific effects.
 For those multiple probe sets matched to the identical gene, the mean log fold change was retained.
-This way fRMA can be seen as a batch effect removal technique for different datasets produced using identical microarray platform.
-Thus, In order to ensure about batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
+This way fRMA can be seen as a batch effect removal technique for different datasets that were produced identical microarray platform.
+Thus, to become ensure about batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
 
@@
 \right.
-$$ where $x_{ij}$ is the normalized intensity of gene $i$ in sample $j$, $C$ is a user-defined parameter, $\sigma^{ne}$ is the standard deviation of the non-expressed distribution and $\mu^{ne}$ is the mean of the non-expressed distribution.
+$$
+
+where $x_{ij}$ is the normalized intensity of gene $i$ in sample $j$, $C$ is a user-defined parameter, $\sigma^{ne}$ is the standard deviation of the non-expressed distribution and $\mu^{ne}$ is the mean of the non-expressed distribution.
 The barcode representation of a sample is a vector of ones and zeros denoting which genes are estimated to be expressed (ones) and unexpressed (zeros).
@@
 
-Also, the same procedure were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs in between them.
+The same procedure were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs in between them.
 
@@
 An adjusted P \< 0.05 was considered to indicate a statistically significant difference.
-Enrichment were counducted on the MI-healthy and CAD-healthy DEGs.
+Enrichment were conducted on the MI-healthy and CAD-healthy DEGs.
 In these analyses, all default parameters were used.
@@
 3.9, numpy [@numpy], pandas [@pandas] and Scikit-Learn packages [@scikit-learn].
-In all ML analysis, the datasets where divided into train and test set by 0.7:0.3 ratio and the reported results are the average of a 10-fold cross-validation.
+When ever hyper-tuning was needed the scikit-opt package [@scikitopt] was used.
+In all ML analysis, the datasets were divided into train and test set by 0.7:0.3 ratio and all reported results are the average of a 10-fold cross-validation.
 
-In the first step, different models has been trained on all microRNAs in DEGs.
-Moreover, since the number of microRNAs in DEGs are limited, in order to reaching the highest predicting value, different models also trained on microRNAs selected by their individual AUC-ROC.
-These approach can provide an informative comparison between the predictive capabilities of set of microRNAs selected with two different above-mentioned approaches.
+Two different approaches were used for selecting microRNAs to train models with.
+The first approach was using the microRNAs that are differentially expressed.
+In the second approach microRNAs were selected by their individual AUC-ROC.
+Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of set of microRNAs selected with different logic.
 
@@
 
-In this approach a two layer architecture has been deployed to the data in order to maximize the prediction values.
-The first layer will predict whether a sample is healthy or not, and the second layer will separates MI from CAD in the samples which were predicted as not healthy in the first layer.
+In this approach a two layer architecture was deployed to the data in order to maximize the prediction values.
+The first layer predicted whether a sample is healthy or not, and the second layer was separated MI from CAD in the samples which were predicted as not healthy in the first layer.
 To this end, a distinct ML model was trained for each layer.
-Since there are limited numbers of microRNAs in DEGs both layers where trained with all microRNAs available.
-For both layers the receiver operating characteristic (ROC) curve were generated and area under curves (AUC) were calculated for each microRNAs in DEGs for further comparison with the models performance.
+Since there are limited numbers of microRNAs in DEGs both layers where trained with all of them.
+For further comparison with the models performance ROC curve of each microRNA for classifying healthy and not-healthy, as well as CAD and MI was generated using a simple Logistic Regression model.
 
@@
 
-An SVM models using RBF kernels were hyper-tunned and trained using all microRNAs in DEGs.
+An SVM model using RBF kernels were hyper-tunned and trained using all microRNAs in DEGs.
 In order to handle the imbalance number of samples in groups (51 for healthy and 157 for not-healthy group), sample weight for not-healthy samples were set to 0.5.
@@
 
-For the sake of reaching the highest classification performance using the set of three microRNAs different models were investigated.
-To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) has been trained using the expression profile of the miRNAs.
+For the sake of reaching the highest classification performance using the set of microRNAs, different models were investigated.
+To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) models were trained.
 All models were trained with their pre-set parameters with 10-fold cross-validation.
 
-Criteria for choosing the best model was the highest accuracy and AUC on the test set. The best model was hyper-tuned with scikit-opt package [@scikitopt] to get the best predictive performance.
+Criteria for choosing the best model was the highest accuracy and AUC on the test set.
+The best model was hyper-tuned with scikit-opt package [@scikitopt] in order to get the best predictive performance.
 
@@
 
-Like the previous approach, two different models will be trained.
+Like the previous approach, a two layers strategy was conducted.
 The first layer for classifying samples to healthy and not-healthy and the second layer for separating MI and CAD samples.
-However, for keeping the number of microRNAs as low as possible microRNAs will be selected from the second layer (which are the microRNAs with the best performance in MI/CAD separation) and then thier performance will be evaluated in the first layer.
-
-#### Second layer for separating MI and CAD:
+However, for keeping the number of microRNAs as low as possible microRNAs will be selected from the second layer (which are the microRNAs with the best performance in MI/CAD separation) and then their performance were evaluated in the first layer.
 
@@
 The set with the highest AUC-ROC has been selected as the set for the following steps.
-
-The selected number of microRNAs has been used to train different algorithms for the sake of finding the best model.
-As previous approach, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) has been trained using the expression profile of the selected miRNAs set.
-All models were trained with their pre-set parameters with 10-fold cross-validation.
-The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned with scikit-opt package [@scikitopt].
+The ROC curves for each selected microRNA for separating healthy samples from not-healthy and MIs from CADs were also plotted for further comparison.
 
@@
 
-The set of microRNAs selected for differentiating MI and CAD samples also used to diagnose not-healthy samples and the ROC curve were ploted for each microRNA.
-Moreover, expression values of the microRNAs in the set were used to train an SVM model with RBF kernel and its ROc curve were plotted also.
+An SVM model with RBF kernel were trained with the selected set of microRNAs.
+Moreover, the model was hyper-tuned for finding the hyper-parameters with the best AUC and accuracy.
+The ROC curve and confusion matrix for the best model was reported.
+
+#### Second layer for separating MI and CAD:
+
+The selected number of microRNAs was used to train different algorithms for finding the best model.
+As previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB and DT were trained.
+All models were trained with their pre-set parameters with 10-fold cross-validation.
+The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned with scikit-opt package [@scikitopt].
+The ROC curve and confusion matrix for the best model was reported.
 
@@
 The PCA plot of the train samples were shown in Fig\@ref(fig:PCA).
-As it is clear, there is a complete separation between healthy samples and CAD and MI samples in primary data which also remains after conducting fRMA on the data.
+As it is clear, there was a complete separation between healthy samples and CAD and MI samples in primary data which also remains after conducting fRMA on the data.
 
 The RLE plot presented in Fig\@ref(fig:PCA) validates batch effect removal.
-For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in RLE plot, and inter-quantile distances would be grater than 0.1 [@BER].
-The mentioned criteria is not met in primary data, but has been met after conducting fRMA algorithm.
+For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in RLE plot, and inter-quantile distances would be greater than 0.1 [@BER].
+The mentioned criteria was not met in primary data, but was met after conducting fRMA algorithm.
 
@@
 
-According to the cutoff criterion of $FDR < 0.05$, there are `r dim(M_H)[1]` DEGs between the MI patients and the healthy controls.
-Among them, `r sum(M_H$logFC > 0)` are up-regulated in MI, and `r sum(M_H$logFC < 0)` are down-regulated in MI compared to the healthy controls.
-For CAD and healthy groups there are `r dim(C_H)[1]` DEGs, `r sum(C_H$logFC > 0)` of them were up-regulated and `r sum(C_H$logFC < 0)` of them were down-regulated in CAD samples in comparison with healthy samples.
-For MI and CAD group the number of DEGs are `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples in comparison with CAD samples are `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively.
+According to the cutoff criterion of $FDR < 0.05$, there were `r dim(M_H)[1]` DEGs between the MI and the healthy samples.
+Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in MI compared to the healthy controls.
+For CAD and healthy groups there were `r dim(C_H)[1]` DEGs, `r sum(C_H$logFC > 0)` of them were up-regulated and `r sum(C_H$logFC < 0)` of them were down-regulated in CAD samples in comparison with healthy samples.
+For MI and CAD group the number of DEGs are `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples in comparison with CAD samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively.
 All these data is summarized in table 2.
@@
 
-: Total DEGs and name of microRNAs in DEGs.
+: Total DEGs and name of differentially expressed microRNAs.
 
@@
 To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
-For both set of DEGs, many biological functions enriched were associated with the immune cells, as expected.
 
-For MI versus healthy samples, GO enrichment analysis in the biological process (BP) category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Fig\@ref(fig:MIHEnrich)A).
-In the cellular component (CC) category the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:MIHEnrich)B).
-In the molecular function (MF) category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Fig\@ref(fig:MIHEnrich)C).
+For MI versus healthy samples, GO enrichment analysis in the BP category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Fig\@ref(fig:MIHEnrich)A).
+In the CC category the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:MIHEnrich)B).
+In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Fig\@ref(fig:MIHEnrich)C).
 KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:MIHEnrich)D).
@@
 
-The enrichment results for DEGs of CAD versus healthy samples are as follows.
-In the biological process (BP) category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Fig\@ref(fig:CADHEnrich)A).
-In the cellular component (CC) category the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:CADHEnrich)B).
-In the molecular function (MF) category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Fig\@ref(fig:CADHEnrich)C).
+The enrichment results for DEGs of CAD versus healthy samples were as follows.
+In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Fig\@ref(fig:CADHEnrich)A).
+In the CC category the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:CADHEnrich)B).
+In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Fig\@ref(fig:CADHEnrich)C).
 KEGG pathway analysis indicated that the  DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:CADHEnrich)D).
@@
 
-Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 are differntially expressed miRNAs.
-The expression profile of these three miRNAs are presented in Fig\@ref(fig:DEMexp).
-The ROC curves of each miRNA for each layer is presented in Fig\@ref(fig:miRROC).
-As it is clear from the Fig\@ref(fig:miRROC)A, all these three microRNAs has an acceptable performance fro separating healthy and not-healthy samples.
-AUC for hsa-miR-21, hsa-miR-32, and hsa-miR-186 is 0.99, 1, and 0.91 respectively.
-Also the accuracy of each microRNA for classifying the samples to healthy and not-healthy groups on test set using a simple logistic regression model is 0.92, 0.98, and 0.89 for hsa-miR-21, hsa-miR-32, and hsa-miR-186 respectively.
+Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 were differntially expressed miRNAs.
+The expression profile of these three miRNAs was presented in Fig\@ref(fig:DEMexp).
+Also, The ROC curves of each miRNA for each layer was presented in Fig\@ref(fig:miRROC).
+AUC for hsa-miR-21, hsa-miR-32, and hsa-miR-186 was 0.99, 1, and 0.91 respectively using a simple logistic regression model (Fig\@ref(fig:miRROC)A).
+Using the same model the accuracy of each microRNA for classifying the samples to healthy and not-healthy groups on test set was 0.92, 0.98, and 0.89 for hsa-miR-21, hsa-miR-32, and hsa-miR-186 respectively.
 Moreover, In Fig\@ref(fig:miRROC)B the ROC curve of each microRNA for classifying MI and CAD samples were presented.
-For  hsa-miR-21, hsa-miR-32, and hsa-miR-186 AUC on test set is 0.85; 0.7; and 0.82, and accuracy on test set is 0.78; 0.67; and 0.74.
+For  hsa-miR-21, hsa-miR-32, and hsa-miR-186 AUC on test set was 0.85; 0.7; and 0.82, and accuracy on test set was 0.78; 0.67; and 0.74.
 
@@
 
-Although single microRNAs have acceptable performance, but their prediction value can improve even more using them as a set.
+Although single microRNAs had acceptable performance, but their prediction value could improve even more using them as a set.
 The ROC curve for the SVM model with rbf kernel trained with all three microRNAs is presented in Fig\@ref(fig:DEMROC)A.
-The model has better performance in classification than single microRNAs.The AUC for the model is 1, and its accuracy on test set is 1 as well.
+The model had better performance in classification than single microRNAs.The AUC for the model is 1, and its accuracy on test set was 1 as well.
 The confusion matrix for the model is presented in Fig\@ref(fig:DEMCF)A.
@@
 
-Different models has been trained using expression values of three differentially microRNAs.
-The models 10-fold cross-validate AUC and accuracy on the test set is reported on Fig\@ref(fig:DEMmodels).
-The best model from both AUC and accuracy point-of-view is an SVM model with linear kernel.
-The AUC and accuracy for this model with its preset values is 0.93 and 0.82 respectively.
-The model is hyper-tuned for C and gamma hyperparameters, and therefore the model showed a better performance.
-The ROC curve of the hyper-tuned model as the best model is presented in Fig\@ref(fig:DEMROC)B.
+Different models were trained using expression values of three differentially microRNAs.
+The models 10-fold cross-validate AUC and accuracy on the test set were reported on Fig\@ref(fig:DEMmodels).
+The best model from both AUC and accuracy point-of-view was the SVM model with linear kernel.
+The AUC and accuracy for this model with its preset values was 0.93 and 0.82 respectively.
+The model was hyper-tuned for C and gamma hyperparameters, and therefore the model showed a better performance.
+The ROC curve of the hyper-tuned model was presented in Fig\@ref(fig:DEMROC)B.
 For this model the AUC reached to 0.95 and the accuracy improved to 0.85.
-Moreover, the sensitivity and specificity for the model on the test set is 0.91 and 0.71 respectively.
+Moreover, the sensitivity and specificity for the model on the test set was 0.91 and 0.71 respectively.
 The confusion matrix for the hyper-tuned model is illustrated in Fig\@ref(fig:DEMCF)B.
@@
 
-#### Second layer; MI form CAD:
-
-After calculating AUC for each microRNA, they are sorted and their performance as a set were investigated.
-The metric for finding the best set was AUC.
-The diagram for AUC against the number of microRNAs in the set is presented in Fig\@ref(fig:AUCset).
-As it is clear, the AUC increase until  the number of microRNAs in the set reaches tho 6 and after that it drops.
-The AUC for separating MI samples from CAD using these microRNAs is 0.93.
-The microRNAs in the set are has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1.
+After calculating each microRNA's AUC for classifying MI and CAD samples, they were sorted and their performance as a set were investigated.
+The metric of choice for finding the best set was AUC.
+As shown in Fig\@ref(fig:AUCset), the AUC increased until  the number of microRNAs in the set reached tho 6 and after that it droped.
+The AUC for separating MI samples from CAD using these microRNAs was 0.93.
+The microRNAs in the set were has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1.
 The expression values of these microRNAs in healthy, CAD, and MI samples is presented in  Fig\@ref(fig:AUCexp).
@@
 
-For finding the most model for training the best set, different models were trained using their pre-set values.
-Their AUC and accuracy results on the test set is presented in Fig\@ref(fig:AUCmodels).
-The best model from AUC point of view is the LR and from accuracy point-of-view is the SVM model with polynomial kernel.
-For LR model the AUC and accuracy were 0.92 and 0.81; and for SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
-Both models were hyper-tuned and ROC curve for their best performance presented in Fig\@ref(fig:AUCROC)A and B.
-The AUC and accuracy for LR model increased to 0.94 and 0.88 respectively; And for SVM model with polynomial kernel these values increased to 0.95 and 0.88 respectively.
-The sensitivity for LR and SVM models are 1 and 0.97; and the specificity for them is 0.57 and 0.64, respectively.
-The confusiom matrix for both models is illustrated in Fig\@ref(fig:AUCCF)A and B.
+#### First layer:
 
-```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) Logistic regression model for CAD and MI samples classification.
-(B) SVM with polynomial kernel for CAD and MI samples classification.
-(C) SVM with RBF kernel for healthy and not-healthy samples classification."}
+Using the selected set, an SVM model with RBF kernel was trained for separating healthy from not-healthy samples.
+The ROC curve for the model is presented in Fig\@ref(fig:AUCROC)A and confusion matrix is illustrated in Fig\@ref(fig:AUCCF)A.
+Both AUC and accuracy for the model on the test set was equal to 1.
+
+```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of microRNAs selected by AUC on test set classification.
+(A) SVM with RBF kernel for healthy and not-healthy samples classification.
+(B) Logistic regression model for CAD and MI samples classification.
+(C) SVM with polynomial kernel for CAD and MI samples classification.
+"}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC best model ROCs.pdf")
+```
+
+```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+(B) Logistic regression model for CAD and MI samples classification.
+(C) SVM with polynomial kernel for CAD and MI samples classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC Conf. matrix.pdf")
 ```
+
+#### Second layer; MI form CAD:
+
+For finding the best model for training the best set, different models were trained using their pre-set values.
+Their AUC and accuracy results on the test set is presented in Fig\@ref(fig:AUCmodels).
+The best model from AUC point-of-view was the LR and from accuracy point-of-view was the SVM model with polynomial kernel.
+For LR model the AUC and accuracy were 0.92 and 0.81; and for SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
+Both models were hyper-tuned and ROC curve for their best performance presented in Fig\@ref(fig:AUCROC)B and C.
+The AUC and accuracy for LR model increased to 0.94 and 0.88 respectively; And for SVM model with polynomial kernel these values increased to 0.95 and 0.88 respectively.
+The sensitivity for LR and SVM models were 1 and 0.97; and the specificity for them were 0.57 and 0.64, respectively.
+The confusion matrix for both models is illustrated in Fig\@ref(fig:AUCCF)B andC.
 
@@
 
-```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of microRNAs selected by AUC on test set classification.
-(A) Logistic regression model for CAD and MI samples classification.
-(B) SVM with polynomial kernel for CAD and MI samples classification.
-(C) SVM with RBF kernel for healthy and not-healthy samples classification."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC best model ROCs.pdf")
-```
-
-#### First layer:
-
-Using the microRNAs set selected for classification of MI and CAD samples, an SVM model with RBF kernel has been trained for separating healthy from not-healthy samples.
-The ROC curve for the model is presented in Fig\@ref(fig:AUCROC)C and confusion matrix is illustrated in Fig\@ref(fig:AUCCF)C.
-Both AUC and accuracy for the model on the test set is equal to 1.
-
 # Refrences

### 2023-02-08 ArticleDraft.Rmd
Grammer check by gmail

@@
 fRMA allows to preprocess individual microarray samples and combine them consistently for analysis.
-For each dataset, background correction was performed using RMA algorithm and then it was quantile normalized based on the reference distribution.
+For each dataset, background correction was performed using the RMA algorithm and then it was quantile normalized based on the reference distribution.
 During summarization, batch effects were removed and variances of the gene expressions were estimated by taking into account these probe-specific effects.
 For those multiple probe sets matched to the identical gene, the mean log fold change was retained.
-This way fRMA can be seen as a batch effect removal technique for different datasets that were produced identical microarray platform.
-Thus, to become ensure about batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
+This way, fRMA can be seen as a batch effect removal technique for different datasets that produced identical microarray platforms.
+Thus, to become sure about batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
 
@@
 Huge sets of samples were collected and normalized using fRMA for several platforms.
-The distribution of the expressed and non-expressed observed intensities for each gene is estimated using these normalized sets.
+The distribution of the expressed and unexpressed observed intensities for each gene is estimated using these normalized sets.
 Genes are deemed expressed (and their value coded to 1) or unexpressed (and their value coded to 0) according to the following equation:
@@
 To determine if the expressed ratios differed in the MI group versus the healthy control group, Fisher's exact test for individual genes was carried out upon the barcode values.
-Genes with a false discovery rate (FDR) of $< 0.05$, which was calculated through the Benjamini-Hochberg (BH) procedure to adjust for multiple testing issue, were considered as differentially expressed genes.
+Genes with a false discovery rate (FDR) of $< 0.05$, which was calculated through the Benjamini-Hochberg (BH) procedure to adjust for multiple testing issues, were considered as differentially expressed genes.
 
-The same procedure were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs in between them.
+The same procedures were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs in between them.
 
@@
 An adjusted P \< 0.05 was considered to indicate a statistically significant difference.
-Enrichment were conducted on the MI-healthy and CAD-healthy DEGs.
+Enrichments were conducted on the MI-healthy and CAD-healthy DEGs.
 In these analyses, all default parameters were used.
@@
 3.9, numpy [@numpy], pandas [@pandas] and Scikit-Learn packages [@scikit-learn].
-When ever hyper-tuning was needed the scikit-opt package [@scikitopt] was used.
-In all ML analysis, the datasets were divided into train and test set by 0.7:0.3 ratio and all reported results are the average of a 10-fold cross-validation.
+Whenever hyper-tuning was needed the scikit-opt package [@scikitopt] was used.
+In all ML analysis, the datasets were divided into train and test sets by 0.7:0.3 ratio and all reported results are the average of a 10-fold cross-validation.
 
@@
 In the second approach microRNAs were selected by their individual AUC-ROC.
-Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of set of microRNAs selected with different logic.
+Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of a set of microRNAs selected with different logic.
 
@@
 In this approach a two layer architecture was deployed to the data in order to maximize the prediction values.
-The first layer predicted whether a sample is healthy or not, and the second layer was separated MI from CAD in the samples which were predicted as not healthy in the first layer.
+The first layer predicted whether a sample is healthy or not, and the second layer separated MI from CAD in the samples which were predicted as not healthy in the first layer.
 To this end, a distinct ML model was trained for each layer.
-Since there are limited numbers of microRNAs in DEGs both layers where trained with all of them.
+Since there are limited numbers of microRNAs in DEGs both layers were trained with all of them.
 For further comparison with the models performance ROC curve of each microRNA for classifying healthy and not-healthy, as well as CAD and MI was generated using a simple Logistic Regression model.
@@
 
-An SVM model using RBF kernels were hyper-tunned and trained using all microRNAs in DEGs.
-In order to handle the imbalance number of samples in groups (51 for healthy and 157 for not-healthy group), sample weight for not-healthy samples were set to 0.5.
+An SVM model using RBF kernels was hyper-tunned and trained using all microRNAs in DEGs.
+In order to handle the imbalance in the number of samples  (51 for healthy and 157 for not-healthy group), sample weight for not-healthy samples was set to 0.5.
 
@@
 Criteria for choosing the best model was the highest accuracy and AUC on the test set.
-The best model was hyper-tuned with scikit-opt package [@scikitopt] in order to get the best predictive performance.
+The best model was hyper-tuned with the scikit-opt package [@scikitopt] in order to get the best predictive performance.
 
@@
 The first layer for classifying samples to healthy and not-healthy and the second layer for separating MI and CAD samples.
-However, for keeping the number of microRNAs as low as possible microRNAs will be selected from the second layer (which are the microRNAs with the best performance in MI/CAD separation) and then their performance were evaluated in the first layer.
+However, to keep the number of microRNAs as low as possible microRNAs will be selected from the second layer (which are the microRNAs with the best performance in MI/CAD separation) and then their performance will be evaluated in the first layer.
 
 AUC-ROC of all microRNAs for classifying MI and CAD samples has been calculated.
-For finding the number of microRNAs with the highest predictive values, the microRNAs with the highest individual AUC-ROC is adding to the set one-by-one and the AUC-ROC for the set is calculated.
+For finding the number of microRNAs with the highest predictive values, the microRNAs with the highest individual AUC-ROC are added to the set one-by-one and the AUC-ROC for the set is calculated.
 The set with the highest AUC-ROC has been selected as the set for the following steps.
@@
 
-An SVM model with RBF kernel were trained with the selected set of microRNAs.
+An SVM model with RBF kernel was trained with the selected set of microRNAs.
 Moreover, the model was hyper-tuned for finding the hyper-parameters with the best AUC and accuracy.
@@
 The selected number of microRNAs was used to train different algorithms for finding the best model.
-As previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB and DT were trained.
+As the previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB and DT were trained.
 All models were trained with their pre-set parameters with 10-fold cross-validation.
-The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned with scikit-opt package [@scikitopt].
+The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned with the scikit-opt package [@scikitopt].
 The ROC curve and confusion matrix for the best model was reported.
@@
 The RLE plot presented in Fig\@ref(fig:PCA) validates batch effect removal.
-For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in RLE plot, and inter-quantile distances would be greater than 0.1 [@BER].
+For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in the RLE plot, and inter-quantile distances would be greater than 0.1 [@BER].
 The mentioned criteria was not met in primary data, but was met after conducting fRMA algorithm.
@@
 
-: Total DEGs and name of differentially expressed microRNAs.
+: Total DEGs and differentially expressed microRNAs.
 
@@
 
-Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 were differntially expressed miRNAs.
+Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 were differentially expressed miRNAs.
 The expression profile of these three miRNAs was presented in Fig\@ref(fig:DEMexp).
@@
 Different models were trained using expression values of three differentially microRNAs.
-The models 10-fold cross-validate AUC and accuracy on the test set were reported on Fig\@ref(fig:DEMmodels).
+The models 10-fold cross-validated AUC and accuracy on the test set were reported on Fig\@ref(fig:DEMmodels).
 The best model from both AUC and accuracy point-of-view was the SVM model with linear kernel.
@@
 The metric of choice for finding the best set was AUC.
-As shown in Fig\@ref(fig:AUCset), the AUC increased until  the number of microRNAs in the set reached tho 6 and after that it droped.
+As shown in Fig\@ref(fig:AUCset), the AUC increased until  the number of microRNAs in the set reached tho 6 and after that it dropped.
 The AUC for separating MI samples from CAD using these microRNAs was 0.93.
@@
 Using the selected set, an SVM model with RBF kernel was trained for separating healthy from not-healthy samples.
-The ROC curve for the model is presented in Fig\@ref(fig:AUCROC)A and confusion matrix is illustrated in Fig\@ref(fig:AUCCF)A.
+The ROC curve for the model is presented in Fig\@ref(fig:AUCROC)A and the confusion matrix is illustrated in Fig\@ref(fig:AUCCF)A.
 Both AUC and accuracy for the model on the test set was equal to 1.
@@
 The best model from AUC point-of-view was the LR and from accuracy point-of-view was the SVM model with polynomial kernel.
-For LR model the AUC and accuracy were 0.92 and 0.81; and for SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
+For the LR model the AUC and accuracy were 0.92 and 0.81; and for the SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
 Both models were hyper-tuned and ROC curve for their best performance presented in Fig\@ref(fig:AUCROC)B and C.
@@
 The sensitivity for LR and SVM models were 1 and 0.97; and the specificity for them were 0.57 and 0.64, respectively.
-The confusion matrix for both models is illustrated in Fig\@ref(fig:AUCCF)B andC.
+The confusion matrix for both models is illustrated in Fig\@ref(fig:AUCCF)B and C.
 

### 2023-04-25 ArticleDraft.Rmd
Final first draft

@@
 link-citation: yes
-bibliography: For Article Refrence.bib
+bibliography: [For Article Refrence.bib, discussion.bib, Introduction.bib]
 output:
+  bookdown::word_document2: default
   bookdown::pdf_book:
-  bookdown::word_document2: default
 #    base_format: rticles::bioinformatics_article
+always_allow_html: true
 ---
@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(knitr, captioner, bundesligR, stringr)
+pacman::p_load(knitr, captioner, bundesligR, stringr, kableExtra)
 ```
 
-```{r , echo=F}
-# What to do next:
+# Introduction
 
-# Generating information related to the machine learning first layer using welcome file.
+At present, cardiovascular diseases (CVDs) are the leading cause of human mortality with 32% of all global deaths.
+It is estimated that about 85% of CVDs mortality were diagnosed with myocardial infarction (MI) [@CVD].
+MI is an acute coronary syndrome with sudden blockage and stenosis of the coronary artery, and subsequent myocardial ischemia, leading to extensive cardiomyocyte damage and necrosis [@Macro].
 
-```
+Over the last 50 years, numerous attempts have been collected to use biomarkers to facilitate diagnosis, assess risk, follow-up therapy, and determine therapeutic efficacy in CVDs candidates.
+Based on the released guidelines, cardiac troponins (cTns) are used as a highly-sensitive and accurate approach for the detection of myocardial ischemia.
+Despite the inherent advantages, the high-rate sensitivity of cTn-based assays has also led to more false positive results [@17], which do necessitate the advent and development of new modalities with pathological values.
+To improve diagnostic value upon existing CVD biomarkers, the combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, is proposed.
+Previous data support the notion that miRNAs exhibit the great potential to be used as alternative biomarkers in CVD detection and follow-up [@miR].
+It is suggested that miRNAs possess 18-22 nucleotides and can play a crucial role in the regulation of gene expression.
+Evidence point to the fact that miRNAs are involved in the pathogenesis of cardiac tissue injury and be as theranostics in terms of CVDs.
+These elements can easily circulate in biofluids with both diagnostic as well as prognostic values [@m1].
+Several biological activities such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque formation, and cardiac rhythm are regulated by miRNAs [@m2].
+It is postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in CVD patients.
+To be specific, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) [@36].
+
+PBMCs are a fraction of white blood cells (WBCs), including monocytes, lymphocytes, macrophages, and other cells belonging to the immune system [@PBMC-miR].
+Emerging data have indicated that PBMCs can be used as a valid source of biomarkers for monitoring various pathological conditions.
+Of note, the alteration of mRNAs and miRNAs under pathological conditions gives us valuable information about different kinds of disorders.
+PBMCs could recapitulate the conditions of the target tissues, thus, providing a highly sensitive and specific source of biomarkers [@Meder6].
+Commensurate with these conditions, these cells are repositories of dysregulated genes and miRNAs expression profiles in CVDs related to control conditions [@PBMC-miR; @Meder6].
+
+In recent years, the advent and use of machine learning (ML) is an exciting prospect for advancing scientific discoveries.
+Although the concept of ML and its initial algorithms were conceived many years ago, recent improvements in computing power and access to vast amounts of data have shown that ML techniques outperform classical statistical methods in various fields.
+Furthermore, the progress made in omics technologies has enabled the analysis of massive and intricate biological data sets, consisting of hundreds to thousands of samples, which makes it possible for ML to extract valuable biological information from such data [@ML-Bio].
+Therefore, ML offers novel techniques to integrate and analyze the various omics data enabling the discovery of de novo biomarkers.
+These biomarkers help us in accurate disease prediction, patient stratification, and finding new therapeutics [@ML].
+
+In this study, we aimed to identify potential miRNA biomarkers for MI patients by combining and analyzing three different microarray datasets from PBMCs.
+It is suggested that the integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring CVDs.
+Besides, this approach can deepen our vision into the underlying mechanisms of CVDs and aid in the development of valid theranostic tools, and patient stratification.
 
@@
 Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database (<https://www.ncbi.nlm.nih.gov/geo/>).
-In order to come by sufficient classification power between MI samples and others, the sample size should be relatively large.
+To obtain sufficient classification power between MI, healthy and CAD samples, a relatively large sample size was required.
 Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475 for healthy samples were selected.
 All samples were produced using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
-Only healthy, stable CAD and early stage MI samples were selected from these datasets for further analysis.
-The basic information for the three datasets evaluated in the current study is provided in Table 1.
+Only healthy, stable CAD and early-stage MI samples were selected from these datasets for further analyses.
+The basic information for the three datasets evaluated in the current study is provided in Table \@ref(tab:datasets).
 
-|          | Platform | Healthy Control | CAD Control | MI  | Reference |
-|----------|----------|-----------------|-------------|-----|-----------|
-| GSE59867 | GPL6244  | \-              | 46          | 111 | [@67]     |
-| GSE56609 | GPL6244  | 46              | \-          | \-  | [@09]     |
-| GSE54475 | GPL6244  | 5               | \-          | \-  | [@75]     |
+```{r, datasets, echo=FALSE}
+df <- data.frame(Dataset = c("GSE59867", "GSE56609", "GSE54475"),
+                 Platform = c("GPL6244", "GPL6244", "GPL6244"),
+                 Healthy = c("-", 46, 5),
+                 CAD = c(46, "-", "-"),
+                 MI = c(111, "-", "-"),
+                 Refrence = c("[@67]", "[@09]", "[@75]"))
 
-: Basic information of the GEO microarray datasets.
+kbl(df, format = "markdown", booktabs = T, align = "c", linesep = c(""), caption = "Basic information of the GEO microarray datasets.")
+```
 
@@
 Raw data (CEL files) of all datasets were downloaded from the GEO and pre-processed using the fRMA package [@fRMA].
-fRMA allows to preprocess individual microarray samples and combine them consistently for analysis.
+fRMA allowed to pre-process of individual microarray samples and combining them consistently for analysis.
 For each dataset, background correction was performed using the RMA algorithm and then it was quantile normalized based on the reference distribution.
@@
 For those multiple probe sets matched to the identical gene, the mean log fold change was retained.
-This way, fRMA can be seen as a batch effect removal technique for different datasets that produced identical microarray platforms.
-Thus, to become sure about batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
+Therefore, fRMA can be seen as a batch effect removal technique for different datasets that are produced by identical microarray platforms.
+Thus, to ensure batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
 
@@
 The barcode algorithm proposed by McCall et al.
-[@Barcode] transforms the actual expression values into binary barcode values.
-Huge sets of samples were collected and normalized using fRMA for several platforms.
+[@Barcode] transformed the actual expression values into binary barcode values.
+Huge sets of samples were collected and normalized using fRMA for several platforms as well as for Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
 The distribution of the expressed and unexpressed observed intensities for each gene is estimated using these normalized sets.
-Genes are deemed expressed (and their value coded to 1) or unexpressed (and their value coded to 0) according to the following equation:
+Genes were considered expressed (and their value coded to 1) or unexpressed (and their value coded to 0) according to the following equation:
 
@@
 
-where $x_{ij}$ is the normalized intensity of gene $i$ in sample $j$, $C$ is a user-defined parameter, $\sigma^{ne}$ is the standard deviation of the non-expressed distribution and $\mu^{ne}$ is the mean of the non-expressed distribution.
+where $x_{ij}$ is the normalized intensity of gene $i$ in sample $j$, $C$ is a user-defined parameter, $\sigma^{ne}$ is the standard deviation of the non-expressed distribution, and $\mu^{ne}$ is the mean of the non-expressed distribution.
 The barcode representation of a sample is a vector of ones and zeros denoting which genes are estimated to be expressed (ones) and unexpressed (zeros).
@@
 Genes with a false discovery rate (FDR) of $< 0.05$, which was calculated through the Benjamini-Hochberg (BH) procedure to adjust for multiple testing issues, were considered as differentially expressed genes.
-
-The same procedures were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs in between them.
+The same procedures were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs between them.
 
@@
 
-Using the R clusterProfilter package, the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation were carried out on the differentially expressed genes.
+Using the R clusterProfiler package, the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation were carried out on the differentially expressed genes.
 The GO analysis included biological process (BP), cellular component (CC) and molecular function (MF) categories.
-An adjusted P \< 0.05 was considered to indicate a statistically significant difference.
+An adjusted p-value of less than 0.05 was considered to indicate a statistically significant difference.
 Enrichments were conducted on the MI-healthy and CAD-healthy DEGs.
@@
 
-## Machine Learning
+## ML procedure
 
-The machine learning analysis was performed using Python software, ver.
-3.9, numpy [@numpy], pandas [@pandas] and Scikit-Learn packages [@scikit-learn].
-Whenever hyper-tuning was needed the scikit-opt package [@scikitopt] was used.
-In all ML analysis, the datasets were divided into train and test sets by 0.7:0.3 ratio and all reported results are the average of a 10-fold cross-validation.
+The ML analysis was performed using Python software, ver.
+3.9, Numpy [@numpy], pandas [@pandas], and Scikit-Learn packages [@scikit-learn].
+Whenever hyper-tuning was needed, the scikit-opt package [@scikitopt] was used.
+In all ML analyses, the datasets were divided into train and test sets by a 0.7:0.3 ratio and all reported results are the average of 10-fold cross-validation.
 
-Two different approaches were used for selecting microRNAs to train models with.
-The first approach was using the microRNAs that are differentially expressed.
-In the second approach microRNAs were selected by their individual AUC-ROC.
-Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of a set of microRNAs selected with different logic.
+Two different approaches were used for selecting miRNAs for model training.
+The first approach was using the miRNAs that are differentially expressed.
+In the second approach, miRNAs were selected by their individual AUC-ROC.
+Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of sets of miRNAs selected with different logics.
 
-### microRNAs in DEGs
+### miRNAs in DEGs
 
-In this approach a two layer architecture was deployed to the data in order to maximize the prediction values.
+In this approach, a two-layer architecture was deployed to the data to maximize the prediction values.
 The first layer predicted whether a sample is healthy or not, and the second layer separated MI from CAD in the samples which were predicted as not healthy in the first layer.
 To this end, a distinct ML model was trained for each layer.
-Since there are limited numbers of microRNAs in DEGs both layers were trained with all of them.
-For further comparison with the models performance ROC curve of each microRNA for classifying healthy and not-healthy, as well as CAD and MI was generated using a simple Logistic Regression model.
+Since there is a limited number of miRNAs in DEGs, both layers were trained with all of them.
+For further comparison with the models' performance, the ROC curve of each miRNA for classifying healthy and not-healthy, as well as CAD and MI, were generated using a Logistic Regression model.
 
-#### First layer for seperating healthy and not-healthy samples:
+#### First layer for isolation of healthy and not-healthy samples:
 
-An SVM model using RBF kernels was hyper-tunned and trained using all microRNAs in DEGs.
-In order to handle the imbalance in the number of samples  (51 for healthy and 157 for not-healthy group), sample weight for not-healthy samples was set to 0.5.
+An SVM model using RBF kernels was trained and  hyper-tuned using all miRNAs in DEGs.
+To handle the severe imbalance in the number of samples (51 for the healthy and 157 for the not-healthy groups), the sample weight for the healthy and the not-healthy samples were set to 1 and 0.5, respectively.
 
@@
 
-For the sake of reaching the highest classification performance using the set of microRNAs, different models were investigated.
+For the sake of reaching the highest classification performance using the set of miRNAs, different models were investigated.
 To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) models were trained.
 All models were trained with their pre-set parameters with 10-fold cross-validation.
+The criteria for choosing the best model were the highest accuracy and AUC on the test set.
+The best model was hyper-tuned with the scikit-opt package [@scikitopt] to get the best predictive performance.
 
-Criteria for choosing the best model was the highest accuracy and AUC on the test set.
-The best model was hyper-tuned with the scikit-opt package [@scikitopt] in order to get the best predictive performance.
+### miRNAs with the highest AUC-ROC
 
-### microRNAs with the highest AUC-ROC
+Like the previous approach, a two-layers strategy was conducted.
+The first layer classified samples into healthy and not-healthy, and the separated MI and CAD samples.
+However, to keep the number of miRNAs as low as possible miRNAs were selected from the second layer (which are the miRNAs with the best performance in MI/CAD separation), and then their performance wase evaluated in the first layer.
+AUC-ROC of all miRNAs for classifying MI and CAD samples were calculated.
+To find the number of miRNAs with the highest predictive values, the miRNAs with the highest individual AUC-ROC were added to the set one by one, and the AUC-ROC for the set was calculated.
+The set with the highest AUC-ROC was selected for the following steps.
+The ROC curves for each selected miRNA for separating healthy samples from not-healthy and MIs from CADs were also plotted for further comparison.
 
-Like the previous approach, a two layers strategy was conducted.
-The first layer for classifying samples to healthy and not-healthy and the second layer for separating MI and CAD samples.
-However, to keep the number of microRNAs as low as possible microRNAs will be selected from the second layer (which are the microRNAs with the best performance in MI/CAD separation) and then their performance will be evaluated in the first layer.
+#### First layer for the detection of healthy and not-healthy samples:
 
-AUC-ROC of all microRNAs for classifying MI and CAD samples has been calculated.
-For finding the number of microRNAs with the highest predictive values, the microRNAs with the highest individual AUC-ROC are added to the set one-by-one and the AUC-ROC for the set is calculated.
-The set with the highest AUC-ROC has been selected as the set for the following steps.
-The ROC curves for each selected microRNA for separating healthy samples from not-healthy and MIs from CADs were also plotted for further comparison.
-
-#### First layer for finding healthy and not-healthy samples:
-
-An SVM model with RBF kernel was trained with the selected set of microRNAs.
-Moreover, the model was hyper-tuned for finding the hyper-parameters with the best AUC and accuracy.
-The ROC curve and confusion matrix for the best model was reported.
+An SVM model with an RBF kernel was trained using the selected set of miRNAs.
+Additionally, the model was hyper-tuned to find the hyper-parameters for the highest AUC and accuracy.
+The ROC curve and confusion matrix for the best model were reported.
 
@@
 
-The selected number of microRNAs was used to train different algorithms for finding the best model.
-As the previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB and DT were trained.
-All models were trained with their pre-set parameters with 10-fold cross-validation.
-The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned with the scikit-opt package [@scikitopt].
-The ROC curve and confusion matrix for the best model was reported.
+The selected miRNAs set was used to train different algorithms to find the best model.
+Similar to the previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT were trained.
+All models were trained with their pre-set parameters using 10-fold cross-validation.
+The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned using the scikit-opt package [@scikitopt].
+The ROC curve and confusion matrix for the best model were reported.
 
-# Results and Discussion
+# Results
 
@@
 
-The PCA plot of the train samples were shown in Fig\@ref(fig:PCA).
-As it is clear, there was a complete separation between healthy samples and CAD and MI samples in primary data which also remains after conducting fRMA on the data.
+The PCA plots of the training samples are shown in Figure \@ref(Fig:PCA)A and B.
+As shown, healthy samples were separated from CAD or MI samples in primary data and also after conducting fRMA.
+In the RLE plot, there was a distinct difference between dataset means for all samples before conducting fRMA  (Figure \@ref(Fig:PCA)C).
+All datasets were rearranged around 0 in the RLE plot after conducting fRMA (Figure \@ref(Fig:PCA)D).
+Moreover, there was a clear change in inter-quantile distances, but the values still were over 0.1.
 
-The RLE plot presented in Fig\@ref(fig:PCA) validates batch effect removal.
-For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in the RLE plot, and inter-quantile distances would be greater than 0.1 [@BER].
-The mentioned criteria was not met in primary data, but was met after conducting fRMA algorithm.
-
-```{r PCA, echo=F, out.width="100%", fig.cap="PCA and RLE plot for all samples before and after fRMA."}
+```{r PCA, echo=F, out.width="100%", Figure .cap="PCA and RLE plot for all samples before and after fRMA."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/PCA and RLE of trains.pdf")
@@
 
-```{r , echo=F}
+```{r , DEGstab, echo=F}
 M_H <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-Healthy.txt", check.names = F)
@@
 C_H <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs CAD-Healthy.txt", check.names = F)
+
+df <- t(data.frame(`MI vs. Healthy` = c(dim(M_H)[1], sum(M_H$logFC > 0), sum(M_H$logFC < 0), "hsa-miR-186, miR-21, miR-32"),
+       `CAD vs. Healthy` =  c(dim(C_H)[1], sum(C_H$logFC > 0), sum(C_H$logFC < 0), "hsa-miR-186, miR-21, miR-32"), 
+       `MI vs. CAD` = c(dim(M_C)[1], sum(M_C$logFC > 0),  sum(M_C$logFC < 0), "hsa-miR-186")))
+
+colnames(df) <- c("DEGs",  "up-regulated DEGs", "down-regulated DEGs", "miRNAs")
+rownames(df) <- c("MI vs. Healthy", "CAD vs. Healthy", "MI vs. CAD")
+
+kbl(df, booktabs = T, align = "c", linesep = c(""), caption = "Total, up-, and down-regulated DEGs and differentially expressed miRNAs.")%>% 
+  kable_styling(full_width = T) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(5, width = "4cm")
 ```
@@
 Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in MI compared to the healthy controls.
-For CAD and healthy groups there were `r dim(C_H)[1]` DEGs, `r sum(C_H$logFC > 0)` of them were up-regulated and `r sum(C_H$logFC < 0)` of them were down-regulated in CAD samples in comparison with healthy samples.
-For MI and CAD group the number of DEGs are `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples in comparison with CAD samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively.
-All these data is summarized in table 2.
+In CAD and healthy groups comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and and `r sum(C_H$logFC < 0)` DEGs were up- and `r down-regulated, respectively in CAD samples.
+In the MI and CAD groups, the number of DEGs was `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively in comparison with CAD samples.
+These data are summarized in Table \@ref(tab:DEGstab).
 
-|                 | No. of DEGs     | no. of up-regulated DEGs | no. of down-regulated DEGs | microRNAs                           |
-|---------------|---------------|---------------|---------------|---------------|
-| MI vs. Healthy  | `r dim(M_H)[1]` | `r sum(M_H$logFC > 0)`   | `r sum(M_H$logFC < 0)`     | hsa-miR-186, hsa-miR-21, hsa-miR-32 |
-| CAD vs. Healthy | `r dim(C_H)[1]` | `r sum(C_H$logFC > 0)`   | `r sum(C_H$logFC < 0)`     | hsa-miR-186, hsa-miR-21, hsa-miR-32 |
-| MI vs. CAD      | `r dim(M_C)[1]` | `r sum(M_C$logFC > 0)`   | `r sum(M_C$logFC < 0)`     | hsa-miR-186                         |
+```{r venn, echo=F, out.width="60%", Figure .align='center', Figure .cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/venny.png")
+```
 
-: Total DEGs and differentially expressed microRNAs.
+The Venn diagram  in Figure \@ref(Fig:venn) shows that CAD and MI samples share a majority of their DEGs.
+From 860 DEGs of MI/healthy and 670 DEGs of CAD/healthy, 531 genes were common which is 62% of MI/healthy DEGs and 79% of CAD/healthy DEGs.
 
@@
 To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
+In MI versus healthy samples, GO enrichment analysis in the BP category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Figure \@ref(Fig:MIHEnrich)A).
+In the CC category, the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Figure \@ref(Fig:MIHEnrich)B).
+In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(Fig:MIHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(Fig:MIHEnrich)D).
 
-For MI versus healthy samples, GO enrichment analysis in the BP category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Fig\@ref(fig:MIHEnrich)A).
-In the CC category the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:MIHEnrich)B).
-In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Fig\@ref(fig:MIHEnrich)C).
-KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:MIHEnrich)D).
-
-```{r MIHEnrich, echo=F, out.width="95%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs.
+```{r MIHEnrich, echo=F, out.width="95%", Figure .cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs.
 (A) Biological process terms.
@@
 The enrichment results for DEGs of CAD versus healthy samples were as follows.
-In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Fig\@ref(fig:CADHEnrich)A).
-In the CC category the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:CADHEnrich)B).
-In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Fig\@ref(fig:CADHEnrich)C).
-KEGG pathway analysis indicated that the  DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:CADHEnrich)D).
+In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Figure \@ref(Fig:CADHEnrich)A).
+In the CC category, the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Figure \@ref(Fig:CADHEnrich)B).
+In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(Fig:CADHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(Fig:CADHEnrich)D).
 
-```{r CADHEnrich, echo=F, out.width="95%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs.
+```{r CADHEnrich, echo=F, out.width="95%", Figure .cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs.
 (A) Biological process terms.
@@
 
-### microRNAs in DEGs
+### miRNAs in DEGs
 
-Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 were differentially expressed miRNAs.
-The expression profile of these three miRNAs was presented in Fig\@ref(fig:DEMexp).
-Also, The ROC curves of each miRNA for each layer was presented in Fig\@ref(fig:miRROC).
-AUC for hsa-miR-21, hsa-miR-32, and hsa-miR-186 was 0.99, 1, and 0.91 respectively using a simple logistic regression model (Fig\@ref(fig:miRROC)A).
-Using the same model the accuracy of each microRNA for classifying the samples to healthy and not-healthy groups on test set was 0.92, 0.98, and 0.89 for hsa-miR-21, hsa-miR-32, and hsa-miR-186 respectively.
-Moreover, In Fig\@ref(fig:miRROC)B the ROC curve of each microRNA for classifying MI and CAD samples were presented.
-For  hsa-miR-21, hsa-miR-32, and hsa-miR-186 AUC on test set was 0.85; 0.7; and 0.82, and accuracy on test set was 0.78; 0.67; and 0.74.
+Among all DEGs, just miR-186, miR-32, and miR-21 were detected as differentially expressed miRNAs.
+The expression profile of these three miRNAs is presented in Figure \@ref(Fig:DEMexp).
+Additionally, The ROC curves of each miRNA for each layer are presented in Figure \@ref(Fig:miRROC).
+Using the logistic regression model the AUC for miR-21, miR-32, and miR-186 was 0.99, 1, and 0.91 respectively  (Figure \@ref(Fig:miRROC)A).
+Besides, the accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set was 0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
+Moreover, the ROC curve of each miRNA for classifying MI and CAD samples was presented in Figure \@ref(Fig:miRROC)B.
+For miR-21, miR-32, and miR-186, the AUC and accuracy on the test set were 0.85; 0.7; and 0.82, and  0.78; 0.67; and 0.74, respectively.
 
-```{r DEMexp, echo=F, out.width="50%", fig.align='center', fig.cap="Expression profile of hsa-miR-186, hsa-miR-21, and hsa-miR32 in Healthy, CAD, and MI samples."}
+```{r, mirExptable, echo=FALSE}
+
+# data
+df <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/microRNAs logFC.txt", sep = "\t", check.names = F, header = T)
+
+kbl(df, booktabs = T, align = "c",linesep = c(""), digits = c(1, 60), caption = "Target miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.") %>%
+  kable_styling(full_width = T) %>%
+  add_header_above(c(" " = 1, "CAD/Healthy" = 2, "MI/Healthy" = 2, "MI/CAD" = 2))
+```
+
+```{r DEMexp, echo=F, out.width="50%", Figure .align='center', Figure .cap="Expression profile of miR-186, miR-21, and miR32 in Healthy, CAD, and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEGs healthy not-healthy/DEMs Expression.pdf")
@@
 
-```{r miRROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for single microRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+```{r miRROC, echo=F, out.width="95%", Figure .align='center', Figure .cap="ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/miRs ROCs.pdf")
@@
 
-#### First layer for healthy not-healthy seperation:
+#### First layer for healthy not-healthy isolation:
 
-Although single microRNAs had acceptable performance, but their prediction value could improve even more using them as a set.
-The ROC curve for the SVM model with rbf kernel trained with all three microRNAs is presented in Fig\@ref(fig:DEMROC)A.
-The model had better performance in classification than single microRNAs.The AUC for the model is 1, and its accuracy on test set was 1 as well.
-The confusion matrix for the model is presented in Fig\@ref(fig:DEMCF)A.
+Although single miRNAs had acceptable performance, their predictive value could be improved even further by using them as a set.
+The ROC curve for the SVM model with an RBF kernel trained with all three miRNAs is presented in Figure \@ref(Fig:DEMROC)A.
+The model had a better performance in classification than single miRNAs.The AUC for the model is 1, and its accuracy on the test set was also 1.
+In Figure \@ref(Fig:DEMCF)A, the confusion matrix for the model is presented.
 
-```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for microRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+```{r DEMROC, echo=F, out.width="95%", Figure .align='center', Figure .cap="ROC curve for miRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs ROCs h not h cad mi.pdf")
@@
 
-```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy samples classification.
-and (B) An SVM model with linear kernel for CAD and MI samples classification."}
+```{r DEMCF, echo=F, out.width="90%", Figure .align='center', Figure .cap="Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs Conf. matrix.pdf")
@@
 
-Different models were trained using expression values of three differentially microRNAs.
-The models 10-fold cross-validated AUC and accuracy on the test set were reported on Fig\@ref(fig:DEMmodels).
+Different models were trained using expression values for three differentially miRNAs.
+The models' 10-fold cross-validated AUC and accuracy on the test set are reported in Figure \@ref(Fig:DEMmodels).
 The best model from both AUC and accuracy point-of-view was the SVM model with linear kernel.
-The AUC and accuracy for this model with its preset values was 0.93 and 0.82 respectively.
-The model was hyper-tuned for C and gamma hyperparameters, and therefore the model showed a better performance.
-The ROC curve of the hyper-tuned model was presented in Fig\@ref(fig:DEMROC)B.
-For this model the AUC reached to 0.95 and the accuracy improved to 0.85.
-Moreover, the sensitivity and specificity for the model on the test set was 0.91 and 0.71 respectively.
-The confusion matrix for the hyper-tuned model is illustrated in Fig\@ref(fig:DEMCF)B.
+The AUC and accuracy for this model with its pre-set values were 0.93 and 0.82 respectively.
+The model was hyper-tuned for C and gamma hyper-parameters, and therefore the model showed better performance.
+The ROC curve of the hyper-tuned model is presented in Figure \@ref(Fig:DEMROC)B.
+For this model the AUC reached  0.95 and the accuracy improved to 0.85 (Table \@ref(tab:DEGsML)).
+Moreover, the sensitivity and specificity for the model on the test set were 0.91 and 0.71 respectively.
+The confusion matrix for the hyper-tuned model is illustrated in Figure \@ref(Fig:DEMCF)B.
 
-```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) and accuracy of different models trained with three microRNAs in DEGs on the test set."}
+```{r DEMmodels, echo=F, out.width="65%", Figure .align='center', Figure .cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models DEGs mirs.pdf")
+```
+
+```{r, DEGsML, echo=FALSE}
+
+# data
+df <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEGsMICAD.txt", sep = "\t", check.names = F, header = T)
+
+kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning") %>%
+  kable_styling(full_width = T) %>%
+  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
 ```
@@
 
-After calculating each microRNA's AUC for classifying MI and CAD samples, they were sorted and their performance as a set were investigated.
-The metric of choice for finding the best set was AUC.
-As shown in Fig\@ref(fig:AUCset), the AUC increased until  the number of microRNAs in the set reached tho 6 and after that it dropped.
-The AUC for separating MI samples from CAD using these microRNAs was 0.93.
-The microRNAs in the set were has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1.
-The expression values of these microRNAs in healthy, CAD, and MI samples is presented in  Fig\@ref(fig:AUCexp).
-The ROC curve of the selected microRNAs for MI and CAD samples classification is illustrated in Fig\@ref(fig:miRROC)B.
+After calculating the AUC for each miRNA for the classification of MI and CAD samples, miRNAs were sorted, and their performance was investigated as a set.
+The metric of choice for selecting the best set was AUC.
+As shown in Figure \@ref(Fig:AUCset), the AUC increased until the number of miRNAs in the set reached six, and after that, it dropped.
+The AUC for separating MI samples from CAD using these miRNAs was 0.93.
+The miRNAs in the set were miR-29a, miR-197, miR-142, miR-21, miR-155, and miR-320C1.
+The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(Fig:AUCexp).
+The ROC curve of the selected miRNAs for MI and CAD sample classification is illustrated in Figure \@ref(Fig:miRROC)B.
 
-```{r AUCset, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) for sets containing increasing number of microRNAs with the highest individual AUC in MI/CAD separation."}
+```{r AUCset, echo=F, out.width="65%", Figure .align='center', Figure .cap="Area under the curve (AUC) for sets containing an increasing number of miRNAs with the highest individual AUC in MI/CAD separation."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC for cad mi mirs.pdf")
@@
 
-```{r AUCexp, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples."}
+```{r AUCexp, echo=F, out.width="90%", Figure .align='center', Figure .cap="Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/ExpressionforAUC.pdf")
@@
 
-#### First layer:
+#### First layer for isolation of healthy and not-healthy samples
 
-Using the selected set, an SVM model with RBF kernel was trained for separating healthy from not-healthy samples.
-The ROC curve for the model is presented in Fig\@ref(fig:AUCROC)A and the confusion matrix is illustrated in Fig\@ref(fig:AUCCF)A.
-Both AUC and accuracy for the model on the test set was equal to 1.
+Using the selected set, an SVM model with an RBF kernel was trained to separate healthy from not-healthy samples.
+The ROC curve for the model is presented in Figure \@ref(Fig:AUCROC)A and the confusion matrix is illustrated in Figure \@ref(Fig:AUCCF)A.
+Both AUC and accuracy for the model on the test set were equal to 1.
 
-```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of microRNAs selected by AUC on test set classification.
+```{r AUCROC, echo=F, out.width="95%", Figure .align='center', Figure .cap="ROC curve for the set of miRNAs selected by AUC on test set classification.
 (A) SVM with RBF kernel for healthy and not-healthy samples classification.
@@
 
-```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+```{r AUCCF, echo=F, out.width="90%", Figure .align='center', Figure .cap="Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
 (B) Logistic regression model for CAD and MI samples classification.
@@
 
-#### Second layer; MI form CAD:
+#### Second layer for isolation of MI samples from CAD samples
 
-For finding the best model for training the best set, different models were trained using their pre-set values.
-Their AUC and accuracy results on the test set is presented in Fig\@ref(fig:AUCmodels).
-The best model from AUC point-of-view was the LR and from accuracy point-of-view was the SVM model with polynomial kernel.
-For the LR model the AUC and accuracy were 0.92 and 0.81; and for the SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
-Both models were hyper-tuned and ROC curve for their best performance presented in Fig\@ref(fig:AUCROC)B and C.
-The AUC and accuracy for LR model increased to 0.94 and 0.88 respectively; And for SVM model with polynomial kernel these values increased to 0.95 and 0.88 respectively.
-The sensitivity for LR and SVM models were 1 and 0.97; and the specificity for them were 0.57 and 0.64, respectively.
-The confusion matrix for both models is illustrated in Fig\@ref(fig:AUCCF)B and C.
+To find the best model for training the best set, different models were trained using their pre-set values.
+Their AUC and accuracy results on the test set are presented in Figure \@ref(Fig:AUCmodels).
+The best model from the AUC point-of-view was the LR and from the accuracy point-of-view, it was the SVM model with a polynomial kernel.
+For the LR model the AUC and accuracy were 0.92 and 0.81, respectively; and for the SVM model with a polynomial kernel, the values were 0.91 and 0.84, respectively.
+Both models were hyper-tuned and the ROC curve for their best performance is presented in Figure \@ref(Fig:AUCROC)B and C.
+The AUC and accuracy for the LR model increased to 0.94 and 0.88, respectively.
+For the SVM model with a polynomial kernel, these values increased to 0.95 and 0.88, respectively (Table \@ref(tab:AUCML)).
+The sensitivity for the LR and SVM models were 1 and 0.97, respectively; and the specificity for them was 0.57 and 0.64, respectively.
+The confusion matrix for both models is illustrated in Figure \@ref(Fig:AUCCF)B and C.
 
-```{r AUCmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) and accuracy of different models trained with three microRNAs in DEGs on the test set."}
+```{r AUCmodels, echo=F, out.width="65%", Figure .align='center', Figure .cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models AUC mirs.pdf")
@@
 
-# Refrences
+```{r, AUCML, echo=FALSE}
+
+# data
+df <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUCMICAD.txt", sep = "\t", check.names = F, header = T)
+
+kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their individual AUC-ROC on the train and test set before and after hyper-tuning") %>%
+  kable_styling(full_width = T) %>%
+  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
+```
+
+# Discussion
+
+The prevalence of acute MI (AMI) can lead to high-rate mortality in the clinical setting.
+However, early diagnosis and application of suitable treatment protocols can reduce mortality and improve AMI prognosis [@CVD; @17; @18].
+Studies have suggested that changes in miRNA expression may play a significant role in the progression of MI and the subsequent remodeling [@23].
+It is believed that the expression of miRNAs is altered during the various biological processes correlated with MI within the myocardium or other related tissues [@24].
+Although several research has been concentrated on examining free circulating miRNAs in the serum samples for the detection of cardiac tissue injuries [@11-15], more information is needed to fully comprehend the miRNAs found in different blood sub-components like plasma, platelets, and PBMCs.
+Based on previous data, PBMCs are critically involved in plaque destabilization and rupture as well as early inflammatory responses during MI [@Meder6; @Meder23].
+Moreover, PBMCs have specific miRNA profile that is altered under certain pathological conditions which are great candidates as disease biomarkers [@Meder6].
+
+PBMCs can respond to several insulting conditions such as MI in the least possible time with prominent changes in their miRNA profile  [@Meder6].
+Considering the regulatory roles, subtle changes in the transcription of miRNAs can be monitored even before alteration in the levels of mRNAs and proteins [@miR].
+These features make the miRNAs an early-stage valid diagnostic tool for the detection of minor and major cell injuries.
+To date, few studies have been performed to compare the miRNA profiles in PBMCs belonging to acute MI patients and other CADs and healthy samples to find a robust set of identical miRNAs to differentiate these pathological conditions.
+
+In this study, we combined three GEO datasets of healthy, CAD, and MI samples.
+Having these samples set alongside bioinformatics analysis and ML means, it is possible to identify potential biomarker sets and also effective therapeutic targets.
+The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(Fig:venn)) are proof of the close relationship between the MI and CAD samples.
+Interestingly, functional enrichment analysis demonstrated that DEGs in both healthy/CAD and healthy/MI were strongly correlated to immune cell response which is a major cellular part of PBMCs.
+Here, two different sets of miRNAs were used as biomarker sets for sample classification.
+miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-21; miR-29a; miR-142; miR155; miR-197; and miR-320c1 were selected according to their AUC values.
+As shown in Figure \@ref(Fig:miRROC), all other miRNAs selected with both approaches had AUC-ROC over 0.9 for the isolation of healthy and not-healthy samples except for miR-142 and miR-29a.
+Data confirmed that the real challenge is to classify CAD and MI samples because of close overlap.
+Of 8 miRNAs under investigation in both approaches except for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of CAD and MI samples.
+Besides, the high AUC-ROC values of miRNAs confirms their high potential as biomarkers.
+
+ML models when trained with miRNA sets selected by both DEG and AUC approaches showed better performance in the classification than each miRNA.
+To avoid unwanted complexity and poor predictive values, a two-layer architecture was also designed.
+The first layer was for for the discrimination of healthy from not-healthy samples, and the second layer separated the CAD from MI candidates.
+As expected in both approaches, a hyper-tuned SVM model could flawlessly separates healthy from not-healthy samples using distinct miRNAs sets.
+The ML models were also capable of effectively separating CAD from MI patients.
+Although both miRNA sets had nearly the same AUC-ROC with their best model, the accuracy, sensitivity, and specificity were different.
+The model trained with DEGs had better specificity, but the one trained with AUC-selected miRNAs had slightly better accuracy and higher sensitivity.
+This difference comes from the different strategies the selection of biomarkers sets.
+
+Numerous studies have reported different biological processes can affect the expression of miRNAs in PBMCs.
+However, there are still controversies regarding the exact role of miRNAs in the function of immune cells and the correlation of specific pathological conditions with miRNA profiles.
+Several studies have proved the activation of specific miRNA types in PBMCs under cardiovascular events  [@21; @186; @32-2; @29aa; @142; @155-3; @197].
+For instance, there is evidence that the elevation of miR-186 suppresses the expression of Cystathionine-$\gamma$-lyase, leading to the subsequent secretion of pro-inflammatory cytokines and cellular lipid accumulation.
+Besides, macrophage-derived miR-186 may promote atherosclerotic plaques [@186].
+In line with this claim, we found that miR-186 is up-regulated in both CAD and MI candidates related to control counterparts.
+Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients in comparison to MI (Figure \@ref(Fig:DEMexp)), To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, indicating its main role in the promotion of atherosclerosis.
+
+As mentioned before, miR-21 was also up-regulated in both MI and CAD patients in comparison to healthy controls.
+Moreover, the expression value of miR-21 was significantly higher in MI than that of the CAD group (Table \@ref(tab:mirExptable)).
+It is thought that the up-regulation of miRNA-21 in PBMCs is a compensatory reaction to reduce T$_{reg}$ lymphocyte number in response to the reduction of TGF$\beta1$ secretion into the plasma through a TGF$\beta1$/smad-independent pathway.
+In line with previous and present data, miR-21 can modulate the activity of PBMCs following the occurrence of cardiovascular diseases [@21].
+
+Recent data have supported the elevation of miR-32 in CAD patients with the calcification of coronary artery.
+Of note, miR-32 stimulates the calcification of mouse vascular smooth muscle through the regulation of bone morphogenetic protein-1, runt-related transcription factor-2 (RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA protein [@32-2].
+Likewise, there are some reports associated with the activity of PBMC miR-32 under several pathologies [@magic; @32].
+The exact role of PBMC miR-32 after cardiovascular events remained to be elucidated.
+
+Molecular analyses have indicated the regulatory role of miRNAs selected by the AUC approach in PBMCs after a cardiovascular event.
+As the only common miRNA of the two sets, molecular function of miR-21 in CVDs was covered in DEGs miRNA set.
+Based on numerous reports mir-29a can be activated in different diseases  [@29aa].Data analysis indicated that miR-29a is significantly up-regulated in CAD patients in comparison to healthy and MI groups (Table \@ref(tab:mirExptable)).
+Increased miR-29a is associated with the progression of atherosclerosis, and the combination of miR-29a and ox-LDL was offered as a valid biomarker set for paraclinical classification [@29a].
+However, the role of miR-29a in the function of PBMCs in CAD patients has not been completely examined.
+
+According to the present data (Table \@ref(tab:mirExptable)), the difference in miR-142 expression between MI/Healthy and CAD/Healthy was not significant, but these values reached statistically significant in MI as compared to CAD.
+Based on different reports, miR-142 is commonly active in PBMCs.
+This miRNA directly targets and inhibits the expression of Adenomatous Polyposis Coli, a negative WNT signaling pathway regulator, contributing to the activation of the WNT signaling pathway and cardiac fibroblast activation after myocardial ischemia [@142].
+
+We also noted that miR-155 is the only down-regulated miRNA among all miRNAs in both CAD/healthy and MI/Healthy comparisons (Table \@ref(tab:mirExptable)).
+There are numerous reports on the down-regulation of miR-155 in PBMCs after MI or in CAD patients.
+Previous data suggested an inverse relationship between miR-155 levels and the severity of the coronary artery condition [@155-2; @155-3].
+Along with current data and previous findings, it is postulated that miR-155 has a protective effect under pathological conditions.
+This genetic element can control inflammation and thus reduce tissue damage through its negative feedback effects on inflammatory factors.
+These features coincide with the reduction of atherosclerosis [@155-2].
+
+Data indicated that miR-197 is also significantly up-regulated in both CAD/healthy and MI/healthy groups.
+In previous studies, it has been shown that miRNA-197 may play a critical role in regulating the anti-inflammatory response of IL-35 by affecting pro/ anti-inflammatory cytokine secretion, M1/M2 macrophage ratio, T$_{reg}$ lymphocyte proliferation, and T cell suppression suggesting the potential diagnostic role of miR-197 in adverse cardiovascular events [@197].
+There are some reports about miR-320 role in the physiopathology of cardiac fibrosis.
+Mechanistically, Molecular analyses revealed that miR-320 might induce various effects by targeting PLEKHM3 and IFITM1 in cardiomyocytes and cardiac fibroblasts, respectively [@320].
+However, there are no reports on miR-320c1 activity in the cardiovascular system or PBMCs.
+
+# References

### 2023-06-02 Complete-V1/Complete-V1.Rmd
The Frist vesrion befor Dr. Kavousi Revision

--- 2023-04-25 ArticleDraft.Rmd
@@
 ---
-title: "The First Draft"
+title: Machine Learning Deriven Set of microRNAs as a Novel Biomarker for Myocardial Infarction Diagnosis
+#titlerunning: Short form of title (if too long for head)
+authorrunning: M. Samadishadlou et al.
+#thanks: | 
+#    Grants or other notes about the article that should go on the front 
+#    page should be placed here. General acknowledgments should be placed at the
+#    end of the article.
+
 authors: 
-  - name: "Mehrdad Samadishadlou"
-    address: "Medical Nanotechnology gruop, Tabriz University of Medical Sciences, Tabriz, Iran"
-date: "`r Sys.Date()`"
-link-citation: yes
-bibliography: [For Article Refrence.bib, discussion.bib, Introduction.bib]
+- name: Mehrdad Samadishadlou
+  address: Medical Nanotechnology gruop, Tabriz University of Medical Sciences, Tabriz, Iran
+  email: samadishadlou@tbzmed.ac.ir
+
+- name: Reza Rahbarghazi
+  address: Medical Nanotechnology gruop, Tabriz University of Medical Sciences, Tabriz, Iran
+  email: rahbarghazir@tbzmed.ac.ir
+
+- name: Farhad Bani
+  address: Medical Nanotechnology gruop, Tabriz University of Medical Sciences, Tabriz, Iran
+  email: bainf@tbzmed.ac.ir
+
+keywords:
+- microRNA
+- Machine Learning
+- Myocardial Infarction
+
+#PACS: 
+#- PAC1
+#- superPAC
+
+#MSC:
+#- MSC code 1
+#- MSC code 2
+
+always_allow_html: true
+
+abstract: |
+  MicroRNAs (miRNAs) play a crucial role in regulating adaptive and maladaptive responses in cardiovascular diseases, making them attractive targets for potential therapeutics.
+However, their potential as novel biomarkers for diagnosing cardiovascular diseases needs to be evaluated systematically.
+In this study, we aimed to identify a key set of miRNA biomarkers using integrated bioinformatics and machine learning analysis.
+We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contained peripheral blood mononuclear cells samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy.
+Additionally, we selected a set of miRNAs based on their area under the curve of the receiver operating characteristic (AUC-ROC) for separating CAD/MI samples.
+We designed a two-layer architecture for sample classification, with the first layer isolating healthy samples from not-healthy ones, and the second layer classifying stable CAD and MI samples.
+We trained different machine learning models using both biomarker sets and evaluated their performance on a test set.
+We identified miR-21, miR-186, and miR-32 as the only miRNAs in the differentially expressed genes, and a set included miR-21, miR-155, miR-142, miR-197, miR-29A, and miR-320C1 as the optimum set of miRNA selected by their AUC-ROC.
+Both biomarker sets were able to distinguish healthy from not-healthy samples with complete accuracy.
+The best performance for classification of CAD and MI was achieved with an SVM model trained using the biomarker set selected by AUC-ROC, with an AUC of 0.95 and an accuracy of 0.88.
+Our study demonstrates that miRNA signatures derived from peripheral blood could serve as valuable novel biomarkers for cardiovascular diseases.
+
 output:
-  bookdown::word_document2: default
-  bookdown::pdf_book:
-#    base_format: rticles::bioinformatics_article
-always_allow_html: true
+  rticles::springer_article
+
+bibliography: [ForArticleRefrence.bib, discussion.bib, Introduction.bib]
+biblio-style: spmpsci
+# bibstyle options spbasic(default), spphys, spmpsci
 ---
@@
 
-```{r, datasets, echo=FALSE}
+```{r datasets, echo=FALSE}
 df <- data.frame(Dataset = c("GSE59867", "GSE56609", "GSE54475"),
@@
 
-The PCA plots of the training samples are shown in Figure \@ref(Fig:PCA)A and B.
+The PCA plots of the training samples are shown in Figure \@ref(fig:PCA)A and B.
 As shown, healthy samples were separated from CAD or MI samples in primary data and also after conducting fRMA.
-In the RLE plot, there was a distinct difference between dataset means for all samples before conducting fRMA  (Figure \@ref(Fig:PCA)C).
-All datasets were rearranged around 0 in the RLE plot after conducting fRMA (Figure \@ref(Fig:PCA)D).
+In the RLE plot, there was a distinct difference between dataset means for all samples before conducting fRMA (Figure \@ref(fig:PCA)C).
+All datasets were rearranged around 0 in the RLE plot after conducting fRMA (Figure \@ref(fig:PCA)D).
 Moreover, there was a clear change in inter-quantile distances, but the values still were over 0.1.
 
-```{r PCA, echo=F, out.width="100%", Figure .cap="PCA and RLE plot for all samples before and after fRMA."}
+```{r PCA, echo=F, out.width="100%", fig.cap="PCA and RLE plot for all samples before and after fRMA."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/PCA and RLE of trains.pdf")
@@
 
-```{r , DEGstab, echo=F}
+```{r DEGstab, echo=F}
 M_H <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-Healthy.txt", check.names = F)
@@
 Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in MI compared to the healthy controls.
-In CAD and healthy groups comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and and `r sum(C_H$logFC < 0)` DEGs were up- and `r down-regulated, respectively in CAD samples.
+In CAD and healthy groups comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and and `r sum(C_H$logFC < 0)` DEGs were up- and down-regulated, respectively in CAD samples.
 In the MI and CAD groups, the number of DEGs was `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively in comparison with CAD samples.
@@
 
-```{r venn, echo=F, out.width="60%", Figure .align='center', Figure .cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
+```{r venn, echo=F, out.width="60%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/venny.png")
@@
 
-The Venn diagram  in Figure \@ref(Fig:venn) shows that CAD and MI samples share a majority of their DEGs.
+The Venn diagram in Figure \@ref(fig:venn) shows that CAD and MI samples share a majority of their DEGs.
 From 860 DEGs of MI/healthy and 670 DEGs of CAD/healthy, 531 genes were common which is 62% of MI/healthy DEGs and 79% of CAD/healthy DEGs.
@@
 To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
-In MI versus healthy samples, GO enrichment analysis in the BP category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Figure \@ref(Fig:MIHEnrich)A).
-In the CC category, the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Figure \@ref(Fig:MIHEnrich)B).
-In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(Fig:MIHEnrich)C).
-KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(Fig:MIHEnrich)D).
+In MI versus healthy samples, GO enrichment analysis in the BP category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Figure \@ref(fig:MIHEnrich)A).
+In the CC category, the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Figure \@ref(fig:MIHEnrich)B).
+In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(fig:MIHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(fig:MIHEnrich)D).
 
-```{r MIHEnrich, echo=F, out.width="95%", Figure .cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs.
+```{r MIHEnrich, echo=F, out.width="95%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs.
 (A) Biological process terms.
@@
 The enrichment results for DEGs of CAD versus healthy samples were as follows.
-In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Figure \@ref(Fig:CADHEnrich)A).
-In the CC category, the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Figure \@ref(Fig:CADHEnrich)B).
-In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(Fig:CADHEnrich)C).
-KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(Fig:CADHEnrich)D).
+In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Figure \@ref(fig:CADHEnrich)A).
+In the CC category, the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Figure \@ref(fig:CADHEnrich)B).
+In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(fig:CADHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(fig:CADHEnrich)D).
 
-```{r CADHEnrich, echo=F, out.width="95%", Figure .cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs.
+```{r CADHEnrich, echo=F, out.width="95%", fig.cap="Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs.
 (A) Biological process terms.
@@
 Among all DEGs, just miR-186, miR-32, and miR-21 were detected as differentially expressed miRNAs.
-The expression profile of these three miRNAs is presented in Figure \@ref(Fig:DEMexp).
-Additionally, The ROC curves of each miRNA for each layer are presented in Figure \@ref(Fig:miRROC).
-Using the logistic regression model the AUC for miR-21, miR-32, and miR-186 was 0.99, 1, and 0.91 respectively  (Figure \@ref(Fig:miRROC)A).
+The expression profile of these three miRNAs is presented in Figure \@ref(fig:DEMexp).
+Additionally, The ROC curves of each miRNA for each layer are presented in Figure \@ref(fig:miRROC).
+Using the logistic regression model the AUC for miR-21, miR-32, and miR-186 was 0.99, 1, and 0.91 respectively (Figure \@ref(fig:miRROC)A).
 Besides, the accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set was 0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
-Moreover, the ROC curve of each miRNA for classifying MI and CAD samples was presented in Figure \@ref(Fig:miRROC)B.
+Moreover, the ROC curve of each miRNA for classifying MI and CAD samples was presented in Figure \@ref(fig:miRROC)B.
 For miR-21, miR-32, and miR-186, the AUC and accuracy on the test set were 0.85; 0.7; and 0.82, and 0.78; 0.67; and 0.74, respectively.
@@
 
-```{r DEMexp, echo=F, out.width="50%", Figure .align='center', Figure .cap="Expression profile of miR-186, miR-21, and miR32 in Healthy, CAD, and MI samples."}
+```{r DEMexp, echo=F, out.width="50%", fig.align='center', fig.cap="Expression profile of miR-186, miR-21, and miR32 in Healthy, CAD, and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEGs healthy not-healthy/DEMs Expression.pdf")
@@
 
-```{r miRROC, echo=F, out.width="95%", Figure .align='center', Figure .cap="ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+```{r miRROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/miRs ROCs.pdf")
@@
 Although single miRNAs had acceptable performance, their predictive value could be improved even further by using them as a set.
-The ROC curve for the SVM model with an RBF kernel trained with all three miRNAs is presented in Figure \@ref(Fig:DEMROC)A.
+The ROC curve for the SVM model with an RBF kernel trained with all three miRNAs is presented in Figure \@ref(fig:DEMROC)A.
 The model had a better performance in classification than single miRNAs.The AUC for the model is 1, and its accuracy on the test set was also 1.
-In Figure \@ref(Fig:DEMCF)A, the confusion matrix for the model is presented.
+In Figure \@ref(fig:DEMCF)A, the confusion matrix for the model is presented.
 
-```{r DEMROC, echo=F, out.width="95%", Figure .align='center', Figure .cap="ROC curve for miRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for miRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs ROCs h not h cad mi.pdf")
@@
 
-```{r DEMCF, echo=F, out.width="90%", Figure .align='center', Figure .cap="Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
+```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs Conf. matrix.pdf")
@@
 Different models were trained using expression values for three differentially miRNAs.
-The models' 10-fold cross-validated AUC and accuracy on the test set are reported in Figure \@ref(Fig:DEMmodels).
+The models' 10-fold cross-validated AUC and accuracy on the test set are reported in Figure \@ref(fig:DEMmodels).
 The best model from both AUC and accuracy point-of-view was the SVM model with linear kernel.
@@
 The model was hyper-tuned for C and gamma hyper-parameters, and therefore the model showed better performance.
-The ROC curve of the hyper-tuned model is presented in Figure \@ref(Fig:DEMROC)B.
+The ROC curve of the hyper-tuned model is presented in Figure \@ref(fig:DEMROC)B.
 For this model the AUC reached 0.95 and the accuracy improved to 0.85 (Table \@ref(tab:DEGsML)).
 Moreover, the sensitivity and specificity for the model on the test set were 0.91 and 0.71 respectively.
-The confusion matrix for the hyper-tuned model is illustrated in Figure \@ref(Fig:DEMCF)B.
+The confusion matrix for the hyper-tuned model is illustrated in Figure \@ref(fig:DEMCF)B.
 
-```{r DEMmodels, echo=F, out.width="65%", Figure .align='center', Figure .cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
+```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models DEGs mirs.pdf")
@@
 The metric of choice for selecting the best set was AUC.
-As shown in Figure \@ref(Fig:AUCset), the AUC increased until the number of miRNAs in the set reached six, and after that, it dropped.
+As shown in Figure \@ref(fig:AUCset), the AUC increased until the number of miRNAs in the set reached six, and after that, it dropped.
 The AUC for separating MI samples from CAD using these miRNAs was 0.93.
 The miRNAs in the set were miR-29a, miR-197, miR-142, miR-21, miR-155, and miR-320C1.
-The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(Fig:AUCexp).
-The ROC curve of the selected miRNAs for MI and CAD sample classification is illustrated in Figure \@ref(Fig:miRROC)B.
+The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:AUCexp).
+The ROC curve of the selected miRNAs for MI and CAD sample classification is illustrated in Figure \@ref(fig:miRROC)B.
 
-```{r AUCset, echo=F, out.width="65%", Figure .align='center', Figure .cap="Area under the curve (AUC) for sets containing an increasing number of miRNAs with the highest individual AUC in MI/CAD separation."}
+```{r AUCset, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the curve (AUC) for sets containing an increasing number of miRNAs with the highest individual AUC in MI/CAD separation."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC for cad mi mirs.pdf")
@@
 
-```{r AUCexp, echo=F, out.width="90%", Figure .align='center', Figure .cap="Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples."}
+```{r AUCexp, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/ExpressionforAUC.pdf")
@@
 Using the selected set, an SVM model with an RBF kernel was trained to separate healthy from not-healthy samples.
-The ROC curve for the model is presented in Figure \@ref(Fig:AUCROC)A and the confusion matrix is illustrated in Figure \@ref(Fig:AUCCF)A.
+The ROC curve for the model is presented in Figure \@ref(fig:AUCROC)A and the confusion matrix is illustrated in Figure \@ref(fig:AUCCF)A.
 Both AUC and accuracy for the model on the test set were equal to 1.
 
-```{r AUCROC, echo=F, out.width="95%", Figure .align='center', Figure .cap="ROC curve for the set of miRNAs selected by AUC on test set classification.
+```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of miRNAs selected by AUC on test set classification.
 (A) SVM with RBF kernel for healthy and not-healthy samples classification.
@@
 
-```{r AUCCF, echo=F, out.width="90%", Figure .align='center', Figure .cap="Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
 (B) Logistic regression model for CAD and MI samples classification.
@@
 To find the best model for training the best set, different models were trained using their pre-set values.
-Their AUC and accuracy results on the test set are presented in Figure \@ref(Fig:AUCmodels).
+Their AUC and accuracy results on the test set are presented in Figure \@ref(fig:AUCmodels).
 The best model from the AUC point-of-view was the LR and from the accuracy point-of-view, it was the SVM model with a polynomial kernel.
 For the LR model the AUC and accuracy were 0.92 and 0.81, respectively; and for the SVM model with a polynomial kernel, the values were 0.91 and 0.84, respectively.
-Both models were hyper-tuned and the ROC curve for their best performance is presented in Figure \@ref(Fig:AUCROC)B and C.
+Both models were hyper-tuned and the ROC curve for their best performance is presented in Figure \@ref(fig:AUCROC)B and C.
 The AUC and accuracy for the LR model increased to 0.94 and 0.88, respectively.
@@
 The sensitivity for the LR and SVM models were 1 and 0.97, respectively; and the specificity for them was 0.57 and 0.64, respectively.
-The confusion matrix for both models is illustrated in Figure \@ref(Fig:AUCCF)B and C.
+The confusion matrix for both models is illustrated in Figure \@ref(fig:AUCCF)B and C.
 
-```{r AUCmodels, echo=F, out.width="65%", Figure .align='center', Figure .cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
+```{r AUCmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models AUC mirs.pdf")
@@
 Having these samples set alongside bioinformatics analysis and ML means, it is possible to identify potential biomarker sets and also effective therapeutic targets.
-The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(Fig:venn)) are proof of the close relationship between the MI and CAD samples.
+The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(fig:venn)) are proof of the close relationship between the MI and CAD samples.
 Interestingly, functional enrichment analysis demonstrated that DEGs in both healthy/CAD and healthy/MI were strongly correlated to immune cell response which is a major cellular part of PBMCs.
@@
 miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-21; miR-29a; miR-142; miR155; miR-197; and miR-320c1 were selected according to their AUC values.
-As shown in Figure \@ref(Fig:miRROC), all other miRNAs selected with both approaches had AUC-ROC over 0.9 for the isolation of healthy and not-healthy samples except for miR-142 and miR-29a.
+As shown in Figure \@ref(fig:miRROC), all other miRNAs selected with both approaches had AUC-ROC over 0.9 for the isolation of healthy and not-healthy samples except for miR-142 and miR-29a.
 Data confirmed that the real challenge is to classify CAD and MI samples because of close overlap.
@@
 In line with this claim, we found that miR-186 is up-regulated in both CAD and MI candidates related to control counterparts.
-Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients in comparison to MI (Figure \@ref(Fig:DEMexp)), To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, indicating its main role in the promotion of atherosclerosis.
+Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients in comparison to MI (Figure \@ref(fig:DEMexp)), To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, indicating its main role in the promotion of atherosclerosis.
 
@@
 
+# Conclusion
+
+In summary, we derived a set of miRNA biomarkers by comparing MI samples to both healthy and CAD samples.
+We found that the SVM model performed best in both the first layer, which separated healthy and not-healthy samples, and the second layer, which classified MI/CAD samples.
+The set of miRNAs selected based on their AUC values had slightly better performance in the second layer.
+Overall, our two-layer structure achieved a weighted accuracy of 0.91.
+This demonstrates the potential for combining bioinformatics and machine learning techniques to identify novel biomarkers and gain a deeper understanding of myocardial infarction.
+
 # References
+
+\begin{align}
+a^2+b^2=c^2
+\end{align}

### 2023-06-02 Complete-V1/Complete-V1.tex
The Frist vesrion befor Dr. Kavousi Revision

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}

\title{Machine Learning Deriven Set of microRNAs as a Novel Biomarker
for Myocardial Infarction Diagnosis }

\author{  Mehrdad Samadishadlou \and  Reza Rahbarghazi \and  Farhad
Bani \and  }

    \authorrunning{ M. Samadishadlou et al. }

\institute{
        Mehrdad Samadishadlou \at
     Medical Nanotechnology gruop, Tabriz University of Medical
Sciences, Tabriz, Iran \\
     \email{\href{mailto:samadishadlou@tbzmed.ac.ir}{\nolinkurl{samadishadlou@tbzmed.ac.ir}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Reza Rahbarghazi \at
     Medical Nanotechnology gruop, Tabriz University of Medical
Sciences, Tabriz, Iran \\
     \email{\href{mailto:rahbarghazir@tbzmed.ac.ir}{\nolinkurl{rahbarghazir@tbzmed.ac.ir}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Farhad Bani \at
     Medical Nanotechnology gruop, Tabriz University of Medical
Sciences, Tabriz, Iran \\
     \email{\href{mailto:bainf@tbzmed.ac.ir}{\nolinkurl{bainf@tbzmed.ac.ir}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
MicroRNAs (miRNAs) play a crucial role in regulating adaptive and
maladaptive responses in cardiovascular diseases, making them attractive
targets for potential therapeutics. However, their potential as novel
biomarkers for diagnosing cardiovascular diseases needs to be evaluated
systematically. In this study, we aimed to identify a key set of miRNA
biomarkers using integrated bioinformatics and machine learning
analysis. We combined and analyzed three gene expression datasets from
the Gene Expression Omnibus (GEO) database, which contained peripheral
blood mononuclear cells samples from individuals with myocardial
infarction (MI), stable coronary artery disease (CAD), and healthy.
Additionally, we selected a set of miRNAs based on their area under the
curve of the receiver operating characteristic (AUC-ROC) for separating
CAD/MI samples. We designed a two-layer architecture for sample
classification, with the first layer isolating healthy samples from
not-healthy ones, and the second layer classifying stable CAD and MI
samples. We trained different machine learning models using both
biomarker sets and evaluated their performance on a test set. We
identified miR-21, miR-186, and miR-32 as the only miRNAs in the
differentially expressed genes, and a set included miR-21, miR-155,
miR-142, miR-197, miR-29A, and miR-320C1 as the optimum set of miRNA
selected by their AUC-ROC. Both biomarker sets were able to distinguish
healthy from not-healthy samples with complete accuracy. The best
performance for classification of CAD and MI was achieved with an SVM
model trained using the biomarker set selected by AUC-ROC, with an AUC
of 0.95 and an accuracy of 0.88. Our study demonstrates that miRNA
signatures derived from peripheral blood could serve as valuable novel
biomarkers for cardiovascular diseases.
\\
\keywords{
        microRNA \and
        Machine Learning \and
        Myocardial Infarction \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

At present, cardiovascular diseases (CVDs) are the leading cause of
human mortality with 32\% of all global deaths. It is estimated that
about 85\% of CVDs mortality were diagnosed with myocardial infarction
(MI) ({``Cardiovascular Diseases ({CVDs})''} n.d.). MI is an acute
coronary syndrome with sudden blockage and stenosis of the coronary
artery, and subsequent myocardial ischemia, leading to extensive
cardiomyocyte damage and necrosis (Yap et al. 2023).

Over the last 50 years, numerous attempts have been collected to use
biomarkers to facilitate diagnosis, assess risk, follow-up therapy, and
determine therapeutic efficacy in CVDs candidates. Based on the released
guidelines, cardiac troponins (cTns) are used as a highly-sensitive and
accurate approach for the detection of myocardial ischemia. Despite the
inherent advantages, the high-rate sensitivity of cTn-based assays has
also led to more false positive results (Thygesen et al. 2018), which do
necessitate the advent and development of new modalities with
pathological values. To improve diagnostic value upon existing CVD
biomarkers, the combination of complementary biological markers, such as
microRNAs (miRNAs) and other genetic factors, is proposed. Previous data
support the notion that miRNAs exhibit the great potential to be used as
alternative biomarkers in CVD detection and follow-up (Schulte et al.
2020). It is suggested that miRNAs possess 18-22 nucleotides and can
play a crucial role in the regulation of gene expression. Evidence point
to the fact that miRNAs are involved in the pathogenesis of cardiac
tissue injury and be as theranostics in terms of CVDs. These elements
can easily circulate in biofluids with both diagnostic as well as
prognostic values (Schulte, Karakas, and Zeller 2017). Several
biological activities such as angiogenesis, cardiomyocyte growth and
contractility, lipid metabolism, plaque formation, and cardiac rhythm
are regulated by miRNAs (Kalayinia et al. 2021). It is postulated that
the function and diagnostic properties of miRNAs are beyond the
myocardium in CVD patients. To be specific, the expression of miRNAs can
vary in different biofluids and cell components such as serum and
peripheral blood mononuclear cells (PBMCs) (Soler-Botija, Gálvez-Montón,
and Bayés-Genís 2019).

PBMCs are a fraction of white blood cells (WBCs), including monocytes,
lymphocytes, macrophages, and other cells belonging to the immune system
(Gao et al. 2020). Emerging data have indicated that PBMCs can be used
as a valid source of biomarkers for monitoring various pathological
conditions. Of note, the alteration of mRNAs and miRNAs under
pathological conditions gives us valuable information about different
kinds of disorders. PBMCs could recapitulate the conditions of the
target tissues, thus, providing a highly sensitive and specific source
of biomarkers (Mosallaei et al. 2022). Commensurate with these
conditions, these cells are repositories of dysregulated genes and
miRNAs expression profiles in CVDs related to control conditions (Gao et
al. 2020; Mosallaei et al. 2022).

In recent years, the advent and use of machine learning (ML) is an
exciting prospect for advancing scientific discoveries. Although the
concept of ML and its initial algorithms were conceived many years ago,
recent improvements in computing power and access to vast amounts of
data have shown that ML techniques outperform classical statistical
methods in various fields. Furthermore, the progress made in omics
technologies has enabled the analysis of massive and intricate
biological data sets, consisting of hundreds to thousands of samples,
which makes it possible for ML to extract valuable biological
information from such data (Torun et al. 2023). Therefore, ML offers
novel techniques to integrate and analyze the various omics data
enabling the discovery of de novo biomarkers. These biomarkers help us
in accurate disease prediction, patient stratification, and finding new
therapeutics (Reel et al. 2021).

In this study, we aimed to identify potential miRNA biomarkers for MI
patients by combining and analyzing three different microarray datasets
from PBMCs. It is suggested that the integration of omics data with
bioinformatics and ML techniques could be a promising tool in the
discovery of new and more accurate biomarkers for monitoring CVDs.
Besides, this approach can deepen our vision into the underlying
mechanisms of CVDs and aid in the development of valid theranostic
tools, and patient stratification.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{microarray-data-collection}{%
\subsection{Microarray data
collection}\label{microarray-data-collection}}

Microarray datasets were obtained from the Gene Expression Omnibus (GEO)
database (\url{https://www.ncbi.nlm.nih.gov/geo/}). To obtain sufficient
classification power between MI, healthy and CAD samples, a relatively
large sample size was required. Therefore, GSE59867 for MI and CAD
samples, and GSE56609 and GSE54475 for healthy samples were selected.
All samples were produced using Affymetrix Human Gene 1.0 ST Array
(GPL6244) platform. Only healthy, stable CAD and early-stage MI samples
were selected from these datasets for further analyses. The basic
information for the three datasets evaluated in the current study is
provided in Table @ref(tab:datasets).

\begin{longtable}[]{@{}cccccc@{}}
\caption{Basic information of the GEO microarray
datasets.}\tabularnewline
\toprule()
Dataset & Platform & Healthy & CAD & MI & Refrence \\
\midrule()
\endfirsthead
\toprule()
Dataset & Platform & Healthy & CAD & MI & Refrence \\
\midrule()
\endhead
GSE59867 & GPL6244 & - & 46 & 111 & (Maciejak et al. 2015) \\
GSE56609 & GPL6244 & 46 & - & - & (Matone et al. 2015) \\
GSE54475 & GPL6244 & 5 & - & - & (Canali et al. 2014) \\
\bottomrule()
\end{longtable}

\hypertarget{pre-processing}{%
\subsection{Pre-processing}\label{pre-processing}}

Raw data (CEL files) of all datasets were downloaded from the GEO and
pre-processed using the fRMA package (M. N. McCall, Bolstad, and
Irizarry 2010). fRMA allowed to pre-process of individual microarray
samples and combining them consistently for analysis. For each dataset,
background correction was performed using the RMA algorithm and then it
was quantile normalized based on the reference distribution. During
summarization, batch effects were removed and variances of the gene
expressions were estimated by taking into account these probe-specific
effects. For those multiple probe sets matched to the identical gene,
the mean log fold change was retained. Therefore, fRMA can be seen as a
batch effect removal technique for different datasets that are produced
by identical microarray platforms. Thus, to ensure batch effect removal,
the principal component analysis and the relative log expression of
train samples were plotted before and after fRMA (Lazar et al. 2013).

\hypertarget{differential-expression-analysis}{%
\subsection{Differential expression
analysis}\label{differential-expression-analysis}}

The barcode algorithm proposed by McCall et al. (Matthew N. McCall et
al. 2011) transformed the actual expression values into binary barcode
values. Huge sets of samples were collected and normalized using fRMA
for several platforms as well as for Affymetrix Human Gene 1.0 ST Array
(GPL6244) platform. The distribution of the expressed and unexpressed
observed intensities for each gene is estimated using these normalized
sets. Genes were considered expressed (and their value coded to 1) or
unexpressed (and their value coded to 0) according to the following
equation:

\[
\hat{x}_{ij} = \left\{
  \begin{array}{ll}
    1 & \mbox{if } x_{ij} >= \mu^{ne} + C \times \sigma^{ne} \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\]

where \(x_{ij}\) is the normalized intensity of gene \(i\) in sample
\(j\), \(C\) is a user-defined parameter, \(\sigma^{ne}\) is the
standard deviation of the non-expressed distribution, and \(\mu^{ne}\)
is the mean of the non-expressed distribution. The barcode
representation of a sample is a vector of ones and zeros denoting which
genes are estimated to be expressed (ones) and unexpressed (zeros). The
barcode algorithm was implemented by the barcode function in the R fRMA
package, and the default value of \(C\) was used.

To determine if the expressed ratios differed in the MI group versus the
healthy control group, Fisher's exact test for individual genes was
carried out upon the barcode values. Genes with a false discovery rate
(FDR) of \(< 0.05\), which was calculated through the Benjamini-Hochberg
(BH) procedure to adjust for multiple testing issues, were considered as
differentially expressed genes. The same procedures were conducted on
CAD versus healthy controls as well as MI versus CAD group to find the
DEGs between them.

\hypertarget{functional-and-pathway-enrichment-analyses}{%
\subsection{Functional and pathway enrichment
analyses}\label{functional-and-pathway-enrichment-analyses}}

Using the R clusterProfiler package, the Kyoto Encyclopedia of Genes and
Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO)
functional annotation were carried out on the differentially expressed
genes. The GO analysis included biological process (BP), cellular
component (CC) and molecular function (MF) categories. An adjusted
p-value of less than 0.05 was considered to indicate a statistically
significant difference. Enrichments were conducted on the MI-healthy and
CAD-healthy DEGs. In these analyses, all default parameters were used.

\hypertarget{ml-procedure}{%
\subsection{ML procedure}\label{ml-procedure}}

The ML analysis was performed using Python software, ver. 3.9, Numpy
(Harris et al. 2020), pandas (McKinney 2010), and Scikit-Learn packages
(Pedregosa et al. 2011). Whenever hyper-tuning was needed, the
scikit-opt package (Head et al. 2021) was used. In all ML analyses, the
datasets were divided into train and test sets by a 0.7:0.3 ratio and
all reported results are the average of 10-fold cross-validation.

Two different approaches were used for selecting miRNAs for model
training. The first approach was using the miRNAs that are
differentially expressed. In the second approach, miRNAs were selected
by their individual AUC-ROC. Having the result of these two different
approaches can provide an informative comparison between the predictive
capabilities of sets of miRNAs selected with different logics.

\hypertarget{mirnas-in-degs}{%
\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs}}

In this approach, a two-layer architecture was deployed to the data to
maximize the prediction values. The first layer predicted whether a
sample is healthy or not, and the second layer separated MI from CAD in
the samples which were predicted as not healthy in the first layer. To
this end, a distinct ML model was trained for each layer. Since there is
a limited number of miRNAs in DEGs, both layers were trained with all of
them. For further comparison with the models' performance, the ROC curve
of each miRNA for classifying healthy and not-healthy, as well as CAD
and MI, were generated using a Logistic Regression model.

\hypertarget{first-layer-for-isolation-of-healthy-and-not-healthy-samples}{%
\paragraph{First layer for isolation of healthy and not-healthy
samples:}\label{first-layer-for-isolation-of-healthy-and-not-healthy-samples}}

An SVM model using RBF kernels was trained and hyper-tuned using all
miRNAs in DEGs. To handle the severe imbalance in the number of samples
(51 for the healthy and 157 for the not-healthy groups), the sample
weight for the healthy and the not-healthy samples were set to 1 and
0.5, respectively.

\hypertarget{second-layer-for-seperating-mi-and-cad-samples}{%
\paragraph{Second layer for seperating MI and CAD
samples:}\label{second-layer-for-seperating-mi-and-cad-samples}}

For the sake of reaching the highest classification performance using
the set of miRNAs, different models were investigated. To do so, SVM
(with linear, polynomial, and RBF kernels), Logistic Regression (LR),
Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB),
XGBoost (XGB) and Decision Tree (DT) models were trained. All models
were trained with their pre-set parameters with 10-fold
cross-validation. The criteria for choosing the best model were the
highest accuracy and AUC on the test set. The best model was hyper-tuned
with the scikit-opt package (Head et al. 2021) to get the best
predictive performance.

\hypertarget{mirnas-with-the-highest-auc-roc}{%
\subsubsection{miRNAs with the highest
AUC-ROC}\label{mirnas-with-the-highest-auc-roc}}

Like the previous approach, a two-layers strategy was conducted. The
first layer classified samples into healthy and not-healthy, and the
separated MI and CAD samples. However, to keep the number of miRNAs as
low as possible miRNAs were selected from the second layer (which are
the miRNAs with the best performance in MI/CAD separation), and then
their performance wase evaluated in the first layer. AUC-ROC of all
miRNAs for classifying MI and CAD samples were calculated. To find the
number of miRNAs with the highest predictive values, the miRNAs with the
highest individual AUC-ROC were added to the set one by one, and the
AUC-ROC for the set was calculated. The set with the highest AUC-ROC was
selected for the following steps. The ROC curves for each selected miRNA
for separating healthy samples from not-healthy and MIs from CADs were
also plotted for further comparison.

\hypertarget{first-layer-for-the-detection-of-healthy-and-not-healthy-samples}{%
\paragraph{First layer for the detection of healthy and not-healthy
samples:}\label{first-layer-for-the-detection-of-healthy-and-not-healthy-samples}}

An SVM model with an RBF kernel was trained using the selected set of
miRNAs. Additionally, the model was hyper-tuned to find the
hyper-parameters for the highest AUC and accuracy. The ROC curve and
confusion matrix for the best model were reported.

\hypertarget{second-layer-for-separating-mi-and-cad}{%
\paragraph{Second layer for separating MI and
CAD:}\label{second-layer-for-separating-mi-and-cad}}

The selected miRNAs set was used to train different algorithms to find
the best model. Similar to the previous approach, SVM (with linear,
polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT were trained.
All models were trained with their pre-set parameters using 10-fold
cross-validation. The models with the highest AUC-ROC and accuracy on
the test set were selected and hyper-tuned using the scikit-opt package
(Head et al. 2021). The ROC curve and confusion matrix for the best
model were reported.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{pre-processing-1}{%
\subsection{Pre-processing}\label{pre-processing-1}}

The PCA plots of the training samples are shown in Figure @ref(fig:PCA)A
and B. As shown, healthy samples were separated from CAD or MI samples
in primary data and also after conducting fRMA. In the RLE plot, there
was a distinct difference between dataset means for all samples before
conducting fRMA (Figure @ref(fig:PCA)C). All datasets were rearranged
around 0 in the RLE plot after conducting fRMA (Figure @ref(fig:PCA)D).
Moreover, there was a clear change in inter-quantile distances, but the
values still were over 0.1.

\begin{figure}
\includegraphics[width=1\linewidth]{../../DEG and DEM/Pics/PCA and RLE of trains} \caption{PCA and RLE plot for all samples before and after fRMA.}\label{fig:PCA}
\end{figure}

\hypertarget{differential-expression-analysis-1}{%
\subsection{Differential expression
analysis}\label{differential-expression-analysis-1}}

\begin{table}

\caption{\label{tab:DEGstab}Total, up-, and down-regulated DEGs and differentially expressed miRNAs.}
\centering
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\centering}X>{\centering}X>{\centering}X>{\centering\arraybackslash}p{4cm}}
\toprule
  & DEGs & up-regulated DEGs & down-regulated DEGs & miRNAs\\
\midrule
MI vs. Healthy & 860 & 323 & 537 & hsa-miR-186, miR-21, miR-32\\
CAD vs. Healthy & 670 & 262 & 408 & hsa-miR-186, miR-21, miR-32\\
MI vs. CAD & 260 & 144 & 116 & hsa-miR-186\\
\bottomrule
\end{tabu}
\end{table}

According to the cutoff criterion of \(FDR < 0.05\), there were 860 DEGs
between the MI and the healthy samples. Among them, 323 were
up-regulated, and 537 were down-regulated in MI compared to the healthy
controls. In CAD and healthy groups comparison, we found 670 DEGs, of
which 262 and and 408 DEGs were up- and down-regulated, respectively in
CAD samples. In the MI and CAD groups, the number of DEGs was 260, and
the number of up- and down-regulated genes in MI samples were 144 and
116, respectively in comparison with CAD samples. These data are
summarized in Table @ref(tab:DEGstab).

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{../../DEG and DEM/Pics/venny}

}

\caption{Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison.}\label{fig:venn}
\end{figure}

The Venn diagram in Figure @ref(fig:venn) shows that CAD and MI samples
share a majority of their DEGs. From 860 DEGs of MI/healthy and 670 DEGs
of CAD/healthy, 531 genes were common which is 62\% of MI/healthy DEGs
and 79\% of CAD/healthy DEGs.

\hypertarget{gene-ontology-go-and-kyoto-encyclopedia-of-genes-and-genomes-kegg-enrichment-analyses-of-the-degs.}{%
\subsection{Gene ontology (GO) and Kyoto Encyclopedia of Genes and
Genomes (KEGG) enrichment analyses of the
DEGs.}\label{gene-ontology-go-and-kyoto-encyclopedia-of-genes-and-genomes-kegg-enrichment-analyses-of-the-degs.}}

To explore the biological classification of the DEGs, we performed GO
and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
In MI versus healthy samples, GO enrichment analysis in the BP category,
suggested that the DEGs were enriched in ``immune response-regulating
signaling pathway'', ``lymphocyte differentiation'', ``immune
response-regulating cell surface receptor signaling pathway'', and
``leukocyte activation involved in immune response'' (Figure
@ref(fig:MIHEnrich)A). In the CC category, the DEGs were enriched in
``secretory granule membrane'', ``azurophil granule'', ``ficolin-1-rich
granule'', ``tertiary granule'', and ``ficolin-1-rich granule membrane''
(Figure @ref(fig:MIHEnrich)B). In the MF category, the DEGs were
involved in ``cadherin binding'' and ``MHC class I protein binding''
(Figure @ref(fig:MIHEnrich)C). KEGG pathway analysis indicated that the
DEGs were related to the following pathways: ``Chemokine signaling
pathway'', ``Lipid and atherosclerosis'', and ``Hematopoietic cell
lineage'' (Figure @ref(fig:MIHEnrich)D).

\begin{figure}
\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/enrich MI-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}\label{fig:MIHEnrich}
\end{figure}

The enrichment results for DEGs of CAD versus healthy samples were as
follows. In the BP category, GO enrichment suggested that the DEGs were
enriched in ``positive regulation of defense response'', ``positive
regulation of innate immune response'', ``mononuclear cell
differentiation'', and ``positive regulation of response to external
stimulus'' (Figure @ref(fig:CADHEnrich)A). In the CC category, the DEGs
were enriched in ``azurophil granule'', ``ficolin-1-rich granule'', and
``ficolin-1-rich granule membrane'' (Figure @ref(fig:CADHEnrich)B). In
the MF category, the DEGs were involved in ``lipoprotein particle
receptor binding'' and ``NF-\(\kappa\)B binding'' (Figure
@ref(fig:CADHEnrich)C). KEGG pathway analysis indicated that the DEGs
were related to the following pathways: ``Chemokine signaling pathway'',
``Lipid and atherosclerosis'', and ``Hematopoietic cell lineage''
(Figure @ref(fig:CADHEnrich)D).

\begin{figure}
\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/enrich CAD-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}\label{fig:CADHEnrich}
\end{figure}

\hypertarget{machine-learning}{%
\subsection{Machine Learning}\label{machine-learning}}

\hypertarget{mirnas-in-degs-1}{%
\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs-1}}

Among all DEGs, just miR-186, miR-32, and miR-21 were detected as
differentially expressed miRNAs. The expression profile of these three
miRNAs is presented in Figure @ref(fig:DEMexp). Additionally, The ROC
curves of each miRNA for each layer are presented in Figure
@ref(fig:miRROC). Using the logistic regression model the AUC for
miR-21, miR-32, and miR-186 was 0.99, 1, and 0.91 respectively (Figure
@ref(fig:miRROC)A). Besides, the accuracy of each miRNA for classifying
the samples into healthy and not-healthy groups on the test set was
0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
Moreover, the ROC curve of each miRNA for classifying MI and CAD samples
was presented in Figure @ref(fig:miRROC)B. For miR-21, miR-32, and
miR-186, the AUC and accuracy on the test set were 0.85; 0.7; and 0.82,
and 0.78; 0.67; and 0.74, respectively.

\begin{table}

\caption{\label{tab:mirExptable}Target miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{CAD/Healthy} & \multicolumn{2}{c}{MI/Healthy} & \multicolumn{2}{c}{MI/CAD} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
  & logFC & adj. p-value & logFC & adj. p-value & logFC & adj. p-value\\
\midrule
miR-186 & 1.4 & 3.60e-25 & 0.9 & 6.76e-20 & -0.5 & 1.05e-09\\
miR-21 & 1.4 & 1.31e-17 & 2.3 & 2.07e-47 & 0.8 & 2.96e-11\\
miR-32 & 2.5 & 8.39e-43 & 2.2 & 3.10e-59 & -0.3 & 7.60e-04\\
miR-155 & -0.6 & 2.49e-12 & -0.9 & 7.59e-33 & -0.2 & 2.68e-07\\
miR-142 & 0.2 & 1.90e-01 & -0.1 & 1.70e-01 & -0.3 & 2.90e-04\\
miR-197 & 0.5 & 2.95e-20 & 0.7 & 1.59e-47 & 0.2 & 8.58e-09\\
miR-29A & 0.7 & 7.76e-29 & 0.1 & 1.70e-01 & -0.5 & 2.14e-10\\
miR-320C1 & 0.8 & 4.72e-22 & 0.5 & 1.89e-30 & -0.2 & 1.84e-05\\
\bottomrule
\end{tabu}
\end{table}

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{../../ML/Pics/DEGs healthy not-healthy/DEMs Expression}

}

\caption{Expression profile of miR-186, miR-21, and miR32 in Healthy, CAD, and MI samples.}\label{fig:DEMexp}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../../ML/Pics/miRs ROCs}

}

\caption{ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples.}\label{fig:miRROC}
\end{figure}

\hypertarget{first-layer-for-healthy-not-healthy-isolation}{%
\paragraph{First layer for healthy not-healthy
isolation:}\label{first-layer-for-healthy-not-healthy-isolation}}

Although single miRNAs had acceptable performance, their predictive
value could be improved even further by using them as a set. The ROC
curve for the SVM model with an RBF kernel trained with all three miRNAs
is presented in Figure @ref(fig:DEMROC)A. The model had a better
performance in classification than single miRNAs.The AUC for the model
is 1, and its accuracy on the test set was also 1. In Figure
@ref(fig:DEMCF)A, the confusion matrix for the model is presented.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../../ML/Pics/DEMs ROCs h not h cad mi}

}

\caption{ROC curve for miRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples.}\label{fig:DEMROC}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/DEMs Conf. matrix}

}

\caption{Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification.}\label{fig:DEMCF}
\end{figure}

\hypertarget{second-layer-for-separating-mi-samples-from-cad}{%
\paragraph{Second layer for separating MI samples from
CAD:}\label{second-layer-for-separating-mi-samples-from-cad}}

Different models were trained using expression values for three
differentially miRNAs. The models' 10-fold cross-validated AUC and
accuracy on the test set are reported in Figure @ref(fig:DEMmodels). The
best model from both AUC and accuracy point-of-view was the SVM model
with linear kernel. The AUC and accuracy for this model with its pre-set
values were 0.93 and 0.82 respectively. The model was hyper-tuned for C
and gamma hyper-parameters, and therefore the model showed better
performance. The ROC curve of the hyper-tuned model is presented in
Figure @ref(fig:DEMROC)B. For this model the AUC reached 0.95 and the
accuracy improved to 0.85 (Table @ref(tab:DEGsML)). Moreover, the
sensitivity and specificity for the model on the test set were 0.91 and
0.71 respectively. The confusion matrix for the hyper-tuned model is
illustrated in Figure @ref(fig:DEMCF)B.

\begin{figure}

{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/Models DEGs mirs}

}

\caption{Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set.}\label{fig:DEMmodels}
\end{figure}

\begin{table}

\caption{\label{tab:DEGsML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning}
\centering
\begin{tabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Pre-set parameters} & \multicolumn{2}{c}{Hyper-tuned} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
Model & Metrics & train & test & train & test\\
\midrule
SVM & AUC-ROC & 0.91 & 0.93 & 0.92 & 0.95\\
(Linear Kernel) & Accuracy & 0.83 & 0.82 & 0.84 & 0.85\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{auc-approach}{%
\subsubsection{AUC approach}\label{auc-approach}}

After calculating the AUC for each miRNA for the classification of MI
and CAD samples, miRNAs were sorted, and their performance was
investigated as a set. The metric of choice for selecting the best set
was AUC. As shown in Figure @ref(fig:AUCset), the AUC increased until
the number of miRNAs in the set reached six, and after that, it dropped.
The AUC for separating MI samples from CAD using these miRNAs was 0.93.
The miRNAs in the set were miR-29a, miR-197, miR-142, miR-21, miR-155,
and miR-320C1. The expression values of these miRNAs in healthy, CAD,
and MI samples are presented in Figure @ref(fig:AUCexp). The ROC curve
of the selected miRNAs for MI and CAD sample classification is
illustrated in Figure @ref(fig:miRROC)B.

\begin{figure}

{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/AUC for cad mi mirs}

}

\caption{Area under the curve (AUC) for sets containing an increasing number of miRNAs with the highest individual AUC in MI/CAD separation.}\label{fig:AUCset}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/ExpressionforAUC}

}

\caption{Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples.}\label{fig:AUCexp}
\end{figure}

\hypertarget{first-layer-for-isolation-of-healthy-and-not-healthy-samples-1}{%
\paragraph{First layer for isolation of healthy and not-healthy
samples}\label{first-layer-for-isolation-of-healthy-and-not-healthy-samples-1}}

Using the selected set, an SVM model with an RBF kernel was trained to
separate healthy from not-healthy samples. The ROC curve for the model
is presented in Figure @ref(fig:AUCROC)A and the confusion matrix is
illustrated in Figure @ref(fig:AUCCF)A. Both AUC and accuracy for the
model on the test set were equal to 1.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../../ML/Pics/AUC best model ROCs}

}

\caption{ROC curve for the set of miRNAs selected by AUC on test set classification. (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI samples classification. (C) SVM with polynomial kernel for CAD and MI samples classification. }\label{fig:AUCROC}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/AUC Conf. matrix}

}

\caption{Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI samples classification. (C) SVM with polynomial kernel for CAD and MI samples classification.}\label{fig:AUCCF}
\end{figure}

\hypertarget{second-layer-for-isolation-of-mi-samples-from-cad-samples}{%
\paragraph{Second layer for isolation of MI samples from CAD
samples}\label{second-layer-for-isolation-of-mi-samples-from-cad-samples}}

To find the best model for training the best set, different models were
trained using their pre-set values. Their AUC and accuracy results on
the test set are presented in Figure @ref(fig:AUCmodels). The best model
from the AUC point-of-view was the LR and from the accuracy
point-of-view, it was the SVM model with a polynomial kernel. For the LR
model the AUC and accuracy were 0.92 and 0.81, respectively; and for the
SVM model with a polynomial kernel, the values were 0.91 and 0.84,
respectively. Both models were hyper-tuned and the ROC curve for their
best performance is presented in Figure @ref(fig:AUCROC)B and C. The AUC
and accuracy for the LR model increased to 0.94 and 0.88, respectively.
For the SVM model with a polynomial kernel, these values increased to
0.95 and 0.88, respectively (Table @ref(tab:AUCML)). The sensitivity for
the LR and SVM models were 1 and 0.97, respectively; and the specificity
for them was 0.57 and 0.64, respectively. The confusion matrix for both
models is illustrated in Figure @ref(fig:AUCCF)B and C.

\begin{figure}

{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/Models AUC mirs}

}

\caption{Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set.}\label{fig:AUCmodels}
\end{figure}

\begin{table}

\caption{\label{tab:AUCML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their individual AUC-ROC on the train and test set before and after hyper-tuning}
\centering
\begin{tabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Pre-set parameters} & \multicolumn{2}{c}{Hyper-tuned} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
Model & Metrics & Train & Test & Train & Test\\
\midrule
LR & AUC-ROC & 0.90 & 0.92 & 0.91 & 0.94\\
 & Accuracy & 0.84 & 0.81 & 0.84 & 0.88\\
SVM & AUC-ROC & 0.90 & 0.91 & 0.91 & 0.95\\
(Polynomial Kernel) & Accuracy & 0.86 & 0.84 & 0.86 & 0.88\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The prevalence of acute MI (AMI) can lead to high-rate mortality in the
clinical setting. However, early diagnosis and application of suitable
treatment protocols can reduce mortality and improve AMI prognosis
({``Cardiovascular Diseases ({CVDs})''} n.d.; Thygesen et al. 2018; Tsao
et al. 2022). Studies have suggested that changes in miRNA expression
may play a significant role in the progression of MI and the subsequent
remodeling (Laggerbauer and Engelhardt 2022). It is believed that the
expression of miRNAs is altered during the various biological processes
correlated with MI within the myocardium or other related tissues (Khan,
Gupta, and Mahapatra 2022). Although several research has been
concentrated on examining free circulating miRNAs in the serum samples
for the detection of cardiac tissue injuries (Kaur et al. 2020), more
information is needed to fully comprehend the miRNAs found in different
blood sub-components like plasma, platelets, and PBMCs. Based on
previous data, PBMCs are critically involved in plaque destabilization
and rupture as well as early inflammatory responses during MI (Mosallaei
et al. 2022; Hapke et al. 2022). Moreover, PBMCs have specific miRNA
profile that is altered under certain pathological conditions which are
great candidates as disease biomarkers (Mosallaei et al. 2022).

PBMCs can respond to several insulting conditions such as MI in the
least possible time with prominent changes in their miRNA profile
(Mosallaei et al. 2022). Considering the regulatory roles, subtle
changes in the transcription of miRNAs can be monitored even before
alteration in the levels of mRNAs and proteins (Schulte et al. 2020).
These features make the miRNAs an early-stage valid diagnostic tool for
the detection of minor and major cell injuries. To date, few studies
have been performed to compare the miRNA profiles in PBMCs belonging to
acute MI patients and other CADs and healthy samples to find a robust
set of identical miRNAs to differentiate these pathological conditions.

In this study, we combined three GEO datasets of healthy, CAD, and MI
samples. Having these samples set alongside bioinformatics analysis and
ML means, it is possible to identify potential biomarker sets and also
effective therapeutic targets. The results of the DEG analysis (Table
@ref(tab:DEGstab) and Figure @ref(fig:venn)) are proof of the close
relationship between the MI and CAD samples. Interestingly, functional
enrichment analysis demonstrated that DEGs in both healthy/CAD and
healthy/MI were strongly correlated to immune cell response which is a
major cellular part of PBMCs. Here, two different sets of miRNAs were
used as biomarker sets for sample classification. miR-21; miR-32; and
miR-186 were selected as differentially expressed miRNAs, and miR-21;
miR-29a; miR-142; miR155; miR-197; and miR-320c1 were selected according
to their AUC values. As shown in Figure @ref(fig:miRROC), all other
miRNAs selected with both approaches had AUC-ROC over 0.9 for the
isolation of healthy and not-healthy samples except for miR-142 and
miR-29a. Data confirmed that the real challenge is to classify CAD and
MI samples because of close overlap. Of 8 miRNAs under investigation in
both approaches except for miR-32, all miRNAs had an AUC-ROC over 0.8
for the discrimination of CAD and MI samples. Besides, the high AUC-ROC
values of miRNAs confirms their high potential as biomarkers.

ML models when trained with miRNA sets selected by both DEG and AUC
approaches showed better performance in the classification than each
miRNA. To avoid unwanted complexity and poor predictive values, a
two-layer architecture was also designed. The first layer was for for
the discrimination of healthy from not-healthy samples, and the second
layer separated the CAD from MI candidates. As expected in both
approaches, a hyper-tuned SVM model could flawlessly separates healthy
from not-healthy samples using distinct miRNAs sets. The ML models were
also capable of effectively separating CAD from MI patients. Although
both miRNA sets had nearly the same AUC-ROC with their best model, the
accuracy, sensitivity, and specificity were different. The model trained
with DEGs had better specificity, but the one trained with AUC-selected
miRNAs had slightly better accuracy and higher sensitivity. This
difference comes from the different strategies the selection of
biomarkers sets.

Numerous studies have reported different biological processes can affect
the expression of miRNAs in PBMCs. However, there are still
controversies regarding the exact role of miRNAs in the function of
immune cells and the correlation of specific pathological conditions
with miRNA profiles. Several studies have proved the activation of
specific miRNA types in PBMCs under cardiovascular events (S. Li et al.
2015; Yao et al. 2016; Liu et al. 2017; Horita, Farquharson, and Stephen
2021; Cai et al. 2020; Zhao et al. 2018; Bhansali et al. 2022). For
instance, there is evidence that the elevation of miR-186 suppresses the
expression of Cystathionine-\(\gamma\)-lyase, leading to the subsequent
secretion of pro-inflammatory cytokines and cellular lipid accumulation.
Besides, macrophage-derived miR-186 may promote atherosclerotic plaques
(Yao et al. 2016). In line with this claim, we found that miR-186 is
up-regulated in both CAD and MI candidates related to control
counterparts. Surprisingly, the obtained data indicated that the
expression of miR-186 is higher in CAD patients in comparison to MI
(Figure @ref(fig:DEMexp)), To be specific, miR-186 is the only
differentially expressed miRNA between CAD and MI, indicating its main
role in the promotion of atherosclerosis.

As mentioned before, miR-21 was also up-regulated in both MI and CAD
patients in comparison to healthy controls. Moreover, the expression
value of miR-21 was significantly higher in MI than that of the CAD
group (Table @ref(tab:mirExptable)). It is thought that the
up-regulation of miRNA-21 in PBMCs is a compensatory reaction to reduce
T\(_{reg}\) lymphocyte number in response to the reduction of
TGF\(\beta1\) secretion into the plasma through a
TGF\(\beta1\)/smad-independent pathway. In line with previous and
present data, miR-21 can modulate the activity of PBMCs following the
occurrence of cardiovascular diseases (S. Li et al. 2015).

Recent data have supported the elevation of miR-32 in CAD patients with
the calcification of coronary artery. Of note, miR-32 stimulates the
calcification of mouse vascular smooth muscle through the regulation of
bone morphogenetic protein-1, runt-related transcription factor-2
(RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA
protein (Liu et al. 2017). Likewise, there are some reports associated
with the activity of PBMC miR-32 under several pathologies (Zeng et al.
2021; Wang et al. 2020). The exact role of PBMC miR-32 after
cardiovascular events remained to be elucidated.

Molecular analyses have indicated the regulatory role of miRNAs selected
by the AUC approach in PBMCs after a cardiovascular event. As the only
common miRNA of the two sets, molecular function of miR-21 in CVDs was
covered in DEGs miRNA set. Based on numerous reports mir-29a can be
activated in different diseases (Horita, Farquharson, and Stephen
2021).Data analysis indicated that miR-29a is significantly up-regulated
in CAD patients in comparison to healthy and MI groups (Table
@ref(tab:mirExptable)). Increased miR-29a is associated with the
progression of atherosclerosis, and the combination of miR-29a and
ox-LDL was offered as a valid biomarker set for paraclinical
classification (Huang et al. 2016). However, the role of miR-29a in the
function of PBMCs in CAD patients has not been completely examined.

According to the present data (Table @ref(tab:mirExptable)), the
difference in miR-142 expression between MI/Healthy and CAD/Healthy was
not significant, but these values reached statistically significant in
MI as compared to CAD. Based on different reports, miR-142 is commonly
active in PBMCs. This miRNA directly targets and inhibits the expression
of Adenomatous Polyposis Coli, a negative WNT signaling pathway
regulator, contributing to the activation of the WNT signaling pathway
and cardiac fibroblast activation after myocardial ischemia (Cai et al.
2020).

We also noted that miR-155 is the only down-regulated miRNA among all
miRNAs in both CAD/healthy and MI/Healthy comparisons (Table
@ref(tab:mirExptable)). There are numerous reports on the
down-regulation of miR-155 in PBMCs after MI or in CAD patients.
Previous data suggested an inverse relationship between miR-155 levels
and the severity of the coronary artery condition (Zhang et al. 2015;
Zhao et al. 2018). Along with current data and previous findings, it is
postulated that miR-155 has a protective effect under pathological
conditions. This genetic element can control inflammation and thus
reduce tissue damage through its negative feedback effects on
inflammatory factors. These features coincide with the reduction of
atherosclerosis (Zhang et al. 2015).

Data indicated that miR-197 is also significantly up-regulated in both
CAD/healthy and MI/healthy groups. In previous studies, it has been
shown that miRNA-197 may play a critical role in regulating the
anti-inflammatory response of IL-35 by affecting pro/ anti-inflammatory
cytokine secretion, M1/M2 macrophage ratio, T\(_{reg}\) lymphocyte
proliferation, and T cell suppression suggesting the potential
diagnostic role of miR-197 in adverse cardiovascular events (Bhansali et
al. 2022). There are some reports about miR-320 role in the
physiopathology of cardiac fibrosis. Mechanistically, Molecular analyses
revealed that miR-320 might induce various effects by targeting PLEKHM3
and IFITM1 in cardiomyocytes and cardiac fibroblasts, respectively (F.
Li et al. 2021). However, there are no reports on miR-320c1 activity in
the cardiovascular system or PBMCs.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In summary, we derived a set of miRNA biomarkers by comparing MI samples
to both healthy and CAD samples. We found that the SVM model performed
best in both the first layer, which separated healthy and not-healthy
samples, and the second layer, which classified MI/CAD samples. The set
of miRNAs selected based on their AUC values had slightly better
performance in the second layer. Overall, our two-layer structure
achieved a weighted accuracy of 0.91. This demonstrates the potential
for combining bioinformatics and machine learning techniques to identify
novel biomarkers and gain a deeper understanding of myocardial
infarction.

\hypertarget{references}{%
\section{References}\label{references}}

\begin{align}
a^2+b^2=c^2
\end{align}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-197}{}}%
Bhansali, Shipra, Amit Kumar Yadav, Chetan Bakshi, and Veena Dhawan.
2022. {``Interleukin-35 {Mitigates} Ox-{LDL}-{Induced} {Proatherogenic}
{Effects} via {Modulating} {miRNAs} {Associated} with {Coronary}
{Artery} {Disease} ({CAD}).''} \emph{Cardiovascular Drugs and Therapy},
April. \url{https://doi.org/10.1007/s10557-022-07335-x}.

\leavevmode\vadjust pre{\hypertarget{ref-142}{}}%
Cai, Lidong, Gong Chao, Weifeng Li, Jumo Zhu, Fangfang Li, Baozhen Qi,
Yong Wei, et al. 2020. {``Activated {Cd4}+ {T} Cells-Derived Exosomal
{miR}-142-3p Boosts Post-Ischemic Ventricular Remodeling by Activating
Myofibroblast.''} \emph{Aging} 12 (8): 7380--96.
\url{https://doi.org/10.18632/aging.103084}.

\leavevmode\vadjust pre{\hypertarget{ref-75}{}}%
Canali, Raffaella, Lucia Natarelli, Guido Leoni, Elena Azzini, Raffaella
Comitato, Oezgur Sancak, Luca Barella, and Fabio Virgili. 2014.
{``Vitamin {C} Supplementation Modulates Gene Expression in Peripheral
Blood Mononuclear Cells Specifically Upon an Inflammatory Stimulus: A
Pilot Study in Healthy Subjects.''} \emph{Genes \& Nutrition} 9 (3):
390. \url{https://doi.org/10.1007/s12263-014-0390-x}.

\leavevmode\vadjust pre{\hypertarget{ref-CVD}{}}%
{``Cardiovascular Diseases ({CVDs}).''} n.d. Accessed March 12, 2023.
\url{https://www.who.int/news-room/fact-sheets/detail/cardiovascular-diseases-(cvds)}.

\leavevmode\vadjust pre{\hypertarget{ref-PBMC-miR}{}}%
Gao, Jie, Jia Liu, Ying Zhang, BaoYi Guan, Hua Qu, Hua Chai, WenTing
Wang, XiaoJuan Ma, and DaZhuo Shi. 2020. {``{PBMCs}-{Derived} {microRNA}
{Signature} as a {Prethrombotic} {Status} {Discriminator} in {Stable}
{Coronary} {Artery} {Disease}.''} \emph{Thrombosis and Haemostasis} 120
(01): 121--31. \url{https://doi.org/10.1055/s-0039-1700518}.

\leavevmode\vadjust pre{\hypertarget{ref-Meder23}{}}%
Hapke, Nils, Margarete Heinrichs, DiyaaElDin Ashour, Elena Vogel, Ulrich
Hofmann, Stefan Frantz, and Gustavo Campos Ramos. 2022.
{``Identification of a Novel Cardiac Epitope Triggering {T}-Cell
Responses in Patients with Myocardial Infarction.''} \emph{Journal of
Molecular and Cellular Cardiology} 173 (December): 25--29.
\url{https://doi.org/10.1016/j.yjmcc.2022.09.001}.

\leavevmode\vadjust pre{\hypertarget{ref-numpy}{}}%
Harris, Charles R., K. Jarrod Millman, Stéfan J. van der Walt, Ralf
Gommers, Pauli Virtanen, David Cournapeau, Eric Wieser, et al. 2020.
{``Array Programming with {NumPy}.''} \emph{Nature} 585 (7825): 357--62.
\url{https://doi.org/10.1038/s41586-020-2649-2}.

\leavevmode\vadjust pre{\hypertarget{ref-scikitopt}{}}%
Head, Tim, Manoj Kumar, Holger Nahrstaedt, Gilles Louppe, and Iaroslav
Shcherbatyi. 2021. \emph{Scikit-Optimize/Scikit-Optimize} (version
v0.9.0). Zenodo. \url{https://doi.org/10.5281/zenodo.5565057}.

\leavevmode\vadjust pre{\hypertarget{ref-29aa}{}}%
Horita, Masahiro, Colin Farquharson, and Louise A Stephen. 2021. {``The
Role of {miR}‐29 Family in Disease.''} \emph{Journal of Cellular
Biochemistry} 122 (7): 696--715.
\url{https://doi.org/10.1002/jcb.29896}.

\leavevmode\vadjust pre{\hypertarget{ref-29a}{}}%
Huang, Yu-Qing, An-Ping Cai, Ji-Yan Chen, Cheng Huang, Jie Li, and
Ying-Qing Feng. 2016. {``The {Relationship} of {Plasma} {miR}-29a and
{Oxidized} {Low} {Density} {Lipoprotein} with {Atherosclerosis}.''}
\emph{Cellular Physiology and Biochemistry} 40 (6): 1521--28.
\url{https://doi.org/10.1159/000453202}.

\leavevmode\vadjust pre{\hypertarget{ref-m2}{}}%
Kalayinia, Samira, Fateme Arjmand, Majid Maleki, Mahshid Malakootian,
and Chandra Pal Singh. 2021. {``{MicroRNAs}: Roles in Cardiovascular
Development and Disease.''} \emph{Cardiovascular Pathology} 50
(January): 107296. \url{https://doi.org/10.1016/j.carpath.2020.107296}.

\leavevmode\vadjust pre{\hypertarget{ref-11-15}{}}%
Kaur, Amanpreet, Sharon T Mackin, Kenny Schlosser, Fui Lin Wong, Malik
Elharram, Christian Delles, Duncan J Stewart, Natalie Dayan, Tara
Landry, and Louise Pilote. 2020. {``Systematic Review of {microRNA}
Biomarkers in Acute Coronary Syndrome and Stable Coronary Artery
Disease.''} \emph{Cardiovascular Research} 116 (6): 1113--24.
\url{https://doi.org/10.1093/cvr/cvz302}.

\leavevmode\vadjust pre{\hypertarget{ref-24}{}}%
Khan, Abrar A., Vinayak Gupta, and Nitish R. Mahapatra. 2022. {``Key
Regulatory {miRNAs} in Lipid Homeostasis: {Implications} for
Cardiometabolic Diseases and Development of Novel Therapeutics.''}
\emph{Drug Discovery Today} 27 (8): 2170--80.
\url{https://doi.org/10.1016/j.drudis.2022.05.003}.

\leavevmode\vadjust pre{\hypertarget{ref-23}{}}%
Laggerbauer, Bernhard, and Stefan Engelhardt. 2022. {``{MicroRNAs} as
Therapeutic Targets in Cardiovascular Disease.''} \emph{Journal of
Clinical Investigation} 132 (11): e159179.
\url{https://doi.org/10.1172/JCI159179}.

\leavevmode\vadjust pre{\hypertarget{ref-BER}{}}%
Lazar, C., S. Meganck, J. Taminau, D. Steenhoff, A. Coletta, C. Molter,
D. Y. Weiss-Solis, R. Duque, H. Bersini, and A. Nowe. 2013. {``Batch
Effect Removal Methods for Microarray Gene Expression Data Integration:
A Survey.''} \emph{Briefings in Bioinformatics} 14 (4): 469--90.
\url{https://doi.org/10.1093/bib/bbs037}.

\leavevmode\vadjust pre{\hypertarget{ref-320}{}}%
Li, Fang, Shan-Shan Li, Hui Chen, Jian-Zhi Zhao, Jie Hao, Jin-Ming Liu,
Xiu-Guang Zu, and Wei Cui. 2021. {``{miR}-320 Accelerates Chronic Heart
Failure with Cardiac Fibrosis Through Activation of the {Il6}/{Stat3}
Axis.''} \emph{Aging} 13 (18): 22516--27.
\url{https://doi.org/10.18632/aging.203562}.

\leavevmode\vadjust pre{\hypertarget{ref-21}{}}%
Li, Sihui, Qian Fan, Shaolin He, Tingting Tang, Yuhua Liao, and
Jiangjiao Xie. 2015. {``{MicroRNA}-21 {Negatively} {Regulates} {Treg}
{Cells} {Through} a {TGF}-\(\beta1\)/{Smad}-{Independent} {Pathway} in
{Patients} with {Coronary} {Heart} {Disease}.''} \emph{Cellular
Physiology and Biochemistry} 37 (3): 866--78.
\url{https://doi.org/10.1159/000430214}.

\leavevmode\vadjust pre{\hypertarget{ref-32-2}{}}%
Liu, Jianghua, Xinhua Xiao, Yingying Shen, Ling Chen, Canxin Xu, Heng
Zhao, Ying Wu, et al. 2017. {``{MicroRNA}-32 Promotes Calcification in
Vascular Smooth Muscle Cells: {Implications} as a Novel Marker for
Coronary Artery Calcification.''} Edited by Yin Tintut. \emph{PLOS ONE}
12 (3): e0174138. \url{https://doi.org/10.1371/journal.pone.0174138}.

\leavevmode\vadjust pre{\hypertarget{ref-67}{}}%
Maciejak, Agata, Marek Kiliszek, Marcin Michalak, Dorota Tulacz,
Grzegorz Opolski, Krzysztof Matlak, Slawomir Dobrzycki, Agnieszka
Segiet, Monika Gora, and Beata Burzynska. 2015. {``Gene Expression
Profiling Reveals Potential Prognostic Biomarkers Associated with the
Progression of Heart Failure.''} \emph{Genome Medicine} 7 (1): 26.
\url{https://doi.org/10.1186/s13073-015-0149-z}.

\leavevmode\vadjust pre{\hypertarget{ref-09}{}}%
Matone, Alice, Colm M. O'Grada, Eugene T. Dillon, Ciara Morris, Miriam
F. Ryan, Marianne Walsh, Eileen R. Gibney, et al. 2015. {``Body Mass
Index Mediates Inflammatory Response to Acute Dietary Challenges.''}
\emph{Molecular Nutrition \& Food Research} 59 (11): 2279--92.
\url{https://doi.org/10.1002/mnfr.201500184}.

\leavevmode\vadjust pre{\hypertarget{ref-fRMA}{}}%
McCall, M. N., B. M. Bolstad, and R. A. Irizarry. 2010. {``Frozen Robust
Multiarray Analysis ({fRMA}).''} \emph{Biostatistics} 11 (2): 242--53.
\url{https://doi.org/10.1093/biostatistics/kxp059}.

\leavevmode\vadjust pre{\hypertarget{ref-Barcode}{}}%
McCall, Matthew N., Karan Uppal, Harris A. Jaffee, Michael J. Zilliox,
and Rafael A. Irizarry. 2011. {``The {Gene} {Expression} {Barcode}:
Leveraging Public Data Repositories to Begin Cataloging the Human and
Murine Transcriptomes.''} \emph{Nucleic Acids Research} 39 (suppl\_1):
D1011--15. \url{https://doi.org/10.1093/nar/gkq1259}.

\leavevmode\vadjust pre{\hypertarget{ref-pandas}{}}%
McKinney, Wes. 2010. {``{D}ata {S}tructures for {S}tatistical
{C}omputing in {P}ython.''} In \emph{{P}roceedings of the 9th {P}ython
in {S}cience {C}onference}, edited by Stéfan van der Walt and Jarrod
Millman, 56--61.
\href{https://doi.org/\%2010.25080/Majora-92bf1922-00a\%20}{https://doi.org/
10.25080/Majora-92bf1922-00a }.

\leavevmode\vadjust pre{\hypertarget{ref-Meder6}{}}%
Mosallaei, Meysam, Naeim Ehtesham, Shima Rahimirad, Mostafa Saghi, Nasim
Vatandoost, and Sharifeh Khosravi. 2022. {``{PBMCs}: A New Source of
Diagnostic and Prognostic Biomarkers.''} \emph{Archives of Physiology
and Biochemistry} 128 (4): 1081--87.
\url{https://doi.org/10.1080/13813455.2020.1752257}.

\leavevmode\vadjust pre{\hypertarget{ref-scikit-learn}{}}%
Pedregosa, F., G. Varoquaux, A. Gramfort, V. Michel, B. Thirion, O.
Grisel, M. Blondel, et al. 2011. {``Scikit-Learn: Machine Learning in
{P}ython.''} \emph{Journal of Machine Learning Research} 12: 2825--30.

\leavevmode\vadjust pre{\hypertarget{ref-ML}{}}%
Reel, Parminder S., Smarti Reel, Ewan Pearson, Emanuele Trucco, and
Emily Jefferson. 2021. {``Using Machine Learning Approaches for
Multi-Omics Data Analysis: {A} Review.''} \emph{Biotechnology Advances}
49 (July): 107739.
\url{https://doi.org/10.1016/j.biotechadv.2021.107739}.

\leavevmode\vadjust pre{\hypertarget{ref-miR}{}}%
Schulte, Christian, Temo Barwari, Abhishek Joshi, Tanja Zeller, and
Manuel Mayr. 2020. {``Noncoding {RNAs} Versus {Protein} {Biomarkers} in
{Cardiovascular} {Disease}.''} \emph{Trends in Molecular Medicine} 26
(6): 583--96. \url{https://doi.org/10.1016/j.molmed.2020.02.001}.

\leavevmode\vadjust pre{\hypertarget{ref-m1}{}}%
Schulte, Christian, Mahir Karakas, and Tanja Zeller. 2017.
{``{microRNAs} in Cardiovascular Disease -- Clinical Application.''}
\emph{Clinical Chemistry and Laboratory Medicine (CCLM)} 55 (5).
\url{https://doi.org/10.1515/cclm-2016-0576}.

\leavevmode\vadjust pre{\hypertarget{ref-36}{}}%
Soler-Botija, Carolina, Carolina Gálvez-Montón, and Antoni Bayés-Genís.
2019. {``Epigenetic {Biomarkers} in {Cardiovascular} {Diseases}.''}
\emph{Frontiers in Genetics} 10 (October): 950.
\url{https://doi.org/10.3389/fgene.2019.00950}.

\leavevmode\vadjust pre{\hypertarget{ref-17}{}}%
Thygesen, Kristian, Joseph S. Alpert, Allan S. Jaffe, Bernard R.
Chaitman, Jeroen J. Bax, David A. Morrow, Harvey D. White, and The
Executive Group on behalf of the Joint European Society of Cardiology
(ESC)/American College of Cardiology (ACC)/American Heart Association
(AHA)/World Heart Federation (WHF) Task Force for the Universal
Definition of Myocardial Infarction. 2018. {``Fourth {Universal}
{Definition} of {Myocardial} {Infarction} (2018).''} \emph{Circulation}
138 (20). \url{https://doi.org/10.1161/CIR.0000000000000617}.

\leavevmode\vadjust pre{\hypertarget{ref-ML-Bio}{}}%
Torun, Furkan M., Sebastian Virreira Winter, Sophia Doll, Felix M.
Riese, Artem Vorobyev, Johannes B. Mueller-Reif, Philipp E. Geyer, and
Maximilian T. Strauss. 2023. {``Transparent {Exploration} of {Machine}
{Learning} for {Biomarker} {Discovery} from {Proteomics} and {Omics}
{Data}.''} \emph{Journal of Proteome Research} 22 (2): 359--67.
\url{https://doi.org/10.1021/acs.jproteome.2c00473}.

\leavevmode\vadjust pre{\hypertarget{ref-18}{}}%
Tsao, Connie W., Aaron W. Aday, Zaid I. Almarzooq, Alvaro Alonso, Andrea
Z. Beaton, Marcio S. Bittencourt, Amelia K. Boehme, et al. 2022.
{``Heart {Disease} and {Stroke} {Statistics}---2022 {Update}: {A}
{Report} {From} the {American} {Heart} {Association}.''}
\emph{Circulation} 145 (8).
\url{https://doi.org/10.1161/CIR.0000000000001052}.

\leavevmode\vadjust pre{\hypertarget{ref-32}{}}%
Wang, Dan, Ting Zeng, Zhi Lin, Lu Yan, Fenglin Wang, Lanlan Tang, Leyuan
Wang, Daolin Tang, Pan Chen, and Minghua Yang. 2020. {``Long Non-Coding
{RNA} {Snhg5} Regulates Chemotherapy Resistance Through the
{miR}-32/{Dnajb9} Axis in Acute Myeloid Leukemia.''} \emph{Biomedicine
\& Pharmacotherapy} 123 (March): 109802.
\url{https://doi.org/10.1016/j.biopha.2019.109802}.

\leavevmode\vadjust pre{\hypertarget{ref-186}{}}%
Yao, Yan, Xin Zhang, Hai-peng Chen, Liang Li, Wei Xie, Gang Lan,
Zhen-wang Zhao, Xi-Long Zheng, Zong-bao Wang, and Chao-ke Tang. 2016.
{``{MicroRNA}-186 Promotes Macrophage Lipid Accumulation and Secretion
of Pro-Inflammatory Cytokines by Targeting Cystathionine
\(\gamma\)-Lyase in {THP}-1 Macrophages.''} \emph{Atherosclerosis} 250
(July): 122--32.
\url{https://doi.org/10.1016/j.atherosclerosis.2016.04.030}.

\leavevmode\vadjust pre{\hypertarget{ref-Macro}{}}%
Yap, Jonathan, Jason Irei, Javier Lozano-Gerona, Selena Vanapruks,
Tianmai Bishop, and William A. Boisvert. 2023. {``Macrophages in Cardiac
Remodelling After Myocardial Infarction.''} \emph{Nature Reviews
Cardiology}, January. \url{https://doi.org/10.1038/s41569-022-00823-5}.

\leavevmode\vadjust pre{\hypertarget{ref-magic}{}}%
Zeng, Zl, Qingyun Zhu, Zhibo Zhao, Xuyu Zu, and Jianghua Liu. 2021.
{``Magic and Mystery of {microRNA}‐32.''} \emph{Journal of Cellular and
Molecular Medicine} 25 (18): 8588--8601.
\url{https://doi.org/10.1111/jcmm.16861}.

\leavevmode\vadjust pre{\hypertarget{ref-155-2}{}}%
Zhang, Yu-Hui, Liang-Hua Xia, Jia-Mei Jin, Ming Zong, Ming Chen, and Bo
Zhang. 2015. {``Expression Level of {miR}--155 in Peripheral Blood.''}
\emph{Asian Pacific Journal of Tropical Medicine} 8 (3): 214--19.
\url{https://doi.org/10.1016/S1995-7645(14)60318-7}.

\leavevmode\vadjust pre{\hypertarget{ref-155-3}{}}%
Zhao, Duo, Jing Zhao, Jiangbin Sun, Yanling Su, Jinfeng Jian, Huaan Ye,
Jiawang Lin, Zongda Yang, Jiatao Feng, and Zhiping Wang. 2018. {``The
Expression Level of {miR}-155 in Plasma and Peripheral Blood Mononuclear
Cells in Coronary Artery Disease Patients and the Associations of These
Levels with the Apoptosis Rate of Peripheral Blood Mononuclear Cells.''}
\emph{Experimental and Therapeutic Medicine}, September.
\url{https://doi.org/10.3892/etm.2018.6797}.

\end{CSLReferences}

\bibliographystyle{spmpsci}
\bibliography{ForArticleRefrence.bib,discussion.bib,Introduction.bib}

\end{document}

### 2023-06-13 ArticleDraft.Rmd
Revised with Ithenticate and Grammerly

--- 2023-02-08 ArticleDraft.Rmd
@@
 ---
-title: "The First Draft"
+title: Machine Learning Driven Set of microRNAs as a Novel Biomarker Set for Myocardial Infarction Diagnosis
 authors:
@@
 link-citation: yes
-bibliography: For Article Refrence.bib
+bibliography: [For Article Refrence.bib, discussion.bib, Introduction.bib, mir296.bib]
+abstract: |
+  MicroRNAs (miRNAs) play a crucial role in regulating adaptive and maladaptive responses in cardiovascular diseases, making them attractive targets for potential biomarkers.
+However, their potential as novel biomarkers for diagnosing cardiovascular diseases requires systematic evaluation.
+In this study, we aimed to identify a key set of miRNA biomarkers using integrated bioinformatics and machine learning analysis.
+We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contains peripheral blood mononuclear cells (PBMCs) samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy individuals.
+Additionally, we selected a set of miRNAs based on their area under the receiver operating characteristic curve (AUC-ROC) for separating the CAD and MI samples.
+We designed a two-layer architecture for sample classification, in which the first layer isolates healthy samples from not-healthy ones, and the second layer classifies stable CAD and MI samples.
+We trained different machine learning models using both biomarker sets and evaluated their performance on a test set.
+We identified miR-21, miR-186, and miR-32 as the only miRNAs among the differentially expressed genes, and a set including miR-186, miR-21, miR-197, miR-29A, and miR-296 as the optimum set of miRNAs selected by their AUC-ROC.
+Both biomarker sets could distinguish healthy from not-healthy samples with complete accuracy.
+The best performance for the classification of CAD and MI was achieved with an SVM model trained using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.96 and an accuracy of 0.94 on the test data.
+Our study demonstrated that miRNA signatures derived from PBMCs could serve as valuable novel biomarkers for cardiovascular diseases.
+
 output: 
@@
   bookdown::word_document2: default
-#    base_format: rticles::bioinformatics_article
+always_allow_html: true
 ---
@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(knitr, captioner, bundesligR, stringr)
+pacman::p_load(knitr, captioner, bundesligR, stringr, kableExtra, ggplot2, ggvenn)
 ```
 
-```{r , echo=F}
-# What to do next:
+# Introduction
 
-# Generating information related to the machine learning first layer using welcome file.
+At present, cardiovascular diseases (CVDs) are the leading cause of human mortality, with 32% of all global deaths.
+It is estimated that about 85% of CVDs mortality were diagnosed with myocardial infarction (MI) [@CVD].
+MI is an acute coronary syndrome with sudden blockage and stenosis of the coronary artery and subsequent myocardial ischemia, leading to extensive cardiomyocyte damage and necrosis [@Macro].
 
-```
+Over the last 50 years, numerous attempts have been collected to use biomarkers to facilitate diagnosis, assess the risk, follow-up therapy, and determine therapeutic efficacy in CVDs candidates.
+Based on the released guidelines, cardiac troponins (cTns) are used as a highly-sensitive and accurate approach for detecting MI.
+Despite the inherent advantages, the high-rate sensitivity of cTn-based assays has also led to more false positive results [@17], necessiating the advent and development of new modalities with pathological values.
+To improve diagnostic value upon existing MI biomarkers, the combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, is proposed.
+Previous researches support the notion that miRNAs exhibit great potential to be used as alternative biomarkers in CVDs detection and follow-up [@miR].
+It is suggested that miRNAs possess 18-22 nucleotides and can play a crucial role in the regulation of gene expression.
+Evidence point to the fact that miRNAs are involved in the pathogenesis of cardiac tissue injury [@m1].
+Several biological activities, such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque formation, and cardiac rhythm, are regulated by miRNAs [@m2].
+These elements can easily circulate in biofluids and could be considered theranostics targets in terms of CVDs [@m1].
+It is postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in CVD patients.
+To be specific, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) [@36].
+
+PBMCs are a fraction of white blood cells, including monocytes, lymphocytes, macrophages, and other cells belonging to the immune system [@PBMC-miR].
+Emerging data have indicated that PBMCs can be used as a valid source of biomarkers for monitoring various pathological conditions.
+Of note, the alteration of mRNAs and miRNAs under pathological conditions gives us valuable information about different kinds of disorders.
+PBMCs could recapitulate the conditions of the target tissues, thus, providing a highly sensitive and specific source of biomarkers [@Meder6].
+Commensurate with these conditions, these cells are repositories of dysregulated genes and miRNAs expression profiles in CVDs [@PBMC-miR; @Meder6].
+
+In recent years, the advent and application of machine learning (ML) is an exciting prospect for advancing scientific discoveries.
+Although the concept of ML and its initial algorithms were conceived many years ago, recent improvements in computing power and access to vast amounts of data have shown that ML techniques outperform classical statistical methods in various fields.
+Furthermore, the progress made in omics technologies has enabled the analysis of massive and intricate biological data sets, consisting of hundreds to thousands of samples, which makes it possible for ML to extract valuable biological insight and information from such data [@ML-Bio].
+As a result, ML provides innovative methods for merging and interpreting diverse types of omics data, leading to the identification of new biomarkers.
+These biomarkers can aid in precise disease prediction, patient stratification, and developing novel therapeutic approaches [@ML].
+
+In this study, we aimed to identify potential miRNA biomarkers for MI patients by combining and analyzing three different microarray datasets from PBMCs.
+It is suggested that the integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring MI.
+Besides, this approach can deepen our vision into the underlying mechanisms of MI and aid in the development of valid diagnostic biomarkers and patient stratification.
 
@@
 Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database (<https://www.ncbi.nlm.nih.gov/geo/>).
-In order to come by sufficient classification power between MI samples and others, the sample size should be relatively large.
-Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475 for healthy samples were selected.
+To obtain sufficient classification power between MI, healthy, and CAD samples, a relatively large sample size was required.
+Therefore, GSE59867 for MI and CAD samples and GSE56609 and GSE54475 for healthy samples were selected.
 All samples were produced using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
-Only healthy, stable CAD and early stage MI samples were selected from these datasets for further analysis.
-The basic information for the three datasets evaluated in the current study is provided in Table 1.
+Only healthy, CAD, and early-stage MI samples were selected from these datasets for further analyses.
+The basic information for the three datasets evaluated in the current study is provided in Table \@ref(tab:datasets).
+The bioinformatics sections were fully conducted on R, ver.
+4.2.0 [@R], using RStudio [@RStudio], and all plots and graphics of these sections were created using the ggplot2 R package [@ggplot].
 
-|          | Platform | Healthy Control | CAD Control | MI  | Reference |
-|----------|----------|-----------------|-------------|-----|-----------|
-| GSE59867 | GPL6244  | \-              | 46          | 111 | [@67]     |
-| GSE56609 | GPL6244  | 46              | \-          | \-  | [@09]     |
-| GSE54475 | GPL6244  | 5               | \-          | \-  | [@75]     |
+```{r, datasets, echo=FALSE}
+df <- data.frame(Dataset = c("GSE59867", "GSE56609", "GSE54475"),
+                 Platform = c("GPL6244", "GPL6244", "GPL6244"),
+                 Healthy = c("-", 46, 5),
+                 CAD = c(46, "-", "-"),
+                 MI = c(111, "-", "-"),
+                 Refrence = c("[@67]", "[@09]", "[@75]"))
 
-: Basic information of the GEO microarray datasets.
+kbl(df, format = "markdown", booktabs = T, align = "c", linesep = c(""), caption = "Samples information on the GEO microarray datasets.")
+```
 
@@
 
-Raw data (CEL files) of all datasets were downloaded from the GEO and pre-processed using the fRMA package [@fRMA].
-fRMA allows to preprocess individual microarray samples and combine them consistently for analysis.
-For each dataset, background correction was performed using RMA algorithm and then it was quantile normalized based on the reference distribution.
-During summarization, batch effects were removed and variances of the gene expressions were estimated by taking into account these probe-specific effects.
-For those multiple probe sets matched to the identical gene, the mean log fold change was retained.
-This way fRMA can be seen as a batch effect removal technique for different datasets that were produced identical microarray platform.
-Thus, to become ensure about batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
+The raw data in the form of CEL files from all datasets were obtained from the GEO.
+To prepare the data for analysis, we utilized the fRMA package [@fRMA], facilitating the pre-processing of individual microarray samples and their consistent combination.
+For each dataset, background correction was applied using the RMA algorithm, followed by quantile normalization based on the reference distribution.
+To account for probe-specific effects, batch effects were eliminated during summarization, and gene expression variances were estimated accordingly.
+In cases where multiple probe sets matched the same gene, the mean log fold change was retained.
+Consequently, fRMA can serves as a technique for removing batch effects across diverse datasets generated by identical microarray platforms [@BER].
+To ensure the effectiveness of batch effect removal, we employed principal component analysis (PCA) and relative log expression (RLE) plots to visualize the data before and after applying fRMA.
 
@@
 
-The barcode algorithm proposed by McCall et al.
-[@Barcode] transforms the actual expression values into binary barcode values.
-Huge sets of samples were collected and normalized using fRMA for several platforms.
-The distribution of the expressed and non-expressed observed intensities for each gene is estimated using these normalized sets.
-Genes are deemed expressed (and their value coded to 1) or unexpressed (and their value coded to 0) according to the following equation:
+The barcode algorithm, as introduced by McCall et al.
+[@Barcode], aimed to convert actual expression values into binary barcode values.
+Extensive sample collections were gathered, and normalization was performed using fRMA across multiple platforms, including the Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
+By utilizing these normalized datasets, the distribution of observed intensities, both for expressed and unexpressed genes, was estimated.
+The determination of whether a gene was expressed or unexpressed was based on the following equation, where a value of 1 indicated expression and a value of 0 indicated non-expression:
 
@@
 
-where $x_{ij}$ is the normalized intensity of gene $i$ in sample $j$, $C$ is a user-defined parameter, $\sigma^{ne}$ is the standard deviation of the non-expressed distribution and $\mu^{ne}$ is the mean of the non-expressed distribution.
-The barcode representation of a sample is a vector of ones and zeros denoting which genes are estimated to be expressed (ones) and unexpressed (zeros).
-The barcode algorithm was implemented by the barcode function in the R fRMA package, and the default value of $C$ was used.
+In the barcode algorithm, the normalized intensity of gene $i$ in sample $j$ is denoted by $x_{ij}$.
+A user-defined parameter, $C$, is introduced, along with the standard deviation ($\sigma^{ne}$) and mean ($\mu^{ne}$) of the non-expressed distribution.
+Based on these values, the barcode representation of a sample is generated as a vector consisting of ones and zeros, representing the estimated expression (ones) and non-expression (zeros) of each gene.
+The barcode function within the R fRMA package was employed for implementing the barcode algorithm, utilizing the default value of $C$.
 
-To determine if the expressed ratios differed in the MI group versus the healthy control group, Fisher's exact test for individual genes was carried out upon the barcode values.
-Genes with a false discovery rate (FDR) of $< 0.05$, which was calculated through the Benjamini-Hochberg (BH) procedure to adjust for multiple testing issue, were considered as differentially expressed genes.
-
-The same procedure were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs in between them.
+To assess the differences in expressed ratios between the MI group and the healthy control group, Fisher's exact test was performed on the barcode values of individual genes.
+Genes that exhibited a false discovery rate (FDR) below $0.05$, calculated using the Benjamini-Hochberg procedure to account for multiple testing issues, were identified as differentially expressed genes (DEGs).
+The same procedures were applied to the CAD versus healthy control comparison, as well as the MI versus CAD group, to identify DEGs specific to each comparison.
 
@@
 
-Using the R clusterProfilter package, the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation were carried out on the differentially expressed genes.
-The GO analysis included biological process (BP), cellular component (CC) and molecular function (MF) categories.
-An adjusted P \< 0.05 was considered to indicate a statistically significant difference.
-Enrichment were conducted on the MI-healthy and CAD-healthy DEGs.
-In these analyses, all default parameters were used.
+The R clusterProfiler package [@cluster] was utilized to perform the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation on the set of DEGs.
+The GO analysis encompassed three categories: biological process (BP), cellular component (CC), and molecular function (MF).
+For statistical significance, an adjusted p-value threshold of less than 0.05 was employed.
+The enrichment analyses were conducted separately for the DEGs specific to the MI-healthy and CAD-healthy comparisons.
+In these analyses, all default parameters provided by the package were used.
 
-## Machine Learning
+## ML procedure
 
-The machine learning analysis was performed using Python software, ver.
-3.9, numpy [@numpy], pandas [@pandas] and Scikit-Learn packages [@scikit-learn].
-When ever hyper-tuning was needed the scikit-opt package [@scikitopt] was used.
-In all ML analysis, the datasets were divided into train and test set by 0.7:0.3 ratio and all reported results are the average of a 10-fold cross-validation.
+The ML analysis was performed using Python software, ver.
+3.9, Numpy [@numpy], pandas [@pandas], and Scikit-Learn packages [@scikit-learn].
+Whenever hyper-tuning was needed, the scikit-opt package [@scikitopt] was used.
+In all ML analyses, the datasets were divided into train and test sets by a 0.7:0.3 ratio and, all reported results are the average of 10-fold cross-validation.
 
-Two different approaches were used for selecting microRNAs to train models with.
-The first approach was using the microRNAs that are differentially expressed.
-In the second approach microRNAs were selected by their individual AUC-ROC.
-Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of set of microRNAs selected with different logic.
+Two different approaches were used for selecting miRNAs for model training.
+The first approach was using the miRNAs that are differentially expressed.
+In the second approach, miRNAs with individual AUC-ROC over 0.8 for separating MI from CAD were selected.
+Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of sets of miRNAs selected with different logics.
 
-### microRNAs in DEGs
+### miRNAs in DEGs
 
-In this approach a two layer architecture was deployed to the data in order to maximize the prediction values.
-The first layer predicted whether a sample is healthy or not, and the second layer was separated MI from CAD in the samples which were predicted as not healthy in the first layer.
+In this approach, a two-layer architecture was deployed to the data to maximize the prediction values.
+The first layer predicted whether a sample was healthy or not, and the second layer separated MI from CAD in the samples which were predicted as not healthy in the first layer.
 To this end, a distinct ML model was trained for each layer.
-Since there are limited numbers of microRNAs in DEGs both layers where trained with all of them.
-For further comparison with the models performance ROC curve of each microRNA for classifying healthy and not-healthy, as well as CAD and MI was generated using a simple Logistic Regression model.
+Since there is a limited number of miRNAs in DEGs, both layers were trained with all of them.
+For further comparison with the models' performance, the ROC curve of each miRNA for classifying healthy and not-healthy, as well as CAD and MI, were generated using a Logistic Regression model.
 
-#### First layer for seperating healthy and not-healthy samples:
+#### First layer for isolation of healthy and not-healthy samples:
 
-An SVM model using RBF kernels were hyper-tunned and trained using all microRNAs in DEGs.
-In order to handle the imbalance number of samples in groups (51 for healthy and 157 for not-healthy group), sample weight for not-healthy samples were set to 0.5.
+A support vector machine (SVM) model using RBF kernels was trained and hyper-tuned using all miRNAs in DEGs.
+To handle the severe imbalance in the number of samples (51 for the healthy and 157 for the not-healthy groups), the sample weight for the healthy and the not-healthy samples was set to 1 and 0.5, respectively.
+The ROC curve and confusion matrix for the model were reported.
 
-#### Second layer for seperating MI and CAD samples:
+#### Second layer for separating MI and CAD samples:
 
-For the sake of reaching the highest classification performance using the set of microRNAs, different models were investigated.
+For the sake of reaching the highest classification performance, different models were investigated.
 To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) models were trained.
 All models were trained with their pre-set parameters with 10-fold cross-validation.
+The criteria for choosing the best model were the highest accuracy and AUC-ROC on the test set.
+The best model was hyper-tuned with the scikit-opt package [@scikitopt] for the best classification performance.
+The ROC curve and confusion matrix for the best model were reported.
 
-Criteria for choosing the best model was the highest accuracy and AUC on the test set.
-The best model was hyper-tuned with scikit-opt package [@scikitopt] in order to get the best predictive performance.
+### miRNAs with the highest AUC-ROC
 
-### microRNAs with the highest AUC-ROC
+Like the previous approach, a two-layers strategy was conducted.
+The first layer classified samples into healthy and not-healthy, and the separated MI and CAD samples.
+However, to keep the number of miRNAs as low as possible, miRNAs were selected from the second layer, and then their performance was evaluated in the first layer.
+The AUC-ROC of all miRNAs for classifying MI and CAD samples were calculated and, the miRNAs with the AUC-ROC over 0.8 were selected.
+The ROC curves for each selected miRNA for separating healthy samples from not-healthy and MI from CAD samples were also plotted for further comparison.
 
-Like the previous approach, a two layers strategy was conducted.
-The first layer for classifying samples to healthy and not-healthy and the second layer for separating MI and CAD samples.
-However, for keeping the number of microRNAs as low as possible microRNAs will be selected from the second layer (which are the microRNAs with the best performance in MI/CAD separation) and then their performance were evaluated in the first layer.
+#### First layer for isolation of healthy and not-healthy samples:
 
-AUC-ROC of all microRNAs for classifying MI and CAD samples has been calculated.
-For finding the number of microRNAs with the highest predictive values, the microRNAs with the highest individual AUC-ROC is adding to the set one-by-one and the AUC-ROC for the set is calculated.
-The set with the highest AUC-ROC has been selected as the set for the following steps.
-The ROC curves for each selected microRNA for separating healthy samples from not-healthy and MIs from CADs were also plotted for further comparison.
+An SVM model with an RBF kernel was trained using the selected set of miRNAs.
+Additionally, the model was hyper-tuned to find the hyper-parameters for the highest AUC-ROC and accuracy.
+The same sample weights as the previous approach (1 for healthy and 0.5 for not-healthy samples) were used.
+The ROC curve and confusion matrix for the model were reported.
 
-#### First layer for finding healthy and not-healthy samples:
+#### Second layer for separating MI and CAD samples:
 
-An SVM model with RBF kernel were trained with the selected set of microRNAs.
-Moreover, the model was hyper-tuned for finding the hyper-parameters with the best AUC and accuracy.
-The ROC curve and confusion matrix for the best model was reported.
+The selected miRNA set was used to train different algorithms to find the best model.
+Similar to the previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT models were trained.
+All models were trained with their pre-set parameters using 10-fold cross-validation.
+The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned using the scikit-opt package [@scikitopt].
+The ROC curve and confusion matrix for the best model were reported.
 
-#### Second layer for separating MI and CAD:
-
-The selected number of microRNAs was used to train different algorithms for finding the best model.
-As previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB and DT were trained.
-All models were trained with their pre-set parameters with 10-fold cross-validation.
-The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned with scikit-opt package [@scikitopt].
-The ROC curve and confusion matrix for the best model was reported.
-
-# Results and Discussion
+# Results
 
@@
 
-The PCA plot of the train samples were shown in Fig\@ref(fig:PCA).
-As it is clear, there was a complete separation between healthy samples and CAD and MI samples in primary data which also remains after conducting fRMA on the data.
+The PCA plots of the samples are shown in Figures \@ref(fig:PCA)A and B.
+As shown, healthy samples were separated from CAD or MI samples in primary data and after conducting fRMA.
+In the RLE plot, there was a distinct difference between dataset means for all samples before performing fRMA (Figure \@ref(fig:PCA)C).
+All datasets were rearranged around 0 in the RLE plot after conducting fRMA (Figure \@ref(fig:PCA)D).
+Moreover, there was an apparent change in inter-quantile distances, but the values still were over 0.1.
 
-The RLE plot presented in Fig\@ref(fig:PCA) validates batch effect removal.
-For an efficient batch effect removal method, the individual boxplots will be all distributed around 0 in RLE plot, and inter-quantile distances would be greater than 0.1 [@BER].
-The mentioned criteria was not met in primary data, but was met after conducting fRMA algorithm.
-
-```{r PCA, echo=F, out.width="100%", fig.cap="PCA and RLE plot for all samples before and after fRMA."}
+```{r PCA, echo=F, out.width="100%", fig.cap="principal component analysis plots for (A) primary data and (B) the data after fRMA, and the relative log expression plots for (C) primary data and (D) the data after fRMA."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/PCA and RLE of trains.pdf")
@@
 C_H <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs CAD-Healthy.txt", check.names = F)
+
+df <- t(data.frame(`MI vs. Healthy` = c(dim(M_H)[1], sum(M_H$logFC > 0), sum(M_H$logFC < 0), "hsa-miR-186, miR-21, miR-32"),
+       `CAD vs. Healthy` =  c(dim(C_H)[1], sum(C_H$logFC > 0), sum(C_H$logFC < 0), "hsa-miR-186, miR-21, miR-32"), 
+       `MI vs. CAD` = c(dim(M_C)[1], sum(M_C$logFC > 0),  sum(M_C$logFC < 0), "hsa-miR-186")))
+
+colnames(df) <- c("Total DEGs",  "Up-regulated DEGs", "Down-regulated DEGs", "miRNAs")
+rownames(df) <- c("MI vs. Healthy", "CAD vs. Healthy", "MI vs. CAD")
+
+kbl(df, booktabs = T, align = "c", linesep = c(""), caption = "Total, up-, and down-regulated DEGs and differentially expressed miRNAs.")%>% 
+  kable_styling(full_width = T) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(5, width = "5cm") %>%
+  column_spec(4, width = "2.5cm") %>%
+  column_spec(3, width = "2.5cm")
 ```
@@
 Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in MI compared to the healthy controls.
-For CAD and healthy groups there were `r dim(C_H)[1]` DEGs, `r sum(C_H$logFC > 0)` of them were up-regulated and `r sum(C_H$logFC < 0)` of them were down-regulated in CAD samples in comparison with healthy samples.
-For MI and CAD group the number of DEGs are `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples in comparison with CAD samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively.
-All these data is summarized in table 2.
+In CAD and healthy group comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and `r sum(C_H$logFC < 0)` DEGs were up- and down-regulated, respectively, in CAD samples.
+In the MI and CAD groups, the number of DEGs was `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)` compared to CAD samples, respectively.
+These data are summarized in Table \@ref(tab:DEGstab).
 
-|                 | No. of DEGs     | no. of up-regulated DEGs | no. of down-regulated DEGs | microRNAs                           |
-|---------------|---------------|---------------|---------------|---------------|
-| MI vs. Healthy  | `r dim(M_H)[1]` | `r sum(M_H$logFC > 0)`   | `r sum(M_H$logFC < 0)`     | hsa-miR-186, hsa-miR-21, hsa-miR-32 |
-| CAD vs. Healthy | `r dim(C_H)[1]` | `r sum(C_H$logFC > 0)`   | `r sum(C_H$logFC < 0)`     | hsa-miR-186, hsa-miR-21, hsa-miR-32 |
-| MI vs. CAD      | `r dim(M_C)[1]` | `r sum(M_C$logFC > 0)`   | `r sum(M_C$logFC < 0)`     | hsa-miR-186                         |
+```{r venn, echo=F, out.width="45%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
 
-: Total DEGs and name of differentially expressed microRNAs.
+ch <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs CAD-Healthy.txt", check.names = F)$DEGs
 
-## Gene ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) enrichment analyses of the DEGs.
+mh <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-Healthy.txt", check.names = F)$DEGs
 
-To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
+mc <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-CAD.txt", check.names = F)$DEGs
 
-For MI versus healthy samples, GO enrichment analysis in the BP category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Fig\@ref(fig:MIHEnrich)A).
-In the CC category the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:MIHEnrich)B).
-In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Fig\@ref(fig:MIHEnrich)C).
-KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:MIHEnrich)D).
+df <- list(`CAD/Healthy` = ch, `MI/Healthy` = mh, `MI/CAD` = mc)
+
+ggvenn(df, show_percentage = F, stroke_alpha = 0.5, fill_alpha = 0.4, text_size = 6, set_name_size = 6) + theme(text = element_text(family = "Times"))
+
+```
+
+The Venn diagram in Figure \@ref(fig:venn) shows that CAD and MI samples shared most of their DEGs.
+From 860 DEGs of MI/healthy and 670 DEGs of CAD/healthy, 531 genes were common, which is 62% of MI/healthy DEGs and 79% of CAD/healthy DEGs.
+
+## GO and KEGG enrichment analyses of the DEGs
+
+To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI/healthy and CAD/healthy DEGs.
+For MI/healthy, GO enrichment analysis in the BP category suggested that the DEGs were enriched in "immune response-regulating signaling pathway," "lymphocyte differentiation," "immune response-regulating cell surface receptor signaling pathway," and "leukocyte activation involved in immune response" (Figure \@ref(fig:MIHEnrich)A).
+In the CC category, the DEGs were enriched in "secretory granule membrane," "azurophil granule," "ficolin-1-rich granule," "tertiary granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:MIHEnrich)B).
+In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(fig:MIHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway," "Lipid and atherosclerosis," and "Hematopoietic cell lineage" (Figure \@ref(fig:MIHEnrich)D).
 
@@
 (D) KEGG analysis."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/enrich MI-Healthy.pdf")
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/Enrichment MI-Healthy.pdf")
 ```
 
-The enrichment results for DEGs of CAD versus healthy samples were as follows.
-In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Fig\@ref(fig:CADHEnrich)A).
-In the CC category the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Fig\@ref(fig:CADHEnrich)B).
-In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Fig\@ref(fig:CADHEnrich)C).
-KEGG pathway analysis indicated that the  DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Fig\@ref(fig:CADHEnrich)D).
+The enrichment results for CAD/healthy DEGs were as follows.
+In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response," "positive regulation of innate immune response," "mononuclear cell differentiation," and "positive regulation of response to external stimulus" (Figure \@ref(fig:CADHEnrich)A).
+In the CC category, the DEGs were enriched in "azurophil granule," "ficolin-1-rich granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:CADHEnrich)B).
+In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(fig:CADHEnrich)C).
+KEGG pathway analysis showed that the DEGs were related to the following pathways: "Chemokine signaling pathway," "Lipid and atherosclerosis," and "Hematopoietic cell lineage" (Figure \@ref(fig:CADHEnrich)D).
 
@@
 (D) KEGG analysis."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/enrich CAD-Healthy.pdf")
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/Enrichment CAD-Healthy.pdf")
 ```
@@
 
-### microRNAs in DEGs
+### miRNAs in DEGs
 
-Among all DEGs, just hsa-miR-186, hsa-miR-32, and hsa-miR-21 were differntially expressed miRNAs.
-The expression profile of these three miRNAs was presented in Fig\@ref(fig:DEMexp).
-Also, The ROC curves of each miRNA for each layer was presented in Fig\@ref(fig:miRROC).
-AUC for hsa-miR-21, hsa-miR-32, and hsa-miR-186 was 0.99, 1, and 0.91 respectively using a simple logistic regression model (Fig\@ref(fig:miRROC)A).
-Using the same model the accuracy of each microRNA for classifying the samples to healthy and not-healthy groups on test set was 0.92, 0.98, and 0.89 for hsa-miR-21, hsa-miR-32, and hsa-miR-186 respectively.
-Moreover, In Fig\@ref(fig:miRROC)B the ROC curve of each microRNA for classifying MI and CAD samples were presented.
-For  hsa-miR-21, hsa-miR-32, and hsa-miR-186 AUC on test set was 0.85; 0.7; and 0.82, and accuracy on test set was 0.78; 0.67; and 0.74.
+Among all DEGs, just miR-186, miR-32, and miR-21 were detected as differentially expressed miRNAs.
+The expression profile of these three miRNAs is presented in Figure \@ref(fig:Expall).
+Additionally, The ROC curves of each miRNA for each layer are presented in Figure \@ref(fig:miRROC).
+Using the logistic regression model, the AUC-ROC of miR-21, miR-32, and miR-186 for separating healthy and not-healthy samples was 0.98, 0.99, and 0.90, respectively (Figure \@ref(fig:miRROC)A).
+Besides, the accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set was 0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
+Moreover, the ROC curve of each miRNA for classifying MI and CAD samples was presented in Figure \@ref(fig:miRROC)B.
+The AUC-ROC and accuracy for miR-21, miR-32, and miR-186 on the test set were 0.85; 0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
 
-```{r DEMexp, echo=F, out.width="50%", fig.align='center', fig.cap="Expression profile of hsa-miR-186, hsa-miR-21, and hsa-miR32 in Healthy, CAD, and MI samples."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEGs healthy not-healthy/DEMs Expression.pdf")
+```{r mirExptable, echo=FALSE}
+
+# data
+df <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/microRNAs logFC.txt", sep = "\t", check.names = F, header = T)
+
+kbl(df, booktabs = T, align = "c",linesep = c(""), digits = c(1, 60), caption = "Investigated miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.") %>%
+  kable_styling(full_width = T) %>%
+  add_header_above(c(" " = 1, "CAD/Healthy" = 2, "MI/Healthy" = 2, "MI/CAD" = 2))
 ```
 
-```{r miRROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for single microRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+```{r Expall, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of all miRNAs in two approaches in different sample classes."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Expressionforall.pdf")
+```
+
+```{r miRROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/miRs ROCs.pdf")
@@
 
-#### First layer for healthy not-healthy seperation:
+#### First layer for healthy not-healthy isolation:
 
-Although single microRNAs had acceptable performance, but their prediction value could improve even more using them as a set.
-The ROC curve for the SVM model with rbf kernel trained with all three microRNAs is presented in Fig\@ref(fig:DEMROC)A.
-The model had better performance in classification than single microRNAs.The AUC for the model is 1, and its accuracy on test set was 1 as well.
-The confusion matrix for the model is presented in Fig\@ref(fig:DEMCF)A.
+Although single miRNAs had an acceptable performance for this layer, their predictive value could be improved even further by using them as a set.
+The ROC curve for the SVM model with an RBF kernel trained with all three miRNAs is presented in Figure \@ref(fig:DEMROC)A.
+The model had a better performance in classification than single miRNAs.The AUC-ROC for the model is 1, and its accuracy on the test set was also 1.
+In Figure \@ref(fig:DEMCF)A, the confusion matrix for the model is presented.
 
-```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for microRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs ROCs h not h cad mi.pdf")
@@
 
-```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy samples classification.
-and (B) An SVM model with linear kernel for CAD and MI samples classification."}
+```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs Conf. matrix.pdf")
@@
 
-Different models were trained using expression values of three differentially microRNAs.
-The models 10-fold cross-validate AUC and accuracy on the test set were reported on Fig\@ref(fig:DEMmodels).
-The best model from both AUC and accuracy point-of-view was the SVM model with linear kernel.
-The AUC and accuracy for this model with its preset values was 0.93 and 0.82 respectively.
-The model was hyper-tuned for C and gamma hyperparameters, and therefore the model showed a better performance.
-The ROC curve of the hyper-tuned model was presented in Fig\@ref(fig:DEMROC)B.
-For this model the AUC reached to 0.95 and the accuracy improved to 0.85.
-Moreover, the sensitivity and specificity for the model on the test set was 0.91 and 0.71 respectively.
-The confusion matrix for the hyper-tuned model is illustrated in Fig\@ref(fig:DEMCF)B.
+Different models were trained using expression values for three differentially expressed miRNAs.
+The models' AUC-ROC and accuracy on the test set are reported in Figure \@ref(fig:DEMmodels).
+The best model from both AUC-ROC and accuracy point-of-view was the SVM model with a linear kernel.
+The AUC-ROC and accuracy for this model with its pre-set values were 0.93 and 0.82, respectively.
+The model was hyper-tuned for C and gamma hyper-parameters, and therefore the model showed better performance.
+The ROC curve of the hyper-tuned model is presented in Figure \@ref(fig:DEMROC)B.
+For this model, the AUC-ROC reached 0.95, and the accuracy improved to 0.85 (Table \@ref(tab:DEGsML)).
+Moreover, the sensitivity and specificity for the model on the test set were 0.91 and 0.71, respectively.
+The confusion matrix for the hyper-tuned model is illustrated in Figure \@ref(fig:DEMCF)B.
 
-```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) and accuracy of different models trained with three microRNAs in DEGs on the test set."}
+```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models DEGs mirs.pdf")
@@
 
-### AUC approach
+```{r DEGsML, echo=FALSE}
 
-After calculating each microRNA's AUC for classifying MI and CAD samples, they were sorted and their performance as a set were investigated.
-The metric of choice for finding the best set was AUC.
-As shown in Fig\@ref(fig:AUCset), the AUC increased until  the number of microRNAs in the set reached tho 6 and after that it droped.
-The AUC for separating MI samples from CAD using these microRNAs was 0.93.
-The microRNAs in the set were has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1.
-The expression values of these microRNAs in healthy, CAD, and MI samples is presented in  Fig\@ref(fig:AUCexp).
-The ROC curve of the selected microRNAs for MI and CAD samples classification is illustrated in Fig\@ref(fig:miRROC)B.
+# data
+df <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEGsMICAD.txt", sep = "\t", check.names = F, header = T)
 
-```{r AUCset, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) for sets containing increasing number of microRNAs with the highest individual AUC in MI/CAD separation."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC for cad mi mirs.pdf")
+kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with a linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning") %>%
+  kable_styling(full_width = T) %>%
+  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
 ```
 
-```{r AUCexp, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/ExpressionforAUC.pdf")
-```
+### AUC-ROC approach
 
-#### First layer:
+After calculating the AUC-ROC for each miRNA to classify of MI and CAD samples, the miRNAs with AUC-ROC over 0.8 were selected.
+The selected miRNAs were miR-29a, miR-197, miR-186, miR-21, and miR-296.
+The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:Expall).
+The ROC curve of the selected miRNAs for both layers is illustrated in Figure \@ref(fig:miRROC).
 
-Using the selected set, an SVM model with RBF kernel was trained for separating healthy from not-healthy samples.
-The ROC curve for the model is presented in Fig\@ref(fig:AUCROC)A and confusion matrix is illustrated in Fig\@ref(fig:AUCCF)A.
-Both AUC and accuracy for the model on the test set was equal to 1.
+#### First layer for healthy not-healthy isolation:
 
-```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of microRNAs selected by AUC on test set classification.
-(A) SVM with RBF kernel for healthy and not-healthy samples classification.
+Using the selected set, an SVM model with an RBF kernel was trained to separate healthy from not-healthy samples.
+The ROC curve for the model is presented in Figure \@ref(fig:AUCROC)A, and the confusion matrix is illustrated in Figure \@ref(fig:AUCCF)A.
+Both AUC-ROC and accuracy for the model on the test set were equal to 1.
+
+```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
 (B) Logistic regression model for CAD and MI samples classification.
@@
 
-```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
 (B) Logistic regression model for CAD and MI samples classification.
@@
 
-#### Second layer; MI form CAD:
+#### Second layer for separating MI samples from CAD:
 
-For finding the best model for training the best set, different models were trained using their pre-set values.
-Their AUC and accuracy results on the test set is presented in Fig\@ref(fig:AUCmodels).
-The best model from AUC point-of-view was the LR and from accuracy point-of-view was the SVM model with polynomial kernel.
-For LR model the AUC and accuracy were 0.92 and 0.81; and for SVM model with polynomial kernel the values were 0.91 and 0.84 respectively.
-Both models were hyper-tuned and ROC curve for their best performance presented in Fig\@ref(fig:AUCROC)B and C.
-The AUC and accuracy for LR model increased to 0.94 and 0.88 respectively; And for SVM model with polynomial kernel these values increased to 0.95 and 0.88 respectively.
-The sensitivity for LR and SVM models were 1 and 0.97; and the specificity for them were 0.57 and 0.64, respectively.
-The confusion matrix for both models is illustrated in Fig\@ref(fig:AUCCF)B andC.
+To find the best model for this set of miRNAs, different models were trained using their pre-set values.
+Their AUC-ROC and accuracy results on the test set are presented in Figure \@ref(fig:AUCmodels).
+The best model from the AUC-ROC point-of-view was the SVM with a linear kernel, and from the accuracy point-of-view, it was the SVM model with an RBF kernel.
+For the SVM-linear model, the AUC-ROC and accuracy were 0.93 and 0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84, respectively.
+Both models were hyper-tuned, and the ROC curve for their best performance is presented in Figure \@ref(fig:AUCROC)B and C.
+The AUC-ROC and accuracy for the SVM-linear model were modified to 0.92 and 0.88, respectively.
+For the SVM-RBF, these values increased to 0.96 and 0.94, respectively (Table \@ref(tab:AUCML)).
+The sensitivity for the SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and the specificity for them was 0.79 and 0.86, respectively.
+The confusion matrix for both models is illustrated in Figure \@ref(fig:AUCCF)B and C.
 
-```{r AUCmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under curve (AUC) and accuracy of different models trained with three microRNAs in DEGs on the test set."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models AUC mirs.pdf")
+```{r AUCmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the receiver operating characteristic curve and accuracy of different models trained with AUC-selected miRNAs."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models AUC 5 mirs CV.pdf")
 ```
 
-# Refrences
+```{r AUCML, echo=FALSE}
+
+# data
+df <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUCMICAD.txt", sep = "\t", check.names = F, header = T)
+
+kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their AUC-ROC on the train and test set before and after hyper-tuning") %>%
+  kable_styling(full_width = T) %>%
+  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
+```
+
+# Discussion
+
+The prevalence of MI can lead to high-rate mortality in the clinical setting.
+However, early diagnosis and application of suitable treatment protocols can reduce mortality and improve MI prognosis [@CVD; @17; @18].
+Studies have suggested that changes in miRNA expression may play a significant role in the progression of MI and the subsequent remodeling [@23].
+It is believed that the expression of miRNAs is altered during the various biological processes correlated with MI within the myocardium or other related tissues [@24].
+Although several researches have concentrated on examining free circulating miRNAs in serum samples for the detection of cardiac tissue injuries [@11-15], more information is needed to fully comprehend the miRNAs found in different blood sub-components like plasma, platelets, and PBMCs.
+Based on previous findings, PBMCs play a crucial role in the destabilization and rupture of plaques and also the initial inflammatory reactions in individuals experiencing a myocardial infarction (MI).
+[@Meder6; @Meder23].
+Moreover, PBMCs have specific miRNA profile that is altered under certain pathological conditions, which are great candidates as disease biomarkers [@Meder6].
+
+PBMCs can respond to several insulting conditions, such as MI, in the least possible time with notable changes in their miRNA profile [@Meder6].
+Considering the regulatory roles, subtle changes in the transcription of miRNAs can be monitored even before alteration in the levels of mRNAs and proteins [@miR].
+These features make the miRNAs an early-stage valid diagnostic tool for the detection of minor and major cell injuries.
+To date, few studies have been performed to compare the miRNA profiles in PBMCs belonging to MI patients and other CADs and healthy samples to find a robust set of identical miRNAs to differentiate these pathological conditions.
+
+In this study, we combined three GEO datasets of healthy, CAD, and MI samples.
+Having these samples set alongside bioinformatics analysis and ML means enabled us to identify potential biomarker sets and effective therapeutic targets.
+The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(fig:venn)) are proof of the close relationship between the MI and CAD samples.
+Interestingly, functional enrichment analysis demonstrated that DEGs in both CAD/healthy and MI/healthy were strongly correlated to immune cell response, which is a major cellular part of PBMCs.
+Here, two sets of miRNAs were selected as biomarker sets for sample classification.
+miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-186, miR-21; miR-29a; miR-197; and miR-296 were selected according to their AUC-ROC values.
+As shown in Figure \@ref(fig:miRROC), all miRNAs selected with both approaches had AUC-ROC over 0.9 for isolating healthy and not-healthy samples except for miR-296 and miR-29a.
+Data confirmed that the real challenge is to classify CAD and MI samples because of close overlap.
+Of 6 miRNAs under investigation in both approaches, except for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of CAD and MI samples.
+As expected, the high AUC-ROC values of miRNAs confirm their high potential as biomarkers.
+
+ML models, trained with miRNA sets selected by both DEG and AUC-ROC approaches, showed better performance in the classification than each miRNA.
+To avoid unwanted complexity and poor predictive values, a two-layer architecture was also designed.
+The first layer was for the discrimination of healthy from not-healthy samples, and the second layer was for separating the CAD from MI candidates.
+As expected, in both approaches, a hyper-tuned SVM model could flawlessly separate healthy from not-healthy samples using distinct miRNA sets.
+The ML models were also capable of effectively separating CAD from MI patients.
+Although both miRNA sets had nearly the same AUC-ROC with their best model, the accuracy, sensitivity, and specificity were different.
+The model trained with AUC-selected miRNAs had better performance in all predictive values, which is logical because of more miRNAs in the set.
+
+Numerous studies have reported different biological processes can affect the expression of miRNAs in PBMCs.
+However, there are still controversies regarding the exact role of miRNAs in the function of immune cells and the correlation of specific pathological conditions with miRNA profiles.
+Several studies have proved the activation of particular miRNA types in PBMCs under cardiovascular events [@21; @186; @32-2; @29aa; @197, @296].
+For instance, there is evidence that the elevation of miR-186 suppresses the expression of Cystathionine-$\gamma$-lyase, leading to the subsequent secretion of pro-inflammatory cytokines and cellular lipid accumulation.
+Besides, macrophage-derived miR-186 may promote atherosclerotic plaques [@186].
+In line with this claim, we found that miR-186 is up-regulated in both CAD and MI candidates related to control counterparts.
+Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients than in MI (Figure \@ref(fig:DEMexp)).
+To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, with a clear up-regulation in CAD, indicating its main role in the promotion of atherosclerosis.
+
+As mentioned before, miR-21 was also up-regulated in both MI and CAD patients in comparison to healthy controls.
+Moreover, the expression value of miR-21 was significantly higher in MI than in the CAD group (Table \@ref(tab:mirExptable)).
+It is thought that the up-regulation of miRNA-21 in PBMCs is a compensatory reaction to reduce T$_{reg}$ lymphocyte number in response to the reduction of TGF$\beta1$ secretion into the plasma through a TGF$\beta1$/smad-independent pathway.
+In line with previous and present data, miR-21 can modulate the activity of PBMCs following the occurrence of cardiovascular diseases [@21].
+
+Recent data have supported the elevation of miR-32 in CAD samples with the calcification of the coronary artery.
+It is worth noting that miR-32 promotes vascular smooth muscle calcification in mice by controlling the activity of several proteins, including bone morphogenetic protein-1, runt-related transcription factor-2 (RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA protein.
+[@32-2].
+Likewise, some reports are associated with the activity of miR-32 in PBMCs under several pathologies [@magic; @32].
+The exact role of miR-32 in PBMCs after cardiovascular events remains to be elucidated.
+
+Molecular analyses have indicated the regulatory role of miRNAs selected by the AUC-ROC approach in PBMCs after a cardiovascular event.
+Two common miRNAs in DEGs and AUC-ROC approaches, miR-21 and miR-186, were covered already.
+Based on numerous reports, miR-29a can be activated in different diseases [@29aa].
+Data analysis indicated that miR-29a is significantly up-regulated in CAD patients compared to healthy and MI groups (Table \@ref(tab:mirExptable)).
+Increased miR-29a is associated with the progression of atherosclerosis, and the combination of miR-29a and ox-LDL was offered as a valid biomarker set for paraclinical classification [@29a].
+However, the role of miR-29a in the function of PBMCs in CAD patients has not been thoroughly examined.
+
+Data indicated that miR-197 is also significantly up-regulated in both CAD/healthy and MI/healthy groups.
+Previous research has demonstrated that miR-197 may play a crucial role in controlling the anti-inflammatory response of IL-35 by influencing the secretion of cytokines that can either promote or suppress inflammation, the ratio of M1/M2 macrophages, and the proliferation of T$_{reg}$ lymphocytes, which are responsible for suppressing immune responses [@197].
+Alongside our findings, it can be concluded that miR-197 could be a useful diagnostic tool for predicting adverse cardiovascular events.
+
+The findings of this study demonstrate the potential of miR-296 as a biomarker with high discriminatory power for distinguishing between samples from individuals with MI and CAD.
+MiR-296 has been identified as a key regulator in the development and advancement of atherosclerosis by controlling the expression of target genes associated with various biological processes, including angiogenesis, cholesterol metabolism, inflammation, cellular proliferation, hypertension, and apoptosis [@296].
+In a study, it has been shown that miR-296 expression levels are significantly increased in the PBMCs of CAD patients compared to healthy controls, suggesting its involvement in regulating pro-inflammatory cytokines like IL-6 and TNF-a [@296a].
+These findings suggest that miR-296 may have a significant impact on the pathogenesis of atherosclerosis and could potentially serve as a diagnostic biomarker for CAD or MI.
+
+# Conclusion
+
+In summary, we derived a set of miRNA biomarkers by comparing MI samples to both healthy and CAD samples.
+We found that the SVM model performed best in both the first layer, which separated healthy and not-healthy samples, and the second layer, which classified MI/CAD samples.
+The set of miRNAs selected based on their AUC-ROC values performed better in the second layer.
+Overall, our two-layer structure achieved an accuracy of 0.96.
+This demonstrates the potential for combining bioinformatics and machine learning techniques to identify novel biomarkers and gain a deeper understanding of myocardial infarction.
+
+# References

### 2023-06-13 Complete-V1/Complete-V1.Rmd
Revised with Ithenticate and Grammerly

@@
 - Myocardial Infarction
+- Bioinformatics
 
@@
   MicroRNAs (miRNAs) play a crucial role in regulating adaptive and maladaptive responses in cardiovascular diseases, making them attractive targets for potential therapeutics.
-However, their potential as novel biomarkers for diagnosing cardiovascular diseases needs to be evaluated systematically.
+However, their potential as novel biomarkers for diagnosing cardiovascular diseases requires systematic evaluation.
 In this study, we aimed to identify a key set of miRNA biomarkers using integrated bioinformatics and machine learning analysis.
-We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contained peripheral blood mononuclear cells samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy.
-Additionally, we selected a set of miRNAs based on their area under the curve of the receiver operating characteristic (AUC-ROC) for separating CAD/MI samples.
+We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contains peripheral blood mononuclear cell samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy individuals.
+Additionally, we selected a set of miRNAs based on their area under the curve of the receiver operating characteristic curve (AUC-ROC) to separate the CAD/MI samples.
 We designed a two-layer architecture for sample classification, with the first layer isolating healthy samples from not-healthy ones, and the second layer classifying stable CAD and MI samples.
 We trained different machine learning models using both biomarker sets and evaluated their performance on a test set.
-We identified miR-21, miR-186, and miR-32 as the only miRNAs in the differentially expressed genes, and a set included miR-21, miR-155, miR-142, miR-197, miR-29A, and miR-320C1 as the optimum set of miRNA selected by their AUC-ROC.
-Both biomarker sets were able to distinguish healthy from not-healthy samples with complete accuracy.
-The best performance for classification of CAD and MI was achieved with an SVM model trained using the biomarker set selected by AUC-ROC, with an AUC of 0.95 and an accuracy of 0.88.
-Our study demonstrates that miRNA signatures derived from peripheral blood could serve as valuable novel biomarkers for cardiovascular diseases.
+We identified miR-21, miR-186, and miR-32 as the only miRNAs among the differentially expressed genes, and a set including miR-186, miR-21, miR-197, miR-29A, and miR-296 as the optimum set of miRNAs selected by their AUC-ROC.
+Both biomarker sets could distinguish healthy from not-healthy samples with complete accuracy.
+The best performance for the classification of CAD and MI was achieved with an SVM model trained using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.95, and an accuracy of 0.88.
+Our study demonstrated that miRNA signatures derived from peripheral blood could serve as valuable novel biomarkers for cardiovascular diseases.
 
@@
 
-bibliography: [ForArticleRefrence.bib, discussion.bib, Introduction.bib]
+bibliography: [ForArticleRefrence.bib, discussion.bib, Introduction.bib, mir296.bib]
 biblio-style: spmpsci
@@
 Over the last 50 years, numerous attempts have been collected to use biomarkers to facilitate diagnosis, assess risk, follow-up therapy, and determine therapeutic efficacy in CVDs candidates.
-Based on the released guidelines, cardiac troponins (cTns) are used as a highly-sensitive and accurate approach for the detection of myocardial ischemia.
+Based on the released guidelines, cardiac troponins (cTns) are used as a highly-sensitive and accurate approach for the detection of MI.
 Despite the inherent advantages, the high-rate sensitivity of cTn-based assays has also led to more false positive results [@17], which do necessitate the advent and development of new modalities with pathological values.
-To improve diagnostic value upon existing CVD biomarkers, the combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, is proposed.
-Previous data support the notion that miRNAs exhibit the great potential to be used as alternative biomarkers in CVD detection and follow-up [@miR].
+To improve diagnostic value upon existing MI biomarkers, the combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, is proposed.
+Previous researches support the notion that miRNAs exhibit the great potential to be used as alternative biomarkers in CVDs detection and follow-up [@miR].
 It is suggested that miRNAs possess 18-22 nucleotides and can play a crucial role in the regulation of gene expression.
-Evidence point to the fact that miRNAs are involved in the pathogenesis of cardiac tissue injury and be as theranostics in terms of CVDs.
+Evidence point to the fact that miRNAs are involved in the pathogenesis of cardiac tissue injury and could be considered as theranostics targets in terms of CVDs.
 These elements can easily circulate in biofluids with both diagnostic as well as prognostic values [@m1].
@@
 PBMCs could recapitulate the conditions of the target tissues, thus, providing a highly sensitive and specific source of biomarkers [@Meder6].
-Commensurate with these conditions, these cells are repositories of dysregulated genes and miRNAs expression profiles in CVDs related to control conditions [@PBMC-miR; @Meder6].
+Commensurate with these conditions, these cells are repositories of dysregulated genes and miRNAs expression profiles in CVDs [@PBMC-miR; @Meder6].
 
@@
 In this study, we aimed to identify potential miRNA biomarkers for MI patients by combining and analyzing three different microarray datasets from PBMCs.
-It is suggested that the integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring CVDs.
-Besides, this approach can deepen our vision into the underlying mechanisms of CVDs and aid in the development of valid theranostic tools, and patient stratification.
+It is suggested that the integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring MI.
+Besides, this approach can deepen our vision into the underlying mechanisms of MI and aid in the development of valid diagnostic biomarkers, and patient stratification.
 
@@
 Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database (<https://www.ncbi.nlm.nih.gov/geo/>).
-To obtain sufficient classification power between MI, healthy and CAD samples, a relatively large sample size was required.
+To obtain sufficient classification power between MI, healthy and stable coronary artery disease (CAD) samples, a relatively large sample size was required.
 Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475 for healthy samples were selected.
@@
 
-```{r datasets, echo=FALSE}
+```{r, datasets, echo=FALSE}
 df <- data.frame(Dataset = c("GSE59867", "GSE56609", "GSE54475"),
@@
 The first approach was using the miRNAs that are differentially expressed.
-In the second approach, miRNAs were selected by their individual AUC-ROC.
+In the second approach, miRNAs with individual area under the curve of the receiver operating characteristic (AUC-ROC) over 0.8 for separating MI from CAD were selected.
 Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of sets of miRNAs selected with different logics.
@@
 All models were trained with their pre-set parameters with 10-fold cross-validation.
-The criteria for choosing the best model were the highest accuracy and AUC on the test set.
+The criteria for choosing the best model were the highest accuracy and AUC-ROC on the test set.
 The best model was hyper-tuned with the scikit-opt package [@scikitopt] to get the best predictive performance.
@@
 The first layer classified samples into healthy and not-healthy, and the separated MI and CAD samples.
-However, to keep the number of miRNAs as low as possible miRNAs were selected from the second layer (which are the miRNAs with the best performance in MI/CAD separation), and then their performance wase evaluated in the first layer.
-AUC-ROC of all miRNAs for classifying MI and CAD samples were calculated.
-To find the number of miRNAs with the highest predictive values, the miRNAs with the highest individual AUC-ROC were added to the set one by one, and the AUC-ROC for the set was calculated.
-The set with the highest AUC-ROC was selected for the following steps.
-The ROC curves for each selected miRNA for separating healthy samples from not-healthy and MIs from CADs were also plotted for further comparison.
+However, to keep the number of miRNAs as low as possible miRNAs were selected from the second layer, and then their performance wase evaluated in the first layer.
+AUC-ROC of all miRNAs for classifying MI and CAD samples were calculated and the miRNAs with the AUC-ROC over 0.8 were selected.
+The ROC curves for each selected miRNA for separating healthy samples from not-healthy and MI from CAD samples were also plotted for further comparison.
 
@@
 An SVM model with an RBF kernel was trained using the selected set of miRNAs.
-Additionally, the model was hyper-tuned to find the hyper-parameters for the highest AUC and accuracy.
+Additionally, the model was hyper-tuned to find the hyper-parameters for the highest AUC-ROC and accuracy.
 The ROC curve and confusion matrix for the best model were reported.
@@
 
-```{r DEGstab, echo=F}
+```{r , DEGstab, echo=F}
 M_H <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-Healthy.txt", check.names = F)
@@
 
-colnames(df) <- c("DEGs",  "up-regulated DEGs", "down-regulated DEGs", "miRNAs")
+colnames(df) <- c("Total DEGs",  "Up-regulated DEGs", "Down-regulated DEGs", "miRNAs")
 rownames(df) <- c("MI vs. Healthy", "CAD vs. Healthy", "MI vs. CAD")
@@
   column_spec(1, width = "3cm") %>%
-  column_spec(5, width = "4cm")
+  column_spec(5, width = "5cm") %>%
+  column_spec(4, width = "2.5cm") %>%
+  column_spec(3, width = "2.5cm")
 ```
@@
 
-```{r venn, echo=F, out.width="60%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
+```{r venn, echo=F, out.width="45%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/venny.png")
@@
 (D) KEGG analysis."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/enrich MI-Healthy.pdf")
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/Enrichment MI-Healthy.pdf")
 ```
@@
 Among all DEGs, just miR-186, miR-32, and miR-21 were detected as differentially expressed miRNAs.
-The expression profile of these three miRNAs is presented in Figure \@ref(fig:DEMexp).
+The expression profile of these three miRNAs is presented in Figure \@ref(fig:Expall).
 Additionally, The ROC curves of each miRNA for each layer are presented in Figure \@ref(fig:miRROC).
-Using the logistic regression model the AUC for miR-21, miR-32, and miR-186 was 0.99, 1, and 0.91 respectively (Figure \@ref(fig:miRROC)A).
+Using the logistic regression model the AUC-ROC of miR-21, miR-32, and miR-186 for separating healthy and not-healthy samples was 0.98, 0.99, and 0.90 respectively (Figure \@ref(fig:miRROC)A).
 Besides, the accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set was 0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
 Moreover, the ROC curve of each miRNA for classifying MI and CAD samples was presented in Figure \@ref(fig:miRROC)B.
-For miR-21, miR-32, and miR-186, the AUC and accuracy on the test set were 0.85; 0.7; and 0.82, and 0.78; 0.67; and 0.74, respectively.
+For miR-21, miR-32, and miR-186, the AUC-ROC and accuracy on the test set were 0.85; 0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
 
-```{r, mirExptable, echo=FALSE}
+```{r mirExptable, echo=FALSE}
 
@@
 
-```{r DEMexp, echo=F, out.width="50%", fig.align='center', fig.cap="Expression profile of miR-186, miR-21, and miR32 in Healthy, CAD, and MI samples."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEGs healthy not-healthy/DEMs Expression.pdf")
+```{r Expall, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of all miRNAs in two approach in different sample classes."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Expressionforall.pdf")
 ```
@@
 The ROC curve for the SVM model with an RBF kernel trained with all three miRNAs is presented in Figure \@ref(fig:DEMROC)A.
-The model had a better performance in classification than single miRNAs.The AUC for the model is 1, and its accuracy on the test set was also 1.
+The model had a better performance in classification than single miRNAs.The AUC-ROC for the model is 1, and its accuracy on the test set was also 1.
 In Figure \@ref(fig:DEMCF)A, the confusion matrix for the model is presented.
@@
 Different models were trained using expression values for three differentially miRNAs.
-The models' 10-fold cross-validated AUC and accuracy on the test set are reported in Figure \@ref(fig:DEMmodels).
-The best model from both AUC and accuracy point-of-view was the SVM model with linear kernel.
-The AUC and accuracy for this model with its pre-set values were 0.93 and 0.82 respectively.
+The models' 10-fold cross-validated AUC-ROC and accuracy on the test set are reported in Figure \@ref(fig:DEMmodels).
+The best model from both AUC-ROC and accuracy point-of-view was the SVM model with linear kernel.
+The AUC-ROC and accuracy for this model with its pre-set values were 0.93 and 0.82 respectively.
 The model was hyper-tuned for C and gamma hyper-parameters, and therefore the model showed better performance.
 The ROC curve of the hyper-tuned model is presented in Figure \@ref(fig:DEMROC)B.
-For this model the AUC reached 0.95 and the accuracy improved to 0.85 (Table \@ref(tab:DEGsML)).
+For this model the AUC-ROC reached 0.95 and the accuracy improved to 0.85 (Table \@ref(tab:DEGsML)).
 Moreover, the sensitivity and specificity for the model on the test set were 0.91 and 0.71 respectively.
@@
 
-```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
+```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs on the test set."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models DEGs mirs.pdf")
@@
 
-### AUC approach
+### AUC-ROC approach
 
-After calculating the AUC for each miRNA for the classification of MI and CAD samples, miRNAs were sorted, and their performance was investigated as a set.
-The metric of choice for selecting the best set was AUC.
-As shown in Figure \@ref(fig:AUCset), the AUC increased until the number of miRNAs in the set reached six, and after that, it dropped.
-The AUC for separating MI samples from CAD using these miRNAs was 0.93.
-The miRNAs in the set were miR-29a, miR-197, miR-142, miR-21, miR-155, and miR-320C1.
-The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:AUCexp).
-The ROC curve of the selected miRNAs for MI and CAD sample classification is illustrated in Figure \@ref(fig:miRROC)B.
-
-```{r AUCset, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the curve (AUC) for sets containing an increasing number of miRNAs with the highest individual AUC in MI/CAD separation."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC for cad mi mirs.pdf")
-```
-
-```{r AUCexp, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/ExpressionforAUC.pdf")
-```
+After calculating the AUC-ROC for each miRNA for the classification of MI and CAD samples, the miRNAs with AUC-ROC over 0.8 were selected.
+The selected miRNAs were miR-29a, miR-197, miR-186, miR-21, and miR-296.
+The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:Expall).
+The ROC curve of the selected miRNAs for both layers are illustrated in Figure \@ref(fig:miRROC).
 
@@
 The ROC curve for the model is presented in Figure \@ref(fig:AUCROC)A and the confusion matrix is illustrated in Figure \@ref(fig:AUCCF)A.
-Both AUC and accuracy for the model on the test set were equal to 1.
+Both AUC-ROCand accuracy for the model on the test set were equal to 1.
 
-```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of miRNAs selected by AUC on test set classification.
+```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of miRNAs selected by AUC-ROC on test set classification.
 (A) SVM with RBF kernel for healthy and not-healthy samples classification.
@@
 To find the best model for training the best set, different models were trained using their pre-set values.
-Their AUC and accuracy results on the test set are presented in Figure \@ref(fig:AUCmodels).
-The best model from the AUC point-of-view was the LR and from the accuracy point-of-view, it was the SVM model with a polynomial kernel.
-For the LR model the AUC and accuracy were 0.92 and 0.81, respectively; and for the SVM model with a polynomial kernel, the values were 0.91 and 0.84, respectively.
+Their AUC-ROC and accuracy results on the test set are presented in Figure \@ref(fig:AUCmodels).
+The best model from the AUC-ROC point-of-view was the SVM with linear kernel and from the accuracy point-of-view, it was the SVM model with an RBF kernel.
+For the SVM-linear model the AUC-ROC and accuracy were 0.93 and 0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84, respectively.
 Both models were hyper-tuned and the ROC curve for their best performance is presented in Figure \@ref(fig:AUCROC)B and C.
-The AUC and accuracy for the LR model increased to 0.94 and 0.88, respectively.
-For the SVM model with a polynomial kernel, these values increased to 0.95 and 0.88, respectively (Table \@ref(tab:AUCML)).
-The sensitivity for the LR and SVM models were 1 and 0.97, respectively; and the specificity for them was 0.57 and 0.64, respectively.
+The AUC-ROC and accuracy for the SVM-linear model modified to 0.92 and 0.88, respectively.
+For the SVM-RBF, these values increased to 0.96 and 0.94, respectively (Table \@ref(tab:AUCML)).
+The sensitivity for the SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and the specificity for them was 0.79 and 0.86, respectively.
 The confusion matrix for both models is illustrated in Figure \@ref(fig:AUCCF)B and C.
 
-```{r AUCmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models AUC mirs.pdf")
+```{r AUCmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the receiver operating characteristic curve and accuracy of different models trained with AUC-selected miRNAs."}
+knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models AUC 5 mirs CV.pdf")
 ```
 
-```{r, AUCML, echo=FALSE}
+```{r AUCML, echo=FALSE}
 
@@
 
-The prevalence of acute MI (AMI) can lead to high-rate mortality in the clinical setting.
+The prevalence of MI can lead to high-rate mortality in the clinical setting.
 However, early diagnosis and application of suitable treatment protocols can reduce mortality and improve AMI prognosis [@CVD; @17; @18].
@@
 Here, two different sets of miRNAs were used as biomarker sets for sample classification.
-miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-21; miR-29a; miR-142; miR155; miR-197; and miR-320c1 were selected according to their AUC values.
-As shown in Figure \@ref(fig:miRROC), all other miRNAs selected with both approaches had AUC-ROC over 0.9 for the isolation of healthy and not-healthy samples except for miR-142 and miR-29a.
+miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-186, miR-21; miR-29a; miR-197; and miR-296 were selected according to their AUC-ROC values.
+As shown in Figure \@ref(fig:miRROC), all miRNAs selected with both approaches had AUC-ROC over 0.9 for the isolation of healthy and not-healthy samples except for miR-296 and miR-29a.
 Data confirmed that the real challenge is to classify CAD and MI samples because of close overlap.
-Of 8 miRNAs under investigation in both approaches except for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of CAD and MI samples.
+Of 6 miRNAs under investigation in both approaches except for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of CAD and MI samples.
 Besides, the high AUC-ROC values of miRNAs confirms their high potential as biomarkers.
 
-ML models when trained with miRNA sets selected by both DEG and AUC approaches showed better performance in the classification than each miRNA.
+ML models when trained with miRNA sets selected by both DEG and AUC-ROC approaches showed better performance in the classification than each miRNA.
 To avoid unwanted complexity and poor predictive values, a two-layer architecture was also designed.
-The first layer was for for the discrimination of healthy from not-healthy samples, and the second layer separated the CAD from MI candidates.
-As expected in both approaches, a hyper-tuned SVM model could flawlessly separates healthy from not-healthy samples using distinct miRNAs sets.
+The first layer was for the discrimination of healthy from not-healthy samples, and the second layer separated the CAD from MI candidates.
+As expected, in both approaches a hyper-tuned SVM model could flawlessly separates healthy from not-healthy samples using distinct miRNAs sets.
 The ML models were also capable of effectively separating CAD from MI patients.
 Although both miRNA sets had nearly the same AUC-ROC with their best model, the accuracy, sensitivity, and specificity were different.
-The model trained with DEGs had better specificity, but the one trained with AUC-selected miRNAs had slightly better accuracy and higher sensitivity.
-This difference comes from the different strategies the selection of biomarkers sets.
+The model trained with AUC-selected miRNAs had better performance in all predictive values.
 
@@
 However, there are still controversies regarding the exact role of miRNAs in the function of immune cells and the correlation of specific pathological conditions with miRNA profiles.
-Several studies have proved the activation of specific miRNA types in PBMCs under cardiovascular events [@21; @186; @32-2; @29aa; @142; @155-3; @197].
+Several studies have proved the activation of specific miRNA types in PBMCs under cardiovascular events [@21; @186; @32-2; @29aa; @197].
 For instance, there is evidence that the elevation of miR-186 suppresses the expression of Cystathionine-$\gamma$-lyase, leading to the subsequent secretion of pro-inflammatory cytokines and cellular lipid accumulation.
@@
 In line with this claim, we found that miR-186 is up-regulated in both CAD and MI candidates related to control counterparts.
-Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients in comparison to MI (Figure \@ref(fig:DEMexp)), To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, indicating its main role in the promotion of atherosclerosis.
+Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients in comparison to MI (Figure \@ref(fig:DEMexp)), To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, with a clear up-regulation in CAD, indicating its main role in the promotion of atherosclerosis.
 
@@
 
-Molecular analyses have indicated the regulatory role of miRNAs selected by the AUC approach in PBMCs after a cardiovascular event.
-As the only common miRNA of the two sets, molecular function of miR-21 in CVDs was covered in DEGs miRNA set.
+Molecular analyses have indicated the regulatory role of miRNAs selected by the AUC-ROC approach in PBMCs after a cardiovascular event.
+Two common miRNAs in DEGs and AUC-ROC approachs, miR-21 and miR-186, were covered in DEGs miRNA set.
 Based on numerous reports mir-29a can be activated in different diseases [@29aa].Data analysis indicated that miR-29a is significantly up-regulated in CAD patients in comparison to healthy and MI groups (Table \@ref(tab:mirExptable)).
@@
 
-According to the present data (Table \@ref(tab:mirExptable)), the difference in miR-142 expression between MI/Healthy and CAD/Healthy was not significant, but these values reached statistically significant in MI as compared to CAD.
-Based on different reports, miR-142 is commonly active in PBMCs.
-This miRNA directly targets and inhibits the expression of Adenomatous Polyposis Coli, a negative WNT signaling pathway regulator, contributing to the activation of the WNT signaling pathway and cardiac fibroblast activation after myocardial ischemia [@142].
-
-We also noted that miR-155 is the only down-regulated miRNA among all miRNAs in both CAD/healthy and MI/Healthy comparisons (Table \@ref(tab:mirExptable)).
-There are numerous reports on the down-regulation of miR-155 in PBMCs after MI or in CAD patients.
-Previous data suggested an inverse relationship between miR-155 levels and the severity of the coronary artery condition [@155-2; @155-3].
-Along with current data and previous findings, it is postulated that miR-155 has a protective effect under pathological conditions.
-This genetic element can control inflammation and thus reduce tissue damage through its negative feedback effects on inflammatory factors.
-These features coincide with the reduction of atherosclerosis [@155-2].
-
 Data indicated that miR-197 is also significantly up-regulated in both CAD/healthy and MI/healthy groups.
 In previous studies, it has been shown that miRNA-197 may play a critical role in regulating the anti-inflammatory response of IL-35 by affecting pro/ anti-inflammatory cytokine secretion, M1/M2 macrophage ratio, T$_{reg}$ lymphocyte proliferation, and T cell suppression suggesting the potential diagnostic role of miR-197 in adverse cardiovascular events [@197].
-There are some reports about miR-320 role in the physiopathology of cardiac fibrosis.
-Mechanistically, Molecular analyses revealed that miR-320 might induce various effects by targeting PLEKHM3 and IFITM1 in cardiomyocytes and cardiac fibroblasts, respectively [@320].
-However, there are no reports on miR-320c1 activity in the cardiovascular system or PBMCs.
+
+The findings of this study demonstrate the potential of miR-296 as a biomarker with high discriminatory power for distinguishing between samples from individuals with MI and CAD.
+MiR-296 has been identified as a key regulator in the development and advancement of atherosclerosis by controlling the expression of target genes associated with various biological processes including angiogenesis, inflammation, cholesterol metabolism, hypertension, cellular proliferation, and apoptosis [@296].
+In a study it has been shown that miR-296 expression levels are significantly increased in the PBMCs of CAD patients compared to healthy controls, suggesting its involvement in regulating pro-inflammatory cytokines like IL-6 and TNF-a [@296a].
+These findings suggest that miR-296 may have a significant impact on the pathogenesis of atherosclerosis and could potentially serve as a diagnostic biomarker for CAD or MI.
 
@@
 We found that the SVM model performed best in both the first layer, which separated healthy and not-healthy samples, and the second layer, which classified MI/CAD samples.
-The set of miRNAs selected based on their AUC values had slightly better performance in the second layer.
-Overall, our two-layer structure achieved a weighted accuracy of 0.91.
+The set of miRNAs selected based on their AUC-ROC values had better performance in the second layer.
+Overall, our two-layer structure achieved a accuracy of 0.96.
 This demonstrates the potential for combining bioinformatics and machine learning techniques to identify novel biomarkers and gain a deeper understanding of myocardial infarction.

### 2023-06-13 Complete-V1/Complete-V1.tex
Revised with Ithenticate and Grammerly

@@
 targets for potential therapeutics. However, their potential as novel
-biomarkers for diagnosing cardiovascular diseases needs to be evaluated
-systematically. In this study, we aimed to identify a key set of miRNA
+biomarkers for diagnosing cardiovascular diseases requires systematic
+evaluation. In this study, we aimed to identify a key set of miRNA
 biomarkers using integrated bioinformatics and machine learning
 analysis. We combined and analyzed three gene expression datasets from
-the Gene Expression Omnibus (GEO) database, which contained peripheral
-blood mononuclear cells samples from individuals with myocardial
-infarction (MI), stable coronary artery disease (CAD), and healthy.
-Additionally, we selected a set of miRNAs based on their area under the
-curve of the receiver operating characteristic (AUC-ROC) for separating
-CAD/MI samples. We designed a two-layer architecture for sample
-classification, with the first layer isolating healthy samples from
-not-healthy ones, and the second layer classifying stable CAD and MI
-samples. We trained different machine learning models using both
-biomarker sets and evaluated their performance on a test set. We
-identified miR-21, miR-186, and miR-32 as the only miRNAs in the
-differentially expressed genes, and a set included miR-21, miR-155,
-miR-142, miR-197, miR-29A, and miR-320C1 as the optimum set of miRNA
-selected by their AUC-ROC. Both biomarker sets were able to distinguish
-healthy from not-healthy samples with complete accuracy. The best
-performance for classification of CAD and MI was achieved with an SVM
-model trained using the biomarker set selected by AUC-ROC, with an AUC
-of 0.95 and an accuracy of 0.88. Our study demonstrates that miRNA
-signatures derived from peripheral blood could serve as valuable novel
-biomarkers for cardiovascular diseases.
+the Gene Expression Omnibus (GEO) database, which contains peripheral
+blood mononuclear cell samples from individuals with myocardial
+infarction (MI), stable coronary artery disease (CAD), and healthy
+individuals. Additionally, we selected a set of miRNAs based on their
+area under the curve of the receiver operating characteristic curve
+(AUC-ROC) to separate the CAD/MI samples. We designed a two-layer
+architecture for sample classification, with the first layer isolating
+healthy samples from not-healthy ones, and the second layer classifying
+stable CAD and MI samples. We trained different machine learning models
+using both biomarker sets and evaluated their performance on a test set.
+We identified miR-21, miR-186, and miR-32 as the only miRNAs among the
+differentially expressed genes, and a set including miR-186, miR-21,
+miR-197, miR-29A, and miR-296 as the optimum set of miRNAs selected by
+their AUC-ROC. Both biomarker sets could distinguish healthy from
+not-healthy samples with complete accuracy. The best performance for the
+classification of CAD and MI was achieved with an SVM model trained
+using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.95,
+and an accuracy of 0.88. Our study demonstrated that miRNA signatures
+derived from peripheral blood could serve as valuable novel biomarkers
+for cardiovascular diseases.
 \\
@@
 guidelines, cardiac troponins (cTns) are used as a highly-sensitive and
-accurate approach for the detection of myocardial ischemia. Despite the
-inherent advantages, the high-rate sensitivity of cTn-based assays has
-also led to more false positive results (Thygesen et al. 2018), which do
+accurate approach for the detection of MI. Despite the inherent
+advantages, the high-rate sensitivity of cTn-based assays has also led
+to more false positive results (Thygesen et al. 2018), which do
 necessitate the advent and development of new modalities with
-pathological values. To improve diagnostic value upon existing CVD
+pathological values. To improve diagnostic value upon existing MI
 biomarkers, the combination of complementary biological markers, such as
-microRNAs (miRNAs) and other genetic factors, is proposed. Previous data
-support the notion that miRNAs exhibit the great potential to be used as
-alternative biomarkers in CVD detection and follow-up (Schulte et al.
-2020). It is suggested that miRNAs possess 18-22 nucleotides and can
-play a crucial role in the regulation of gene expression. Evidence point
-to the fact that miRNAs are involved in the pathogenesis of cardiac
-tissue injury and be as theranostics in terms of CVDs. These elements
-can easily circulate in biofluids with both diagnostic as well as
-prognostic values (Schulte, Karakas, and Zeller 2017). Several
-biological activities such as angiogenesis, cardiomyocyte growth and
-contractility, lipid metabolism, plaque formation, and cardiac rhythm
-are regulated by miRNAs (Kalayinia et al. 2021). It is postulated that
-the function and diagnostic properties of miRNAs are beyond the
-myocardium in CVD patients. To be specific, the expression of miRNAs can
-vary in different biofluids and cell components such as serum and
-peripheral blood mononuclear cells (PBMCs) (Soler-Botija, Gálvez-Montón,
-and Bayés-Genís 2019).
+microRNAs (miRNAs) and other genetic factors, is proposed. Previous
+researches support the notion that miRNAs exhibit the great potential to
+be used as alternative biomarkers in CVDs detection and follow-up
+(Schulte et al. 2020). It is suggested that miRNAs possess 18-22
+nucleotides and can play a crucial role in the regulation of gene
+expression. Evidence point to the fact that miRNAs are involved in the
+pathogenesis of cardiac tissue injury and could be considered as
+theranostics targets in terms of CVDs. These elements can easily
+circulate in biofluids with both diagnostic as well as prognostic values
+(Schulte, Karakas, and Zeller 2017). Several biological activities such
+as angiogenesis, cardiomyocyte growth and contractility, lipid
+metabolism, plaque formation, and cardiac rhythm are regulated by miRNAs
+(Kalayinia et al. 2021). It is postulated that the function and
+diagnostic properties of miRNAs are beyond the myocardium in CVD
+patients. To be specific, the expression of miRNAs can vary in different
+biofluids and cell components such as serum and peripheral blood
+mononuclear cells (PBMCs) (Soler-Botija, Gálvez-Montón, and Bayés-Genís
+2019).
 
@@
 conditions, these cells are repositories of dysregulated genes and
-miRNAs expression profiles in CVDs related to control conditions (Gao et
-al. 2020; Mosallaei et al. 2022).
+miRNAs expression profiles in CVDs (Gao et al. 2020; Mosallaei et al.
+2022).
 
@@
 bioinformatics and ML techniques could be a promising tool in the
-discovery of new and more accurate biomarkers for monitoring CVDs.
+discovery of new and more accurate biomarkers for monitoring MI.
 Besides, this approach can deepen our vision into the underlying
-mechanisms of CVDs and aid in the development of valid theranostic
-tools, and patient stratification.
+mechanisms of MI and aid in the development of valid diagnostic
+biomarkers, and patient stratification.
 
@@
 database (\url{https://www.ncbi.nlm.nih.gov/geo/}). To obtain sufficient
-classification power between MI, healthy and CAD samples, a relatively
-large sample size was required. Therefore, GSE59867 for MI and CAD
-samples, and GSE56609 and GSE54475 for healthy samples were selected.
-All samples were produced using Affymetrix Human Gene 1.0 ST Array
-(GPL6244) platform. Only healthy, stable CAD and early-stage MI samples
-were selected from these datasets for further analyses. The basic
-information for the three datasets evaluated in the current study is
-provided in Table @ref(tab:datasets).
+classification power between MI, healthy and stable coronary artery
+disease (CAD) samples, a relatively large sample size was required.
+Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475
+for healthy samples were selected. All samples were produced using
+Affymetrix Human Gene 1.0 ST Array (GPL6244) platform. Only healthy,
+stable CAD and early-stage MI samples were selected from these datasets
+for further analyses. The basic information for the three datasets
+evaluated in the current study is provided in Table @ref(tab:datasets).
 
@@
 training. The first approach was using the miRNAs that are
-differentially expressed. In the second approach, miRNAs were selected
-by their individual AUC-ROC. Having the result of these two different
-approaches can provide an informative comparison between the predictive
-capabilities of sets of miRNAs selected with different logics.
+differentially expressed. In the second approach, miRNAs with individual
+area under the curve of the receiver operating characteristic (AUC-ROC)
+over 0.8 for separating MI from CAD were selected. Having the result of
+these two different approaches can provide an informative comparison
+between the predictive capabilities of sets of miRNAs selected with
+different logics.
 
@@
 cross-validation. The criteria for choosing the best model were the
-highest accuracy and AUC on the test set. The best model was hyper-tuned
-with the scikit-opt package (Head et al. 2021) to get the best
-predictive performance.
+highest accuracy and AUC-ROC on the test set. The best model was
+hyper-tuned with the scikit-opt package (Head et al. 2021) to get the
+best predictive performance.
 
@@
 separated MI and CAD samples. However, to keep the number of miRNAs as
-low as possible miRNAs were selected from the second layer (which are
-the miRNAs with the best performance in MI/CAD separation), and then
+low as possible miRNAs were selected from the second layer, and then
 their performance wase evaluated in the first layer. AUC-ROC of all
-miRNAs for classifying MI and CAD samples were calculated. To find the
-number of miRNAs with the highest predictive values, the miRNAs with the
-highest individual AUC-ROC were added to the set one by one, and the
-AUC-ROC for the set was calculated. The set with the highest AUC-ROC was
-selected for the following steps. The ROC curves for each selected miRNA
-for separating healthy samples from not-healthy and MIs from CADs were
-also plotted for further comparison.
+miRNAs for classifying MI and CAD samples were calculated and the miRNAs
+with the AUC-ROC over 0.8 were selected. The ROC curves for each
+selected miRNA for separating healthy samples from not-healthy and MI
+from CAD samples were also plotted for further comparison.
 
@@
 miRNAs. Additionally, the model was hyper-tuned to find the
-hyper-parameters for the highest AUC and accuracy. The ROC curve and
+hyper-parameters for the highest AUC-ROC and accuracy. The ROC curve and
 confusion matrix for the best model were reported.
@@
 \centering
-\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\centering}X>{\centering}X>{\centering}X>{\centering\arraybackslash}p{4cm}}
+\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\centering}X>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{5cm}}
 \toprule
-  & DEGs & up-regulated DEGs & down-regulated DEGs & miRNAs\\
+  & Total DEGs & Up-regulated DEGs & Down-regulated DEGs & miRNAs\\
 \midrule
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{../../DEG and DEM/Pics/venny}
+{\centering \includegraphics[width=0.45\linewidth]{../../DEG and DEM/Pics/venny}
 
@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/enrich MI-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs.
+\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/Enrichment MI-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs.
 (A) Biological process terms.
@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/enrich CAD-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs.
+\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/Enrichment CAD-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs.
 (A) Biological process terms.
@@
 differentially expressed miRNAs. The expression profile of these three
-miRNAs is presented in Figure @ref(fig:DEMexp). Additionally, The ROC
+miRNAs is presented in Figure @ref(fig:Expall). Additionally, The ROC
 curves of each miRNA for each layer are presented in Figure
-@ref(fig:miRROC). Using the logistic regression model the AUC for
-miR-21, miR-32, and miR-186 was 0.99, 1, and 0.91 respectively (Figure
+@ref(fig:miRROC). Using the logistic regression model the AUC-ROC of
+miR-21, miR-32, and miR-186 for separating healthy and not-healthy
+samples was 0.98, 0.99, and 0.90 respectively (Figure
 @ref(fig:miRROC)A). Besides, the accuracy of each miRNA for classifying
@@
 was presented in Figure @ref(fig:miRROC)B. For miR-21, miR-32, and
-miR-186, the AUC and accuracy on the test set were 0.85; 0.7; and 0.82,
-and 0.78; 0.67; and 0.74, respectively.
+miR-186, the AUC-ROC and accuracy on the test set were 0.85; 0.70; and
+0.86, and 0.78; 0.67; and 0.74, respectively.
 
@@
 miR-32 & 2.5 & 8.39e-43 & 2.2 & 3.10e-59 & -0.3 & 7.60e-04\\
-miR-155 & -0.6 & 2.49e-12 & -0.9 & 7.59e-33 & -0.2 & 2.68e-07\\
-miR-142 & 0.2 & 1.90e-01 & -0.1 & 1.70e-01 & -0.3 & 2.90e-04\\
 miR-197 & 0.5 & 2.95e-20 & 0.7 & 1.59e-47 & 0.2 & 8.58e-09\\
 miR-29A & 0.7 & 7.76e-29 & 0.1 & 1.70e-01 & -0.5 & 2.14e-10\\
-miR-320C1 & 0.8 & 4.72e-22 & 0.5 & 1.89e-30 & -0.2 & 1.84e-05\\
+miR-296 & -0.1 & 5.00e-02 & 0.1 & 2.00e-02 & 0.2 & 6.15e-06\\
 \bottomrule
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{../../ML/Pics/DEGs healthy not-healthy/DEMs Expression}
+{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/Expressionforall}
 
@@
 
-\caption{Expression profile of miR-186, miR-21, and miR32 in Healthy, CAD, and MI samples.}\label{fig:DEMexp}
+\caption{Expression profile of all miRNAs in two approach in different sample classes.}\label{fig:Expall}
 \end{figure}
@@
 is presented in Figure @ref(fig:DEMROC)A. The model had a better
-performance in classification than single miRNAs.The AUC for the model
-is 1, and its accuracy on the test set was also 1. In Figure
+performance in classification than single miRNAs.The AUC-ROC for the
+model is 1, and its accuracy on the test set was also 1. In Figure
 @ref(fig:DEMCF)A, the confusion matrix for the model is presented.
@@
 Different models were trained using expression values for three
-differentially miRNAs. The models' 10-fold cross-validated AUC and
+differentially miRNAs. The models' 10-fold cross-validated AUC-ROC and
 accuracy on the test set are reported in Figure @ref(fig:DEMmodels). The
-best model from both AUC and accuracy point-of-view was the SVM model
-with linear kernel. The AUC and accuracy for this model with its pre-set
-values were 0.93 and 0.82 respectively. The model was hyper-tuned for C
-and gamma hyper-parameters, and therefore the model showed better
-performance. The ROC curve of the hyper-tuned model is presented in
-Figure @ref(fig:DEMROC)B. For this model the AUC reached 0.95 and the
-accuracy improved to 0.85 (Table @ref(tab:DEGsML)). Moreover, the
-sensitivity and specificity for the model on the test set were 0.91 and
-0.71 respectively. The confusion matrix for the hyper-tuned model is
-illustrated in Figure @ref(fig:DEMCF)B.
+best model from both AUC-ROC and accuracy point-of-view was the SVM
+model with linear kernel. The AUC-ROC and accuracy for this model with
+its pre-set values were 0.93 and 0.82 respectively. The model was
+hyper-tuned for C and gamma hyper-parameters, and therefore the model
+showed better performance. The ROC curve of the hyper-tuned model is
+presented in Figure @ref(fig:DEMROC)B. For this model the AUC-ROC
+reached 0.95 and the accuracy improved to 0.85 (Table @ref(tab:DEGsML)).
+Moreover, the sensitivity and specificity for the model on the test set
+were 0.91 and 0.71 respectively. The confusion matrix for the
+hyper-tuned model is illustrated in Figure @ref(fig:DEMCF)B.
 
@@
 
-\caption{Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set.}\label{fig:DEMmodels}
+\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs on the test set.}\label{fig:DEMmodels}
 \end{figure}
@@
 \midrule
-SVM & AUC-ROC & 0.91 & 0.93 & 0.92 & 0.95\\
-(Linear Kernel) & Accuracy & 0.83 & 0.82 & 0.84 & 0.85\\
+SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.92 & 0.95\\
+ & Accuracy & 0.83 & 0.82 & 0.84 & 0.85\\
 \bottomrule
@@
 
-\hypertarget{auc-approach}{%
-\subsubsection{AUC approach}\label{auc-approach}}
-
-After calculating the AUC for each miRNA for the classification of MI
-and CAD samples, miRNAs were sorted, and their performance was
-investigated as a set. The metric of choice for selecting the best set
-was AUC. As shown in Figure @ref(fig:AUCset), the AUC increased until
-the number of miRNAs in the set reached six, and after that, it dropped.
-The AUC for separating MI samples from CAD using these miRNAs was 0.93.
-The miRNAs in the set were miR-29a, miR-197, miR-142, miR-21, miR-155,
-and miR-320C1. The expression values of these miRNAs in healthy, CAD,
-and MI samples are presented in Figure @ref(fig:AUCexp). The ROC curve
-of the selected miRNAs for MI and CAD sample classification is
-illustrated in Figure @ref(fig:miRROC)B.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/AUC for cad mi mirs}
-
-}
-
-\caption{Area under the curve (AUC) for sets containing an increasing number of miRNAs with the highest individual AUC in MI/CAD separation.}\label{fig:AUCset}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/ExpressionforAUC}
-
-}
-
-\caption{Expression profile of has-miR-29A, has-miR-197, has-miR-142, has-miR-21, has-miR-155, and has-miR-320C1 in Healthy, CAD, and MI samples.}\label{fig:AUCexp}
-\end{figure}
+\hypertarget{auc-roc-approach}{%
+\subsubsection{AUC-ROC approach}\label{auc-roc-approach}}
+
+After calculating the AUC-ROC for each miRNA for the classification of
+MI and CAD samples, the miRNAs with AUC-ROC over 0.8 were selected. The
+selected miRNAs were miR-29a, miR-197, miR-186, miR-21, and miR-296. The
+expression values of these miRNAs in healthy, CAD, and MI samples are
+presented in Figure @ref(fig:Expall). The ROC curve of the selected
+miRNAs for both layers are illustrated in Figure @ref(fig:miRROC).
 
@@
 is presented in Figure @ref(fig:AUCROC)A and the confusion matrix is
-illustrated in Figure @ref(fig:AUCCF)A. Both AUC and accuracy for the
+illustrated in Figure @ref(fig:AUCCF)A. Both AUC-ROCand accuracy for the
 model on the test set were equal to 1.
@@
 
-\caption{ROC curve for the set of miRNAs selected by AUC on test set classification.
+\caption{ROC curve for the set of miRNAs selected by AUC-ROC on test set classification.
 (A) SVM with RBF kernel for healthy and not-healthy samples classification.
@@
 To find the best model for training the best set, different models were
-trained using their pre-set values. Their AUC and accuracy results on
-the test set are presented in Figure @ref(fig:AUCmodels). The best model
-from the AUC point-of-view was the LR and from the accuracy
-point-of-view, it was the SVM model with a polynomial kernel. For the LR
-model the AUC and accuracy were 0.92 and 0.81, respectively; and for the
-SVM model with a polynomial kernel, the values were 0.91 and 0.84,
+trained using their pre-set values. Their AUC-ROC and accuracy results
+on the test set are presented in Figure @ref(fig:AUCmodels). The best
+model from the AUC-ROC point-of-view was the SVM with linear kernel and
+from the accuracy point-of-view, it was the SVM model with an RBF
+kernel. For the SVM-linear model the AUC-ROC and accuracy were 0.93 and
+0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84,
 respectively. Both models were hyper-tuned and the ROC curve for their
-best performance is presented in Figure @ref(fig:AUCROC)B and C. The AUC
-and accuracy for the LR model increased to 0.94 and 0.88, respectively.
-For the SVM model with a polynomial kernel, these values increased to
-0.95 and 0.88, respectively (Table @ref(tab:AUCML)). The sensitivity for
-the LR and SVM models were 1 and 0.97, respectively; and the specificity
-for them was 0.57 and 0.64, respectively. The confusion matrix for both
+best performance is presented in Figure @ref(fig:AUCROC)B and C. The
+AUC-ROC and accuracy for the SVM-linear model modified to 0.92 and 0.88,
+respectively. For the SVM-RBF, these values increased to 0.96 and 0.94,
+respectively (Table @ref(tab:AUCML)). The sensitivity for the SVM-linear
+and SVM-RBF models were 0.91 and 0.97, respectively; and the specificity
+for them was 0.79 and 0.86, respectively. The confusion matrix for both
 models is illustrated in Figure @ref(fig:AUCCF)B and C.
@@
 
-{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/Models AUC mirs}
+{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/Models AUC 5 mirs CV}
 
@@
 
-\caption{Area under the curve (AUC) and accuracy of different models trained with three miRNAs in DEGs on the test set.}\label{fig:AUCmodels}
+\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with AUC-selected miRNAs.}\label{fig:AUCmodels}
 \end{figure}
@@
 \midrule
-LR & AUC-ROC & 0.90 & 0.92 & 0.91 & 0.94\\
- & Accuracy & 0.84 & 0.81 & 0.84 & 0.88\\
-SVM & AUC-ROC & 0.90 & 0.91 & 0.91 & 0.95\\
-(Polynomial Kernel) & Accuracy & 0.86 & 0.84 & 0.86 & 0.88\\
+SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.93 & 0.92\\
+ & Accuracy & 0.85 & 0.82 & 0.90 & 0.88\\
+SVM-RBF & AUC-ROC & 0.90 & 0.92 & 0.96 & 0.96\\
+ & Accuracy & 0.86 & 0.84 & 0.96 & 0.94\\
 \bottomrule
@@
 
-The prevalence of acute MI (AMI) can lead to high-rate mortality in the
-clinical setting. However, early diagnosis and application of suitable
-treatment protocols can reduce mortality and improve AMI prognosis
+The prevalence of MI can lead to high-rate mortality in the clinical
+setting. However, early diagnosis and application of suitable treatment
+protocols can reduce mortality and improve AMI prognosis
 ({``Cardiovascular Diseases ({CVDs})''} n.d.; Thygesen et al. 2018; Tsao
@@
 used as biomarker sets for sample classification. miR-21; miR-32; and
-miR-186 were selected as differentially expressed miRNAs, and miR-21;
-miR-29a; miR-142; miR155; miR-197; and miR-320c1 were selected according
-to their AUC values. As shown in Figure @ref(fig:miRROC), all other
-miRNAs selected with both approaches had AUC-ROC over 0.9 for the
-isolation of healthy and not-healthy samples except for miR-142 and
-miR-29a. Data confirmed that the real challenge is to classify CAD and
-MI samples because of close overlap. Of 8 miRNAs under investigation in
-both approaches except for miR-32, all miRNAs had an AUC-ROC over 0.8
-for the discrimination of CAD and MI samples. Besides, the high AUC-ROC
-values of miRNAs confirms their high potential as biomarkers.
-
-ML models when trained with miRNA sets selected by both DEG and AUC
+miR-186 were selected as differentially expressed miRNAs, and miR-186,
+miR-21; miR-29a; miR-197; and miR-296 were selected according to their
+AUC-ROC values. As shown in Figure @ref(fig:miRROC), all miRNAs selected
+with both approaches had AUC-ROC over 0.9 for the isolation of healthy
+and not-healthy samples except for miR-296 and miR-29a. Data confirmed
+that the real challenge is to classify CAD and MI samples because of
+close overlap. Of 6 miRNAs under investigation in both approaches except
+for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of
+CAD and MI samples. Besides, the high AUC-ROC values of miRNAs confirms
+their high potential as biomarkers.
+
+ML models when trained with miRNA sets selected by both DEG and AUC-ROC
 approaches showed better performance in the classification than each
 miRNA. To avoid unwanted complexity and poor predictive values, a
-two-layer architecture was also designed. The first layer was for for
-the discrimination of healthy from not-healthy samples, and the second
-layer separated the CAD from MI candidates. As expected in both
-approaches, a hyper-tuned SVM model could flawlessly separates healthy
-from not-healthy samples using distinct miRNAs sets. The ML models were
-also capable of effectively separating CAD from MI patients. Although
-both miRNA sets had nearly the same AUC-ROC with their best model, the
+two-layer architecture was also designed. The first layer was for the
+discrimination of healthy from not-healthy samples, and the second layer
+separated the CAD from MI candidates. As expected, in both approaches a
+hyper-tuned SVM model could flawlessly separates healthy from
+not-healthy samples using distinct miRNAs sets. The ML models were also
+capable of effectively separating CAD from MI patients. Although both
+miRNA sets had nearly the same AUC-ROC with their best model, the
 accuracy, sensitivity, and specificity were different. The model trained
-with DEGs had better specificity, but the one trained with AUC-selected
-miRNAs had slightly better accuracy and higher sensitivity. This
-difference comes from the different strategies the selection of
-biomarkers sets.
+with AUC-selected miRNAs had better performance in all predictive
+values.
 
@@
 2015; Yao et al. 2016; Liu et al. 2017; Horita, Farquharson, and Stephen
-2021; Cai et al. 2020; Zhao et al. 2018; Bhansali et al. 2022). For
-instance, there is evidence that the elevation of miR-186 suppresses the
-expression of Cystathionine-\(\gamma\)-lyase, leading to the subsequent
-secretion of pro-inflammatory cytokines and cellular lipid accumulation.
-Besides, macrophage-derived miR-186 may promote atherosclerotic plaques
-(Yao et al. 2016). In line with this claim, we found that miR-186 is
+2021; Bhansali et al. 2022). For instance, there is evidence that the
+elevation of miR-186 suppresses the expression of
+Cystathionine-\(\gamma\)-lyase, leading to the subsequent secretion of
+pro-inflammatory cytokines and cellular lipid accumulation. Besides,
+macrophage-derived miR-186 may promote atherosclerotic plaques (Yao et
+al. 2016). In line with this claim, we found that miR-186 is
 up-regulated in both CAD and MI candidates related to control
@@
 (Figure @ref(fig:DEMexp)), To be specific, miR-186 is the only
-differentially expressed miRNA between CAD and MI, indicating its main
-role in the promotion of atherosclerosis.
+differentially expressed miRNA between CAD and MI, with a clear
+up-regulation in CAD, indicating its main role in the promotion of
+atherosclerosis.
 
@@
 Molecular analyses have indicated the regulatory role of miRNAs selected
-by the AUC approach in PBMCs after a cardiovascular event. As the only
-common miRNA of the two sets, molecular function of miR-21 in CVDs was
+by the AUC-ROC approach in PBMCs after a cardiovascular event. Two
+common miRNAs in DEGs and AUC-ROC approachs, miR-21 and miR-186, were
 covered in DEGs miRNA set. Based on numerous reports mir-29a can be
@@
 
-According to the present data (Table @ref(tab:mirExptable)), the
-difference in miR-142 expression between MI/Healthy and CAD/Healthy was
-not significant, but these values reached statistically significant in
-MI as compared to CAD. Based on different reports, miR-142 is commonly
-active in PBMCs. This miRNA directly targets and inhibits the expression
-of Adenomatous Polyposis Coli, a negative WNT signaling pathway
-regulator, contributing to the activation of the WNT signaling pathway
-and cardiac fibroblast activation after myocardial ischemia (Cai et al.
-2020).
-
-We also noted that miR-155 is the only down-regulated miRNA among all
-miRNAs in both CAD/healthy and MI/Healthy comparisons (Table
-@ref(tab:mirExptable)). There are numerous reports on the
-down-regulation of miR-155 in PBMCs after MI or in CAD patients.
-Previous data suggested an inverse relationship between miR-155 levels
-and the severity of the coronary artery condition (Zhang et al. 2015;
-Zhao et al. 2018). Along with current data and previous findings, it is
-postulated that miR-155 has a protective effect under pathological
-conditions. This genetic element can control inflammation and thus
-reduce tissue damage through its negative feedback effects on
-inflammatory factors. These features coincide with the reduction of
-atherosclerosis (Zhang et al. 2015).
-
 Data indicated that miR-197 is also significantly up-regulated in both
@@
 diagnostic role of miR-197 in adverse cardiovascular events (Bhansali et
-al. 2022). There are some reports about miR-320 role in the
-physiopathology of cardiac fibrosis. Mechanistically, Molecular analyses
-revealed that miR-320 might induce various effects by targeting PLEKHM3
-and IFITM1 in cardiomyocytes and cardiac fibroblasts, respectively (F.
-Li et al. 2021). However, there are no reports on miR-320c1 activity in
-the cardiovascular system or PBMCs.
+al. 2022).
+
+The findings of this study demonstrate the potential of miR-296 as a
+biomarker with high discriminatory power for distinguishing between
+samples from individuals with MI and CAD. MiR-296 has been identified as
+a key regulator in the development and advancement of atherosclerosis by
+controlling the expression of target genes associated with various
+biological processes including angiogenesis, inflammation, cholesterol
+metabolism, hypertension, cellular proliferation, and apoptosis (H. Li
+et al. 2018). In a study it has been shown that miR-296 expression
+levels are significantly increased in the PBMCs of CAD patients compared
+to healthy controls, suggesting its involvement in regulating
+pro-inflammatory cytokines like IL-6 and TNF-a (Fard et al. 2020). These
+findings suggest that miR-296 may have a significant impact on the
+pathogenesis of atherosclerosis and could potentially serve as a
+diagnostic biomarker for CAD or MI.
 
@@
 samples, and the second layer, which classified MI/CAD samples. The set
-of miRNAs selected based on their AUC values had slightly better
-performance in the second layer. Overall, our two-layer structure
-achieved a weighted accuracy of 0.91. This demonstrates the potential
-for combining bioinformatics and machine learning techniques to identify
-novel biomarkers and gain a deeper understanding of myocardial
-infarction.
+of miRNAs selected based on their AUC-ROC values had better performance
+in the second layer. Overall, our two-layer structure achieved a
+accuracy of 0.96. This demonstrates the potential for combining
+bioinformatics and machine learning techniques to identify novel
+biomarkers and gain a deeper understanding of myocardial infarction.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-142}{}}%
-Cai, Lidong, Gong Chao, Weifeng Li, Jumo Zhu, Fangfang Li, Baozhen Qi,
-Yong Wei, et al. 2020. {``Activated {Cd4}+ {T} Cells-Derived Exosomal
-{miR}-142-3p Boosts Post-Ischemic Ventricular Remodeling by Activating
-Myofibroblast.''} \emph{Aging} 12 (8): 7380--96.
-\url{https://doi.org/10.18632/aging.103084}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-75}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-296a}{}}%
+Fard, Toktam Kazemi, Samareh Tavakoli, Reza Ahmadi, Nariman Moradi, Reza
+Fadaei, Asghar Mohammadi, and Soudabeh Fallah. 2020. {``Evaluation of
+{Ip10} and {miRNA} 296-a {Expression} {Levels} in {Peripheral} {Blood}
+{Mononuclear} {Cell} of {Coronary} {Artery} {Disease} {Patients} and
+{Controls}.''} \emph{DNA and Cell Biology} 39 (9): 1678--84.
+\url{https://doi.org/10.1089/dna.2020.5650}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-PBMC-miR}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-320}{}}%
-Li, Fang, Shan-Shan Li, Hui Chen, Jian-Zhi Zhao, Jie Hao, Jin-Ming Liu,
-Xiu-Guang Zu, and Wei Cui. 2021. {``{miR}-320 Accelerates Chronic Heart
-Failure with Cardiac Fibrosis Through Activation of the {Il6}/{Stat3}
-Axis.''} \emph{Aging} 13 (18): 22516--27.
-\url{https://doi.org/10.18632/aging.203562}.
+\leavevmode\vadjust pre{\hypertarget{ref-296}{}}%
+Li, Heng, Heng Li, Xin-Ping Ouyang, Ting Jiang, Xi-Long Zheng, Xi-Long
+Zheng, Ping-Ping He, and Guo-Jun Zhao. 2018. {``{MicroRNA}-296: A
+Promising Target in the Pathogenesis of Atherosclerosis?''}
+\emph{Molecular Medicine} 24 (1): 12--12.
+\url{https://doi.org/10.1186/s10020-018-0012-y}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-155-2}{}}%
-Zhang, Yu-Hui, Liang-Hua Xia, Jia-Mei Jin, Ming Zong, Ming Chen, and Bo
-Zhang. 2015. {``Expression Level of {miR}--155 in Peripheral Blood.''}
-\emph{Asian Pacific Journal of Tropical Medicine} 8 (3): 214--19.
-\url{https://doi.org/10.1016/S1995-7645(14)60318-7}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-155-3}{}}%
-Zhao, Duo, Jing Zhao, Jiangbin Sun, Yanling Su, Jinfeng Jian, Huaan Ye,
-Jiawang Lin, Zongda Yang, Jiatao Feng, and Zhiping Wang. 2018. {``The
-Expression Level of {miR}-155 in Plasma and Peripheral Blood Mononuclear
-Cells in Coronary Artery Disease Patients and the Associations of These
-Levels with the Apoptosis Rate of Peripheral Blood Mononuclear Cells.''}
-\emph{Experimental and Therapeutic Medicine}, September.
-\url{https://doi.org/10.3892/etm.2018.6797}.
-
 \end{CSLReferences}
@@
 \bibliographystyle{spmpsci}
-\bibliography{ForArticleRefrence.bib,discussion.bib,Introduction.bib}
+\bibliography{ForArticleRefrence.bib,discussion.bib,Introduction.bib,mir296.bib}
 

### 2023-08-07 ArticleDraft.Rmd
The version after revision

@@
 In this study, we aimed to identify a key set of miRNA biomarkers using integrated bioinformatics and machine learning analysis.
-We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contains peripheral blood mononuclear cells (PBMCs) samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy individuals.
+We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contains peripheral blood mononuclear cell (PBMC) samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy individuals.
 Additionally, we selected a set of miRNAs based on their area under the receiver operating characteristic curve (AUC-ROC) for separating the CAD and MI samples.
-We designed a two-layer architecture for sample classification, in which the first layer isolates healthy samples from not-healthy ones, and the second layer classifies stable CAD and MI samples.
+We designed a two-layer architecture for sample classification, in which the first layer isolates healthy samples from unhealthy samples, and the second layer classifies stable CAD and MI samples.
 We trained different machine learning models using both biomarker sets and evaluated their performance on a test set.
@@
 output: 
+  bookdown::word_document2: default
   bookdown::pdf_book:
-  bookdown::word_document2: default
 always_allow_html: true
@@
 
-At present, cardiovascular diseases (CVDs) are the leading cause of human mortality, with 32% of all global deaths.
-It is estimated that about 85% of CVDs mortality were diagnosed with myocardial infarction (MI) [@CVD].
-MI is an acute coronary syndrome with sudden blockage and stenosis of the coronary artery and subsequent myocardial ischemia, leading to extensive cardiomyocyte damage and necrosis [@Macro].
+Cardiovascular diseases (CVDs) are the leading cause of human mortality, accounting for 32% of all global deaths.
+It is estimated that approximately 85% of CVD mortality is due to myocardial infarction (MI) [@CVD].
+MI is an acute coronary syndrome characterized by sudden blockage and stenosis of the coronary artery and subsequent myocardial ischemia, leading to extensive cardiomyocyte damage and necrosis [@Macro].
 
-Over the last 50 years, numerous attempts have been collected to use biomarkers to facilitate diagnosis, assess the risk, follow-up therapy, and determine therapeutic efficacy in CVDs candidates.
-Based on the released guidelines, cardiac troponins (cTns) are used as a highly-sensitive and accurate approach for detecting MI.
-Despite the inherent advantages, the high-rate sensitivity of cTn-based assays has also led to more false positive results [@17], necessiating the advent and development of new modalities with pathological values.
-To improve diagnostic value upon existing MI biomarkers, the combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, is proposed.
-Previous researches support the notion that miRNAs exhibit great potential to be used as alternative biomarkers in CVDs detection and follow-up [@miR].
-It is suggested that miRNAs possess 18-22 nucleotides and can play a crucial role in the regulation of gene expression.
-Evidence point to the fact that miRNAs are involved in the pathogenesis of cardiac tissue injury [@m1].
-Several biological activities, such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque formation, and cardiac rhythm, are regulated by miRNAs [@m2].
-These elements can easily circulate in biofluids and could be considered theranostics targets in terms of CVDs [@m1].
-It is postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in CVD patients.
-To be specific, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) [@36].
+Over the last 50 years, numerous attempts have been made to use biomarkers to facilitate diagnosis, assess the risk, follow-up therapy, and determine therapeutic efficacy in CVD candidates.
+Based on released guidelines, cardiac troponins (cTns) are used as a highly sensitive and accurate approach for detecting MI.
+Despite these inherent advantages, the high sensitivity of cTn-based assays has also led to more false-positive results [@17], necessitating the advent and development of new modalities with pathological value.
+To improve the diagnostic value of  existing MI biomarkers, a combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, has been proposed.
+Previous research supports the notion that miRNAs exhibit great potential as alternative biomarkers for CVD detection and follow-up [@miR].
+It has been suggested that miRNAs possess 18-22 nucleotides and play a crucial role in the regulation of gene expression.
+Evidence indicates that miRNAs are involved in the pathogenesis of cardiac tissue injury [@m1].
+Several biological processes, such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque formation, and cardiac rhythm, are regulated by miRNAs [@m2].
+These elements can easily circulate in biofluids and could be considered theranostic targets in terms of CVDs [@m1].
+It has been postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in patients with CVD.
+Specifically, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) [@36].
 
-PBMCs are a fraction of white blood cells, including monocytes, lymphocytes, macrophages, and other cells belonging to the immune system [@PBMC-miR].
-Emerging data have indicated that PBMCs can be used as a valid source of biomarkers for monitoring various pathological conditions.
-Of note, the alteration of mRNAs and miRNAs under pathological conditions gives us valuable information about different kinds of disorders.
-PBMCs could recapitulate the conditions of the target tissues, thus, providing a highly sensitive and specific source of biomarkers [@Meder6].
-Commensurate with these conditions, these cells are repositories of dysregulated genes and miRNAs expression profiles in CVDs [@PBMC-miR; @Meder6].
+PBMCs are a fraction of white blood cells, including monocytes, lymphocytes, macrophages, and other cells of the immune system [@PBMC-miR].
+Emerging data indicate that PBMCs can be used as a valid source of biomarkers for monitoring various pathological conditions.
+Of note, the alteration of mRNAs and miRNAs under pathological conditions provides valuable information about different kinds of disorders.
+PBMCs can recapitulate the conditions of target tissues, thus providing a highly sensitive and specific source of biomarkers [@Meder6].
+Combined with these conditions, these cells are repositories of dysregulated genes and miRNA expression profiles in CVDs [@PBMC-miR; @Meder6].
 
-In recent years, the advent and application of machine learning (ML) is an exciting prospect for advancing scientific discoveries.
-Although the concept of ML and its initial algorithms were conceived many years ago, recent improvements in computing power and access to vast amounts of data have shown that ML techniques outperform classical statistical methods in various fields.
-Furthermore, the progress made in omics technologies has enabled the analysis of massive and intricate biological data sets, consisting of hundreds to thousands of samples, which makes it possible for ML to extract valuable biological insight and information from such data [@ML-Bio].
-As a result, ML provides innovative methods for merging and interpreting diverse types of omics data, leading to the identification of new biomarkers.
-These biomarkers can aid in precise disease prediction, patient stratification, and developing novel therapeutic approaches [@ML].
+In recent years, the advent and application of machine learning (ML) has been an exciting prospect for advancing scientific research.
+Although the concept of ML and its initial algorithms were conceived many years ago, recent improvements in computing power and access to vast amounts of data have demonstrated that ML techniques outperform classical statistical methods in various fields.
+Furthermore, the progress made in omics technologies has enabled the analysis of massive and intricate biological datasets, consisting of hundreds to thousands of samples, which makes it possible for ML to extract valuable biological insights and information from such data [@ML-Bio].
+Consequently, ML provides innovative methods for merging and interpreting diverse types of omics data, leading to the identification of new biomarkers.
+These biomarkers can aid in precise disease prediction, patient stratification, and the development of novel therapeutic approaches [@ML].
 
-In this study, we aimed to identify potential miRNA biomarkers for MI patients by combining and analyzing three different microarray datasets from PBMCs.
-It is suggested that the integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring MI.
-Besides, this approach can deepen our vision into the underlying mechanisms of MI and aid in the development of valid diagnostic biomarkers and patient stratification.
+In this study, we aimed to identify potential miRNA biomarkers in patients with MI by combining and analyzing three different microarray datasets from PBMCs.
+The integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring MI.
+Additionally, this approach can deepen our understanding of the underlying mechanisms of MI and aid in the development of valid diagnostic biomarkers and patient stratification.
 
@@
 Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database (<https://www.ncbi.nlm.nih.gov/geo/>).
-To obtain sufficient classification power between MI, healthy, and CAD samples, a relatively large sample size was required.
-Therefore, GSE59867 for MI and CAD samples and GSE56609 and GSE54475 for healthy samples were selected.
-All samples were produced using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
-Only healthy, CAD, and early-stage MI samples were selected from these datasets for further analyses.
-The basic information for the three datasets evaluated in the current study is provided in Table \@ref(tab:datasets).
-The bioinformatics sections were fully conducted on R, ver.
-4.2.0 [@R], using RStudio [@RStudio], and all plots and graphics of these sections were created using the ggplot2 R package [@ggplot].
+To obtain sufficient classification power between MI, healthy, and CAD samples, a relatively large sample size is needed Therefore, GSE59867 for the MI and CAD samples and GSE56609 and GSE54475 for the healthy samples were selected.
+All samples were produced using Affymetrix Human Gene 1.0 ST Array platform (GPL6244).
+Only healthy, CAD, and early-stage MI samples were selected from these datasets for further analysis.
+The basic information for the three datasets evaluated in this study is provided in Table \@ref(tab:datasets).
+Bioinformatics analyses were conducted using R, ver.
+4.2.0 [@R], RStudio [@RStudio].
+All plots and graphics of these sections were created using the ggplot2 R package [@ggplot].
 
@@
 
-kbl(df, format = "markdown", booktabs = T, align = "c", linesep = c(""), caption = "Samples information on the GEO microarray datasets.")
+kbl(df, format = "markdown", booktabs = T, align = "c", linesep = c(""), caption = "Sample information on the GEO microarray datasets.")
 ```
 
-## Pre-processing
+## Preprocessing
 
-The raw data in the form of CEL files from all datasets were obtained from the GEO.
-To prepare the data for analysis, we utilized the fRMA package [@fRMA], facilitating the pre-processing of individual microarray samples and their consistent combination.
+The raw data in the form of CEL files from all datasets were obtained from GEO.
+To prepare the data for analysis, we utilized the fRMA package [@fRMA] to facilitate preprocessing of individual microarray samples and their consistent combination.
 For each dataset, background correction was applied using the RMA algorithm, followed by quantile normalization based on the reference distribution.
-To account for probe-specific effects, batch effects were eliminated during summarization, and gene expression variances were estimated accordingly.
-In cases where multiple probe sets matched the same gene, the mean log fold change was retained.
-Consequently, fRMA can serves as a technique for removing batch effects across diverse datasets generated by identical microarray platforms [@BER].
-To ensure the effectiveness of batch effect removal, we employed principal component analysis (PCA) and relative log expression (RLE) plots to visualize the data before and after applying fRMA.
+To account for probe-specific effects, batch effects were eliminated during summarization and gene expression variances were estimated accordingly.
+In cases where multiple probe sets matched the same gene, the mean log-fold change was retained.
+Consequently, fRMA can serve as a technique to remove batch effects across diverse datasets generated by identical microarray platforms [@BER].
+To ensure the effectiveness of the batch effect removal, we employed principal component analysis (PCA) and relative log expression (RLE) plots to visualize the data before and after applying fRMA.
 
@@
 
-The barcode algorithm, as introduced by McCall et al.
+The barcode algorithm was introduced by McCall et al.
 [@Barcode], aimed to convert actual expression values into binary barcode values.
-Extensive sample collections were gathered, and normalization was performed using fRMA across multiple platforms, including the Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
-By utilizing these normalized datasets, the distribution of observed intensities, both for expressed and unexpressed genes, was estimated.
-The determination of whether a gene was expressed or unexpressed was based on the following equation, where a value of 1 indicated expression and a value of 0 indicated non-expression:
+Extensive sample collections were gathered and normalization was performed using fRMA across multiple platforms, including the Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
+By utilizing these normalized datasets, the distribution of the observed intensities for both the expressed and unexpressed genes was estimated.
+The determination of whether a gene was expressed or not was based on the following equation, where a value of 1 indicates expression and a value of 0 indicates non-expression:
 
@@
 
-In the barcode algorithm, the normalized intensity of gene $i$ in sample $j$ is denoted by $x_{ij}$.
-A user-defined parameter, $C$, is introduced, along with the standard deviation ($\sigma^{ne}$) and mean ($\mu^{ne}$) of the non-expressed distribution.
-Based on these values, the barcode representation of a sample is generated as a vector consisting of ones and zeros, representing the estimated expression (ones) and non-expression (zeros) of each gene.
-The barcode function within the R fRMA package was employed for implementing the barcode algorithm, utilizing the default value of $C$.
+In the barcode algorithm, the normalized intensity of gene $i$ in sample $j$ is denoted as $x_{ij}$.
+A user-defined parameter, $C$, was introduced along with the standard deviation ($\sigma^{ne}$) and mean ($\mu^{ne}$) of the non-expressed distribution.
+Based on these values, the barcode representation of a sample was generated as a vector consisting of ones and zeros, representing the estimated expression (ones) and non-expression (zeros) of each gene.
+The barcode function within the R fRMA package was employed to implement the barcode algorithm, utilizing the default value of $C$.
 
-To assess the differences in expressed ratios between the MI group and the healthy control group, Fisher's exact test was performed on the barcode values of individual genes.
-Genes that exhibited a false discovery rate (FDR) below $0.05$, calculated using the Benjamini-Hochberg procedure to account for multiple testing issues, were identified as differentially expressed genes (DEGs).
-The same procedures were applied to the CAD versus healthy control comparison, as well as the MI versus CAD group, to identify DEGs specific to each comparison.
+To assess the differences in expressed ratios between the MI and healthy control groups, Fisher's exact test was performed on the barcode values of individual genes.
+Genes that exhibited a false discovery rate (FDR) below $0.05$, calculated using the Benjamini-Hochberg procedure to account for multiple testing issues were identified as differentially expressed genes (DEGs).
+The same procedures were applied to the CAD versus healthy control comparison, as well as to the MI versus CAD group, to identify DEGs specific to each comparison.
 
@@
 
-The R clusterProfiler package [@cluster] was utilized to perform the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation on the set of DEGs.
-The GO analysis encompassed three categories: biological process (BP), cellular component (CC), and molecular function (MF).
+The R clusterProfiler package [@cluster] was utilized to perform Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation on the set of DEGs.
+GO analysis encompassed three categories: biological process (BP), cellular component (CC), and molecular function (MF).
 For statistical significance, an adjusted p-value threshold of less than 0.05 was employed.
-The enrichment analyses were conducted separately for the DEGs specific to the MI-healthy and CAD-healthy comparisons.
-In these analyses, all default parameters provided by the package were used.
+Enrichment analyses were conducted separately for DEGs specific to the MI-healthy and CAD-healthy comparisons.
+All the default parameters provided by the package were used in the analyses.
 
@@
 
-The ML analysis was performed using Python software, ver.
-3.9, Numpy [@numpy], pandas [@pandas], and Scikit-Learn packages [@scikit-learn].
-Whenever hyper-tuning was needed, the scikit-opt package [@scikitopt] was used.
-In all ML analyses, the datasets were divided into train and test sets by a 0.7:0.3 ratio and, all reported results are the average of 10-fold cross-validation.
+ML analysis was performed using Python software, ver.
+3.9, Numpy [@numpy], Pandas [@pandas], and Scikit-Learn packages [@scikit-learn].
+Whenever hypertuning was needed, the Scikit-opt package [@scikitopt] was used.
+In all ML analyses, the datasets were divided into training and test sets at a 0.7:0.3 ratio, and all reported results are the average of 10-fold cross-validation.
 
-Two different approaches were used for selecting miRNAs for model training.
-The first approach was using the miRNAs that are differentially expressed.
-In the second approach, miRNAs with individual AUC-ROC over 0.8 for separating MI from CAD were selected.
-Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of sets of miRNAs selected with different logics.
+Two different approaches were used to select miRNAs for model training.
+The first approach was to use differentially expressed miRNAs.
+In the second approach, miRNAs with individual AUC-ROCs over 0.8 for separating MI from CAD were selected.
+The results of these two approaches can provide an informative comparison between the predictive capabilities of sets of miRNAs selected with different logics.
 
@@
 
-In this approach, a two-layer architecture was deployed to the data to maximize the prediction values.
-The first layer predicted whether a sample was healthy or not, and the second layer separated MI from CAD in the samples which were predicted as not healthy in the first layer.
+In this approach, a two-layer architecture is deployed to the data to maximize the prediction values.
+The first layer predicted whether a sample was healthy or not, and the second layer separated MI from CAD in the samples that were predicted as not healthy in the first layer.
 To this end, a distinct ML model was trained for each layer.
-Since there is a limited number of miRNAs in DEGs, both layers were trained with all of them.
-For further comparison with the models' performance, the ROC curve of each miRNA for classifying healthy and not-healthy, as well as CAD and MI, were generated using a Logistic Regression model.
+Because, there were a limited number of miRNAs in the DEGs, both layers were trained with all of them.
+For further comparison with the models' performance, the ROC curve of each miRNA for classifying healthy and not-healthy, as well as CAD and MI, was generated using a logistic regression model.
 
-#### First layer for isolation of healthy and not-healthy samples:
+#### First layer for the isolation of healthy and not-healthy samples:
 
-A support vector machine (SVM) model using RBF kernels was trained and hyper-tuned using all miRNAs in DEGs.
-To handle the severe imbalance in the number of samples (51 for the healthy and 157 for the not-healthy groups), the sample weight for the healthy and the not-healthy samples was set to 1 and 0.5, respectively.
-The ROC curve and confusion matrix for the model were reported.
+A support vector machine (SVM) model using RBF kernels was trained and hypertuned using all miRNAs in the DEGs.
+To handle the severe imbalance in the number of samples (51 for the healthy group and 157 for the not-healthy group), the sample weights for the healthy and the not-healthy samples were set to 1 and 0.5, respectively.
+The ROC curve and confusion matrix for the model are reported.
 
-#### Second layer for separating MI and CAD samples:
+#### Second layer for separating the MI and CAD samples:
 
-For the sake of reaching the highest classification performance, different models were investigated.
-To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) models were trained.
-All models were trained with their pre-set parameters with 10-fold cross-validation.
-The criteria for choosing the best model were the highest accuracy and AUC-ROC on the test set.
-The best model was hyper-tuned with the scikit-opt package [@scikitopt] for the best classification performance.
-The ROC curve and confusion matrix for the best model were reported.
+Different models were investigated to achieve the highest classification performance.
+To do so, SVM (with linear, polynomial, and RBF kernels), logistic regression (LR), random forests (RF), k-nearest neighbor (kNN), gradient boosting (GB), XGBoost (XGB) and decision tree (DT) models were trained.
+All models were trained with their preset parameters using 10-fold cross-validation.
+The criteria for selecting the best model were the highest accuracy and AUC-ROC for the test set.
+The best model was hypertuned using the scikit-opt package [@scikitopt] for the best classification performance.
+The ROC curve and confusion matrix for the best model are reported.
 
@@
 
-Like the previous approach, a two-layers strategy was conducted.
-The first layer classified samples into healthy and not-healthy, and the separated MI and CAD samples.
-However, to keep the number of miRNAs as low as possible, miRNAs were selected from the second layer, and then their performance was evaluated in the first layer.
-The AUC-ROC of all miRNAs for classifying MI and CAD samples were calculated and, the miRNAs with the AUC-ROC over 0.8 were selected.
-The ROC curves for each selected miRNA for separating healthy samples from not-healthy and MI from CAD samples were also plotted for further comparison.
+As in the previous approach, a two-layer strategy was employed.
+The first layer classified samples into healthy and not-healthy, and the second layer separated the MI and CAD samples.
+However, to keep the number of miRNAs as low as possible, miRNAs were selected from the second layer and their performance was evaluated in the first layer.
+The AUC-ROC of all miRNAs for classifying MI and CAD samples was calculated, and miRNAs with AUC-ROC $> 0.8$ were selected.
+ROC curves for each selected miRNA for separating healthy samples from not-healthy samples and MI from CAD samples were also plotted for further comparison.
 
-#### First layer for isolation of healthy and not-healthy samples:
+#### First layer for the isolation of healthy and not-healthy samples:
 
-An SVM model with an RBF kernel was trained using the selected set of miRNAs.
-Additionally, the model was hyper-tuned to find the hyper-parameters for the highest AUC-ROC and accuracy.
-The same sample weights as the previous approach (1 for healthy and 0.5 for not-healthy samples) were used.
+An SVM model with an RBF kernel is trained using the selected set of miRNAs.
+Additionally, the model was hypertuned to find the hyperparameters for the highest AUC-ROC and accuracy.
+The same sample weights as in the previous approach (1 for healthy and 0.5 for not-healthy samples) were used.
 The ROC curve and confusion matrix for the model were reported.
 
-#### Second layer for separating MI and CAD samples:
+#### Second layer for separating the MI and CAD samples:
 
-The selected miRNA set was used to train different algorithms to find the best model.
-Similar to the previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT models were trained.
-All models were trained with their pre-set parameters using 10-fold cross-validation.
-The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned using the scikit-opt package [@scikitopt].
+The selected miRNA set was used to train different algorithms to determine the best model.
+Similar to the previous approach, the SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT models were trained.
+All models were trained with their preset parameters using 10-fold cross-validation.
+The models with the highest AUC-ROC and accuracy on the test set were selected and hypertuned using the scikit-opt package [@scikitopt].
 The ROC curve and confusion matrix for the best model were reported.
@@
 
-## Pre-processing
+## Preprocessing
 
 The PCA plots of the samples are shown in Figures \@ref(fig:PCA)A and B.
-As shown, healthy samples were separated from CAD or MI samples in primary data and after conducting fRMA.
-In the RLE plot, there was a distinct difference between dataset means for all samples before performing fRMA (Figure \@ref(fig:PCA)C).
-All datasets were rearranged around 0 in the RLE plot after conducting fRMA (Figure \@ref(fig:PCA)D).
-Moreover, there was an apparent change in inter-quantile distances, but the values still were over 0.1.
+Healthy samples were separated from the CAD or MI samples in the primary data and after conducting fRMA.
+In the RLE plot, there was a distinct difference between the dataset means for all samples before fRMA was performed (Figure \@ref(fig:PCA)C).
+All datasets were rearranged to approximately 0 in the RLE plot after fRMA was conducted (Figure \@ref(fig:PCA)D).
+Moreover, there was an apparent change in the interquantile distances, but the values were still greater than 0.1.
 
-```{r PCA, echo=F, out.width="100%", fig.cap="principal component analysis plots for (A) primary data and (B) the data after fRMA, and the relative log expression plots for (C) primary data and (D) the data after fRMA."}
+```{r PCA, echo=F, out.width="100%", fig.cap="Principal component analysis plots for (A) primary data and (B) the data after fRMA, and the relative log expression plots for (C) primary data and (D) the data after fRMA."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/PCA and RLE of trains.pdf")
@@
 
-According to the cutoff criterion of $FDR < 0.05$, there were `r dim(M_H)[1]` DEGs between the MI and the healthy samples.
-Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in MI compared to the healthy controls.
-In CAD and healthy group comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and `r sum(C_H$logFC < 0)` DEGs were up- and down-regulated, respectively, in CAD samples.
-In the MI and CAD groups, the number of DEGs was `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)` compared to CAD samples, respectively.
-These data are summarized in Table \@ref(tab:DEGstab).
+According to the cutoff criterion of $FDR < 0.05$, there were `r dim(M_H)[1]` DEGs between MI and healthy samples.
+Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in the MI group compared to the healthy group.
+In the CAD and healthy group comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and `r sum(C_H$logFC < 0)` DEGs were up- and down-regulated, respectively, in CAD samples.
+In the MI and CAD groups, the number of DEGs was `r dim(M_C)[1]`, and the numbers of up- and down-regulated genes in MI samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively, compared to CAD samples.
+The data are summarized in Table \@ref(tab:DEGstab).
 
-```{r venn, echo=F, out.width="45%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
+```{r venn, echo=F, out.width="45%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD."}
 
@@
 
-The Venn diagram in Figure \@ref(fig:venn) shows that CAD and MI samples shared most of their DEGs.
+The Venn diagram in Figure \@ref(fig:venn) shows that the CAD and MI samples shared most of their DEGs.
 From 860 DEGs of MI/healthy and 670 DEGs of CAD/healthy, 531 genes were common, which is 62% of MI/healthy DEGs and 79% of CAD/healthy DEGs.
@@
 
-To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI/healthy and CAD/healthy DEGs.
+To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on the MI/healthy and CAD/healthy DEGs.
 For MI/healthy, GO enrichment analysis in the BP category suggested that the DEGs were enriched in "immune response-regulating signaling pathway," "lymphocyte differentiation," "immune response-regulating cell surface receptor signaling pathway," and "leukocyte activation involved in immune response" (Figure \@ref(fig:MIHEnrich)A).
-In the CC category, the DEGs were enriched in "secretory granule membrane," "azurophil granule," "ficolin-1-rich granule," "tertiary granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:MIHEnrich)B).
-In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(fig:MIHEnrich)C).
+In the CC category, DEGs were enriched in "secretory granule membrane," "azurophil granule," "ficolin-1-rich granule," "tertiary granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:MIHEnrich)B).
+In the MF category, DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(fig:MIHEnrich)C).
 KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway," "Lipid and atherosclerosis," and "Hematopoietic cell lineage" (Figure \@ref(fig:MIHEnrich)D).
@@
 
-The enrichment results for CAD/healthy DEGs were as follows.
+The enrichment results for the CAD/healthy DEGs were as follows.
 In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response," "positive regulation of innate immune response," "mononuclear cell differentiation," and "positive regulation of response to external stimulus" (Figure \@ref(fig:CADHEnrich)A).
-In the CC category, the DEGs were enriched in "azurophil granule," "ficolin-1-rich granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:CADHEnrich)B).
-In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(fig:CADHEnrich)C).
+In the CC category, DEGs were enriched in "azurophil granule," "ficolin-1-rich granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:CADHEnrich)B).
+In the MF category, DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(fig:CADHEnrich)C).
 KEGG pathway analysis showed that the DEGs were related to the following pathways: "Chemokine signaling pathway," "Lipid and atherosclerosis," and "Hematopoietic cell lineage" (Figure \@ref(fig:CADHEnrich)D).
@@
 
-Among all DEGs, just miR-186, miR-32, and miR-21 were detected as differentially expressed miRNAs.
-The expression profile of these three miRNAs is presented in Figure \@ref(fig:Expall).
-Additionally, The ROC curves of each miRNA for each layer are presented in Figure \@ref(fig:miRROC).
-Using the logistic regression model, the AUC-ROC of miR-21, miR-32, and miR-186 for separating healthy and not-healthy samples was 0.98, 0.99, and 0.90, respectively (Figure \@ref(fig:miRROC)A).
-Besides, the accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set was 0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
-Moreover, the ROC curve of each miRNA for classifying MI and CAD samples was presented in Figure \@ref(fig:miRROC)B.
-The AUC-ROC and accuracy for miR-21, miR-32, and miR-186 on the test set were 0.85; 0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
+Among the DEGs, miR-186, miR-32, and miR-21 were identified as differentially expressed miRNAs.
+The expression profiles of the three miRNAs are shown in Figure \@ref(fig:Expall).
+The ROC curves of each miRNA in each layer are presented in Figure \@ref(fig:miRROC).
+Using the logistic regression model, the AUC-ROC values of miR-21, miR-32, and miR-186 for separating healthy and not-healthy samples were 0.98, 0.99, and 0.90, respectively (Figure \@ref(fig:miRROC)A).
+The accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set for miR-21, miR-32, and miR-186 was 0.92, 0.98, and 0.89, respectively.
+The ROC curve of each miRNA for classifying MI and CAD samples is presented in Figure \@ref(fig:miRROC)B.
+The AUC-ROC and accuracy for miR-21, miR-32, and miR-186 in the test set were 0.85; 0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
 
@@
 
-#### First layer for healthy not-healthy isolation:
+#### First layer for the isolation of healthy and not-healthy samples:
 
-Although single miRNAs had an acceptable performance for this layer, their predictive value could be improved even further by using them as a set.
+Although single miRNAs had an acceptable performance for this layer, their predictive value could be  further improved by using them as a set.
 The ROC curve for the SVM model with an RBF kernel trained with all three miRNAs is presented in Figure \@ref(fig:DEMROC)A.
-The model had a better performance in classification than single miRNAs.The AUC-ROC for the model is 1, and its accuracy on the test set was also 1.
+The model had a better performance in classification than single miRNAs.The AUC-ROC for the model was 1, and its accuracy on the test set was also 1.
 In Figure \@ref(fig:DEMCF)A, the confusion matrix for the model is presented.
 
-```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
+```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs ROCs h not h cad mi.pdf")
@@
 
-```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
+```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs Conf. matrix.pdf")
@@
 
-#### Second layer for separating MI samples from CAD:
+#### Second layer for separating the MI and CAD samples:
 
-Different models were trained using expression values for three differentially expressed miRNAs.
-The models' AUC-ROC and accuracy on the test set are reported in Figure \@ref(fig:DEMmodels).
-The best model from both AUC-ROC and accuracy point-of-view was the SVM model with a linear kernel.
-The AUC-ROC and accuracy for this model with its pre-set values were 0.93 and 0.82, respectively.
-The model was hyper-tuned for C and gamma hyper-parameters, and therefore the model showed better performance.
-The ROC curve of the hyper-tuned model is presented in Figure \@ref(fig:DEMROC)B.
-For this model, the AUC-ROC reached 0.95, and the accuracy improved to 0.85 (Table \@ref(tab:DEGsML)).
+Different models were trained using the expression values of three differentially expressed miRNAs.
+The models' AUC-ROC and the accuracy of the test set are shown in Figure \@ref(fig:DEMmodels).
+The best model from both the AUC-ROC and accuracy points of view was the SVM model with a linear kernel.
+The AUC-ROC and accuracy for this model with its preset values were 0.93 and 0.82, respectively.
+The model was hypertuned for C and gamma hyperparameters, and therefore the model showed better performance.
+The ROC curve of the hypertuned model is presented in Figure \@ref(fig:DEMROC)B.
+For this model, the AUC-ROC reached 0.95, and the accuracy was improved to 0.85 (Table \@ref(tab:DEGsML)).
 Moreover, the sensitivity and specificity for the model on the test set were 0.91 and 0.71, respectively.
-The confusion matrix for the hyper-tuned model is illustrated in Figure \@ref(fig:DEMCF)B.
+The confusion matrix for the hypertuned model is illustrated in Figure \@ref(fig:DEMCF)B.
 
@@
 
-kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with a linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning") %>%
+kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with a linear kernel as the best model trained with differentially expressed miRNAs on the training and test sets before and after hypertuning") %>%
   kable_styling(full_width = T) %>%
-  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
+  add_header_above(c(" " = 2, "Preset parameters" = 2, "Hypertuned" = 2))
 ```
@@
 
-After calculating the AUC-ROC for each miRNA to classify of MI and CAD samples, the miRNAs with AUC-ROC over 0.8 were selected.
-The selected miRNAs were miR-29a, miR-197, miR-186, miR-21, and miR-296.
-The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:Expall).
-The ROC curve of the selected miRNAs for both layers is illustrated in Figure \@ref(fig:miRROC).
+After calculating the AUC-ROC for each miRNA to classify of MI and CAD samples, the miRNAs with AUC-ROC  $> 0.8$ were selected.
+The miRNAs selected were miR-29a, miR-197, miR-186, miR-21, and miR-296.
+The expression levels of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:Expall).
+The ROC curves of the selected miRNAs in both layers are shown in Figure \@ref(fig:miRROC).
 
-#### First layer for healthy not-healthy isolation:
+#### First layer for the isolation of healthy and not-healthy samples:
 
-Using the selected set, an SVM model with an RBF kernel was trained to separate healthy from not-healthy samples.
+Using the selected set, an SVM model with an RBF kernel was trained to separate healthy and not-healthy samples.
 The ROC curve for the model is presented in Figure \@ref(fig:AUCROC)A, and the confusion matrix is illustrated in Figure \@ref(fig:AUCCF)A.
-Both AUC-ROC and accuracy for the model on the test set were equal to 1.
+Both the AUC-ROC and accuracy of the model on the test set were 1.
 
 ```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
-(B) Logistic regression model for CAD and MI samples classification.
-(C) SVM with polynomial kernel for CAD and MI samples classification.
+(B) Logistic regression model for CAD and MI sample classification.
+(C) SVM with polynomial kernel for CAD and MI sample classification.
 "}
@@
 ```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
-(B) Logistic regression model for CAD and MI samples classification.
-(C) SVM with polynomial kernel for CAD and MI samples classification."}
+(B) Logistic regression model for CAD and MI sample classification.
+(C) SVM with polynomial kernel for CAD and MI sample classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC Conf. matrix.pdf")
@@
 
-#### Second layer for separating MI samples from CAD:
+#### Second layer for separating the MI and CAD samples:
 
-To find the best model for this set of miRNAs, different models were trained using their pre-set values.
-Their AUC-ROC and accuracy results on the test set are presented in Figure \@ref(fig:AUCmodels).
-The best model from the AUC-ROC point-of-view was the SVM with a linear kernel, and from the accuracy point-of-view, it was the SVM model with an RBF kernel.
+To find the best model for this set of miRNAs, different models were trained using their preset values.
+The AUC-ROC and accuracy results for the test set are presented in Figure \@ref(fig:AUCmodels).
+The best model from the AUC-ROC point of view was the SVM with a linear kernel, and from the accuracy point of view, it was the SVM model with an RBF kernel.
 For the SVM-linear model, the AUC-ROC and accuracy were 0.93 and 0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84, respectively.
@@
 For the SVM-RBF, these values increased to 0.96 and 0.94, respectively (Table \@ref(tab:AUCML)).
-The sensitivity for the SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and the specificity for them was 0.79 and 0.86, respectively.
+The sensitivities for the SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and their specificities were 0.79 and 0.86, respectively.
 The confusion matrix for both models is illustrated in Figure \@ref(fig:AUCCF)B and C.
@@
   kable_styling(full_width = T) %>%
-  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
+  add_header_above(c(" " = 2, "Preset parameters" = 2, "Hypertuned" = 2))
 ```
@@
 
-The prevalence of MI can lead to high-rate mortality in the clinical setting.
-However, early diagnosis and application of suitable treatment protocols can reduce mortality and improve MI prognosis [@CVD; @17; @18].
+The prevalence of MI can lead to high mortality rates in the clinical setting.
+However, early diagnosis and the application of suitable treatment protocols can reduce mortality and improve MI prognosis [@CVD; @17; @18].
 Studies have suggested that changes in miRNA expression may play a significant role in the progression of MI and the subsequent remodeling [@23].
-It is believed that the expression of miRNAs is altered during the various biological processes correlated with MI within the myocardium or other related tissues [@24].
-Although several researches have concentrated on examining free circulating miRNAs in serum samples for the detection of cardiac tissue injuries [@11-15], more information is needed to fully comprehend the miRNAs found in different blood sub-components like plasma, platelets, and PBMCs.
-Based on previous findings, PBMCs play a crucial role in the destabilization and rupture of plaques and also the initial inflammatory reactions in individuals experiencing a myocardial infarction (MI).
+It is believed that miRNA expression is altered during the various biological processes correlated with MI within the myocardium or other related tissues [@24].
+Although several studies have focused on examining free circulating miRNAs in serum samples for the detection of cardiac tissue injuries [@11-15], more information is needed to fully comprehend the miRNAs found in different blood subcomponents, such as plasma, platelets, and PBMCs.
+Based on previous findings, PBMCs play a crucial role in the destabilization and rupture of plaques as well as in the initial inflammatory reactions in individuals experiencing myocardial infarction (MI).
 [@Meder6; @Meder23].
-Moreover, PBMCs have specific miRNA profile that is altered under certain pathological conditions, which are great candidates as disease biomarkers [@Meder6].
+Moreover, PBMCs have specific miRNA profiles that are altered under certain pathological conditions,  making them great candidates as disease biomarkers [@Meder6].
 
-PBMCs can respond to several insulting conditions, such as MI, in the least possible time with notable changes in their miRNA profile [@Meder6].
-Considering the regulatory roles, subtle changes in the transcription of miRNAs can be monitored even before alteration in the levels of mRNAs and proteins [@miR].
-These features make the miRNAs an early-stage valid diagnostic tool for the detection of minor and major cell injuries.
-To date, few studies have been performed to compare the miRNA profiles in PBMCs belonging to MI patients and other CADs and healthy samples to find a robust set of identical miRNAs to differentiate these pathological conditions.
+PBMCs can respond to several insulting conditions, such as MI, in the shortest possible time with notable changes in their miRNA profile [@Meder6].
+Considering their regulatory roles, subtle changes in the transcription of miRNAs can be monitored even before alterations in mRNA and protein levels[@miR].
+These features make miRNAs a valid early-stage diagnostic tool for the detection of minor and major cell injuries.
+To date, few studies have compared the miRNA profiles in PBMCs from patients with MI and other CADs and healthy samples to find a robust set of identical miRNAs to differentiate these pathological conditions.
 
-In this study, we combined three GEO datasets of healthy, CAD, and MI samples.
-Having these samples set alongside bioinformatics analysis and ML means enabled us to identify potential biomarker sets and effective therapeutic targets.
-The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(fig:venn)) are proof of the close relationship between the MI and CAD samples.
-Interestingly, functional enrichment analysis demonstrated that DEGs in both CAD/healthy and MI/healthy were strongly correlated to immune cell response, which is a major cellular part of PBMCs.
-Here, two sets of miRNAs were selected as biomarker sets for sample classification.
-miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-186, miR-21; miR-29a; miR-197; and miR-296 were selected according to their AUC-ROC values.
-As shown in Figure \@ref(fig:miRROC), all miRNAs selected with both approaches had AUC-ROC over 0.9 for isolating healthy and not-healthy samples except for miR-296 and miR-29a.
-Data confirmed that the real challenge is to classify CAD and MI samples because of close overlap.
-Of 6 miRNAs under investigation in both approaches, except for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of CAD and MI samples.
-As expected, the high AUC-ROC values of miRNAs confirm their high potential as biomarkers.
+In this study, we combined three GEO datasets for healthy, CAD, and MI samples.
+Having these sample sets alongside bioinformatics analysis and ML methods enabled us to identify potential biomarker sets and effective therapeutic targets.
+The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(fig:venn)) prove the close relationship between the MI and CAD samples.
+Interestingly, functional enrichment analysis demonstrated that DEGs in both CAD/healthy and MI/healthy were strongly correlated with the immune cell response, which is a major part of PBMCs.
+Two sets of miRNAs were selected as biomarker sets for sample classification.
+miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-186; miR-21; miR-29a; miR-197; and miR-296 were selected based on their AUC-ROC values.
+As shown in Figure \@ref(fig:miRROC), all miRNAs selected with both approaches had AUC-ROCs $> 0.9$ for isolating healthy and not-healthy samples except for miR-296 and miR-29a.
+The data confirmed that the real challenge was to classify CAD and MI samples because of the close overlap.
+Of the six miRNAs under investigation in both approaches, except for miR-32, all miRNAs had an AUC-ROC $> 0.8$ for the discrimination of CAD and MI samples.
+As expected, the high AUC-ROC values of the miRNAs confirmed their high potential as biomarkers.
 
-ML models, trained with miRNA sets selected by both DEG and AUC-ROC approaches, showed better performance in the classification than each miRNA.
-To avoid unwanted complexity and poor predictive values, a two-layer architecture was also designed.
-The first layer was for the discrimination of healthy from not-healthy samples, and the second layer was for separating the CAD from MI candidates.
-As expected, in both approaches, a hyper-tuned SVM model could flawlessly separate healthy from not-healthy samples using distinct miRNA sets.
-The ML models were also capable of effectively separating CAD from MI patients.
-Although both miRNA sets had nearly the same AUC-ROC with their best model, the accuracy, sensitivity, and specificity were different.
-The model trained with AUC-selected miRNAs had better performance in all predictive values, which is logical because of more miRNAs in the set.
+ML models trained with miRNA sets selected by both DEG and AUC-ROC approaches, showed better  classification performance than each miRNA.
+To avoid unwanted complexity and poor predictive values, a two-layer architecture was designed.
+The first layer was used to discriminate between healthy and not-healthy samples, and the second layer was was used to separate CAD from MI candidates.
+As expected, in both approaches, a hypertuned SVM model could flawlessly separate healthy and not-healthy samples using distinct miRNA sets.
+ML models are also capable of effectively separating CAD from MI patients.
+Although both miRNA sets had nearly the same AUC-ROC using the best model, their accuracy, sensitivity, and specificity were different.
+The model trained with AUC-selected miRNAs showed better performance in all predictive values, which is logical because of the higher number of miRNAs in the set.
 
-Numerous studies have reported different biological processes can affect the expression of miRNAs in PBMCs.
-However, there are still controversies regarding the exact role of miRNAs in the function of immune cells and the correlation of specific pathological conditions with miRNA profiles.
-Several studies have proved the activation of particular miRNA types in PBMCs under cardiovascular events [@21; @186; @32-2; @29aa; @197, @296].
-For instance, there is evidence that the elevation of miR-186 suppresses the expression of Cystathionine-$\gamma$-lyase, leading to the subsequent secretion of pro-inflammatory cytokines and cellular lipid accumulation.
-Besides, macrophage-derived miR-186 may promote atherosclerotic plaques [@186].
-In line with this claim, we found that miR-186 is up-regulated in both CAD and MI candidates related to control counterparts.
-Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients than in MI (Figure \@ref(fig:DEMexp)).
-To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, with a clear up-regulation in CAD, indicating its main role in the promotion of atherosclerosis.
+Numerous studies have reported that different biological processes can affect the miRNA expression in PBMCs.
+However, the exact role of miRNAs in the function of immune cells and the correlation between specific pathological conditions and miRNA profiles remain controversial.
+Several studies have proven the activation of particular miRNA types in PBMCs under cardiovascular events [@21; @186; @32-2; @29aa; @197, @296].
+For instance, there is evidence that elevation of miR-186 suppresses the expression of cystathionine-$\gamma$-lyase, leading to the subsequent secretion of pro-inflammatory cytokines and cellular lipid accumulation.
+In addition, macrophage-derived miR-186 may promote atherosclerotic plaque formation [@186].
+In line with this claim, we found that miR-186 was up-regulated in both CAD and MI candidates compared to their control counterparts.
+Surprisingly, the obtained data indicated that the expression of miR-186 was higher in patients with CAD than in patients with MI (Figure \@ref(fig:Expall)).
+Specifically, miR-186 was the only differentially expressed miRNA between CAD and MI, with a clear up-regulation in CAD, indicating its main role in the promotion of atherosclerosis.
 
-As mentioned before, miR-21 was also up-regulated in both MI and CAD patients in comparison to healthy controls.
-Moreover, the expression value of miR-21 was significantly higher in MI than in the CAD group (Table \@ref(tab:mirExptable)).
-It is thought that the up-regulation of miRNA-21 in PBMCs is a compensatory reaction to reduce T$_{reg}$ lymphocyte number in response to the reduction of TGF$\beta1$ secretion into the plasma through a TGF$\beta1$/smad-independent pathway.
-In line with previous and present data, miR-21 can modulate the activity of PBMCs following the occurrence of cardiovascular diseases [@21].
+As mentioned before, miR-21 was also up-regulated in both MI and CAD patients compared to healthy controls.
+Moreover, the expression value of miR-21 was significantly higher in the MI group than in the CAD group (Table \@ref(tab:mirExptable)).
+It is thought that the up-regulation of miR-21 in PBMCs is a compensatory reaction to reduce the T$_{reg}$ lymphocyte number in response to the reduction in TGF$\beta1$ secretion into the plasma through a TGF$\beta1$/smad-independent pathway.
+In line with the previous and present data, miR-21 can modulate the activity of PBMCs following the occurrence of cardiovascular diseases [@21].
 
-Recent data have supported the elevation of miR-32 in CAD samples with the calcification of the coronary artery.
-It is worth noting that miR-32 promotes vascular smooth muscle calcification in mice by controlling the activity of several proteins, including bone morphogenetic protein-1, runt-related transcription factor-2 (RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA protein.
+Recent data have supported the elevation of miR-32 levels in CAD samples with calcification of the coronary artery.
+Notably, miR-32 promotes vascular smooth muscle calcification in mice by controlling the activity of several proteins, including bone morphogenetic protein-1, runt-related transcription factor-2 (RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA protein.
 [@32-2].
-Likewise, some reports are associated with the activity of miR-32 in PBMCs under several pathologies [@magic; @32].
-The exact role of miR-32 in PBMCs after cardiovascular events remains to be elucidated.
+Likewise, some reports are associated with the activity of miR-32 in PBMCs in several pathologies [@magic; @32].
+The exact role of miR-32 in PBMCs after cardiovascular events remains unclear.
 
-Molecular analyses have indicated the regulatory role of miRNAs selected by the AUC-ROC approach in PBMCs after a cardiovascular event.
-Two common miRNAs in DEGs and AUC-ROC approaches, miR-21 and miR-186, were covered already.
+Molecular analyses have indicated the regulatory role of miRNAs selected using the AUC-ROC approach in PBMCs after a cardiovascular event.
+The biological importance of two common miRNAs in the DEGs and AUC-ROC approaches, miR-21 and miR-186, have already been discussed.
 Based on numerous reports, miR-29a can be activated in different diseases [@29aa].
-Data analysis indicated that miR-29a is significantly up-regulated in CAD patients compared to healthy and MI groups (Table \@ref(tab:mirExptable)).
-Increased miR-29a is associated with the progression of atherosclerosis, and the combination of miR-29a and ox-LDL was offered as a valid biomarker set for paraclinical classification [@29a].
-However, the role of miR-29a in the function of PBMCs in CAD patients has not been thoroughly examined.
+Data analysis indicated that miR-29a was significantly up-regulated in CAD patients compared to the healthy and MI groups (Table \@ref(tab:mirExptable)).
+Increased miR-29a is associated with the progression of atherosclerosis, and the combination of miR-29a and ox-LDL has been suggested as a valid biomarker set for paraclinical classification [@29a].
+However, the role of miR-29a in the function of PBMCs from patients with CAD has not been thoroughly examined.
 
-Data indicated that miR-197 is also significantly up-regulated in both CAD/healthy and MI/healthy groups.
-Previous research has demonstrated that miR-197 may play a crucial role in controlling the anti-inflammatory response of IL-35 by influencing the secretion of cytokines that can either promote or suppress inflammation, the ratio of M1/M2 macrophages, and the proliferation of T$_{reg}$ lymphocytes, which are responsible for suppressing immune responses [@197].
+The data indicated that miR-197 was significantly up-regulated in both the CAD/healthy and MI/healthy groups.
+Previous studies have demonstrated that miR-197 may play a crucial role in controlling the anti-inflammatory response of IL-35 by influencing the secretion of cytokines that can either promote or suppress inflammation, the ratio of M1/M2 macrophages, and the proliferation of T$_{reg}$ lymphocytes, which are responsible for suppressing immune responses [@197].
 Alongside our findings, it can be concluded that miR-197 could be a useful diagnostic tool for predicting adverse cardiovascular events.
 
-The findings of this study demonstrate the potential of miR-296 as a biomarker with high discriminatory power for distinguishing between samples from individuals with MI and CAD.
+The findings of this study demonstrate the potential of miR-296 as a biomarker with high discriminatory power to distinguish between samples from individuals with MI and CAD.
 MiR-296 has been identified as a key regulator in the development and advancement of atherosclerosis by controlling the expression of target genes associated with various biological processes, including angiogenesis, cholesterol metabolism, inflammation, cellular proliferation, hypertension, and apoptosis [@296].
-In a study, it has been shown that miR-296 expression levels are significantly increased in the PBMCs of CAD patients compared to healthy controls, suggesting its involvement in regulating pro-inflammatory cytokines like IL-6 and TNF-a [@296a].
-These findings suggest that miR-296 may have a significant impact on the pathogenesis of atherosclerosis and could potentially serve as a diagnostic biomarker for CAD or MI.
+In a previous study, miR-296 expression levels were found to be significantly increased in the PBMCs of CAD patients compared to healthy controls, suggesting its involvement in regulating proinflammatory cytokines such as IL-6 and TNF-$\alpha$ [@296a].
+These findings suggested that miR-296 may have a significant impact on the pathogenesis of atherosclerosis and could potentially serve as a diagnostic biomarker for CAD or MI.
 
@@
 
-In summary, we derived a set of miRNA biomarkers by comparing MI samples to both healthy and CAD samples.
-We found that the SVM model performed best in both the first layer, which separated healthy and not-healthy samples, and the second layer, which classified MI/CAD samples.
+In summary, we derived a set of miRNA biomarkers by comparing MI samples with both healthy and CAD samples.
+We found that the SVM model performed best in both the first layer, which separated healthy and unhealthy samples, and the second layer, which classified the MI/CAD samples.
 The set of miRNAs selected based on their AUC-ROC values performed better in the second layer.
 Overall, our two-layer structure achieved an accuracy of 0.96.
-This demonstrates the potential for combining bioinformatics and machine learning techniques to identify novel biomarkers and gain a deeper understanding of myocardial infarction.
+This demonstrates the potential of combining bioinformatics and machine learning techniques to identify novel biomarkers and gain a deeper understanding of myocardial infarction.
 

### 2023-08-07 ArticleDraft.tex
The version after revision

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Machine Learning Driven Set of microRNAs as a Novel Biomarker Set for Myocardial Infarction Diagnosis},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Machine Learning Driven Set of microRNAs as a Novel Biomarker Set
for Myocardial Infarction Diagnosis}
\author{}
\date{\vspace{-2.5em}2023-06-20}

\begin{document}
\maketitle
\begin{abstract}
MicroRNAs (miRNAs) play a crucial role in regulating adaptive and
maladaptive responses in cardiovascular diseases, making them attractive
targets for potential biomarkers. However, their potential as novel
biomarkers for diagnosing cardiovascular diseases requires systematic
evaluation. In this study, we aimed to identify a key set of miRNA
biomarkers using integrated bioinformatics and machine learning
analysis. We combined and analyzed three gene expression datasets from
the Gene Expression Omnibus (GEO) database, which contains peripheral
blood mononuclear cells (PBMCs) samples from individuals with myocardial
infarction (MI), stable coronary artery disease (CAD), and healthy
individuals. Additionally, we selected a set of miRNAs based on their
area under the receiver operating characteristic curve (AUC-ROC) for
separating the CAD and MI samples. We designed a two-layer architecture
for sample classification, in which the first layer isolates healthy
samples from not-healthy ones, and the second layer classifies stable
CAD and MI samples. We trained different machine learning models using
both biomarker sets and evaluated their performance on a test set. We
identified miR-21, miR-186, and miR-32 as the only miRNAs among the
differentially expressed genes, and a set including miR-186, miR-21,
miR-197, miR-29A, and miR-296 as the optimum set of miRNAs selected by
their AUC-ROC. Both biomarker sets could distinguish healthy from
not-healthy samples with complete accuracy. The best performance for the
classification of CAD and MI was achieved with an SVM model trained
using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.96 and
an accuracy of 0.94 on the test data. Our study demonstrated that miRNA
signatures derived from PBMCs could serve as valuable novel biomarkers
for cardiovascular diseases.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Cardiovascular diseases (CVDs) are the leading cause of human mortality,
accounting for 32\% of all global deaths. It is estimated that about
85\% of CVDs mortality is due to myocardial infarction (MI)
({``Cardiovascular Diseases ({CVDs})''} n.d.). MI is an acute coronary
syndrome characterized by sudden blockage and stenosis of the coronary
artery and subsequent myocardial ischemia, leading to extensive
cardiomyocyte damage and necrosis (Yap et al. 2023).

Over the last 50 years, numerous attempts have been made to use
biomarkers to facilitate diagnosis, assess the risk, follow-up therapy,
and determine therapeutic efficacy in CVDs candidates. Based on released
guidelines, cardiac troponins (cTns) are used as a highly sensitive and
accurate approach for detecting MI. Despite these inherent advantages,
the high-rate sensitivity of cTn-based assays has also led to more
false-positive results (Thygesen et al. 2018), necessitating the advent
and development of new modalities with pathological values. To improve
the diagnostic value of existing MI biomarkers, a combination of
complementary biological markers, such as microRNAs (miRNAs) and other
genetic factors, has been proposed. Previous research supports the
notion that miRNAs exhibit great potential as alternative biomarkers for
CVDs detection and follow-up (Schulte et al. 2020). It has been
suggested that miRNAs possess 18-22 nucleotides and play a crucial role
in the regulation of gene expression. Evidence point to the fact that
miRNAs are involved in the pathogenesis of cardiac tissue injury
(Schulte, Karakas, and Zeller 2017). Several biological processes, such
as angiogenesis, cardiomyocyte growth and contractility, lipid
metabolism, plaque formation, and cardiac rhythm, are regulated by
miRNAs (Kalayinia et al. 2021). These elements can easily circulate in
biofluids and could be considered theranostics targets in terms of CVDs
(Schulte, Karakas, and Zeller 2017). It has been postulated that the
function and diagnostic properties of miRNAs are beyond the myocardium
in patients with CVD. To be specific, the expression of miRNAs can vary
in different biofluids and cell components such as serum and peripheral
blood mononuclear cells (PBMCs) (Soler-Botija, Gálvez-Montón, and
Bayés-Genís 2019).

PBMCs are a fraction of the white blood cells, including monocytes,
lymphocytes, macrophages, and other cells of the immune system (Gao et
al. 2020). Emerging data indicate that PBMCs can be used as a valid
source of biomarkers for monitoring various pathological conditions. Of
note, the alteration of mRNAs and miRNAs under pathological conditions
provides valuable information about different kinds of disorders. PBMCs
can recapitulate the conditions of target tissues, thus providing a
highly sensitive and specific source of biomarkers (Mosallaei et al.
2022). Combines with these conditions, these cells are repositories of
dysregulated genes and miRNA expression profiles in CVDs (Gao et al.
2020; Mosallaei et al. 2022).

In recent years, the advent and application of machine learning (ML) has
been an exciting prospect for advancing scientific research. Although
the concept of ML and its initial algorithms were conceived many years
ago, recent improvements in computing power and access to vast amounts
of data have demonstrated that ML techniques outperform classical
statistical methods in various fields. Furthermore, the progress made in
omics technologies has enabled the analysis of massive and intricate
biological datasets, consisting of hundreds to thousands of samples,
which makes it possible for ML to extract valuable biological insights
and information from such data (Torun et al. 2023). Consequently, ML
provides innovative methods for merging and interpreting diverse types
of omics data, leading to the identification of new biomarkers. These
biomarkers can aid in precise disease prediction, patient
stratification, and development of novel therapeutic approaches (Reel et
al. 2021).

In this study, we aimed to identify potential miRNA biomarkers in
patients with MI by combining and analyzing three different microarray
datasets from PBMCs. The integration of omics data with bioinformatics
and ML techniques could be a promising tool in the discovery of new and
more accurate biomarkers for monitoring MI. Additionally, this approach
can deepen our understanding of the underlying mechanisms of MI and aid
in the development of valid diagnostic biomarkers and patient
stratification.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{microarray-data-collection}{%
\subsection{Microarray data
collection}\label{microarray-data-collection}}

Microarray datasets were obtained from the Gene Expression Omnibus (GEO)
database (\url{https://www.ncbi.nlm.nih.gov/geo/}). To obtain sufficient
classification power between MI, healthy, and CAD samples, a relatively
large sample size is required. Therefore, GSE59867 for the MI and CAD
samples and GSE56609 and GSE54475 for the healthy samples were selected.
All samples were produced using Affymetrix Human Gene 1.0 ST Array
platform (GPL6244). Only healthy, CAD, and early stage MI samples were
selected from these datasets for further analysis. The basic information
for the three datasets evaluated in this study is provided in Table
@ref(tab:datasets). Bioinformatics sections was fully conducted using R,
ver. 4.2.0 (R Core Team 2022), RStudio (RStudio Team 2020). All plots
and graphics of these sections were created using the ggplot2 R package
(Wickham 2016).

\begin{longtable}[]{@{}cccccc@{}}
\caption{Samples information on the GEO microarray
datasets.}\tabularnewline
\toprule()
Dataset & Platform & Healthy & CAD & MI & Refrence \\
\midrule()
\endfirsthead
\toprule()
Dataset & Platform & Healthy & CAD & MI & Refrence \\
\midrule()
\endhead
GSE59867 & GPL6244 & - & 46 & 111 & (Maciejak et al. 2015) \\
GSE56609 & GPL6244 & 46 & - & - & (Matone et al. 2015) \\
GSE54475 & GPL6244 & 5 & - & - & (Canali et al. 2014) \\
\bottomrule()
\end{longtable}

\hypertarget{pre-processing}{%
\subsection{Pre-processing}\label{pre-processing}}

The raw data in the form of CEL files from all datasets were obtained
from GEO. To prepare the data for analysis, we utilized the fRMA package
(M. N. McCall, Bolstad, and Irizarry 2010) to facilitate pre-processing
of individual microarray samples and their consistent combination. For
each dataset, background correction was applied using the RMA algorithm,
followed by quantile normalization based on the reference distribution.
To account for probe-specific effects, batch effects were eliminated
during summarization and gene expression variances were estimated
accordingly. In cases where multiple probe sets matched the same gene,
the mean log-fold change was retained. Consequently, fRMA can serve as a
technique to remove batch effects across diverse datasets generated by
identical microarray platforms (Lazar et al. 2013). To ensure the
effectiveness of the batch effect removal, we employed principal
component analysis (PCA) and relative log expression (RLE) plots to
visualize the data before and after applying fRMA.

\hypertarget{differential-expression-analysis}{%
\subsection{Differential expression
analysis}\label{differential-expression-analysis}}

The barcode algorithm was introduced by McCall et al. (Matthew N. McCall
et al. 2011), aimed to convert actual expression values into binary
barcode values. Extensive sample collections were gathered and
normalization was performed using fRMA across multiple platforms,
including the Affymetrix Human Gene 1.0 ST Array (GPL6244) platform. By
utilizing these normalized datasets, the distribution of the observed
intensities for both the expressed and unexpressed genes was estimated.
The determination of whether a gene was expressed or not was based on
the following equation, where a value of 1 indicates expression and a
value of 0 indicates non-expression:

\[
\hat{x}_{ij} = \left\{
  \begin{array}{ll}
    1 & \mbox{if } x_{ij} >= \mu^{ne} + C \times \sigma^{ne} \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\]

In the barcode algorithm, the normalized intensity of gene \(i\) in
sample \(j\) is denoted as \(x_{ij}\). A user-defined parameter, \(C\),
was introduced along with the standard deviation (\(\sigma^{ne}\)) and
mean (\(\mu^{ne}\)) of the non-expressed distribution. Based on these
values, the barcode representation of a sample was generated as a vector
consisting of ones and zeros, representing the estimated expression
(ones) and non-expression (zeros) of each gene. The barcode function
within the R fRMA package was employed to implement the barcode
algorithm, utilizing the default value of \(C\).

To assess the differences in expressed ratios between the MI and healthy
control groups, Fisher's exact test was performed on the barcode values
of individual genes. Genes that exhibited a false discovery rate (FDR)
below \(0.05\), calculated using the Benjamini-Hochberg procedure to
account for multiple testing issues were identified as differentially
expressed genes (DEGs). The same procedures were applied to the CAD
versus healthy control comparison, as well as to the MI versus CAD
group, to identify DEGs specific to each comparison.

\hypertarget{functional-and-pathway-enrichment-analyses}{%
\subsection{Functional and pathway enrichment
analyses}\label{functional-and-pathway-enrichment-analyses}}

The R clusterProfiler package (Yu et al. 2012) was utilized to perform
the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment
analysis and Gene Ontology (GO) functional annotation on the set of
DEGs. GO analysis encompassed three categories: biological process (BP),
cellular component (CC), and molecular function (MF). For statistical
significance, an adjusted p-value threshold of less than 0.05 was
employed. Enrichment analyses were conducted separately for DEGs
specific to the MI-healthy and CAD-healthy comparisons. All the default
parameters provided by the package were used in the analyses.

\hypertarget{ml-procedure}{%
\subsection{ML procedure}\label{ml-procedure}}

ML analysis was performed using Python software, ver. 3.9, Numpy (Harris
et al. 2020), Pandas (McKinney 2010), and Scikit-Learn packages
(Pedregosa et al. 2011). Whenever hyper-tuning was needed, the
Scikit-opt package (Head et al. 2021) was used. In all ML analyses, the
datasets were divided into training and test sets at a 0.7:0.3 ratio,
and all reported results are the average of 10-fold cross-validation.

Two different approaches were used to select miRNAs for model training.
The first approach was to use differentially expressed miRNAs. In the
second approach, miRNAs with individual AUC-ROC over 0.8 for separating
MI from CAD were selected. The results of these two approaches can
provide an informative comparison between the predictive capabilities of
sets of miRNAs selected with different logics.

\hypertarget{mirnas-in-degs}{%
\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs}}

In this approach, a two-layer architecture is deployed to the data to
maximize the prediction values. The first layer predicted whether a
sample was healthy or not, and the second layer separated MI from CAD in
the samples that were predicted as not healthy in the first layer. To
this end, a distinct ML model was trained for each layer. Becuase, there
were a limited number of miRNAs in the DEGs, both layers were trained
with all of them. For further comparison with the models' performance,
the ROC curve of each miRNA for classifying healthy and not-healthy, as
well as CAD and MI, was generated using a Logistic Regression model.

\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples}{%
\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples}}

A support vector machine (SVM) model using RBF kernels was trained and
hyper-tuned using all miRNAs in the DEGs. To handle the severe imbalance
in the number of samples (51 for the healthy group and 157 for the
not-healthy group), the sample weights for the healthy and the
not-healthy samples were set to 1 and 0.5, respectively. The ROC curve
and confusion matrix for the model are reported.

\hypertarget{second-layer-for-separating-the-mi-and-cad-samples}{%
\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples}}

Different models were investigated to achieve the highest classification
performance. To do so, SVM (with linear, polynomial, and RBF kernels),
Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN),
Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) models were
trained. All models were trained with their preset parameters using
10-fold cross-validation. The criteria for selecting the best model were
the highest accuracy and AUC-ROC for the test set. The best model was
hyper-tuned using the scikit-opt package (Head et al. 2021) for the best
classification performance. The ROC curve and confusion matrix for the
best model are reported.

\hypertarget{mirnas-with-the-highest-auc-roc}{%
\subsubsection{miRNAs with the highest
AUC-ROC}\label{mirnas-with-the-highest-auc-roc}}

As in the previous approach, a two-layer strategy was employed. The
first layer classified samples into healthy and not-healthy, and the
second layer separated the MI and CAD samples. However, to keep the
number of miRNAs as low as possible, miRNAs were selected from the
second layer and their performance was evaluated in the first layer. The
AUC-ROC of all miRNAs for classifying MI and CAD samples was calculated,
and miRNAs with AUC-ROC \(> 0.8\) were selected. ROC curves for each
selected miRNA for separating healthy samples from not-healthy and MI
from CAD samples were also plotted for further comparison.

\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-1}{%
\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-1}}

An SVM model with an RBF kernel is trained using the selected set of
miRNAs. Additionally, the model was hyper-tuned to find the
hyper-parameters for the highest AUC-ROC and accuracy. The same sample
weights as in the previous approach (1 for healthy and 0.5 for
not-healthy samples) were used. The ROC curve and confusion matrix for
the model were reported.

\hypertarget{second-layer-for-separating-the-mi-and-cad-samples-1}{%
\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-1}}

The selected miRNA set was used to train different algorithms to
determine the best model. Similar to the previous approach, the SVM
(with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT
models were trained. All models were trained with their pre-set
parameters using 10-fold cross-validation. The models with the highest
AUC-ROC and accuracy on the test set were selected and hyper-tuned using
the scikit-opt package (Head et al. 2021). The ROC curve and confusion
matrix for the best model were reported.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{pre-processing-1}{%
\subsection{Pre-processing}\label{pre-processing-1}}

The PCA plots of the samples are shown in Figures @ref(fig:PCA)A and B.
Healthy samples were separated from the CAD or MI samples in the primary
data and after conducting fRMA. In the RLE plot, there was a distinct
difference between the dataset means for all samples before fRMA was
performed (Figure @ref(fig:PCA)C). All datasets were rearranged to
approximately 0 in the RLE plot after fRMA was conducted (Figure
@ref(fig:PCA)D). Moreover, there was an apparent change in the
inter-quantile distances, but the values were still greater then 0.1.

\begin{figure}
\includegraphics[width=1\linewidth]{../DEG and DEM/Pics/PCA and RLE of trains} \caption{principal component analysis plots for (A) primary data and (B) the data after fRMA, and the relative log expression plots for (C) primary data and (D) the data after fRMA.}\label{fig:PCA}
\end{figure}

\hypertarget{differential-expression-analysis-1}{%
\subsection{Differential expression
analysis}\label{differential-expression-analysis-1}}

\begin{table}

\caption{\label{tab:DEGstab}Total, up-, and down-regulated DEGs and differentially expressed miRNAs.}
\centering
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\centering}X>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{5cm}}
\toprule
  & Total DEGs & Up-regulated DEGs & Down-regulated DEGs & miRNAs\\
\midrule
MI vs. Healthy & 860 & 323 & 537 & hsa-miR-186, miR-21, miR-32\\
CAD vs. Healthy & 670 & 262 & 408 & hsa-miR-186, miR-21, miR-32\\
MI vs. CAD & 260 & 144 & 116 & hsa-miR-186\\
\bottomrule
\end{tabu}
\end{table}

According to the cutoff criterion of \(FDR < 0.05\), there were 860 DEGs
between MI and healthy samples. Among them, 323 were up-regulated, and
537 were down-regulated in the MI group compared to those in healthy
group. In the CAD and healthy group comparison, we found 670 DEGs, of
which 262 and 408 DEGs were up- and down-regulated, respectively, in CAD
samples. In the MI and CAD groups, the number of DEGs was 260, and the
number of up- and down-regulated genes in MI samples were 144 and 116,
respectively, compared to CAD samples. The data are summarized in Table
@ref(tab:DEGstab).

\begin{figure}

{\centering \includegraphics[width=0.45\linewidth]{ArticleDraft_files/figure-latex/venn-1}

}

\caption{Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison.}\label{fig:venn}
\end{figure}

The Venn diagram in Figure @ref(fig:venn) shows that the CAD and MI
samples shared most of their DEGs. From 860 DEGs of MI/healthy and 670
DEGs of CAD/healthy, 531 genes were common, which is 62\% of MI/healthy
DEGs and 79\% of CAD/healthy DEGs.

\hypertarget{go-and-kegg-enrichment-analyses-of-the-degs}{%
\subsection{GO and KEGG enrichment analyses of the
DEGs}\label{go-and-kegg-enrichment-analyses-of-the-degs}}

To explore the biological classification of the DEGs, we performed GO
and KEGG pathway enrichment analyses on the MI/healthy and CAD/healthy
DEGs. For MI/healthy, GO enrichment analysis in the BP category
suggested that the DEGs were enriched in ``immune response-regulating
signaling pathway,'' ``lymphocyte differentiation,'' ``immune
response-regulating cell surface receptor signaling pathway,'' and
``leukocyte activation involved in immune response'' (Figure
@ref(fig:MIHEnrich)A). In the CC category, DEGs were enriched in
``secretory granule membrane,'' ``azurophil granule,'' ``ficolin-1-rich
granule,'' ``tertiary granule,'' and ``ficolin-1-rich granule membrane''
(Figure @ref(fig:MIHEnrich)B). In the MF category, DEGs were involved in
``cadherin binding'' and ``MHC class I protein binding'' (Figure
@ref(fig:MIHEnrich)C). KEGG pathway analysis indicated that the DEGs
were related to the following pathways: ``Chemokine signaling pathway,''
``Lipid and atherosclerosis,'' and ``Hematopoietic cell lineage''
(Figure @ref(fig:MIHEnrich)D).

\begin{figure}
\includegraphics[width=0.95\linewidth]{../DEG and DEM/Pics/Enrichment MI-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}\label{fig:MIHEnrich}
\end{figure}

The enrichment results for the CAD/healthy DEGs were as follows. In the
BP category, GO enrichment suggested that the DEGs were enriched in
``positive regulation of defense response,'' ``positive regulation of
innate immune response,'' ``mononuclear cell differentiation,'' and
``positive regulation of response to external stimulus'' (Figure
@ref(fig:CADHEnrich)A). In the CC category, DEGs were enriched in
``azurophil granule,'' ``ficolin-1-rich granule,'' and ``ficolin-1-rich
granule membrane'' (Figure @ref(fig:CADHEnrich)B). In the MF category,
DEGs were involved in ``lipoprotein particle receptor binding'' and
``NF-\(\kappa\)B binding'' (Figure @ref(fig:CADHEnrich)C). KEGG pathway
analysis showed that the DEGs were related to the following pathways:
``Chemokine signaling pathway,'' ``Lipid and atherosclerosis,'' and
``Hematopoietic cell lineage'' (Figure @ref(fig:CADHEnrich)D).

\begin{figure}
\includegraphics[width=0.95\linewidth]{../DEG and DEM/Pics/Enrichment CAD-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}\label{fig:CADHEnrich}
\end{figure}

\hypertarget{machine-learning}{%
\subsection{Machine Learning}\label{machine-learning}}

\hypertarget{mirnas-in-degs-1}{%
\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs-1}}

Among the DEGs, miR-186, miR-32, and miR-21 were identified as
differentially expressed miRNAs. The expression profiles of the three
miRNAs are shown in Figure @ref(fig:Expall). The ROC curves of each
miRNA in each layer are presented in Figure @ref(fig:miRROC). Using the
logistic regression model, the AUC-ROC values of miR-21, miR-32, and
miR-186 for separating healthy and not-healthy samples were 0.98, 0.99,
and 0.90, respectively (Figure @ref(fig:miRROC)A). The accuracy of each
miRNA for classifying the samples into healthy and not-healthy groups on
the test set for miR-21, miR-32, and miR-186 was 0.92, 0.98, and 0.89,
respectively. The ROC curve of each miRNA for classifying MI and CAD
samples is presented in Figure @ref(fig:miRROC)B. The AUC-ROC and
accuracy for miR-21, miR-32, and miR-186 in the test set were 0.85;
0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.

\begin{table}

\caption{\label{tab:mirExptable}Investigated miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{CAD/Healthy} & \multicolumn{2}{c}{MI/Healthy} & \multicolumn{2}{c}{MI/CAD} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
  & logFC & adj. p-value & logFC & adj. p-value & logFC & adj. p-value\\
\midrule
miR-186 & 1.4 & 3.60e-25 & 0.9 & 6.76e-20 & -0.5 & 1.05e-09\\
miR-21 & 1.4 & 1.31e-17 & 2.3 & 2.07e-47 & 0.8 & 2.96e-11\\
miR-32 & 2.5 & 8.39e-43 & 2.2 & 3.10e-59 & -0.3 & 7.60e-04\\
miR-197 & 0.5 & 2.95e-20 & 0.7 & 1.59e-47 & 0.2 & 8.58e-09\\
miR-29A & 0.7 & 7.76e-29 & 0.1 & 1.70e-01 & -0.5 & 2.14e-10\\
miR-296 & -0.1 & 5.00e-02 & 0.1 & 2.00e-02 & 0.2 & 6.15e-06\\
\bottomrule
\end{tabu}
\end{table}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../ML/Pics/Expressionforall}

}

\caption{Expression profile of all miRNAs in two approaches in different sample classes.}\label{fig:Expall}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../ML/Pics/miRs ROCs}

}

\caption{ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples.}\label{fig:miRROC}
\end{figure}

\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-2}{%
\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-2}}

Although single miRNAs had an acceptable performance for this layer,
their predictive value could be further improved by using them as a set.
The ROC curve for the SVM model with an RBF kernel trained with all
three miRNAs is presented in Figure @ref(fig:DEMROC)A. The model had a
better performance in classification than single miRNAs.The AUC-ROC for
the model is 1, and its accuracy on the test set was also 1. In Figure
@ref(fig:DEMCF)A, the confusion matrix for the model is presented.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../ML/Pics/DEMs ROCs h not h cad mi}

}

\caption{ROC curve for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification.}\label{fig:DEMROC}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../ML/Pics/DEMs Conf. matrix}

}

\caption{Confusion matrix for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification.}\label{fig:DEMCF}
\end{figure}

\hypertarget{second-layer-for-separating-the-mi-and-cad-samples-2}{%
\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-2}}

Different models were trained using the expression values of three
differentially expressed miRNAs. The models' AUC-ROC and the accuracy of
the test set are shown in Figure @ref(fig:DEMmodels). The best model
from both AUC-ROC and accuracy point of view was the SVM model with a
linear kernel. The AUC-ROC and accuracy for this model with its preset
values were 0.93 and 0.82, respectively. The model was hyper-tuned for C
and gamma hyper-parameters, and therefore the model showed better
performance. The ROC curve of the hyper-tuned model is presented in
Figure @ref(fig:DEMROC)B. For this model, the AUC-ROC reached 0.95, and
the accuracy was improved to 0.85 (Table @ref(tab:DEGsML)). Moreover,
the sensitivity and specificity for the model on the test set were 0.91
and 0.71, respectively. The confusion matrix for the hyper-tuned model
is illustrated in Figure @ref(fig:DEMCF)B.

\begin{figure}

{\centering \includegraphics[width=0.65\linewidth]{../ML/Pics/Models DEGs mirs}

}

\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs.}\label{fig:DEMmodels}
\end{figure}

\begin{table}

\caption{\label{tab:DEGsML}AUC-ROC and accuracy for SVM with a linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning}
\centering
\begin{tabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Pre-set parameters} & \multicolumn{2}{c}{Hyper-tuned} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
Model & Metrics & train & test & train & test\\
\midrule
SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.92 & 0.95\\
 & Accuracy & 0.83 & 0.82 & 0.84 & 0.85\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{auc-roc-approach}{%
\subsubsection{AUC-ROC approach}\label{auc-roc-approach}}

After calculating the AUC-ROC for each miRNA to classify of MI and CAD
samples, the miRNAs with AUC-ROC \(> 0.8\) were selected. The miRNAs
selected were miR-29a, miR-197, miR-186, miR-21, and miR-296. The
expression levels of these miRNAs in healthy, CAD, and MI samples are
presented in Figure @ref(fig:Expall). The ROC curves of the selected
miRNAs in both layers is shown in Figure @ref(fig:miRROC).

\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-3}{%
\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-3}}

Using the selected set, an SVM model with an RBF kernel was trained to
separate healthy and not-healthy samples. The ROC curve for the model is
presented in Figure @ref(fig:AUCROC)A, and the confusion matrix is
illustrated in Figure @ref(fig:AUCCF)A. Both the AUC-ROC and accuracy of
the model on the test set were 1.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../ML/Pics/AUC best model ROCs}

}

\caption{ROC curve for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI samples classification. (C) SVM with polynomial kernel for CAD and MI samples classification. }\label{fig:AUCROC}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../ML/Pics/AUC Conf. matrix}

}

\caption{Confusion matrix for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI samples classification. (C) SVM with polynomial kernel for CAD and MI samples classification.}\label{fig:AUCCF}
\end{figure}

\hypertarget{second-layer-for-separating-the-mi-and-cad-samples-3}{%
\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-3}}

To find the best model for this set of miRNAs, different models were
trained using their preset values. The AUC-ROC and accuracy results for
the test set are presented in Figure @ref(fig:AUCmodels). The best model
from the AUC-ROC point of view was the SVM with a linear kernel, and
from the accuracy point of view, it was the SVM model with an RBF
kernel. For the SVM-linear model, the AUC-ROC and accuracy were 0.93 and
0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84,
respectively. Both models were hyper-tuned, and the ROC curve for their
best performance is presented in Figure @ref(fig:AUCROC)B and C. The
AUC-ROC and accuracy for the SVM-linear model were modified to 0.92 and
0.88, respectively. For the SVM-RBF, these values increased to 0.96 and
0.94, respectively (Table @ref(tab:AUCML)). The sensitivity for the
SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and the
specificity for them was 0.79 and 0.86, respectively. The confusion
matrix for both models is illustrated in Figure @ref(fig:AUCCF)B and C.

\begin{figure}

{\centering \includegraphics[width=0.65\linewidth]{../ML/Pics/Models AUC 5 mirs CV}

}

\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with AUC-selected miRNAs.}\label{fig:AUCmodels}
\end{figure}

\begin{table}

\caption{\label{tab:AUCML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their AUC-ROC on the train and test set before and after hyper-tuning}
\centering
\begin{tabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Pre-set parameters} & \multicolumn{2}{c}{Hyper-tuned} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
Model & Metrics & Train & Test & Train & Test\\
\midrule
SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.93 & 0.92\\
 & Accuracy & 0.85 & 0.82 & 0.90 & 0.88\\
SVM-RBF & AUC-ROC & 0.90 & 0.92 & 0.96 & 0.96\\
 & Accuracy & 0.86 & 0.84 & 0.96 & 0.94\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The prevalence of MI can lead to high mortality rates in the clinical
setting. However, early diagnosis and the application of suitable
treatment protocols can reduce mortality and improve MI prognosis
({``Cardiovascular Diseases ({CVDs})''} n.d.; Thygesen et al. 2018; Tsao
et al. 2022). Studies have suggested that changes in miRNA expression
may play a significant role in the progression of MI and the subsequent
remodeling (Laggerbauer and Engelhardt 2022). It is believed that miRNA
expression is altered during the various biological processes correlated
with MI within the myocardium or other related tissues (Khan, Gupta, and
Mahapatra 2022). Although several studies have focused on examining free
circulating miRNAs in serum samples for the detection of cardiac tissue
injuries (Kaur et al. 2020), more information is needed to fully
comprehend the miRNAs found in different blood sub-components, such as
plasma, platelets, and PBMCs. Based on previous findings, PBMCs play a
crucial role in the destabilization and rupture of plaques as well as in
the initial inflammatory reactions in individuals experiencing
myocardial infarction (MI). (Mosallaei et al. 2022; Hapke et al. 2022).
Moreover, PBMCs have specific miRNA profiles that are altered under
certain pathological conditions, making them great candidates as disease
biomarkers (Mosallaei et al. 2022).

PBMCs can respond to several insulting conditions, such as MI, in the
least possible time with notable changes in their miRNA profile
(Mosallaei et al. 2022). Considering their regulatory roles, subtle
changes in the transcription of miRNAs can be monitored even before
alterations in mRNA and protein levels(Schulte et al. 2020). These
features make miRNAs a valid early stage diagnostic tool for the
detection of minor and major cell injuries. To date, few studies have
compared the miRNA profiles in PBMCs from patients with MI and other
CADs and healthy samples to find a robust set of identical miRNAs to
differentiate these pathological conditions.

In this study, we combined three GEO datasets for healthy, CAD, and MI
samples. Having these sample set alongside bioinformatics analysis and
ML means enabled us to identify potential biomarker sets and effective
therapeutic targets. The results of the DEG analysis (Table
@ref(tab:DEGstab) and Figure @ref(fig:venn)) prove the close
relationship between the MI and CAD samples. Interestingly, functional
enrichment analysis demonstrated that DEGs in both CAD/healthy and
MI/healthy were strongly correlated with the immune cell response, which
is a major part of PBMCs. Two sets of miRNAs were selected as biomarker
sets for sample classification. miR-21; miR-32; and miR-186 were
selected as differentially expressed miRNAs, and miR-186; miR-21;
miR-29a; miR-197; and miR-296 were selected based on their AUC-ROC
values. As shown in Figure @ref(fig:miRROC), all miRNAs selected with
both approaches had AUC-ROC over 0.9 for isolating healthy and
not-healthy samples except for miR-296 and miR-29a. The data confirmed
that the real challenge was to classify CAD and MI samples because of
the close overlap. Of the six miRNAs under investigation in both
approaches, except for miR-32, all miRNAs had an AUC-ROC \(> 0.8\) for
the discrimination of CAD and MI samples. As expected, the high AUC-ROC
values of the miRNAs confirmed their high potential as biomarkers.

ML models trained with miRNA sets selected by both DEG and AUC-ROC
approaches, showed better classification performance than each miRNA. To
avoid unwanted complexity and poor predictive values, a two-layer
architecture was designed. The first layer was used to discriminate
between healthy and not-healthy samples, and the second layer was was
used to separate CAD from MI candidates. As expected, in both
approaches, a hyper-tuned SVM model could flawlessly separate healthy
and not-healthy samples using distinct miRNA sets. ML models are also
capable of effectively separating CAD from MI patients. Although both
miRNA sets had nearly the same AUC-ROC using the best model, their
accuracy, sensitivity, and specificity were different. The model trained
with AUC-selected miRNAs showed better performance in all predictive
values, which is logical because of the higher number of miRNAs in the
set.

Numerous studies have reported that different biological processes can
affect the miRNA expression in PBMCs. However, the exact role of miRNAs
in the function of immune cells and the correlation between specific
pathological conditions and miRNA profiles remain controversial. Several
studies have proved the activation of particular miRNA types in PBMCs
under cardiovascular events H. Li et al. (2018). For instance, there is
evidence that elevation of miR-186 suppresses the expression of
cystathionine-\(\gamma\)-lyase, leading to the subsequent secretion of
pro-inflammatory cytokines and cellular lipid accumulation. In addition,
macrophage-derived miR-186 may promote atherosclerotic plaque formation
(Yao et al. 2016). In line with this claim, we found that miR-186 was
up-regulated in both CAD and MI candidates compared to their control
counterparts. Surprisingly, the obtained data indicated that the
expression of miR-186 was higher in patients with CAD than in patients
with MI (Figure @ref(fig:Expall)). Specifically, miR-186 was the only
differentially expressed miRNA between CAD and MI, with a clear
up-regulation in CAD, indicating its main role in the promotion of
atherosclerosis.

As mentioned before, miR-21 was also up-regulated in both MI and CAD
patients compared to healthy controls. Moreover, the expression value of
miR-21 was significantly higher in the MI group than in the CAD group
(Table @ref(tab:mirExptable)). It is thought that the up-regulation of
miR-21 in PBMCs is a compensatory reaction to reduce the T\(_{reg}\)
lymphocyte number in response to the reduction of TGF\(\beta1\)
secretion into the plasma through a TGF\(\beta1\)/smad-independent
pathway. In line with the previous and present data, miR-21 can modulate
the activity of PBMCs following the occurrence of cardiovascular
diseases (S. Li et al. 2015).

Recent data have supported the elevation of miR-32 levels in CAD samples
with calcification of the coronary artery. Notably, miR-32 promotes
vascular smooth muscle calcification in mice by controlling the activity
of several proteins, including bone morphogenetic protein-1,
runt-related transcription factor-2 (RUNX2), osteopontin, and
bone-specific phosphoprotein matrix GLA protein. (Liu et al. 2017).
Likewise, some reports are associated with the activity of miR-32 in
PBMCs in several pathologies (Zeng et al. 2021; Wang et al. 2020). The
exact role of miR-32 in PBMCs after cardiovascular events remains
unclear.

Molecular analyses have indicated the regulatory role of miRNAs selected
using the AUC-ROC approach in PBMCs after a cardiovascular event. The
biological importance of two common miRNAs in the DEGs and AUC-ROC
approaches, miR-21 and miR-186, have already been discussed. Based on
numerous reports, miR-29a can be activated in different diseases
(Horita, Farquharson, and Stephen 2021). Data analysis indicated that
miR-29a was significantly up-regulated in CAD patients compared to the
healthy and MI groups (Table @ref(tab:mirExptable)). Increased miR-29a
is associated with the progression of atherosclerosis, and the
combination of miR-29a and ox-LDL has been suggested as a valid
biomarker set for paraclinical classification (Huang et al. 2016).
However, the role of miR-29a in the function of PBMCs from patients with
CAD has not been thoroughly examined.

The data indicated that miR-197 was significantly up-regulated in both
the CAD/healthy and MI/healthy groups. Previous studies have
demonstrated that miR-197 may play a crucial role in controlling the
anti-inflammatory response of IL-35 by influencing the secretion of
cytokines that can either promote or suppress inflammation, the ratio of
M1/M2 macrophages, and the proliferation of T\(_{reg}\) lymphocytes,
which are responsible for suppressing immune responses (Bhansali et al.
2022). Alongside our findings, it can be concluded that miR-197 could be
a useful diagnostic tool for predicting adverse cardiovascular events.

The findings of this study demonstrate the potential of miR-296 as a
biomarker with high discriminatory power for to distinguish between
samples from individuals with MI and CAD. MiR-296 has been identified as
a key regulator in the development and advancement of atherosclerosis by
controlling the expression of target genes associated with various
biological processes, including angiogenesis, cholesterol metabolism,
inflammation, cellular proliferation, hypertension, and apoptosis (H. Li
et al. 2018). In a previous study, miR-296 expression levels were found
to be significantly increased in the PBMCs of CAD patients compared to
healthy controls, suggesting its involvement in regulating
pro-inflammatory cytokines such as IL-6 and TNF-\(\alpha\) (Fard et al.
2020). These findings suggested that miR-296 may have a significant
impact on the pathogenesis of atherosclerosis and could potentially
serve as a diagnostic biomarker for CAD or MI.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In summary, we derived a set of miRNA biomarkers by comparing MI samples
with both healthy and CAD samples. We found that the SVM model performed
best in both the first layer, which separated healthy and not-healthy
samples, and the second layer, which classified the MI/CAD samples. The
set of miRNAs selected based on their AUC-ROC values performed better in
the second layer. Overall, our two-layer structure achieved an accuracy
of 0.96. This demonstrates the potential of combining bioinformatics and
machine learning techniques to identify novel biomarkers and gain a
deeper understanding of myocardial infarction.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-197}{}}%
Bhansali, Shipra, Amit Kumar Yadav, Chetan Bakshi, and Veena Dhawan.
2022. {``Interleukin-35 {Mitigates} Ox-{LDL}-{Induced} {Proatherogenic}
{Effects} via {Modulating} {miRNAs} {Associated} with {Coronary}
{Artery} {Disease} ({CAD}).''} \emph{Cardiovascular Drugs and Therapy},
April. \url{https://doi.org/10.1007/s10557-022-07335-x}.

\leavevmode\vadjust pre{\hypertarget{ref-75}{}}%
Canali, Raffaella, Lucia Natarelli, Guido Leoni, Elena Azzini, Raffaella
Comitato, Oezgur Sancak, Luca Barella, and Fabio Virgili. 2014.
{``Vitamin {C} Supplementation Modulates Gene Expression in Peripheral
Blood Mononuclear Cells Specifically Upon an Inflammatory Stimulus: A
Pilot Study in Healthy Subjects.''} \emph{Genes \& Nutrition} 9 (3):
390. \url{https://doi.org/10.1007/s12263-014-0390-x}.

\leavevmode\vadjust pre{\hypertarget{ref-CVD}{}}%
{``Cardiovascular Diseases ({CVDs}).''} n.d. Accessed March 12, 2023.
\url{https://www.who.int/news-room/fact-sheets/detail/cardiovascular-diseases-(cvds)}.

\leavevmode\vadjust pre{\hypertarget{ref-296a}{}}%
Fard, Toktam Kazemi, Samareh Tavakoli, Reza Ahmadi, Nariman Moradi, Reza
Fadaei, Asghar Mohammadi, and Soudabeh Fallah. 2020. {``Evaluation of
{Ip10} and {miRNA} 296-a {Expression} {Levels} in {Peripheral} {Blood}
{Mononuclear} {Cell} of {Coronary} {Artery} {Disease} {Patients} and
{Controls}.''} \emph{DNA and Cell Biology} 39 (9): 1678--84.
\url{https://doi.org/10.1089/dna.2020.5650}.

\leavevmode\vadjust pre{\hypertarget{ref-PBMC-miR}{}}%
Gao, Jie, Jia Liu, Ying Zhang, BaoYi Guan, Hua Qu, Hua Chai, WenTing
Wang, XiaoJuan Ma, and DaZhuo Shi. 2020. {``{PBMCs}-{Derived} {microRNA}
{Signature} as a {Prethrombotic} {Status} {Discriminator} in {Stable}
{Coronary} {Artery} {Disease}.''} \emph{Thrombosis and Haemostasis} 120
(01): 121--31. \url{https://doi.org/10.1055/s-0039-1700518}.

\leavevmode\vadjust pre{\hypertarget{ref-Meder23}{}}%
Hapke, Nils, Margarete Heinrichs, DiyaaElDin Ashour, Elena Vogel, Ulrich
Hofmann, Stefan Frantz, and Gustavo Campos Ramos. 2022.
{``Identification of a Novel Cardiac Epitope Triggering {T}-Cell
Responses in Patients with Myocardial Infarction.''} \emph{Journal of
Molecular and Cellular Cardiology} 173 (December): 25--29.
\url{https://doi.org/10.1016/j.yjmcc.2022.09.001}.

\leavevmode\vadjust pre{\hypertarget{ref-numpy}{}}%
Harris, Charles R., K. Jarrod Millman, Stéfan J. van der Walt, Ralf
Gommers, Pauli Virtanen, David Cournapeau, Eric Wieser, et al. 2020.
{``Array Programming with {NumPy}.''} \emph{Nature} 585 (7825): 357--62.
\url{https://doi.org/10.1038/s41586-020-2649-2}.

\leavevmode\vadjust pre{\hypertarget{ref-scikitopt}{}}%
Head, Tim, Manoj Kumar, Holger Nahrstaedt, Gilles Louppe, and Iaroslav
Shcherbatyi. 2021. \emph{Scikit-Optimize/Scikit-Optimize} (version
v0.9.0). Zenodo. \url{https://doi.org/10.5281/zenodo.5565057}.

\leavevmode\vadjust pre{\hypertarget{ref-29aa}{}}%
Horita, Masahiro, Colin Farquharson, and Louise A Stephen. 2021. {``The
Role of {miR}‐29 Family in Disease.''} \emph{Journal of Cellular
Biochemistry} 122 (7): 696--715.
\url{https://doi.org/10.1002/jcb.29896}.

\leavevmode\vadjust pre{\hypertarget{ref-29a}{}}%
Huang, Yu-Qing, An-Ping Cai, Ji-Yan Chen, Cheng Huang, Jie Li, and
Ying-Qing Feng. 2016. {``The {Relationship} of {Plasma} {miR}-29a and
{Oxidized} {Low} {Density} {Lipoprotein} with {Atherosclerosis}.''}
\emph{Cellular Physiology and Biochemistry} 40 (6): 1521--28.
\url{https://doi.org/10.1159/000453202}.

\leavevmode\vadjust pre{\hypertarget{ref-m2}{}}%
Kalayinia, Samira, Fateme Arjmand, Majid Maleki, Mahshid Malakootian,
and Chandra Pal Singh. 2021. {``{MicroRNAs}: Roles in Cardiovascular
Development and Disease.''} \emph{Cardiovascular Pathology} 50
(January): 107296. \url{https://doi.org/10.1016/j.carpath.2020.107296}.

\leavevmode\vadjust pre{\hypertarget{ref-11-15}{}}%
Kaur, Amanpreet, Sharon T Mackin, Kenny Schlosser, Fui Lin Wong, Malik
Elharram, Christian Delles, Duncan J Stewart, Natalie Dayan, Tara
Landry, and Louise Pilote. 2020. {``Systematic Review of {microRNA}
Biomarkers in Acute Coronary Syndrome and Stable Coronary Artery
Disease.''} \emph{Cardiovascular Research} 116 (6): 1113--24.
\url{https://doi.org/10.1093/cvr/cvz302}.

\leavevmode\vadjust pre{\hypertarget{ref-24}{}}%
Khan, Abrar A., Vinayak Gupta, and Nitish R. Mahapatra. 2022. {``Key
Regulatory {miRNAs} in Lipid Homeostasis: {Implications} for
Cardiometabolic Diseases and Development of Novel Therapeutics.''}
\emph{Drug Discovery Today} 27 (8): 2170--80.
\url{https://doi.org/10.1016/j.drudis.2022.05.003}.

\leavevmode\vadjust pre{\hypertarget{ref-23}{}}%
Laggerbauer, Bernhard, and Stefan Engelhardt. 2022. {``{MicroRNAs} as
Therapeutic Targets in Cardiovascular Disease.''} \emph{Journal of
Clinical Investigation} 132 (11): e159179.
\url{https://doi.org/10.1172/JCI159179}.

\leavevmode\vadjust pre{\hypertarget{ref-BER}{}}%
Lazar, C., S. Meganck, J. Taminau, D. Steenhoff, A. Coletta, C. Molter,
D. Y. Weiss-Solis, R. Duque, H. Bersini, and A. Nowe. 2013. {``Batch
Effect Removal Methods for Microarray Gene Expression Data Integration:
A Survey.''} \emph{Briefings in Bioinformatics} 14 (4): 469--90.
\url{https://doi.org/10.1093/bib/bbs037}.

\leavevmode\vadjust pre{\hypertarget{ref-296}{}}%
Li, Heng, Heng Li, Xin-Ping Ouyang, Ting Jiang, Xi-Long Zheng, Xi-Long
Zheng, Ping-Ping He, and Guo-Jun Zhao. 2018. {``{MicroRNA}-296: A
Promising Target in the Pathogenesis of Atherosclerosis?''}
\emph{Molecular Medicine} 24 (1): 12--12.
\url{https://doi.org/10.1186/s10020-018-0012-y}.

\leavevmode\vadjust pre{\hypertarget{ref-21}{}}%
Li, Sihui, Qian Fan, Shaolin He, Tingting Tang, Yuhua Liao, and
Jiangjiao Xie. 2015. {``{MicroRNA}-21 {Negatively} {Regulates} {Treg}
{Cells} {Through} a {TGF}-\(\beta1\)/{Smad}-{Independent} {Pathway} in
{Patients} with {Coronary} {Heart} {Disease}.''} \emph{Cellular
Physiology and Biochemistry} 37 (3): 866--78.
\url{https://doi.org/10.1159/000430214}.

\leavevmode\vadjust pre{\hypertarget{ref-32-2}{}}%
Liu, Jianghua, Xinhua Xiao, Yingying Shen, Ling Chen, Canxin Xu, Heng
Zhao, Ying Wu, et al. 2017. {``{MicroRNA}-32 Promotes Calcification in
Vascular Smooth Muscle Cells: {Implications} as a Novel Marker for
Coronary Artery Calcification.''} Edited by Yin Tintut. \emph{PLOS ONE}
12 (3): e0174138. \url{https://doi.org/10.1371/journal.pone.0174138}.

\leavevmode\vadjust pre{\hypertarget{ref-67}{}}%
Maciejak, Agata, Marek Kiliszek, Marcin Michalak, Dorota Tulacz,
Grzegorz Opolski, Krzysztof Matlak, Slawomir Dobrzycki, Agnieszka
Segiet, Monika Gora, and Beata Burzynska. 2015. {``Gene Expression
Profiling Reveals Potential Prognostic Biomarkers Associated with the
Progression of Heart Failure.''} \emph{Genome Medicine} 7 (1): 26.
\url{https://doi.org/10.1186/s13073-015-0149-z}.

\leavevmode\vadjust pre{\hypertarget{ref-09}{}}%
Matone, Alice, Colm M. O'Grada, Eugene T. Dillon, Ciara Morris, Miriam
F. Ryan, Marianne Walsh, Eileen R. Gibney, et al. 2015. {``Body Mass
Index Mediates Inflammatory Response to Acute Dietary Challenges.''}
\emph{Molecular Nutrition \& Food Research} 59 (11): 2279--92.
\url{https://doi.org/10.1002/mnfr.201500184}.

\leavevmode\vadjust pre{\hypertarget{ref-fRMA}{}}%
McCall, M. N., B. M. Bolstad, and R. A. Irizarry. 2010. {``Frozen Robust
Multiarray Analysis ({fRMA}).''} \emph{Biostatistics} 11 (2): 242--53.
\url{https://doi.org/10.1093/biostatistics/kxp059}.

\leavevmode\vadjust pre{\hypertarget{ref-Barcode}{}}%
McCall, Matthew N., Karan Uppal, Harris A. Jaffee, Michael J. Zilliox,
and Rafael A. Irizarry. 2011. {``The {Gene} {Expression} {Barcode}:
Leveraging Public Data Repositories to Begin Cataloging the Human and
Murine Transcriptomes.''} \emph{Nucleic Acids Research} 39 (suppl\_1):
D1011--15. \url{https://doi.org/10.1093/nar/gkq1259}.

\leavevmode\vadjust pre{\hypertarget{ref-pandas}{}}%
McKinney, Wes. 2010. {``{D}ata {S}tructures for {S}tatistical
{C}omputing in {P}ython.''} In \emph{{P}roceedings of the 9th {P}ython
in {S}cience {C}onference}, edited by Stéfan van der Walt and Jarrod
Millman, 56--61.
\href{https://doi.org/\%2010.25080/Majora-92bf1922-00a\%20}{https://doi.org/
10.25080/Majora-92bf1922-00a }.

\leavevmode\vadjust pre{\hypertarget{ref-Meder6}{}}%
Mosallaei, Meysam, Naeim Ehtesham, Shima Rahimirad, Mostafa Saghi, Nasim
Vatandoost, and Sharifeh Khosravi. 2022. {``{PBMCs}: A New Source of
Diagnostic and Prognostic Biomarkers.''} \emph{Archives of Physiology
and Biochemistry} 128 (4): 1081--87.
\url{https://doi.org/10.1080/13813455.2020.1752257}.

\leavevmode\vadjust pre{\hypertarget{ref-scikit-learn}{}}%
Pedregosa, F., G. Varoquaux, A. Gramfort, V. Michel, B. Thirion, O.
Grisel, M. Blondel, et al. 2011. {``Scikit-Learn: Machine Learning in
{P}ython.''} \emph{Journal of Machine Learning Research} 12: 2825--30.

\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
R Core Team. 2022. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-ML}{}}%
Reel, Parminder S., Smarti Reel, Ewan Pearson, Emanuele Trucco, and
Emily Jefferson. 2021. {``Using Machine Learning Approaches for
Multi-Omics Data Analysis: {A} Review.''} \emph{Biotechnology Advances}
49 (July): 107739.
\url{https://doi.org/10.1016/j.biotechadv.2021.107739}.

\leavevmode\vadjust pre{\hypertarget{ref-RStudio}{}}%
RStudio Team. 2020. \emph{RStudio: Integrated Development Environment
for r}. Boston, MA: RStudio, PBC. \url{http://www.rstudio.com/}.

\leavevmode\vadjust pre{\hypertarget{ref-miR}{}}%
Schulte, Christian, Temo Barwari, Abhishek Joshi, Tanja Zeller, and
Manuel Mayr. 2020. {``Noncoding {RNAs} Versus {Protein} {Biomarkers} in
{Cardiovascular} {Disease}.''} \emph{Trends in Molecular Medicine} 26
(6): 583--96. \url{https://doi.org/10.1016/j.molmed.2020.02.001}.

\leavevmode\vadjust pre{\hypertarget{ref-m1}{}}%
Schulte, Christian, Mahir Karakas, and Tanja Zeller. 2017.
{``{microRNAs} in Cardiovascular Disease -- Clinical Application.''}
\emph{Clinical Chemistry and Laboratory Medicine (CCLM)} 55 (5).
\url{https://doi.org/10.1515/cclm-2016-0576}.

\leavevmode\vadjust pre{\hypertarget{ref-36}{}}%
Soler-Botija, Carolina, Carolina Gálvez-Montón, and Antoni Bayés-Genís.
2019. {``Epigenetic {Biomarkers} in {Cardiovascular} {Diseases}.''}
\emph{Frontiers in Genetics} 10 (October): 950.
\url{https://doi.org/10.3389/fgene.2019.00950}.

\leavevmode\vadjust pre{\hypertarget{ref-17}{}}%
Thygesen, Kristian, Joseph S. Alpert, Allan S. Jaffe, Bernard R.
Chaitman, Jeroen J. Bax, David A. Morrow, Harvey D. White, and The
Executive Group on behalf of the Joint European Society of Cardiology
(ESC)/American College of Cardiology (ACC)/American Heart Association
(AHA)/World Heart Federation (WHF) Task Force for the Universal
Definition of Myocardial Infarction. 2018. {``Fourth {Universal}
{Definition} of {Myocardial} {Infarction} (2018).''} \emph{Circulation}
138 (20). \url{https://doi.org/10.1161/CIR.0000000000000617}.

\leavevmode\vadjust pre{\hypertarget{ref-ML-Bio}{}}%
Torun, Furkan M., Sebastian Virreira Winter, Sophia Doll, Felix M.
Riese, Artem Vorobyev, Johannes B. Mueller-Reif, Philipp E. Geyer, and
Maximilian T. Strauss. 2023. {``Transparent {Exploration} of {Machine}
{Learning} for {Biomarker} {Discovery} from {Proteomics} and {Omics}
{Data}.''} \emph{Journal of Proteome Research} 22 (2): 359--67.
\url{https://doi.org/10.1021/acs.jproteome.2c00473}.

\leavevmode\vadjust pre{\hypertarget{ref-18}{}}%
Tsao, Connie W., Aaron W. Aday, Zaid I. Almarzooq, Alvaro Alonso, Andrea
Z. Beaton, Marcio S. Bittencourt, Amelia K. Boehme, et al. 2022.
{``Heart {Disease} and {Stroke} {Statistics}---2022 {Update}: {A}
{Report} {From} the {American} {Heart} {Association}.''}
\emph{Circulation} 145 (8).
\url{https://doi.org/10.1161/CIR.0000000000001052}.

\leavevmode\vadjust pre{\hypertarget{ref-32}{}}%
Wang, Dan, Ting Zeng, Zhi Lin, Lu Yan, Fenglin Wang, Lanlan Tang, Leyuan
Wang, Daolin Tang, Pan Chen, and Minghua Yang. 2020. {``Long Non-Coding
{RNA} {Snhg5} Regulates Chemotherapy Resistance Through the
{miR}-32/{Dnajb9} Axis in Acute Myeloid Leukemia.''} \emph{Biomedicine
\& Pharmacotherapy} 123 (March): 109802.
\url{https://doi.org/10.1016/j.biopha.2019.109802}.

\leavevmode\vadjust pre{\hypertarget{ref-ggplot}{}}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\leavevmode\vadjust pre{\hypertarget{ref-186}{}}%
Yao, Yan, Xin Zhang, Hai-peng Chen, Liang Li, Wei Xie, Gang Lan,
Zhen-wang Zhao, Xi-Long Zheng, Zong-bao Wang, and Chao-ke Tang. 2016.
{``{MicroRNA}-186 Promotes Macrophage Lipid Accumulation and Secretion
of Pro-Inflammatory Cytokines by Targeting Cystathionine
\(\gamma\)-Lyase in {THP}-1 Macrophages.''} \emph{Atherosclerosis} 250
(July): 122--32.
\url{https://doi.org/10.1016/j.atherosclerosis.2016.04.030}.

\leavevmode\vadjust pre{\hypertarget{ref-Macro}{}}%
Yap, Jonathan, Jason Irei, Javier Lozano-Gerona, Selena Vanapruks,
Tianmai Bishop, and William A. Boisvert. 2023. {``Macrophages in Cardiac
Remodelling After Myocardial Infarction.''} \emph{Nature Reviews
Cardiology}, January. \url{https://doi.org/10.1038/s41569-022-00823-5}.

\leavevmode\vadjust pre{\hypertarget{ref-cluster}{}}%
Yu, Guangchuang, Li-Gen Wang, Yanyan Han, and Qing-Yu He. 2012.
{``clusterProfiler: An r Package for Comparing Biological Themes Among
Gene Clusters.''} \emph{OMICS: A Journal of Integrative Biology} 16 (5):
284--87. \url{https://doi.org/10.1089/omi.2011.0118}.

\leavevmode\vadjust pre{\hypertarget{ref-magic}{}}%
Zeng, Zl, Qingyun Zhu, Zhibo Zhao, Xuyu Zu, and Jianghua Liu. 2021.
{``Magic and Mystery of {microRNA}‐32.''} \emph{Journal of Cellular and
Molecular Medicine} 25 (18): 8588--8601.
\url{https://doi.org/10.1111/jcmm.16861}.

\end{CSLReferences}

\end{document}

### 2023-08-07 Complete-V1/Complete-V1.Rmd
The version after revision

@@
 ---
-title: Machine Learning Deriven Set of microRNAs as a Novel Biomarker for Myocardial Infarction Diagnosis
+title: "Unlocking the Potential of microRNAs: Machine Learning Identifies Key Biomarkers for Myocardial Infarction Diagnosis"
 #titlerunning: Short form of title (if too long for head)
@@
   address: Medical Nanotechnology gruop, Tabriz University of Medical Sciences, Tabriz, Iran
-  email: samadishadlou@tbzmed.ac.ir
 
@@
   address: Medical Nanotechnology gruop, Tabriz University of Medical Sciences, Tabriz, Iran
-  email: rahbarghazir@tbzmed.ac.ir
+
+- name: Zeynab Piryaei 
+  address: Laboratory of Complex Biological Systems and Bioinformatics (CBB), Department of Bioinformatics, Institute of Biochemistry and Biophysics (IBB), University of Tehran, Iran.
+
+- name: Mahdad Esmaeili
+  address: Medical Bioengineering Department, School of Advanced Medical Sciences, Tabriz University of Medical Sciences, Tabriz, Iran
+
+- name: Çığır Biray Avcı
+  address: Medical Biology Department, School of Medicine, Ege University, İzmir, Turkey
 
@@
   email: bainf@tbzmed.ac.ir
+
+- name: Kaveh Kavousi
+  address: Laboratory of Complex Biological Systems and Bioinformatics (CBB), Department of Bioinformatics, Institute of Biochemistry and Biophysics (IBB), University of Tehran, Iran
+  email: kkavousi@ut.ac.ir
 
@@
 - Bioinformatics
+- Biomarker
 
@@
 abstract: |
-  MicroRNAs (miRNAs) play a crucial role in regulating adaptive and maladaptive responses in cardiovascular diseases, making them attractive targets for potential therapeutics.
+  MicroRNAs (miRNAs) play a crucial role in regulating adaptive and maladaptive responses in cardiovascular diseases, making them attractive targets for potential biomarkers.
 However, their potential as novel biomarkers for diagnosing cardiovascular diseases requires systematic evaluation.
 In this study, we aimed to identify a key set of miRNA biomarkers using integrated bioinformatics and machine learning analysis.
-We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contains peripheral blood mononuclear cell samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy individuals.
-Additionally, we selected a set of miRNAs based on their area under the curve of the receiver operating characteristic curve (AUC-ROC) to separate the CAD/MI samples.
-We designed a two-layer architecture for sample classification, with the first layer isolating healthy samples from not-healthy ones, and the second layer classifying stable CAD and MI samples.
+We combined and analyzed three gene expression datasets from the Gene Expression Omnibus (GEO) database, which contains peripheral blood mononuclear cell (PBMC) samples from individuals with myocardial infarction (MI), stable coronary artery disease (CAD), and healthy individuals.
+Additionally, we selected a set of miRNAs based on their area under the receiver operating characteristic curve (AUC-ROC) for separating the CAD and MI samples.
+We designed a two-layer architecture for sample classification, in which the first layer isolates healthy samples from unhealthy samples, and the second layer classifies stable CAD and MI samples.
 We trained different machine learning models using both biomarker sets and evaluated their performance on a test set.
@@
 Both biomarker sets could distinguish healthy from not-healthy samples with complete accuracy.
-The best performance for the classification of CAD and MI was achieved with an SVM model trained using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.95, and an accuracy of 0.88.
-Our study demonstrated that miRNA signatures derived from peripheral blood could serve as valuable novel biomarkers for cardiovascular diseases.
+The best performance for the classification of CAD and MI was achieved with an SVM model trained using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.96 and an accuracy of 0.94 on the test data.
+Our study demonstrated that miRNA signatures derived from PBMCs could serve as valuable novel biomarkers for cardiovascular diseases.
 
@@
 
-bibliography: [ForArticleRefrence.bib, discussion.bib, Introduction.bib, mir296.bib]
-biblio-style: spmpsci
+bibliography: [For Article Refrence.bib, discussion.bib, Introduction.bib, mir296.bib]
+biblio-style: spbasic
 # bibstyle options spbasic(default), spphys, spmpsci
@@
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load(knitr, captioner, bundesligR, stringr, kableExtra)
+pacman::p_load(knitr, captioner, bundesligR, stringr, kableExtra, ggplot2, ggvenn)
 ```
@@
 
-At present, cardiovascular diseases (CVDs) are the leading cause of human mortality with 32% of all global deaths.
-It is estimated that about 85% of CVDs mortality were diagnosed with myocardial infarction (MI) [@CVD].
-MI is an acute coronary syndrome with sudden blockage and stenosis of the coronary artery, and subsequent myocardial ischemia, leading to extensive cardiomyocyte damage and necrosis [@Macro].
+Cardiovascular diseases (CVDs) are the leading cause of human mortality, accounting for 32% of all global deaths.
+It is estimated that approximately 85% of CVD mortality is due to myocardial infarction (MI) [@CVD].
+MI is an acute coronary syndrome characterized by sudden blockage and stenosis of the coronary artery and subsequent myocardial ischemia, leading to extensive cardiomyocyte damage and necrosis [@Macro].
 
-Over the last 50 years, numerous attempts have been collected to use biomarkers to facilitate diagnosis, assess risk, follow-up therapy, and determine therapeutic efficacy in CVDs candidates.
-Based on the released guidelines, cardiac troponins (cTns) are used as a highly-sensitive and accurate approach for the detection of MI.
-Despite the inherent advantages, the high-rate sensitivity of cTn-based assays has also led to more false positive results [@17], which do necessitate the advent and development of new modalities with pathological values.
-To improve diagnostic value upon existing MI biomarkers, the combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, is proposed.
-Previous researches support the notion that miRNAs exhibit the great potential to be used as alternative biomarkers in CVDs detection and follow-up [@miR].
-It is suggested that miRNAs possess 18-22 nucleotides and can play a crucial role in the regulation of gene expression.
-Evidence point to the fact that miRNAs are involved in the pathogenesis of cardiac tissue injury and could be considered as theranostics targets in terms of CVDs.
-These elements can easily circulate in biofluids with both diagnostic as well as prognostic values [@m1].
-Several biological activities such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque formation, and cardiac rhythm are regulated by miRNAs [@m2].
-It is postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in CVD patients.
-To be specific, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) [@36].
+Over the last 50 years, numerous attempts have been made to use biomarkers to facilitate diagnosis, assess the risk, follow-up therapy, and determine therapeutic efficacy in CVD candidates.
+Based on released guidelines, cardiac troponins (cTns) are used as a highly sensitive and accurate approach for detecting MI.
+Despite these inherent advantages, the high sensitivity of cTn-based assays has also led to more false-positive results [@17], necessitating the advent and development of new modalities with pathological value.
+To improve the diagnostic value of  existing MI biomarkers, a combination of complementary biological markers, such as microRNAs (miRNAs) and other genetic factors, has been proposed.
+Previous research supports the notion that miRNAs exhibit great potential as alternative biomarkers for CVD detection and follow-up [@miR].
+It has been suggested that miRNAs possess 18-22 nucleotides and play a crucial role in the regulation of gene expression.
+Evidence indicates that miRNAs are involved in the pathogenesis of cardiac tissue injury [@m1].
+Several biological processes, such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque formation, and cardiac rhythm, are regulated by miRNAs [@m2].
+These elements can easily circulate in biofluids and could be considered theranostic targets in terms of CVDs [@m1].
+It has been postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in patients with CVD.
+Specifically, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) [@36].
 
-PBMCs are a fraction of white blood cells (WBCs), including monocytes, lymphocytes, macrophages, and other cells belonging to the immune system [@PBMC-miR].
-Emerging data have indicated that PBMCs can be used as a valid source of biomarkers for monitoring various pathological conditions.
-Of note, the alteration of mRNAs and miRNAs under pathological conditions gives us valuable information about different kinds of disorders.
-PBMCs could recapitulate the conditions of the target tissues, thus, providing a highly sensitive and specific source of biomarkers [@Meder6].
-Commensurate with these conditions, these cells are repositories of dysregulated genes and miRNAs expression profiles in CVDs [@PBMC-miR; @Meder6].
+PBMCs are a fraction of white blood cells, including monocytes, lymphocytes, macrophages, and other cells of the immune system [@PBMC-miR].
+Emerging data indicate that PBMCs can be used as a valid source of biomarkers for monitoring various pathological conditions.
+Of note, the alteration of mRNAs and miRNAs under pathological conditions provides valuable information about different kinds of disorders.
+PBMCs can recapitulate the conditions of target tissues, thus providing a highly sensitive and specific source of biomarkers [@Meder6].
+Combined with these conditions, these cells are repositories of dysregulated genes and miRNA expression profiles in CVDs [@PBMC-miR; @Meder6].
 
-In recent years, the advent and use of machine learning (ML) is an exciting prospect for advancing scientific discoveries.
-Although the concept of ML and its initial algorithms were conceived many years ago, recent improvements in computing power and access to vast amounts of data have shown that ML techniques outperform classical statistical methods in various fields.
-Furthermore, the progress made in omics technologies has enabled the analysis of massive and intricate biological data sets, consisting of hundreds to thousands of samples, which makes it possible for ML to extract valuable biological information from such data [@ML-Bio].
-Therefore, ML offers novel techniques to integrate and analyze the various omics data enabling the discovery of de novo biomarkers.
-These biomarkers help us in accurate disease prediction, patient stratification, and finding new therapeutics [@ML].
+In recent years, the advent and application of machine learning (ML) has been an exciting prospect for advancing scientific research.
+Although the concept of ML and its initial algorithms were conceived many years ago, recent improvements in computing power and access to vast amounts of data have demonstrated that ML techniques outperform classical statistical methods in various fields.
+Furthermore, the progress made in omics technologies has enabled the analysis of massive and intricate biological datasets, consisting of hundreds to thousands of samples, which makes it possible for ML to extract valuable biological insights and information from such data [@ML-Bio].
+Consequently, ML provides innovative methods for merging and interpreting diverse types of omics data, leading to the identification of new biomarkers.
+These biomarkers can aid in precise disease prediction, patient stratification, and the development of novel therapeutic approaches [@ML].
 
-In this study, we aimed to identify potential miRNA biomarkers for MI patients by combining and analyzing three different microarray datasets from PBMCs.
-It is suggested that the integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring MI.
-Besides, this approach can deepen our vision into the underlying mechanisms of MI and aid in the development of valid diagnostic biomarkers, and patient stratification.
+In this study, we aimed to identify potential miRNA biomarkers in patients with MI by combining and analyzing three different microarray datasets from PBMCs.
+The integration of omics data with bioinformatics and ML techniques could be a promising tool in the discovery of new and more accurate biomarkers for monitoring MI.
+Additionally, this approach can deepen our understanding of the underlying mechanisms of MI and aid in the development of valid diagnostic biomarkers and patient stratification.
 
@@
 Microarray datasets were obtained from the Gene Expression Omnibus (GEO) database (<https://www.ncbi.nlm.nih.gov/geo/>).
-To obtain sufficient classification power between MI, healthy and stable coronary artery disease (CAD) samples, a relatively large sample size was required.
-Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475 for healthy samples were selected.
-All samples were produced using Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
-Only healthy, stable CAD and early-stage MI samples were selected from these datasets for further analyses.
-The basic information for the three datasets evaluated in the current study is provided in Table \@ref(tab:datasets).
+To obtain sufficient classification power between MI, healthy, and CAD samples, a relatively large sample size is needed Therefore, GSE59867 for the MI and CAD samples and GSE56609 and GSE54475 for the healthy samples were selected.
+All samples were produced using Affymetrix Human Gene 1.0 ST Array platform (GPL6244).
+Only healthy, CAD, and early-stage MI samples were selected from these datasets for further analysis.
+The basic information for the three datasets evaluated in this study is provided in Table \@ref(tab:datasets).
+Bioinformatics analyses were conducted using R, ver.
+4.2.0 [@R], RStudio [@RStudio].
+All plots and graphics of these sections were created using the ggplot2 R package [@ggplot].
 
@@
 
-kbl(df, format = "markdown", booktabs = T, align = "c", linesep = c(""), caption = "Basic information of the GEO microarray datasets.")
+kbl(df, format = "markdown", booktabs = T, align = "c", linesep = c(""), caption = "Sample information on the GEO microarray datasets.")
 ```
 
-## Pre-processing
+## Preprocessing
 
-Raw data (CEL files) of all datasets were downloaded from the GEO and pre-processed using the fRMA package [@fRMA].
-fRMA allowed to pre-process of individual microarray samples and combining them consistently for analysis.
-For each dataset, background correction was performed using the RMA algorithm and then it was quantile normalized based on the reference distribution.
-During summarization, batch effects were removed and variances of the gene expressions were estimated by taking into account these probe-specific effects.
-For those multiple probe sets matched to the identical gene, the mean log fold change was retained.
-Therefore, fRMA can be seen as a batch effect removal technique for different datasets that are produced by identical microarray platforms.
-Thus, to ensure batch effect removal, the principal component analysis and the relative log expression of train samples were plotted before and after fRMA [@BER].
+The raw data in the form of CEL files from all datasets were obtained from GEO.
+To prepare the data for analysis, we utilized the fRMA package [@fRMA] to facilitate preprocessing of individual microarray samples and their consistent combination.
+For each dataset, background correction was applied using the RMA algorithm, followed by quantile normalization based on the reference distribution.
+To account for probe-specific effects, batch effects were eliminated during summarization and gene expression variances were estimated accordingly.
+In cases where multiple probe sets matched the same gene, the mean log-fold change was retained.
+Consequently, fRMA can serve as a technique to remove batch effects across diverse datasets generated by identical microarray platforms [@BER].
+To ensure the effectiveness of the batch effect removal, we employed principal component analysis (PCA) and relative log expression (RLE) plots to visualize the data before and after applying fRMA.
 
@@
 
-The barcode algorithm proposed by McCall et al.
-[@Barcode] transformed the actual expression values into binary barcode values.
-Huge sets of samples were collected and normalized using fRMA for several platforms as well as for Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
-The distribution of the expressed and unexpressed observed intensities for each gene is estimated using these normalized sets.
-Genes were considered expressed (and their value coded to 1) or unexpressed (and their value coded to 0) according to the following equation:
+The barcode algorithm was introduced by McCall et al.
+[@Barcode], aimed to convert actual expression values into binary barcode values.
+Extensive sample collections were gathered and normalization was performed using fRMA across multiple platforms, including the Affymetrix Human Gene 1.0 ST Array (GPL6244) platform.
+By utilizing these normalized datasets, the distribution of the observed intensities for both the expressed and unexpressed genes was estimated.
+The determination of whether a gene was expressed or not was based on the following equation, where a value of 1 indicates expression and a value of 0 indicates non-expression:
 
@@
 
-where $x_{ij}$ is the normalized intensity of gene $i$ in sample $j$, $C$ is a user-defined parameter, $\sigma^{ne}$ is the standard deviation of the non-expressed distribution, and $\mu^{ne}$ is the mean of the non-expressed distribution.
-The barcode representation of a sample is a vector of ones and zeros denoting which genes are estimated to be expressed (ones) and unexpressed (zeros).
-The barcode algorithm was implemented by the barcode function in the R fRMA package, and the default value of $C$ was used.
+In the barcode algorithm, the normalized intensity of gene $i$ in sample $j$ is denoted as $x_{ij}$.
+A user-defined parameter, $C$, was introduced along with the standard deviation ($\sigma^{ne}$) and mean ($\mu^{ne}$) of the non-expressed distribution.
+Based on these values, the barcode representation of a sample was generated as a vector consisting of ones and zeros, representing the estimated expression (ones) and non-expression (zeros) of each gene.
+The barcode function within the R fRMA package was employed to implement the barcode algorithm, utilizing the default value of $C$.
 
-To determine if the expressed ratios differed in the MI group versus the healthy control group, Fisher's exact test for individual genes was carried out upon the barcode values.
-Genes with a false discovery rate (FDR) of $< 0.05$, which was calculated through the Benjamini-Hochberg (BH) procedure to adjust for multiple testing issues, were considered as differentially expressed genes.
-The same procedures were conducted on CAD versus healthy controls as well as MI versus CAD group to find the DEGs between them.
+To assess the differences in expressed ratios between the MI and healthy control groups, Fisher's exact test was performed on the barcode values of individual genes.
+Genes that exhibited a false discovery rate (FDR) below $0.05$, calculated using the Benjamini-Hochberg procedure to account for multiple testing issues were identified as differentially expressed genes (DEGs).
+The same procedures were applied to the CAD versus healthy control comparison, as well as to the MI versus CAD group, to identify DEGs specific to each comparison.
 
@@
 
-Using the R clusterProfiler package, the Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation were carried out on the differentially expressed genes.
-The GO analysis included biological process (BP), cellular component (CC) and molecular function (MF) categories.
-An adjusted p-value of less than 0.05 was considered to indicate a statistically significant difference.
-Enrichments were conducted on the MI-healthy and CAD-healthy DEGs.
-In these analyses, all default parameters were used.
+The R clusterProfiler package [@cluster] was utilized to perform Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO) functional annotation on the set of DEGs.
+GO analysis encompassed three categories: biological process (BP), cellular component (CC), and molecular function (MF).
+For statistical significance, an adjusted p-value threshold of less than 0.05 was employed.
+Enrichment analyses were conducted separately for DEGs specific to the MI-healthy and CAD-healthy comparisons.
+All the default parameters provided by the package were used in the analyses.
 
@@
 
-The ML analysis was performed using Python software, ver.
-3.9, Numpy [@numpy], pandas [@pandas], and Scikit-Learn packages [@scikit-learn].
-Whenever hyper-tuning was needed, the scikit-opt package [@scikitopt] was used.
-In all ML analyses, the datasets were divided into train and test sets by a 0.7:0.3 ratio and all reported results are the average of 10-fold cross-validation.
+ML analysis was performed using Python software, ver.
+3.9, Numpy [@numpy], Pandas [@pandas], and Scikit-Learn packages [@scikit-learn].
+Whenever hypertuning was needed, the Scikit-opt package [@scikitopt] was used.
+In all ML analyses, the datasets were divided into training and test sets at a 0.7:0.3 ratio, and all reported results are the average of 10-fold cross-validation.
 
-Two different approaches were used for selecting miRNAs for model training.
-The first approach was using the miRNAs that are differentially expressed.
-In the second approach, miRNAs with individual area under the curve of the receiver operating characteristic (AUC-ROC) over 0.8 for separating MI from CAD were selected.
-Having the result of these two different approaches can provide an informative comparison between the predictive capabilities of sets of miRNAs selected with different logics.
+Two different approaches were used to select miRNAs for model training.
+The first approach was to use differentially expressed miRNAs.
+In the second approach, miRNAs with individual AUC-ROCs over 0.8 for separating MI from CAD were selected.
+The results of these two approaches can provide an informative comparison between the predictive capabilities of sets of miRNAs selected with different logics.
 
@@
 
-In this approach, a two-layer architecture was deployed to the data to maximize the prediction values.
-The first layer predicted whether a sample is healthy or not, and the second layer separated MI from CAD in the samples which were predicted as not healthy in the first layer.
+In this approach, a two-layer architecture is deployed to the data to maximize the prediction values.
+The first layer predicted whether a sample was healthy or not, and the second layer separated MI from CAD in the samples that were predicted as not healthy in the first layer.
 To this end, a distinct ML model was trained for each layer.
-Since there is a limited number of miRNAs in DEGs, both layers were trained with all of them.
-For further comparison with the models' performance, the ROC curve of each miRNA for classifying healthy and not-healthy, as well as CAD and MI, were generated using a Logistic Regression model.
+Because, there were a limited number of miRNAs in the DEGs, both layers were trained with all of them.
+For further comparison with the models' performance, the ROC curve of each miRNA for classifying healthy and not-healthy, as well as CAD and MI, was generated using a logistic regression model.
 
-#### First layer for isolation of healthy and not-healthy samples:
+#### First layer for the isolation of healthy and not-healthy samples:
 
-An SVM model using RBF kernels was trained and hyper-tuned using all miRNAs in DEGs.
-To handle the severe imbalance in the number of samples (51 for the healthy and 157 for the not-healthy groups), the sample weight for the healthy and the not-healthy samples were set to 1 and 0.5, respectively.
+A support vector machine (SVM) model using RBF kernels was trained and hypertuned using all miRNAs in the DEGs.
+To handle the severe imbalance in the number of samples (51 for the healthy group and 157 for the not-healthy group), the sample weights for the healthy and the not-healthy samples were set to 1 and 0.5, respectively.
+The ROC curve and confusion matrix for the model are reported.
 
-#### Second layer for seperating MI and CAD samples:
+#### Second layer for separating the MI and CAD samples:
 
-For the sake of reaching the highest classification performance using the set of miRNAs, different models were investigated.
-To do so, SVM (with linear, polynomial, and RBF kernels), Logistic Regression (LR), Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB), XGBoost (XGB) and Decision Tree (DT) models were trained.
-All models were trained with their pre-set parameters with 10-fold cross-validation.
-The criteria for choosing the best model were the highest accuracy and AUC-ROC on the test set.
-The best model was hyper-tuned with the scikit-opt package [@scikitopt] to get the best predictive performance.
+Different models were investigated to achieve the highest classification performance.
+To do so, SVM (with linear, polynomial, and RBF kernels), logistic regression (LR), random forests (RF), k-nearest neighbor (kNN), gradient boosting (GB), XGBoost (XGB) and decision tree (DT) models were trained.
+All models were trained with their preset parameters using 10-fold cross-validation.
+The criteria for selecting the best model were the highest accuracy and AUC-ROC for the test set.
+The best model was hypertuned using the scikit-opt package [@scikitopt] for the best classification performance.
+The ROC curve and confusion matrix for the best model are reported.
 
@@
 
-Like the previous approach, a two-layers strategy was conducted.
-The first layer classified samples into healthy and not-healthy, and the separated MI and CAD samples.
-However, to keep the number of miRNAs as low as possible miRNAs were selected from the second layer, and then their performance wase evaluated in the first layer.
-AUC-ROC of all miRNAs for classifying MI and CAD samples were calculated and the miRNAs with the AUC-ROC over 0.8 were selected.
-The ROC curves for each selected miRNA for separating healthy samples from not-healthy and MI from CAD samples were also plotted for further comparison.
+As in the previous approach, a two-layer strategy was employed.
+The first layer classified samples into healthy and not-healthy, and the second layer separated the MI and CAD samples.
+However, to keep the number of miRNAs as low as possible, miRNAs were selected from the second layer and their performance was evaluated in the first layer.
+The AUC-ROC of all miRNAs for classifying MI and CAD samples was calculated, and miRNAs with AUC-ROC $> 0.8$ were selected.
+ROC curves for each selected miRNA for separating healthy samples from not-healthy samples and MI from CAD samples were also plotted for further comparison.
 
-#### First layer for the detection of healthy and not-healthy samples:
+#### First layer for the isolation of healthy and not-healthy samples:
 
-An SVM model with an RBF kernel was trained using the selected set of miRNAs.
-Additionally, the model was hyper-tuned to find the hyper-parameters for the highest AUC-ROC and accuracy.
-The ROC curve and confusion matrix for the best model were reported.
+An SVM model with an RBF kernel is trained using the selected set of miRNAs.
+Additionally, the model was hypertuned to find the hyperparameters for the highest AUC-ROC and accuracy.
+The same sample weights as in the previous approach (1 for healthy and 0.5 for not-healthy samples) were used.
+The ROC curve and confusion matrix for the model were reported.
 
-#### Second layer for separating MI and CAD:
+#### Second layer for separating the MI and CAD samples:
 
-The selected miRNAs set was used to train different algorithms to find the best model.
-Similar to the previous approach, SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT were trained.
-All models were trained with their pre-set parameters using 10-fold cross-validation.
-The models with the highest AUC-ROC and accuracy on the test set were selected and hyper-tuned using the scikit-opt package [@scikitopt].
+The selected miRNA set was used to train different algorithms to determine the best model.
+Similar to the previous approach, the SVM (with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT models were trained.
+All models were trained with their preset parameters using 10-fold cross-validation.
+The models with the highest AUC-ROC and accuracy on the test set were selected and hypertuned using the scikit-opt package [@scikitopt].
 The ROC curve and confusion matrix for the best model were reported.
@@
 
-## Pre-processing
+## Preprocessing
 
-The PCA plots of the training samples are shown in Figure \@ref(fig:PCA)A and B.
-As shown, healthy samples were separated from CAD or MI samples in primary data and also after conducting fRMA.
-In the RLE plot, there was a distinct difference between dataset means for all samples before conducting fRMA (Figure \@ref(fig:PCA)C).
-All datasets were rearranged around 0 in the RLE plot after conducting fRMA (Figure \@ref(fig:PCA)D).
-Moreover, there was a clear change in inter-quantile distances, but the values still were over 0.1.
+The PCA plots of the samples are shown in Figures \@ref(fig:PCA)A and B.
+Healthy samples were separated from the CAD or MI samples in the primary data and after conducting fRMA.
+In the RLE plot, there was a distinct difference between the dataset means for all samples before fRMA was performed (Figure \@ref(fig:PCA)C).
+All datasets were rearranged to approximately 0 in the RLE plot after fRMA was conducted (Figure \@ref(fig:PCA)D).
+Moreover, there was an apparent change in the interquantile distances, but the values were still greater than 0.1.
 
-```{r PCA, echo=F, out.width="100%", fig.cap="PCA and RLE plot for all samples before and after fRMA."}
+```{r PCA, echo=F, out.width="100%", fig.cap="Principal component analysis plots for (A) primary data and (B) the data after fRMA, and the relative log expression plots for (C) primary data and (D) the data after fRMA."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/PCA and RLE of trains.pdf")
@@
 
-According to the cutoff criterion of $FDR < 0.05$, there were `r dim(M_H)[1]` DEGs between the MI and the healthy samples.
-Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in MI compared to the healthy controls.
-In CAD and healthy groups comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and and `r sum(C_H$logFC < 0)` DEGs were up- and down-regulated, respectively in CAD samples.
-In the MI and CAD groups, the number of DEGs was `r dim(M_C)[1]`, and the number of up- and down-regulated genes in MI samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively in comparison with CAD samples.
-These data are summarized in Table \@ref(tab:DEGstab).
+According to the cutoff criterion of $FDR < 0.05$, there were `r dim(M_H)[1]` DEGs between MI and healthy samples.
+Among them, `r sum(M_H$logFC > 0)` were up-regulated, and `r sum(M_H$logFC < 0)` were down-regulated in the MI group compared to the healthy group.
+In the CAD and healthy group comparison, we found `r dim(C_H)[1]` DEGs, of which `r sum(C_H$logFC > 0)` and `r sum(C_H$logFC < 0)` DEGs were up- and down-regulated, respectively, in CAD samples.
+In the MI and CAD groups, the number of DEGs was `r dim(M_C)[1]`, and the numbers of up- and down-regulated genes in MI samples were `r sum(M_C$logFC > 0)` and `r sum(M_C$logFC < 0)`, respectively, compared to CAD samples.
+The data are summarized in Table \@ref(tab:DEGstab).
 
-```{r venn, echo=F, out.width="45%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison."}
-knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Pics/venny.png")
+```{r venn, echo=F, out.width="45%", fig.align='center', fig.cap="Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD."}
+
+ch <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs CAD-Healthy.txt", check.names = F)$DEGs
+
+mh <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-Healthy.txt", check.names = F)$DEGs
+
+mc <- read.table("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/DEG and DEM/Fisher results/DEGs MI-CAD.txt", check.names = F)$DEGs
+
+df <- list(`CAD/Healthy` = ch, `MI/Healthy` = mh, `MI/CAD` = mc)
+
+ggvenn(df, show_percentage = F, stroke_alpha = 0.5, fill_alpha = 0.4, text_size = 6, set_name_size = 6) + theme(text = element_text(family = "Times"))
+
 ```
 
-The Venn diagram in Figure \@ref(fig:venn) shows that CAD and MI samples share a majority of their DEGs.
-From 860 DEGs of MI/healthy and 670 DEGs of CAD/healthy, 531 genes were common which is 62% of MI/healthy DEGs and 79% of CAD/healthy DEGs.
+The Venn diagram in Figure \@ref(fig:venn) shows that the CAD and MI samples shared most of their DEGs.
+From 860 DEGs of MI/healthy and 670 DEGs of CAD/healthy, 531 genes were common, which is 62% of MI/healthy DEGs and 79% of CAD/healthy DEGs.
 
-## Gene ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) enrichment analyses of the DEGs.
+## GO and KEGG enrichment analyses of the DEGs
 
-To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
-In MI versus healthy samples, GO enrichment analysis in the BP category, suggested that the DEGs were enriched in "immune response-regulating signaling pathway", "lymphocyte differentiation", "immune response-regulating cell surface receptor signaling pathway", and "leukocyte activation involved in immune response" (Figure \@ref(fig:MIHEnrich)A).
-In the CC category, the DEGs were enriched in "secretory granule membrane", "azurophil granule", "ficolin-1-rich granule", "tertiary granule", and "ficolin-1-rich granule membrane" (Figure \@ref(fig:MIHEnrich)B).
-In the MF category, the DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(fig:MIHEnrich)C).
-KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(fig:MIHEnrich)D).
+To explore the biological classification of the DEGs, we performed GO and KEGG pathway enrichment analyses on the MI/healthy and CAD/healthy DEGs.
+For MI/healthy, GO enrichment analysis in the BP category suggested that the DEGs were enriched in "immune response-regulating signaling pathway," "lymphocyte differentiation," "immune response-regulating cell surface receptor signaling pathway," and "leukocyte activation involved in immune response" (Figure \@ref(fig:MIHEnrich)A).
+In the CC category, DEGs were enriched in "secretory granule membrane," "azurophil granule," "ficolin-1-rich granule," "tertiary granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:MIHEnrich)B).
+In the MF category, DEGs were involved in "cadherin binding" and "MHC class I protein binding" (Figure \@ref(fig:MIHEnrich)C).
+KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway," "Lipid and atherosclerosis," and "Hematopoietic cell lineage" (Figure \@ref(fig:MIHEnrich)D).
 
@@
 
-The enrichment results for DEGs of CAD versus healthy samples were as follows.
-In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response", "positive regulation of innate immune response", "mononuclear cell differentiation", and "positive regulation of response to external stimulus" (Figure \@ref(fig:CADHEnrich)A).
-In the CC category, the DEGs were enriched in "azurophil granule", "ficolin-1-rich granule", and "ficolin-1-rich granule membrane" (Figure \@ref(fig:CADHEnrich)B).
-In the MF category, the DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(fig:CADHEnrich)C).
-KEGG pathway analysis indicated that the DEGs were related to the following pathways: "Chemokine signaling pathway", "Lipid and atherosclerosis", and "Hematopoietic cell lineage" (Figure \@ref(fig:CADHEnrich)D).
+The enrichment results for the CAD/healthy DEGs were as follows.
+In the BP category, GO enrichment suggested that the DEGs were enriched in "positive regulation of defense response," "positive regulation of innate immune response," "mononuclear cell differentiation," and "positive regulation of response to external stimulus" (Figure \@ref(fig:CADHEnrich)A).
+In the CC category, DEGs were enriched in "azurophil granule," "ficolin-1-rich granule," and "ficolin-1-rich granule membrane" (Figure \@ref(fig:CADHEnrich)B).
+In the MF category, DEGs were involved in "lipoprotein particle receptor binding" and "NF-$\kappa$B binding" (Figure \@ref(fig:CADHEnrich)C).
+KEGG pathway analysis showed that the DEGs were related to the following pathways: "Chemokine signaling pathway," "Lipid and atherosclerosis," and "Hematopoietic cell lineage" (Figure \@ref(fig:CADHEnrich)D).
 
@@
 
-Among all DEGs, just miR-186, miR-32, and miR-21 were detected as differentially expressed miRNAs.
-The expression profile of these three miRNAs is presented in Figure \@ref(fig:Expall).
-Additionally, The ROC curves of each miRNA for each layer are presented in Figure \@ref(fig:miRROC).
-Using the logistic regression model the AUC-ROC of miR-21, miR-32, and miR-186 for separating healthy and not-healthy samples was 0.98, 0.99, and 0.90 respectively (Figure \@ref(fig:miRROC)A).
-Besides, the accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set was 0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
-Moreover, the ROC curve of each miRNA for classifying MI and CAD samples was presented in Figure \@ref(fig:miRROC)B.
-For miR-21, miR-32, and miR-186, the AUC-ROC and accuracy on the test set were 0.85; 0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
+Among the DEGs, miR-186, miR-32, and miR-21 were identified as differentially expressed miRNAs.
+The expression profiles of the three miRNAs are shown in Figure \@ref(fig:Expall).
+The ROC curves of each miRNA in each layer are presented in Figure \@ref(fig:miRROC).
+Using the logistic regression model, the AUC-ROC values of miR-21, miR-32, and miR-186 for separating healthy and not-healthy samples were 0.98, 0.99, and 0.90, respectively (Figure \@ref(fig:miRROC)A).
+The accuracy of each miRNA for classifying the samples into healthy and not-healthy groups on the test set for miR-21, miR-32, and miR-186 was 0.92, 0.98, and 0.89, respectively.
+The ROC curve of each miRNA for classifying MI and CAD samples is presented in Figure \@ref(fig:miRROC)B.
+The AUC-ROC and accuracy for miR-21, miR-32, and miR-186 in the test set were 0.85; 0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
 
@@
 
-kbl(df, booktabs = T, align = "c",linesep = c(""), digits = c(1, 60), caption = "Target miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.") %>%
+kbl(df, booktabs = T, align = "c",linesep = c(""), digits = c(1, 60), caption = "Investigated miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.") %>%
   kable_styling(full_width = T) %>%
@@
 
-```{r Expall, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of all miRNAs in two approach in different sample classes."}
+```{r Expall, echo=F, out.width="90%", fig.align='center', fig.cap="Expression profile of all miRNAs in two approaches in different sample classes."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Expressionforall.pdf")
@@
 
-#### First layer for healthy not-healthy isolation:
+#### First layer for the isolation of healthy and not-healthy samples:
 
-Although single miRNAs had acceptable performance, their predictive value could be improved even further by using them as a set.
+Although single miRNAs had an acceptable performance for this layer, their predictive value could be  further improved by using them as a set.
 The ROC curve for the SVM model with an RBF kernel trained with all three miRNAs is presented in Figure \@ref(fig:DEMROC)A.
-The model had a better performance in classification than single miRNAs.The AUC-ROC for the model is 1, and its accuracy on the test set was also 1.
+The model had a better performance in classification than single miRNAs.The AUC-ROC for the model was 1, and its accuracy on the test set was also 1.
 In Figure \@ref(fig:DEMCF)A, the confusion matrix for the model is presented.
 
-```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for miRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples."}
+```{r DEMROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs ROCs h not h cad mi.pdf")
@@
 
-```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification."}
+```{r DEMCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/DEMs Conf. matrix.pdf")
@@
 
-#### Second layer for separating MI samples from CAD:
+#### Second layer for separating the MI and CAD samples:
 
-Different models were trained using expression values for three differentially miRNAs.
-The models' 10-fold cross-validated AUC-ROC and accuracy on the test set are reported in Figure \@ref(fig:DEMmodels).
-The best model from both AUC-ROC and accuracy point-of-view was the SVM model with linear kernel.
-The AUC-ROC and accuracy for this model with its pre-set values were 0.93 and 0.82 respectively.
-The model was hyper-tuned for C and gamma hyper-parameters, and therefore the model showed better performance.
-The ROC curve of the hyper-tuned model is presented in Figure \@ref(fig:DEMROC)B.
-For this model the AUC-ROC reached 0.95 and the accuracy improved to 0.85 (Table \@ref(tab:DEGsML)).
-Moreover, the sensitivity and specificity for the model on the test set were 0.91 and 0.71 respectively.
-The confusion matrix for the hyper-tuned model is illustrated in Figure \@ref(fig:DEMCF)B.
+Different models were trained using the expression values of three differentially expressed miRNAs.
+The models' AUC-ROC and the accuracy of the test set are shown in Figure \@ref(fig:DEMmodels).
+The best model from both the AUC-ROC and accuracy points of view was the SVM model with a linear kernel.
+The AUC-ROC and accuracy for this model with its preset values were 0.93 and 0.82, respectively.
+The model was hypertuned for C and gamma hyperparameters, and therefore the model showed better performance.
+The ROC curve of the hypertuned model is presented in Figure \@ref(fig:DEMROC)B.
+For this model, the AUC-ROC reached 0.95, and the accuracy was improved to 0.85 (Table \@ref(tab:DEGsML)).
+Moreover, the sensitivity and specificity for the model on the test set were 0.91 and 0.71, respectively.
+The confusion matrix for the hypertuned model is illustrated in Figure \@ref(fig:DEMCF)B.
 
-```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs on the test set."}
+```{r DEMmodels, echo=F, out.width="65%", fig.align='center', fig.cap="Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/Models DEGs mirs.pdf")
@@
 
-```{r, DEGsML, echo=FALSE}
+```{r DEGsML, echo=FALSE}
 
@@
 
-kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning") %>%
+kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with a linear kernel as the best model trained with differentially expressed miRNAs on the training and test sets before and after hypertuning") %>%
   kable_styling(full_width = T) %>%
-  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
+  add_header_above(c(" " = 2, "Preset parameters" = 2, "Hypertuned" = 2))
 ```
@@
 
-After calculating the AUC-ROC for each miRNA for the classification of MI and CAD samples, the miRNAs with AUC-ROC over 0.8 were selected.
-The selected miRNAs were miR-29a, miR-197, miR-186, miR-21, and miR-296.
-The expression values of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:Expall).
-The ROC curve of the selected miRNAs for both layers are illustrated in Figure \@ref(fig:miRROC).
+After calculating the AUC-ROC for each miRNA to classify of MI and CAD samples, the miRNAs with AUC-ROC  $> 0.8$ were selected.
+The miRNAs selected were miR-29a, miR-197, miR-186, miR-21, and miR-296.
+The expression levels of these miRNAs in healthy, CAD, and MI samples are presented in Figure \@ref(fig:Expall).
+The ROC curves of the selected miRNAs in both layers are shown in Figure \@ref(fig:miRROC).
 
-#### First layer for isolation of healthy and not-healthy samples
+#### First layer for the isolation of healthy and not-healthy samples:
 
-Using the selected set, an SVM model with an RBF kernel was trained to separate healthy from not-healthy samples.
-The ROC curve for the model is presented in Figure \@ref(fig:AUCROC)A and the confusion matrix is illustrated in Figure \@ref(fig:AUCCF)A.
-Both AUC-ROCand accuracy for the model on the test set were equal to 1.
+Using the selected set, an SVM model with an RBF kernel was trained to separate healthy and not-healthy samples.
+The ROC curve for the model is presented in Figure \@ref(fig:AUCROC)A, and the confusion matrix is illustrated in Figure \@ref(fig:AUCCF)A.
+Both the AUC-ROC and accuracy of the model on the test set were 1.
 
-```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for the set of miRNAs selected by AUC-ROC on test set classification.
-(A) SVM with RBF kernel for healthy and not-healthy samples classification.
-(B) Logistic regression model for CAD and MI samples classification.
-(C) SVM with polynomial kernel for CAD and MI samples classification.
+```{r AUCROC, echo=F, out.width="95%", fig.align='center', fig.cap="ROC curve for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+(B) Logistic regression model for CAD and MI sample classification.
+(C) SVM with polynomial kernel for CAD and MI sample classification.
 "}
@@
 
-```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
-(B) Logistic regression model for CAD and MI samples classification.
-(C) SVM with polynomial kernel for CAD and MI samples classification."}
+```{r AUCCF, echo=F, out.width="90%", fig.align='center', fig.cap="Confusion matrix for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+(B) Logistic regression model for CAD and MI sample classification.
+(C) SVM with polynomial kernel for CAD and MI sample classification."}
 knitr::include_graphics("/Users/mehrdad/Projects/TBZ.Med/PhD. Project/Bioinformatics and ML/Article Materials/ML/Pics/AUC Conf. matrix.pdf")
@@
 
-#### Second layer for isolation of MI samples from CAD samples
+#### Second layer for separating the MI and CAD samples:
 
-To find the best model for training the best set, different models were trained using their pre-set values.
-Their AUC-ROC and accuracy results on the test set are presented in Figure \@ref(fig:AUCmodels).
-The best model from the AUC-ROC point-of-view was the SVM with linear kernel and from the accuracy point-of-view, it was the SVM model with an RBF kernel.
-For the SVM-linear model the AUC-ROC and accuracy were 0.93 and 0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84, respectively.
-Both models were hyper-tuned and the ROC curve for their best performance is presented in Figure \@ref(fig:AUCROC)B and C.
-The AUC-ROC and accuracy for the SVM-linear model modified to 0.92 and 0.88, respectively.
+To find the best model for this set of miRNAs, different models were trained using their preset values.
+The AUC-ROC and accuracy results for the test set are presented in Figure \@ref(fig:AUCmodels).
+The best model from the AUC-ROC point of view was the SVM with a linear kernel, and from the accuracy point of view, it was the SVM model with an RBF kernel.
+For the SVM-linear model, the AUC-ROC and accuracy were 0.93 and 0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84, respectively.
+Both models were hyper-tuned, and the ROC curve for their best performance is presented in Figure \@ref(fig:AUCROC)B and C.
+The AUC-ROC and accuracy for the SVM-linear model were modified to 0.92 and 0.88, respectively.
 For the SVM-RBF, these values increased to 0.96 and 0.94, respectively (Table \@ref(tab:AUCML)).
-The sensitivity for the SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and the specificity for them was 0.79 and 0.86, respectively.
+The sensitivities for the SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and their specificities were 0.79 and 0.86, respectively.
 The confusion matrix for both models is illustrated in Figure \@ref(fig:AUCCF)B and C.
@@
 
-kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their individual AUC-ROC on the train and test set before and after hyper-tuning") %>%
+kbl(df, booktabs = T, align = "c",linesep = c(""), caption = "AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their AUC-ROC on the train and test set before and after hyper-tuning") %>%
   kable_styling(full_width = T) %>%
-  add_header_above(c(" " = 2, "Pre-set parameters" = 2, "Hyper-tuned" = 2))
+  add_header_above(c(" " = 2, "Preset parameters" = 2, "Hypertuned" = 2))
 ```
@@
 
-The prevalence of MI can lead to high-rate mortality in the clinical setting.
-However, early diagnosis and application of suitable treatment protocols can reduce mortality and improve AMI prognosis [@CVD; @17; @18].
+The prevalence of MI can lead to high mortality rates in the clinical setting.
+However, early diagnosis and the application of suitable treatment protocols can reduce mortality and improve MI prognosis [@CVD; @17; @18].
 Studies have suggested that changes in miRNA expression may play a significant role in the progression of MI and the subsequent remodeling [@23].
-It is believed that the expression of miRNAs is altered during the various biological processes correlated with MI within the myocardium or other related tissues [@24].
-Although several research has been concentrated on examining free circulating miRNAs in the serum samples for the detection of cardiac tissue injuries [@11-15], more information is needed to fully comprehend the miRNAs found in different blood sub-components like plasma, platelets, and PBMCs.
-Based on previous data, PBMCs are critically involved in plaque destabilization and rupture as well as early inflammatory responses during MI [@Meder6; @Meder23].
-Moreover, PBMCs have specific miRNA profile that is altered under certain pathological conditions which are great candidates as disease biomarkers [@Meder6].
+It is believed that miRNA expression is altered during the various biological processes correlated with MI within the myocardium or other related tissues [@24].
+Although several studies have focused on examining free circulating miRNAs in serum samples for the detection of cardiac tissue injuries [@11-15], more information is needed to fully comprehend the miRNAs found in different blood subcomponents, such as plasma, platelets, and PBMCs.
+Based on previous findings, PBMCs play a crucial role in the destabilization and rupture of plaques as well as in the initial inflammatory reactions in individuals experiencing myocardial infarction (MI).
+[@Meder6; @Meder23].
+Moreover, PBMCs have specific miRNA profiles that are altered under certain pathological conditions,  making them great candidates as disease biomarkers [@Meder6].
 
-PBMCs can respond to several insulting conditions such as MI in the least possible time with prominent changes in their miRNA profile [@Meder6].
-Considering the regulatory roles, subtle changes in the transcription of miRNAs can be monitored even before alteration in the levels of mRNAs and proteins [@miR].
-These features make the miRNAs an early-stage valid diagnostic tool for the detection of minor and major cell injuries.
-To date, few studies have been performed to compare the miRNA profiles in PBMCs belonging to acute MI patients and other CADs and healthy samples to find a robust set of identical miRNAs to differentiate these pathological conditions.
+PBMCs can respond to several insulting conditions, such as MI, in the shortest possible time with notable changes in their miRNA profile [@Meder6].
+Considering their regulatory roles, subtle changes in the transcription of miRNAs can be monitored even before alterations in mRNA and protein levels[@miR].
+These features make miRNAs a valid early-stage diagnostic tool for the detection of minor and major cell injuries.
+To date, few studies have compared the miRNA profiles in PBMCs from patients with MI and other CADs and healthy samples to find a robust set of identical miRNAs to differentiate these pathological conditions.
 
-In this study, we combined three GEO datasets of healthy, CAD, and MI samples.
-Having these samples set alongside bioinformatics analysis and ML means, it is possible to identify potential biomarker sets and also effective therapeutic targets.
-The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(fig:venn)) are proof of the close relationship between the MI and CAD samples.
-Interestingly, functional enrichment analysis demonstrated that DEGs in both healthy/CAD and healthy/MI were strongly correlated to immune cell response which is a major cellular part of PBMCs.
-Here, two different sets of miRNAs were used as biomarker sets for sample classification.
-miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-186, miR-21; miR-29a; miR-197; and miR-296 were selected according to their AUC-ROC values.
-As shown in Figure \@ref(fig:miRROC), all miRNAs selected with both approaches had AUC-ROC over 0.9 for the isolation of healthy and not-healthy samples except for miR-296 and miR-29a.
-Data confirmed that the real challenge is to classify CAD and MI samples because of close overlap.
-Of 6 miRNAs under investigation in both approaches except for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of CAD and MI samples.
-Besides, the high AUC-ROC values of miRNAs confirms their high potential as biomarkers.
+In this study, we combined three GEO datasets for healthy, CAD, and MI samples.
+Having these sample sets alongside bioinformatics analysis and ML methods enabled us to identify potential biomarker sets and effective therapeutic targets.
+The results of the DEG analysis (Table \@ref(tab:DEGstab) and Figure \@ref(fig:venn)) prove the close relationship between the MI and CAD samples.
+Interestingly, functional enrichment analysis demonstrated that DEGs in both CAD/healthy and MI/healthy were strongly correlated with the immune cell response, which is a major part of PBMCs.
+Two sets of miRNAs were selected as biomarker sets for sample classification.
+miR-21; miR-32; and miR-186 were selected as differentially expressed miRNAs, and miR-186; miR-21; miR-29a; miR-197; and miR-296 were selected based on their AUC-ROC values.
+As shown in Figure \@ref(fig:miRROC), all miRNAs selected with both approaches had AUC-ROCs $> 0.9$ for isolating healthy and not-healthy samples except for miR-296 and miR-29a.
+The data confirmed that the real challenge was to classify CAD and MI samples because of the close overlap.
+Of the six miRNAs under investigation in both approaches, except for miR-32, all miRNAs had an AUC-ROC $> 0.8$ for the discrimination of CAD and MI samples.
+As expected, the high AUC-ROC values of the miRNAs confirmed their high potential as biomarkers.
 
-ML models when trained with miRNA sets selected by both DEG and AUC-ROC approaches showed better performance in the classification than each miRNA.
-To avoid unwanted complexity and poor predictive values, a two-layer architecture was also designed.
-The first layer was for the discrimination of healthy from not-healthy samples, and the second layer separated the CAD from MI candidates.
-As expected, in both approaches a hyper-tuned SVM model could flawlessly separates healthy from not-healthy samples using distinct miRNAs sets.
-The ML models were also capable of effectively separating CAD from MI patients.
-Although both miRNA sets had nearly the same AUC-ROC with their best model, the accuracy, sensitivity, and specificity were different.
-The model trained with AUC-selected miRNAs had better performance in all predictive values.
+ML models trained with miRNA sets selected by both DEG and AUC-ROC approaches, showed better  classification performance than each miRNA.
+To avoid unwanted complexity and poor predictive values, a two-layer architecture was designed.
+The first layer was used to discriminate between healthy and not-healthy samples, and the second layer was was used to separate CAD from MI candidates.
+As expected, in both approaches, a hypertuned SVM model could flawlessly separate healthy and not-healthy samples using distinct miRNA sets.
+ML models are also capable of effectively separating CAD from MI patients.
+Although both miRNA sets had nearly the same AUC-ROC using the best model, their accuracy, sensitivity, and specificity were different.
+The model trained with AUC-selected miRNAs showed better performance in all predictive values, which is logical because of the higher number of miRNAs in the set.
 
-Numerous studies have reported different biological processes can affect the expression of miRNAs in PBMCs.
-However, there are still controversies regarding the exact role of miRNAs in the function of immune cells and the correlation of specific pathological conditions with miRNA profiles.
-Several studies have proved the activation of specific miRNA types in PBMCs under cardiovascular events [@21; @186; @32-2; @29aa; @197].
-For instance, there is evidence that the elevation of miR-186 suppresses the expression of Cystathionine-$\gamma$-lyase, leading to the subsequent secretion of pro-inflammatory cytokines and cellular lipid accumulation.
-Besides, macrophage-derived miR-186 may promote atherosclerotic plaques [@186].
-In line with this claim, we found that miR-186 is up-regulated in both CAD and MI candidates related to control counterparts.
-Surprisingly, the obtained data indicated that the expression of miR-186 is higher in CAD patients in comparison to MI (Figure \@ref(fig:DEMexp)), To be specific, miR-186 is the only differentially expressed miRNA between CAD and MI, with a clear up-regulation in CAD, indicating its main role in the promotion of atherosclerosis.
+Numerous studies have reported that different biological processes can affect the miRNA expression in PBMCs.
+However, the exact role of miRNAs in the function of immune cells and the correlation between specific pathological conditions and miRNA profiles remain controversial.
+Several studies have proven the activation of particular miRNA types in PBMCs under cardiovascular events [@21; @186; @32-2; @29aa; @197, @296].
+For instance, there is evidence that elevation of miR-186 suppresses the expression of cystathionine-$\gamma$-lyase, leading to the subsequent secretion of pro-inflammatory cytokines and cellular lipid accumulation.
+In addition, macrophage-derived miR-186 may promote atherosclerotic plaque formation [@186].
+In line with this claim, we found that miR-186 was up-regulated in both CAD and MI candidates compared to their control counterparts.
+Surprisingly, the obtained data indicated that the expression of miR-186 was higher in patients with CAD than in patients with MI (Figure \@ref(fig:Expall)).
+Specifically, miR-186 was the only differentially expressed miRNA between CAD and MI, with a clear up-regulation in CAD, indicating its main role in the promotion of atherosclerosis.
 
-As mentioned before, miR-21 was also up-regulated in both MI and CAD patients in comparison to healthy controls.
-Moreover, the expression value of miR-21 was significantly higher in MI than that of the CAD group (Table \@ref(tab:mirExptable)).
-It is thought that the up-regulation of miRNA-21 in PBMCs is a compensatory reaction to reduce T$_{reg}$ lymphocyte number in response to the reduction of TGF$\beta1$ secretion into the plasma through a TGF$\beta1$/smad-independent pathway.
-In line with previous and present data, miR-21 can modulate the activity of PBMCs following the occurrence of cardiovascular diseases [@21].
+As mentioned before, miR-21 was also up-regulated in both MI and CAD patients compared to healthy controls.
+Moreover, the expression value of miR-21 was significantly higher in the MI group than in the CAD group (Table \@ref(tab:mirExptable)).
+It is thought that the up-regulation of miR-21 in PBMCs is a compensatory reaction to reduce the T$_{reg}$ lymphocyte number in response to the reduction in TGF$\beta1$ secretion into the plasma through a TGF$\beta1$/smad-independent pathway.
+In line with the previous and present data, miR-21 can modulate the activity of PBMCs following the occurrence of cardiovascular diseases [@21].
 
-Recent data have supported the elevation of miR-32 in CAD patients with the calcification of coronary artery.
-Of note, miR-32 stimulates the calcification of mouse vascular smooth muscle through the regulation of bone morphogenetic protein-1, runt-related transcription factor-2 (RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA protein [@32-2].
-Likewise, there are some reports associated with the activity of PBMC miR-32 under several pathologies [@magic; @32].
-The exact role of PBMC miR-32 after cardiovascular events remained to be elucidated.
+Recent data have supported the elevation of miR-32 levels in CAD samples with calcification of the coronary artery.
+Notably, miR-32 promotes vascular smooth muscle calcification in mice by controlling the activity of several proteins, including bone morphogenetic protein-1, runt-related transcription factor-2 (RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA protein.
+[@32-2].
+Likewise, some reports are associated with the activity of miR-32 in PBMCs in several pathologies [@magic; @32].
+The exact role of miR-32 in PBMCs after cardiovascular events remains unclear.
 
-Molecular analyses have indicated the regulatory role of miRNAs selected by the AUC-ROC approach in PBMCs after a cardiovascular event.
-Two common miRNAs in DEGs and AUC-ROC approachs, miR-21 and miR-186, were covered in DEGs miRNA set.
-Based on numerous reports mir-29a can be activated in different diseases [@29aa].Data analysis indicated that miR-29a is significantly up-regulated in CAD patients in comparison to healthy and MI groups (Table \@ref(tab:mirExptable)).
-Increased miR-29a is associated with the progression of atherosclerosis, and the combination of miR-29a and ox-LDL was offered as a valid biomarker set for paraclinical classification [@29a].
-However, the role of miR-29a in the function of PBMCs in CAD patients has not been completely examined.
+Molecular analyses have indicated the regulatory role of miRNAs selected using the AUC-ROC approach in PBMCs after a cardiovascular event.
+The biological importance of two common miRNAs in the DEGs and AUC-ROC approaches, miR-21 and miR-186, have already been discussed.
+Based on numerous reports, miR-29a can be activated in different diseases [@29aa].
+Data analysis indicated that miR-29a was significantly up-regulated in CAD patients compared to the healthy and MI groups (Table \@ref(tab:mirExptable)).
+Increased miR-29a is associated with the progression of atherosclerosis, and the combination of miR-29a and ox-LDL has been suggested as a valid biomarker set for paraclinical classification [@29a].
+However, the role of miR-29a in the function of PBMCs from patients with CAD has not been thoroughly examined.
 
-Data indicated that miR-197 is also significantly up-regulated in both CAD/healthy and MI/healthy groups.
-In previous studies, it has been shown that miRNA-197 may play a critical role in regulating the anti-inflammatory response of IL-35 by affecting pro/ anti-inflammatory cytokine secretion, M1/M2 macrophage ratio, T$_{reg}$ lymphocyte proliferation, and T cell suppression suggesting the potential diagnostic role of miR-197 in adverse cardiovascular events [@197].
+The data indicated that miR-197 was significantly up-regulated in both the CAD/healthy and MI/healthy groups.
+Previous studies have demonstrated that miR-197 may play a crucial role in controlling the anti-inflammatory response of IL-35 by influencing the secretion of cytokines that can either promote or suppress inflammation, the ratio of M1/M2 macrophages, and the proliferation of T$_{reg}$ lymphocytes, which are responsible for suppressing immune responses [@197].
+Alongside our findings, it can be concluded that miR-197 could be a useful diagnostic tool for predicting adverse cardiovascular events.
 
-The findings of this study demonstrate the potential of miR-296 as a biomarker with high discriminatory power for distinguishing between samples from individuals with MI and CAD.
-MiR-296 has been identified as a key regulator in the development and advancement of atherosclerosis by controlling the expression of target genes associated with various biological processes including angiogenesis, inflammation, cholesterol metabolism, hypertension, cellular proliferation, and apoptosis [@296].
-In a study it has been shown that miR-296 expression levels are significantly increased in the PBMCs of CAD patients compared to healthy controls, suggesting its involvement in regulating pro-inflammatory cytokines like IL-6 and TNF-a [@296a].
-These findings suggest that miR-296 may have a significant impact on the pathogenesis of atherosclerosis and could potentially serve as a diagnostic biomarker for CAD or MI.
+The findings of this study demonstrate the potential of miR-296 as a biomarker with high discriminatory power to distinguish between samples from individuals with MI and CAD.
+MiR-296 has been identified as a key regulator in the development and advancement of atherosclerosis by controlling the expression of target genes associated with various biological processes, including angiogenesis, cholesterol metabolism, inflammation, cellular proliferation, hypertension, and apoptosis [@296].
+In a previous study, miR-296 expression levels were found to be significantly increased in the PBMCs of CAD patients compared to healthy controls, suggesting its involvement in regulating proinflammatory cytokines such as IL-6 and TNF-$\alpha$ [@296a].
+These findings suggested that miR-296 may have a significant impact on the pathogenesis of atherosclerosis and could potentially serve as a diagnostic biomarker for CAD or MI.
 
@@
 
-In summary, we derived a set of miRNA biomarkers by comparing MI samples to both healthy and CAD samples.
-We found that the SVM model performed best in both the first layer, which separated healthy and not-healthy samples, and the second layer, which classified MI/CAD samples.
-The set of miRNAs selected based on their AUC-ROC values had better performance in the second layer.
-Overall, our two-layer structure achieved a accuracy of 0.96.
-This demonstrates the potential for combining bioinformatics and machine learning techniques to identify novel biomarkers and gain a deeper understanding of myocardial infarction.
+In summary, we derived a set of miRNA biomarkers by comparing MI samples with both healthy and CAD samples.
+We found that the SVM model performed best in both the first layer, which separated healthy and unhealthy samples, and the second layer, which classified the MI/CAD samples.
+The set of miRNAs selected based on their AUC-ROC values performed better in the second layer.
+Overall, our two-layer structure achieved an accuracy of 0.96.
+This demonstrates the potential of combining bioinformatics and machine learning techniques to identify novel biomarkers and gain a deeper understanding of myocardial infarction.
 
 # References
-
-\begin{align}
-a^2+b^2=c^2
-\end{align}

### 2023-08-07 Complete-V1/Complete-V1.tex
The version after revision

@@
 
-\title{Machine Learning Deriven Set of microRNAs as a Novel Biomarker
-for Myocardial Infarction Diagnosis }
-
-\author{  Mehrdad Samadishadlou \and  Reza Rahbarghazi \and  Farhad
-Bani \and  }
+\title{Unlocking the Potential of microRNAs: Machine Learning Identifies
+Key Biomarkers for Myocardial Infarction Diagnosis }
+
+\author{  Mehrdad Samadishadlou \and  Reza Rahbarghazi \and  Zeynab
+Piryaei \and  Mahdad Esmaeili \and  Çığır Biray Avcı \and  Farhad
+Bani \and  Kaveh Kavousi \and  }
 
@@
 Sciences, Tabriz, Iran \\
-     \email{\href{mailto:samadishadlou@tbzmed.ac.ir}{\nolinkurl{samadishadlou@tbzmed.ac.ir}}}  %  \\
+     %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
 Sciences, Tabriz, Iran \\
-     \email{\href{mailto:rahbarghazir@tbzmed.ac.ir}{\nolinkurl{rahbarghazir@tbzmed.ac.ir}}}  %  \\
+     %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+    \and
+        Zeynab Piryaei \at
+     Laboratory of Complex Biological Systems and Bioinformatics (CBB),
+Department of Bioinformatics, Institute of Biochemistry and Biophysics
+(IBB), University of Tehran, Iran. \\
+     %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+    \and
+        Mahdad Esmaeili \at
+     Medical Bioengineering Department, School of Advanced Medical
+Sciences, Tabriz University of Medical Sciences, Tabriz, Iran \\
+     %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+    \and
+        Çığır Biray Avcı \at
+     Medical Biology Department, School of Medicine, Ege University,
+İzmir, Turkey \\
+     %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
     \and
+        Kaveh Kavousi \at
+     Laboratory of Complex Biological Systems and Bioinformatics (CBB),
+Department of Bioinformatics, Institute of Biochemistry and Biophysics
+(IBB), University of Tehran, Iran \\
+     \email{\href{mailto:kkavousi@ut.ac.ir}{\nolinkurl{kkavousi@ut.ac.ir}}}  %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+    \and
     }
@@
 maladaptive responses in cardiovascular diseases, making them attractive
-targets for potential therapeutics. However, their potential as novel
+targets for potential biomarkers. However, their potential as novel
 biomarkers for diagnosing cardiovascular diseases requires systematic
@@
 the Gene Expression Omnibus (GEO) database, which contains peripheral
-blood mononuclear cell samples from individuals with myocardial
+blood mononuclear cell (PBMC) samples from individuals with myocardial
 infarction (MI), stable coronary artery disease (CAD), and healthy
 individuals. Additionally, we selected a set of miRNAs based on their
-area under the curve of the receiver operating characteristic curve
-(AUC-ROC) to separate the CAD/MI samples. We designed a two-layer
-architecture for sample classification, with the first layer isolating
-healthy samples from not-healthy ones, and the second layer classifying
-stable CAD and MI samples. We trained different machine learning models
-using both biomarker sets and evaluated their performance on a test set.
-We identified miR-21, miR-186, and miR-32 as the only miRNAs among the
+area under the receiver operating characteristic curve (AUC-ROC) for
+separating the CAD and MI samples. We designed a two-layer architecture
+for sample classification, in which the first layer isolates healthy
+samples from unhealthy samples, and the second layer classifies stable
+CAD and MI samples. We trained different machine learning models using
+both biomarker sets and evaluated their performance on a test set. We
+identified miR-21, miR-186, and miR-32 as the only miRNAs among the
 differentially expressed genes, and a set including miR-186, miR-21,
@@
 classification of CAD and MI was achieved with an SVM model trained
-using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.95,
-and an accuracy of 0.88. Our study demonstrated that miRNA signatures
-derived from peripheral blood could serve as valuable novel biomarkers
+using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.96 and
+an accuracy of 0.94 on the test data. Our study demonstrated that miRNA
+signatures derived from PBMCs could serve as valuable novel biomarkers
 for cardiovascular diseases.
@@
         Myocardial Infarction \and
+        Bioinformatics \and
+        Biomarker \and
     }
@@
 
-At present, cardiovascular diseases (CVDs) are the leading cause of
-human mortality with 32\% of all global deaths. It is estimated that
-about 85\% of CVDs mortality were diagnosed with myocardial infarction
-(MI) ({``Cardiovascular Diseases ({CVDs})''} n.d.). MI is an acute
-coronary syndrome with sudden blockage and stenosis of the coronary
-artery, and subsequent myocardial ischemia, leading to extensive
+Cardiovascular diseases (CVDs) are the leading cause of human mortality,
+accounting for 32\% of all global deaths. It is estimated that
+approximately 85\% of CVD mortality is due to myocardial infarction (MI)
+({``Cardiovascular Diseases ({CVDs})''} n.d.). MI is an acute coronary
+syndrome characterized by sudden blockage and stenosis of the coronary
+artery and subsequent myocardial ischemia, leading to extensive
 cardiomyocyte damage and necrosis (Yap et al. 2023).
 
-Over the last 50 years, numerous attempts have been collected to use
-biomarkers to facilitate diagnosis, assess risk, follow-up therapy, and
-determine therapeutic efficacy in CVDs candidates. Based on the released
-guidelines, cardiac troponins (cTns) are used as a highly-sensitive and
-accurate approach for the detection of MI. Despite the inherent
-advantages, the high-rate sensitivity of cTn-based assays has also led
-to more false positive results (Thygesen et al. 2018), which do
-necessitate the advent and development of new modalities with
-pathological values. To improve diagnostic value upon existing MI
-biomarkers, the combination of complementary biological markers, such as
-microRNAs (miRNAs) and other genetic factors, is proposed. Previous
-researches support the notion that miRNAs exhibit the great potential to
-be used as alternative biomarkers in CVDs detection and follow-up
-(Schulte et al. 2020). It is suggested that miRNAs possess 18-22
-nucleotides and can play a crucial role in the regulation of gene
-expression. Evidence point to the fact that miRNAs are involved in the
-pathogenesis of cardiac tissue injury and could be considered as
-theranostics targets in terms of CVDs. These elements can easily
-circulate in biofluids with both diagnostic as well as prognostic values
-(Schulte, Karakas, and Zeller 2017). Several biological activities such
-as angiogenesis, cardiomyocyte growth and contractility, lipid
-metabolism, plaque formation, and cardiac rhythm are regulated by miRNAs
-(Kalayinia et al. 2021). It is postulated that the function and
-diagnostic properties of miRNAs are beyond the myocardium in CVD
-patients. To be specific, the expression of miRNAs can vary in different
-biofluids and cell components such as serum and peripheral blood
-mononuclear cells (PBMCs) (Soler-Botija, Gálvez-Montón, and Bayés-Genís
-2019).
-
-PBMCs are a fraction of white blood cells (WBCs), including monocytes,
-lymphocytes, macrophages, and other cells belonging to the immune system
-(Gao et al. 2020). Emerging data have indicated that PBMCs can be used
-as a valid source of biomarkers for monitoring various pathological
-conditions. Of note, the alteration of mRNAs and miRNAs under
-pathological conditions gives us valuable information about different
-kinds of disorders. PBMCs could recapitulate the conditions of the
-target tissues, thus, providing a highly sensitive and specific source
-of biomarkers (Mosallaei et al. 2022). Commensurate with these
-conditions, these cells are repositories of dysregulated genes and
-miRNAs expression profiles in CVDs (Gao et al. 2020; Mosallaei et al.
-2022).
-
-In recent years, the advent and use of machine learning (ML) is an
-exciting prospect for advancing scientific discoveries. Although the
-concept of ML and its initial algorithms were conceived many years ago,
-recent improvements in computing power and access to vast amounts of
-data have shown that ML techniques outperform classical statistical
-methods in various fields. Furthermore, the progress made in omics
-technologies has enabled the analysis of massive and intricate
-biological data sets, consisting of hundreds to thousands of samples,
-which makes it possible for ML to extract valuable biological
-information from such data (Torun et al. 2023). Therefore, ML offers
-novel techniques to integrate and analyze the various omics data
-enabling the discovery of de novo biomarkers. These biomarkers help us
-in accurate disease prediction, patient stratification, and finding new
-therapeutics (Reel et al. 2021).
-
-In this study, we aimed to identify potential miRNA biomarkers for MI
-patients by combining and analyzing three different microarray datasets
-from PBMCs. It is suggested that the integration of omics data with
-bioinformatics and ML techniques could be a promising tool in the
-discovery of new and more accurate biomarkers for monitoring MI.
-Besides, this approach can deepen our vision into the underlying
-mechanisms of MI and aid in the development of valid diagnostic
-biomarkers, and patient stratification.
+Over the last 50 years, numerous attempts have been made to use
+biomarkers to facilitate diagnosis, assess the risk, follow-up therapy,
+and determine therapeutic efficacy in CVD candidates. Based on released
+guidelines, cardiac troponins (cTns) are used as a highly sensitive and
+accurate approach for detecting MI. Despite these inherent advantages,
+the high sensitivity of cTn-based assays has also led to more
+false-positive results (Thygesen et al. 2018), necessitating the advent
+and development of new modalities with pathological value. To improve
+the diagnostic value of existing MI biomarkers, a combination of
+complementary biological markers, such as microRNAs (miRNAs) and other
+genetic factors, has been proposed. Previous research supports the
+notion that miRNAs exhibit great potential as alternative biomarkers for
+CVD detection and follow-up (Schulte et al. 2020). It has been suggested
+that miRNAs possess 18-22 nucleotides and play a crucial role in the
+regulation of gene expression. Evidence indicates that miRNAs are
+involved in the pathogenesis of cardiac tissue injury (Schulte, Karakas,
+and Zeller 2017). Several biological processes, such as angiogenesis,
+cardiomyocyte growth and contractility, lipid metabolism, plaque
+formation, and cardiac rhythm, are regulated by miRNAs (Kalayinia et al.
+2021). These elements can easily circulate in biofluids and could be
+considered theranostic targets in terms of CVDs (Schulte, Karakas, and
+Zeller 2017). It has been postulated that the function and diagnostic
+properties of miRNAs are beyond the myocardium in patients with CVD.
+Specifically, the expression of miRNAs can vary in different biofluids
+and cell components such as serum and peripheral blood mononuclear cells
+(PBMCs) (Soler-Botija, Gálvez-Montón, and Bayés-Genís 2019).
+
+PBMCs are a fraction of white blood cells, including monocytes,
+lymphocytes, macrophages, and other cells of the immune system (Gao et
+al. 2020). Emerging data indicate that PBMCs can be used as a valid
+source of biomarkers for monitoring various pathological conditions. Of
+note, the alteration of mRNAs and miRNAs under pathological conditions
+provides valuable information about different kinds of disorders. PBMCs
+can recapitulate the conditions of target tissues, thus providing a
+highly sensitive and specific source of biomarkers (Mosallaei et al.
+2022). Combined with these conditions, these cells are repositories of
+dysregulated genes and miRNA expression profiles in CVDs (Gao et al.
+2020; Mosallaei et al. 2022).
+
+In recent years, the advent and application of machine learning (ML) has
+been an exciting prospect for advancing scientific research. Although
+the concept of ML and its initial algorithms were conceived many years
+ago, recent improvements in computing power and access to vast amounts
+of data have demonstrated that ML techniques outperform classical
+statistical methods in various fields. Furthermore, the progress made in
+omics technologies has enabled the analysis of massive and intricate
+biological datasets, consisting of hundreds to thousands of samples,
+which makes it possible for ML to extract valuable biological insights
+and information from such data (Torun et al. 2023). Consequently, ML
+provides innovative methods for merging and interpreting diverse types
+of omics data, leading to the identification of new biomarkers. These
+biomarkers can aid in precise disease prediction, patient
+stratification, and the development of novel therapeutic approaches
+(Reel et al. 2021).
+
+In this study, we aimed to identify potential miRNA biomarkers in
+patients with MI by combining and analyzing three different microarray
+datasets from PBMCs. The integration of omics data with bioinformatics
+and ML techniques could be a promising tool in the discovery of new and
+more accurate biomarkers for monitoring MI. Additionally, this approach
+can deepen our understanding of the underlying mechanisms of MI and aid
+in the development of valid diagnostic biomarkers and patient
+stratification.
 
@@
 database (\url{https://www.ncbi.nlm.nih.gov/geo/}). To obtain sufficient
-classification power between MI, healthy and stable coronary artery
-disease (CAD) samples, a relatively large sample size was required.
-Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475
-for healthy samples were selected. All samples were produced using
-Affymetrix Human Gene 1.0 ST Array (GPL6244) platform. Only healthy,
-stable CAD and early-stage MI samples were selected from these datasets
-for further analyses. The basic information for the three datasets
-evaluated in the current study is provided in Table @ref(tab:datasets).
+classification power between MI, healthy, and CAD samples, a relatively
+large sample size is needed Therefore, GSE59867 for the MI and CAD
+samples and GSE56609 and GSE54475 for the healthy samples were selected.
+All samples were produced using Affymetrix Human Gene 1.0 ST Array
+platform (GPL6244). Only healthy, CAD, and early-stage MI samples were
+selected from these datasets for further analysis. The basic information
+for the three datasets evaluated in this study is provided in Table
+@ref(tab:datasets). Bioinformatics analyses were conducted using R, ver.
+4.2.0 (R Core Team 2022), RStudio (RStudio Team 2020). All plots and
+graphics of these sections were created using the ggplot2 R package
+(Wickham 2016).
 
 \begin{longtable}[]{@{}cccccc@{}}
-\caption{Basic information of the GEO microarray
+\caption{Sample information on the GEO microarray
 datasets.}\tabularnewline
@@
 
-\hypertarget{pre-processing}{%
-\subsection{Pre-processing}\label{pre-processing}}
-
-Raw data (CEL files) of all datasets were downloaded from the GEO and
-pre-processed using the fRMA package (M. N. McCall, Bolstad, and
-Irizarry 2010). fRMA allowed to pre-process of individual microarray
-samples and combining them consistently for analysis. For each dataset,
-background correction was performed using the RMA algorithm and then it
-was quantile normalized based on the reference distribution. During
-summarization, batch effects were removed and variances of the gene
-expressions were estimated by taking into account these probe-specific
-effects. For those multiple probe sets matched to the identical gene,
-the mean log fold change was retained. Therefore, fRMA can be seen as a
-batch effect removal technique for different datasets that are produced
-by identical microarray platforms. Thus, to ensure batch effect removal,
-the principal component analysis and the relative log expression of
-train samples were plotted before and after fRMA (Lazar et al. 2013).
+\hypertarget{preprocessing}{%
+\subsection{Preprocessing}\label{preprocessing}}
+
+The raw data in the form of CEL files from all datasets were obtained
+from GEO. To prepare the data for analysis, we utilized the fRMA package
+(M. N. McCall, Bolstad, and Irizarry 2010) to facilitate preprocessing
+of individual microarray samples and their consistent combination. For
+each dataset, background correction was applied using the RMA algorithm,
+followed by quantile normalization based on the reference distribution.
+To account for probe-specific effects, batch effects were eliminated
+during summarization and gene expression variances were estimated
+accordingly. In cases where multiple probe sets matched the same gene,
+the mean log-fold change was retained. Consequently, fRMA can serve as a
+technique to remove batch effects across diverse datasets generated by
+identical microarray platforms (Lazar et al. 2013). To ensure the
+effectiveness of the batch effect removal, we employed principal
+component analysis (PCA) and relative log expression (RLE) plots to
+visualize the data before and after applying fRMA.
 
@@
 
-The barcode algorithm proposed by McCall et al. (Matthew N. McCall et
-al. 2011) transformed the actual expression values into binary barcode
-values. Huge sets of samples were collected and normalized using fRMA
-for several platforms as well as for Affymetrix Human Gene 1.0 ST Array
-(GPL6244) platform. The distribution of the expressed and unexpressed
-observed intensities for each gene is estimated using these normalized
-sets. Genes were considered expressed (and their value coded to 1) or
-unexpressed (and their value coded to 0) according to the following
-equation:
+The barcode algorithm was introduced by McCall et al. (Matthew N. McCall
+et al. 2011), aimed to convert actual expression values into binary
+barcode values. Extensive sample collections were gathered and
+normalization was performed using fRMA across multiple platforms,
+including the Affymetrix Human Gene 1.0 ST Array (GPL6244) platform. By
+utilizing these normalized datasets, the distribution of the observed
+intensities for both the expressed and unexpressed genes was estimated.
+The determination of whether a gene was expressed or not was based on
+the following equation, where a value of 1 indicates expression and a
+value of 0 indicates non-expression:
 
@@
 
-where \(x_{ij}\) is the normalized intensity of gene \(i\) in sample
-\(j\), \(C\) is a user-defined parameter, \(\sigma^{ne}\) is the
-standard deviation of the non-expressed distribution, and \(\mu^{ne}\)
-is the mean of the non-expressed distribution. The barcode
-representation of a sample is a vector of ones and zeros denoting which
-genes are estimated to be expressed (ones) and unexpressed (zeros). The
-barcode algorithm was implemented by the barcode function in the R fRMA
-package, and the default value of \(C\) was used.
-
-To determine if the expressed ratios differed in the MI group versus the
-healthy control group, Fisher's exact test for individual genes was
-carried out upon the barcode values. Genes with a false discovery rate
-(FDR) of \(< 0.05\), which was calculated through the Benjamini-Hochberg
-(BH) procedure to adjust for multiple testing issues, were considered as
-differentially expressed genes. The same procedures were conducted on
-CAD versus healthy controls as well as MI versus CAD group to find the
-DEGs between them.
+In the barcode algorithm, the normalized intensity of gene \(i\) in
+sample \(j\) is denoted as \(x_{ij}\). A user-defined parameter, \(C\),
+was introduced along with the standard deviation (\(\sigma^{ne}\)) and
+mean (\(\mu^{ne}\)) of the non-expressed distribution. Based on these
+values, the barcode representation of a sample was generated as a vector
+consisting of ones and zeros, representing the estimated expression
+(ones) and non-expression (zeros) of each gene. The barcode function
+within the R fRMA package was employed to implement the barcode
+algorithm, utilizing the default value of \(C\).
+
+To assess the differences in expressed ratios between the MI and healthy
+control groups, Fisher's exact test was performed on the barcode values
+of individual genes. Genes that exhibited a false discovery rate (FDR)
+below \(0.05\), calculated using the Benjamini-Hochberg procedure to
+account for multiple testing issues were identified as differentially
+expressed genes (DEGs). The same procedures were applied to the CAD
+versus healthy control comparison, as well as to the MI versus CAD
+group, to identify DEGs specific to each comparison.
 
@@
 
-Using the R clusterProfiler package, the Kyoto Encyclopedia of Genes and
-Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO)
-functional annotation were carried out on the differentially expressed
-genes. The GO analysis included biological process (BP), cellular
-component (CC) and molecular function (MF) categories. An adjusted
-p-value of less than 0.05 was considered to indicate a statistically
-significant difference. Enrichments were conducted on the MI-healthy and
-CAD-healthy DEGs. In these analyses, all default parameters were used.
+The R clusterProfiler package (Yu et al. 2012) was utilized to perform
+Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment
+analysis and Gene Ontology (GO) functional annotation on the set of
+DEGs. GO analysis encompassed three categories: biological process (BP),
+cellular component (CC), and molecular function (MF). For statistical
+significance, an adjusted p-value threshold of less than 0.05 was
+employed. Enrichment analyses were conducted separately for DEGs
+specific to the MI-healthy and CAD-healthy comparisons. All the default
+parameters provided by the package were used in the analyses.
 
@@
 
-The ML analysis was performed using Python software, ver. 3.9, Numpy
-(Harris et al. 2020), pandas (McKinney 2010), and Scikit-Learn packages
-(Pedregosa et al. 2011). Whenever hyper-tuning was needed, the
-scikit-opt package (Head et al. 2021) was used. In all ML analyses, the
-datasets were divided into train and test sets by a 0.7:0.3 ratio and
-all reported results are the average of 10-fold cross-validation.
-
-Two different approaches were used for selecting miRNAs for model
-training. The first approach was using the miRNAs that are
-differentially expressed. In the second approach, miRNAs with individual
-area under the curve of the receiver operating characteristic (AUC-ROC)
-over 0.8 for separating MI from CAD were selected. Having the result of
-these two different approaches can provide an informative comparison
-between the predictive capabilities of sets of miRNAs selected with
-different logics.
+ML analysis was performed using Python software, ver. 3.9, Numpy (Harris
+et al. 2020), Pandas (McKinney 2010), and Scikit-Learn packages
+(Pedregosa et al. 2011). Whenever hypertuning was needed, the Scikit-opt
+package (Head et al. 2021) was used. In all ML analyses, the datasets
+were divided into training and test sets at a 0.7:0.3 ratio, and all
+reported results are the average of 10-fold cross-validation.
+
+Two different approaches were used to select miRNAs for model training.
+The first approach was to use differentially expressed miRNAs. In the
+second approach, miRNAs with individual AUC-ROCs over 0.8 for separating
+MI from CAD were selected. The results of these two approaches can
+provide an informative comparison between the predictive capabilities of
+sets of miRNAs selected with different logics.
 
@@
 
-In this approach, a two-layer architecture was deployed to the data to
+In this approach, a two-layer architecture is deployed to the data to
 maximize the prediction values. The first layer predicted whether a
-sample is healthy or not, and the second layer separated MI from CAD in
-the samples which were predicted as not healthy in the first layer. To
-this end, a distinct ML model was trained for each layer. Since there is
-a limited number of miRNAs in DEGs, both layers were trained with all of
-them. For further comparison with the models' performance, the ROC curve
-of each miRNA for classifying healthy and not-healthy, as well as CAD
-and MI, were generated using a Logistic Regression model.
-
-\hypertarget{first-layer-for-isolation-of-healthy-and-not-healthy-samples}{%
-\paragraph{First layer for isolation of healthy and not-healthy
-samples:}\label{first-layer-for-isolation-of-healthy-and-not-healthy-samples}}
-
-An SVM model using RBF kernels was trained and hyper-tuned using all
-miRNAs in DEGs. To handle the severe imbalance in the number of samples
-(51 for the healthy and 157 for the not-healthy groups), the sample
-weight for the healthy and the not-healthy samples were set to 1 and
-0.5, respectively.
-
-\hypertarget{second-layer-for-seperating-mi-and-cad-samples}{%
-\paragraph{Second layer for seperating MI and CAD
-samples:}\label{second-layer-for-seperating-mi-and-cad-samples}}
-
-For the sake of reaching the highest classification performance using
-the set of miRNAs, different models were investigated. To do so, SVM
-(with linear, polynomial, and RBF kernels), Logistic Regression (LR),
-Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB),
-XGBoost (XGB) and Decision Tree (DT) models were trained. All models
-were trained with their pre-set parameters with 10-fold
-cross-validation. The criteria for choosing the best model were the
-highest accuracy and AUC-ROC on the test set. The best model was
-hyper-tuned with the scikit-opt package (Head et al. 2021) to get the
-best predictive performance.
+sample was healthy or not, and the second layer separated MI from CAD in
+the samples that were predicted as not healthy in the first layer. To
+this end, a distinct ML model was trained for each layer. Because, there
+were a limited number of miRNAs in the DEGs, both layers were trained
+with all of them. For further comparison with the models' performance,
+the ROC curve of each miRNA for classifying healthy and not-healthy, as
+well as CAD and MI, was generated using a logistic regression model.
+
+\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples}{%
+\paragraph{First layer for the isolation of healthy and not-healthy
+samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples}}
+
+A support vector machine (SVM) model using RBF kernels was trained and
+hypertuned using all miRNAs in the DEGs. To handle the severe imbalance
+in the number of samples (51 for the healthy group and 157 for the
+not-healthy group), the sample weights for the healthy and the
+not-healthy samples were set to 1 and 0.5, respectively. The ROC curve
+and confusion matrix for the model are reported.
+
+\hypertarget{second-layer-for-separating-the-mi-and-cad-samples}{%
+\paragraph{Second layer for separating the MI and CAD
+samples:}\label{second-layer-for-separating-the-mi-and-cad-samples}}
+
+Different models were investigated to achieve the highest classification
+performance. To do so, SVM (with linear, polynomial, and RBF kernels),
+logistic regression (LR), random forests (RF), k-nearest neighbor (kNN),
+gradient boosting (GB), XGBoost (XGB) and decision tree (DT) models were
+trained. All models were trained with their preset parameters using
+10-fold cross-validation. The criteria for selecting the best model were
+the highest accuracy and AUC-ROC for the test set. The best model was
+hypertuned using the scikit-opt package (Head et al. 2021) for the best
+classification performance. The ROC curve and confusion matrix for the
+best model are reported.
 
@@
 
-Like the previous approach, a two-layers strategy was conducted. The
+As in the previous approach, a two-layer strategy was employed. The
 first layer classified samples into healthy and not-healthy, and the
-separated MI and CAD samples. However, to keep the number of miRNAs as
-low as possible miRNAs were selected from the second layer, and then
-their performance wase evaluated in the first layer. AUC-ROC of all
-miRNAs for classifying MI and CAD samples were calculated and the miRNAs
-with the AUC-ROC over 0.8 were selected. The ROC curves for each
-selected miRNA for separating healthy samples from not-healthy and MI
-from CAD samples were also plotted for further comparison.
-
-\hypertarget{first-layer-for-the-detection-of-healthy-and-not-healthy-samples}{%
-\paragraph{First layer for the detection of healthy and not-healthy
-samples:}\label{first-layer-for-the-detection-of-healthy-and-not-healthy-samples}}
-
-An SVM model with an RBF kernel was trained using the selected set of
-miRNAs. Additionally, the model was hyper-tuned to find the
-hyper-parameters for the highest AUC-ROC and accuracy. The ROC curve and
-confusion matrix for the best model were reported.
-
-\hypertarget{second-layer-for-separating-mi-and-cad}{%
-\paragraph{Second layer for separating MI and
-CAD:}\label{second-layer-for-separating-mi-and-cad}}
-
-The selected miRNAs set was used to train different algorithms to find
-the best model. Similar to the previous approach, SVM (with linear,
-polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT were trained.
-All models were trained with their pre-set parameters using 10-fold
-cross-validation. The models with the highest AUC-ROC and accuracy on
-the test set were selected and hyper-tuned using the scikit-opt package
-(Head et al. 2021). The ROC curve and confusion matrix for the best
-model were reported.
+second layer separated the MI and CAD samples. However, to keep the
+number of miRNAs as low as possible, miRNAs were selected from the
+second layer and their performance was evaluated in the first layer. The
+AUC-ROC of all miRNAs for classifying MI and CAD samples was calculated,
+and miRNAs with AUC-ROC \(> 0.8\) were selected. ROC curves for each
+selected miRNA for separating healthy samples from not-healthy samples
+and MI from CAD samples were also plotted for further comparison.
+
+\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-1}{%
+\paragraph{First layer for the isolation of healthy and not-healthy
+samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-1}}
+
+An SVM model with an RBF kernel is trained using the selected set of
+miRNAs. Additionally, the model was hypertuned to find the
+hyperparameters for the highest AUC-ROC and accuracy. The same sample
+weights as in the previous approach (1 for healthy and 0.5 for
+not-healthy samples) were used. The ROC curve and confusion matrix for
+the model were reported.
+
+\hypertarget{second-layer-for-separating-the-mi-and-cad-samples-1}{%
+\paragraph{Second layer for separating the MI and CAD
+samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-1}}
+
+The selected miRNA set was used to train different algorithms to
+determine the best model. Similar to the previous approach, the SVM
+(with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT
+models were trained. All models were trained with their preset
+parameters using 10-fold cross-validation. The models with the highest
+AUC-ROC and accuracy on the test set were selected and hypertuned using
+the scikit-opt package (Head et al. 2021). The ROC curve and confusion
+matrix for the best model were reported.
 
@@
 
-\hypertarget{pre-processing-1}{%
-\subsection{Pre-processing}\label{pre-processing-1}}
-
-The PCA plots of the training samples are shown in Figure @ref(fig:PCA)A
-and B. As shown, healthy samples were separated from CAD or MI samples
-in primary data and also after conducting fRMA. In the RLE plot, there
-was a distinct difference between dataset means for all samples before
-conducting fRMA (Figure @ref(fig:PCA)C). All datasets were rearranged
-around 0 in the RLE plot after conducting fRMA (Figure @ref(fig:PCA)D).
-Moreover, there was a clear change in inter-quantile distances, but the
-values still were over 0.1.
+\hypertarget{preprocessing-1}{%
+\subsection{Preprocessing}\label{preprocessing-1}}
+
+The PCA plots of the samples are shown in Figures @ref(fig:PCA)A and B.
+Healthy samples were separated from the CAD or MI samples in the primary
+data and after conducting fRMA. In the RLE plot, there was a distinct
+difference between the dataset means for all samples before fRMA was
+performed (Figure @ref(fig:PCA)C). All datasets were rearranged to
+approximately 0 in the RLE plot after fRMA was conducted (Figure
+@ref(fig:PCA)D). Moreover, there was an apparent change in the
+interquantile distances, but the values were still greater than 0.1.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../../DEG and DEM/Pics/PCA and RLE of trains} \caption{PCA and RLE plot for all samples before and after fRMA.}\label{fig:PCA}
+\includegraphics[width=1\linewidth]{../../DEG and DEM/Pics/PCA and RLE of trains} \caption{Principal component analysis plots for (A) primary data and (B) the data after fRMA, and the relative log expression plots for (C) primary data and (D) the data after fRMA.}\label{fig:PCA}
 \end{figure}
@@
 According to the cutoff criterion of \(FDR < 0.05\), there were 860 DEGs
-between the MI and the healthy samples. Among them, 323 were
-up-regulated, and 537 were down-regulated in MI compared to the healthy
-controls. In CAD and healthy groups comparison, we found 670 DEGs, of
-which 262 and and 408 DEGs were up- and down-regulated, respectively in
-CAD samples. In the MI and CAD groups, the number of DEGs was 260, and
-the number of up- and down-regulated genes in MI samples were 144 and
-116, respectively in comparison with CAD samples. These data are
-summarized in Table @ref(tab:DEGstab).
+between MI and healthy samples. Among them, 323 were up-regulated, and
+537 were down-regulated in the MI group compared to the healthy group.
+In the CAD and healthy group comparison, we found 670 DEGs, of which 262
+and 408 DEGs were up- and down-regulated, respectively, in CAD samples.
+In the MI and CAD groups, the number of DEGs was 260, and the numbers of
+up- and down-regulated genes in MI samples were 144 and 116,
+respectively, compared to CAD samples. The data are summarized in Table
+@ref(tab:DEGstab).
 
@@
 
-{\centering \includegraphics[width=0.45\linewidth]{../../DEG and DEM/Pics/venny}
+{\centering \includegraphics[width=0.45\linewidth]{Complete-V1_files/figure-latex/venn-1}
 
@@
 
-\caption{Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison.}\label{fig:venn}
+\caption{Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD.}\label{fig:venn}
 \end{figure}
 
-The Venn diagram in Figure @ref(fig:venn) shows that CAD and MI samples
-share a majority of their DEGs. From 860 DEGs of MI/healthy and 670 DEGs
-of CAD/healthy, 531 genes were common which is 62\% of MI/healthy DEGs
-and 79\% of CAD/healthy DEGs.
-
-\hypertarget{gene-ontology-go-and-kyoto-encyclopedia-of-genes-and-genomes-kegg-enrichment-analyses-of-the-degs.}{%
-\subsection{Gene ontology (GO) and Kyoto Encyclopedia of Genes and
-Genomes (KEGG) enrichment analyses of the
-DEGs.}\label{gene-ontology-go-and-kyoto-encyclopedia-of-genes-and-genomes-kegg-enrichment-analyses-of-the-degs.}}
+The Venn diagram in Figure @ref(fig:venn) shows that the CAD and MI
+samples shared most of their DEGs. From 860 DEGs of MI/healthy and 670
+DEGs of CAD/healthy, 531 genes were common, which is 62\% of MI/healthy
+DEGs and 79\% of CAD/healthy DEGs.
+
+\hypertarget{go-and-kegg-enrichment-analyses-of-the-degs}{%
+\subsection{GO and KEGG enrichment analyses of the
+DEGs}\label{go-and-kegg-enrichment-analyses-of-the-degs}}
 
 To explore the biological classification of the DEGs, we performed GO
-and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
-In MI versus healthy samples, GO enrichment analysis in the BP category,
+and KEGG pathway enrichment analyses on the MI/healthy and CAD/healthy
+DEGs. For MI/healthy, GO enrichment analysis in the BP category
 suggested that the DEGs were enriched in ``immune response-regulating
-signaling pathway'', ``lymphocyte differentiation'', ``immune
-response-regulating cell surface receptor signaling pathway'', and
+signaling pathway,'' ``lymphocyte differentiation,'' ``immune
+response-regulating cell surface receptor signaling pathway,'' and
 ``leukocyte activation involved in immune response'' (Figure
-@ref(fig:MIHEnrich)A). In the CC category, the DEGs were enriched in
-``secretory granule membrane'', ``azurophil granule'', ``ficolin-1-rich
-granule'', ``tertiary granule'', and ``ficolin-1-rich granule membrane''
-(Figure @ref(fig:MIHEnrich)B). In the MF category, the DEGs were
-involved in ``cadherin binding'' and ``MHC class I protein binding''
-(Figure @ref(fig:MIHEnrich)C). KEGG pathway analysis indicated that the
-DEGs were related to the following pathways: ``Chemokine signaling
-pathway'', ``Lipid and atherosclerosis'', and ``Hematopoietic cell
-lineage'' (Figure @ref(fig:MIHEnrich)D).
+@ref(fig:MIHEnrich)A). In the CC category, DEGs were enriched in
+``secretory granule membrane,'' ``azurophil granule,'' ``ficolin-1-rich
+granule,'' ``tertiary granule,'' and ``ficolin-1-rich granule membrane''
+(Figure @ref(fig:MIHEnrich)B). In the MF category, DEGs were involved in
+``cadherin binding'' and ``MHC class I protein binding'' (Figure
+@ref(fig:MIHEnrich)C). KEGG pathway analysis indicated that the DEGs
+were related to the following pathways: ``Chemokine signaling pathway,''
+``Lipid and atherosclerosis,'' and ``Hematopoietic cell lineage''
+(Figure @ref(fig:MIHEnrich)D).
 
@@
 
-The enrichment results for DEGs of CAD versus healthy samples were as
-follows. In the BP category, GO enrichment suggested that the DEGs were
-enriched in ``positive regulation of defense response'', ``positive
-regulation of innate immune response'', ``mononuclear cell
-differentiation'', and ``positive regulation of response to external
-stimulus'' (Figure @ref(fig:CADHEnrich)A). In the CC category, the DEGs
-were enriched in ``azurophil granule'', ``ficolin-1-rich granule'', and
-``ficolin-1-rich granule membrane'' (Figure @ref(fig:CADHEnrich)B). In
-the MF category, the DEGs were involved in ``lipoprotein particle
-receptor binding'' and ``NF-\(\kappa\)B binding'' (Figure
-@ref(fig:CADHEnrich)C). KEGG pathway analysis indicated that the DEGs
-were related to the following pathways: ``Chemokine signaling pathway'',
-``Lipid and atherosclerosis'', and ``Hematopoietic cell lineage''
-(Figure @ref(fig:CADHEnrich)D).
+The enrichment results for the CAD/healthy DEGs were as follows. In the
+BP category, GO enrichment suggested that the DEGs were enriched in
+``positive regulation of defense response,'' ``positive regulation of
+innate immune response,'' ``mononuclear cell differentiation,'' and
+``positive regulation of response to external stimulus'' (Figure
+@ref(fig:CADHEnrich)A). In the CC category, DEGs were enriched in
+``azurophil granule,'' ``ficolin-1-rich granule,'' and ``ficolin-1-rich
+granule membrane'' (Figure @ref(fig:CADHEnrich)B). In the MF category,
+DEGs were involved in ``lipoprotein particle receptor binding'' and
+``NF-\(\kappa\)B binding'' (Figure @ref(fig:CADHEnrich)C). KEGG pathway
+analysis showed that the DEGs were related to the following pathways:
+``Chemokine signaling pathway,'' ``Lipid and atherosclerosis,'' and
+``Hematopoietic cell lineage'' (Figure @ref(fig:CADHEnrich)D).
 
@@
 
-Among all DEGs, just miR-186, miR-32, and miR-21 were detected as
-differentially expressed miRNAs. The expression profile of these three
-miRNAs is presented in Figure @ref(fig:Expall). Additionally, The ROC
-curves of each miRNA for each layer are presented in Figure
-@ref(fig:miRROC). Using the logistic regression model the AUC-ROC of
-miR-21, miR-32, and miR-186 for separating healthy and not-healthy
-samples was 0.98, 0.99, and 0.90 respectively (Figure
-@ref(fig:miRROC)A). Besides, the accuracy of each miRNA for classifying
-the samples into healthy and not-healthy groups on the test set was
-0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
-Moreover, the ROC curve of each miRNA for classifying MI and CAD samples
-was presented in Figure @ref(fig:miRROC)B. For miR-21, miR-32, and
-miR-186, the AUC-ROC and accuracy on the test set were 0.85; 0.70; and
-0.86, and 0.78; 0.67; and 0.74, respectively.
+Among the DEGs, miR-186, miR-32, and miR-21 were identified as
+differentially expressed miRNAs. The expression profiles of the three
+miRNAs are shown in Figure @ref(fig:Expall). The ROC curves of each
+miRNA in each layer are presented in Figure @ref(fig:miRROC). Using the
+logistic regression model, the AUC-ROC values of miR-21, miR-32, and
+miR-186 for separating healthy and not-healthy samples were 0.98, 0.99,
+and 0.90, respectively (Figure @ref(fig:miRROC)A). The accuracy of each
+miRNA for classifying the samples into healthy and not-healthy groups on
+the test set for miR-21, miR-32, and miR-186 was 0.92, 0.98, and 0.89,
+respectively. The ROC curve of each miRNA for classifying MI and CAD
+samples is presented in Figure @ref(fig:miRROC)B. The AUC-ROC and
+accuracy for miR-21, miR-32, and miR-186 in the test set were 0.85;
+0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
 
@@
 
-\caption{\label{tab:mirExptable}Target miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.}
+\caption{\label{tab:mirExptable}Investigated miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.}
 \centering
@@
 
-\caption{Expression profile of all miRNAs in two approach in different sample classes.}\label{fig:Expall}
+\caption{Expression profile of all miRNAs in two approaches in different sample classes.}\label{fig:Expall}
 \end{figure}
@@
 
-\hypertarget{first-layer-for-healthy-not-healthy-isolation}{%
-\paragraph{First layer for healthy not-healthy
-isolation:}\label{first-layer-for-healthy-not-healthy-isolation}}
-
-Although single miRNAs had acceptable performance, their predictive
-value could be improved even further by using them as a set. The ROC
-curve for the SVM model with an RBF kernel trained with all three miRNAs
-is presented in Figure @ref(fig:DEMROC)A. The model had a better
-performance in classification than single miRNAs.The AUC-ROC for the
-model is 1, and its accuracy on the test set was also 1. In Figure
+\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-2}{%
+\paragraph{First layer for the isolation of healthy and not-healthy
+samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-2}}
+
+Although single miRNAs had an acceptable performance for this layer,
+their predictive value could be further improved by using them as a set.
+The ROC curve for the SVM model with an RBF kernel trained with all
+three miRNAs is presented in Figure @ref(fig:DEMROC)A. The model had a
+better performance in classification than single miRNAs.The AUC-ROC for
+the model was 1, and its accuracy on the test set was also 1. In Figure
 @ref(fig:DEMCF)A, the confusion matrix for the model is presented.
@@
 
-\caption{ROC curve for miRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples.}\label{fig:DEMROC}
+\caption{ROC curve for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMROC}
 \end{figure}
@@
 
-\caption{Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification.}\label{fig:DEMCF}
+\caption{Confusion matrix for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMCF}
 \end{figure}
 
-\hypertarget{second-layer-for-separating-mi-samples-from-cad}{%
-\paragraph{Second layer for separating MI samples from
-CAD:}\label{second-layer-for-separating-mi-samples-from-cad}}
-
-Different models were trained using expression values for three
-differentially miRNAs. The models' 10-fold cross-validated AUC-ROC and
-accuracy on the test set are reported in Figure @ref(fig:DEMmodels). The
-best model from both AUC-ROC and accuracy point-of-view was the SVM
-model with linear kernel. The AUC-ROC and accuracy for this model with
-its pre-set values were 0.93 and 0.82 respectively. The model was
-hyper-tuned for C and gamma hyper-parameters, and therefore the model
-showed better performance. The ROC curve of the hyper-tuned model is
-presented in Figure @ref(fig:DEMROC)B. For this model the AUC-ROC
-reached 0.95 and the accuracy improved to 0.85 (Table @ref(tab:DEGsML)).
-Moreover, the sensitivity and specificity for the model on the test set
-were 0.91 and 0.71 respectively. The confusion matrix for the
-hyper-tuned model is illustrated in Figure @ref(fig:DEMCF)B.
+\hypertarget{second-layer-for-separating-the-mi-and-cad-samples-2}{%
+\paragraph{Second layer for separating the MI and CAD
+samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-2}}
+
+Different models were trained using the expression values of three
+differentially expressed miRNAs. The models' AUC-ROC and the accuracy of
+the test set are shown in Figure @ref(fig:DEMmodels). The best model
+from both the AUC-ROC and accuracy points of view was the SVM model with
+a linear kernel. The AUC-ROC and accuracy for this model with its preset
+values were 0.93 and 0.82, respectively. The model was hypertuned for C
+and gamma hyperparameters, and therefore the model showed better
+performance. The ROC curve of the hypertuned model is presented in
+Figure @ref(fig:DEMROC)B. For this model, the AUC-ROC reached 0.95, and
+the accuracy was improved to 0.85 (Table @ref(tab:DEGsML)). Moreover,
+the sensitivity and specificity for the model on the test set were 0.91
+and 0.71, respectively. The confusion matrix for the hypertuned model is
+illustrated in Figure @ref(fig:DEMCF)B.
 
@@
 
-\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs on the test set.}\label{fig:DEMmodels}
+\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs.}\label{fig:DEMmodels}
 \end{figure}
@@
 
-\caption{\label{tab:DEGsML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning}
+\caption{\label{tab:DEGsML}AUC-ROC and accuracy for SVM with a linear kernel as the best model trained with differentially expressed miRNAs on the training and test sets before and after hypertuning}
 \centering
@@
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Pre-set parameters} & \multicolumn{2}{c}{Hyper-tuned} \\
+\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Preset parameters} & \multicolumn{2}{c}{Hypertuned} \\
 \cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
@@
 
-After calculating the AUC-ROC for each miRNA for the classification of
-MI and CAD samples, the miRNAs with AUC-ROC over 0.8 were selected. The
-selected miRNAs were miR-29a, miR-197, miR-186, miR-21, and miR-296. The
-expression values of these miRNAs in healthy, CAD, and MI samples are
-presented in Figure @ref(fig:Expall). The ROC curve of the selected
-miRNAs for both layers are illustrated in Figure @ref(fig:miRROC).
-
-\hypertarget{first-layer-for-isolation-of-healthy-and-not-healthy-samples-1}{%
-\paragraph{First layer for isolation of healthy and not-healthy
-samples}\label{first-layer-for-isolation-of-healthy-and-not-healthy-samples-1}}
+After calculating the AUC-ROC for each miRNA to classify of MI and CAD
+samples, the miRNAs with AUC-ROC \(> 0.8\) were selected. The miRNAs
+selected were miR-29a, miR-197, miR-186, miR-21, and miR-296. The
+expression levels of these miRNAs in healthy, CAD, and MI samples are
+presented in Figure @ref(fig:Expall). The ROC curves of the selected
+miRNAs in both layers are shown in Figure @ref(fig:miRROC).
+
+\hypertarget{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-3}{%
+\paragraph{First layer for the isolation of healthy and not-healthy
+samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-3}}
 
 Using the selected set, an SVM model with an RBF kernel was trained to
-separate healthy from not-healthy samples. The ROC curve for the model
-is presented in Figure @ref(fig:AUCROC)A and the confusion matrix is
-illustrated in Figure @ref(fig:AUCCF)A. Both AUC-ROCand accuracy for the
-model on the test set were equal to 1.
+separate healthy and not-healthy samples. The ROC curve for the model is
+presented in Figure @ref(fig:AUCROC)A, and the confusion matrix is
+illustrated in Figure @ref(fig:AUCCF)A. Both the AUC-ROC and accuracy of
+the model on the test set were 1.
 
@@
 
-\caption{ROC curve for the set of miRNAs selected by AUC-ROC on test set classification.
-(A) SVM with RBF kernel for healthy and not-healthy samples classification.
-(B) Logistic regression model for CAD and MI samples classification.
-(C) SVM with polynomial kernel for CAD and MI samples classification.
+\caption{ROC curve for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+(B) Logistic regression model for CAD and MI sample classification.
+(C) SVM with polynomial kernel for CAD and MI sample classification.
 }\label{fig:AUCROC}
@@
 
-\caption{Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification.
-(B) Logistic regression model for CAD and MI samples classification.
-(C) SVM with polynomial kernel for CAD and MI samples classification.}\label{fig:AUCCF}
+\caption{Confusion matrix for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification.
+(B) Logistic regression model for CAD and MI sample classification.
+(C) SVM with polynomial kernel for CAD and MI sample classification.}\label{fig:AUCCF}
 \end{figure}
 
-\hypertarget{second-layer-for-isolation-of-mi-samples-from-cad-samples}{%
-\paragraph{Second layer for isolation of MI samples from CAD
-samples}\label{second-layer-for-isolation-of-mi-samples-from-cad-samples}}
-
-To find the best model for training the best set, different models were
-trained using their pre-set values. Their AUC-ROC and accuracy results
-on the test set are presented in Figure @ref(fig:AUCmodels). The best
-model from the AUC-ROC point-of-view was the SVM with linear kernel and
-from the accuracy point-of-view, it was the SVM model with an RBF
-kernel. For the SVM-linear model the AUC-ROC and accuracy were 0.93 and
+\hypertarget{second-layer-for-separating-the-mi-and-cad-samples-3}{%
+\paragraph{Second layer for separating the MI and CAD
+samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-3}}
+
+To find the best model for this set of miRNAs, different models were
+trained using their preset values. The AUC-ROC and accuracy results for
+the test set are presented in Figure @ref(fig:AUCmodels). The best model
+from the AUC-ROC point of view was the SVM with a linear kernel, and
+from the accuracy point of view, it was the SVM model with an RBF
+kernel. For the SVM-linear model, the AUC-ROC and accuracy were 0.93 and
 0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84,
-respectively. Both models were hyper-tuned and the ROC curve for their
+respectively. Both models were hyper-tuned, and the ROC curve for their
 best performance is presented in Figure @ref(fig:AUCROC)B and C. The
-AUC-ROC and accuracy for the SVM-linear model modified to 0.92 and 0.88,
-respectively. For the SVM-RBF, these values increased to 0.96 and 0.94,
-respectively (Table @ref(tab:AUCML)). The sensitivity for the SVM-linear
-and SVM-RBF models were 0.91 and 0.97, respectively; and the specificity
-for them was 0.79 and 0.86, respectively. The confusion matrix for both
-models is illustrated in Figure @ref(fig:AUCCF)B and C.
+AUC-ROC and accuracy for the SVM-linear model were modified to 0.92 and
+0.88, respectively. For the SVM-RBF, these values increased to 0.96 and
+0.94, respectively (Table @ref(tab:AUCML)). The sensitivities for the
+SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and
+their specificities were 0.79 and 0.86, respectively. The confusion
+matrix for both models is illustrated in Figure @ref(fig:AUCCF)B and C.
 
@@
 
-\caption{\label{tab:AUCML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their individual AUC-ROC on the train and test set before and after hyper-tuning}
+\caption{\label{tab:AUCML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their AUC-ROC on the train and test set before and after hyper-tuning}
 \centering
@@
 
-The prevalence of MI can lead to high-rate mortality in the clinical
-setting. However, early diagnosis and application of suitable treatment
-protocols can reduce mortality and improve AMI prognosis
+The prevalence of MI can lead to high mortality rates in the clinical
+setting. However, early diagnosis and the application of suitable
+treatment protocols can reduce mortality and improve MI prognosis
 ({``Cardiovascular Diseases ({CVDs})''} n.d.; Thygesen et al. 2018; Tsao
@@
 may play a significant role in the progression of MI and the subsequent
-remodeling (Laggerbauer and Engelhardt 2022). It is believed that the
-expression of miRNAs is altered during the various biological processes
-correlated with MI within the myocardium or other related tissues (Khan,
-Gupta, and Mahapatra 2022). Although several research has been
-concentrated on examining free circulating miRNAs in the serum samples
-for the detection of cardiac tissue injuries (Kaur et al. 2020), more
-information is needed to fully comprehend the miRNAs found in different
-blood sub-components like plasma, platelets, and PBMCs. Based on
-previous data, PBMCs are critically involved in plaque destabilization
-and rupture as well as early inflammatory responses during MI (Mosallaei
-et al. 2022; Hapke et al. 2022). Moreover, PBMCs have specific miRNA
-profile that is altered under certain pathological conditions which are
-great candidates as disease biomarkers (Mosallaei et al. 2022).
-
-PBMCs can respond to several insulting conditions such as MI in the
-least possible time with prominent changes in their miRNA profile
-(Mosallaei et al. 2022). Considering the regulatory roles, subtle
+remodeling (Laggerbauer and Engelhardt 2022). It is believed that miRNA
+expression is altered during the various biological processes correlated
+with MI within the myocardium or other related tissues (Khan, Gupta, and
+Mahapatra 2022). Although several studies have focused on examining free
+circulating miRNAs in serum samples for the detection of cardiac tissue
+injuries (Kaur et al. 2020), more information is needed to fully
+comprehend the miRNAs found in different blood subcomponents, such as
+plasma, platelets, and PBMCs. Based on previous findings, PBMCs play a
+crucial role in the destabilization and rupture of plaques as well as in
+the initial inflammatory reactions in individuals experiencing
+myocardial infarction (MI). (Mosallaei et al. 2022; Hapke et al. 2022).
+Moreover, PBMCs have specific miRNA profiles that are altered under
+certain pathological conditions, making them great candidates as disease
+biomarkers (Mosallaei et al. 2022).
+
+PBMCs can respond to several insulting conditions, such as MI, in the
+shortest possible time with notable changes in their miRNA profile
+(Mosallaei et al. 2022). Considering their regulatory roles, subtle
 changes in the transcription of miRNAs can be monitored even before
-alteration in the levels of mRNAs and proteins (Schulte et al. 2020).
-These features make the miRNAs an early-stage valid diagnostic tool for
-the detection of minor and major cell injuries. To date, few studies
-have been performed to compare the miRNA profiles in PBMCs belonging to
-acute MI patients and other CADs and healthy samples to find a robust
-set of identical miRNAs to differentiate these pathological conditions.
-
-In this study, we combined three GEO datasets of healthy, CAD, and MI
-samples. Having these samples set alongside bioinformatics analysis and
-ML means, it is possible to identify potential biomarker sets and also
-effective therapeutic targets. The results of the DEG analysis (Table
-@ref(tab:DEGstab) and Figure @ref(fig:venn)) are proof of the close
+alterations in mRNA and protein levels(Schulte et al. 2020). These
+features make miRNAs a valid early-stage diagnostic tool for the
+detection of minor and major cell injuries. To date, few studies have
+compared the miRNA profiles in PBMCs from patients with MI and other
+CADs and healthy samples to find a robust set of identical miRNAs to
+differentiate these pathological conditions.
+
+In this study, we combined three GEO datasets for healthy, CAD, and MI
+samples. Having these sample sets alongside bioinformatics analysis and
+ML methods enabled us to identify potential biomarker sets and effective
+therapeutic targets. The results of the DEG analysis (Table
+@ref(tab:DEGstab) and Figure @ref(fig:venn)) prove the close
 relationship between the MI and CAD samples. Interestingly, functional
-enrichment analysis demonstrated that DEGs in both healthy/CAD and
-healthy/MI were strongly correlated to immune cell response which is a
-major cellular part of PBMCs. Here, two different sets of miRNAs were
-used as biomarker sets for sample classification. miR-21; miR-32; and
-miR-186 were selected as differentially expressed miRNAs, and miR-186,
-miR-21; miR-29a; miR-197; and miR-296 were selected according to their
-AUC-ROC values. As shown in Figure @ref(fig:miRROC), all miRNAs selected
-with both approaches had AUC-ROC over 0.9 for the isolation of healthy
-and not-healthy samples except for miR-296 and miR-29a. Data confirmed
-that the real challenge is to classify CAD and MI samples because of
-close overlap. Of 6 miRNAs under investigation in both approaches except
-for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of
-CAD and MI samples. Besides, the high AUC-ROC values of miRNAs confirms
-their high potential as biomarkers.
-
-ML models when trained with miRNA sets selected by both DEG and AUC-ROC
-approaches showed better performance in the classification than each
-miRNA. To avoid unwanted complexity and poor predictive values, a
-two-layer architecture was also designed. The first layer was for the
-discrimination of healthy from not-healthy samples, and the second layer
-separated the CAD from MI candidates. As expected, in both approaches a
-hyper-tuned SVM model could flawlessly separates healthy from
-not-healthy samples using distinct miRNAs sets. The ML models were also
+enrichment analysis demonstrated that DEGs in both CAD/healthy and
+MI/healthy were strongly correlated with the immune cell response, which
+is a major part of PBMCs. Two sets of miRNAs were selected as biomarker
+sets for sample classification. miR-21; miR-32; and miR-186 were
+selected as differentially expressed miRNAs, and miR-186; miR-21;
+miR-29a; miR-197; and miR-296 were selected based on their AUC-ROC
+values. As shown in Figure @ref(fig:miRROC), all miRNAs selected with
+both approaches had AUC-ROCs \(> 0.9\) for isolating healthy and
+not-healthy samples except for miR-296 and miR-29a. The data confirmed
+that the real challenge was to classify CAD and MI samples because of
+the close overlap. Of the six miRNAs under investigation in both
+approaches, except for miR-32, all miRNAs had an AUC-ROC \(> 0.8\) for
+the discrimination of CAD and MI samples. As expected, the high AUC-ROC
+values of the miRNAs confirmed their high potential as biomarkers.
+
+ML models trained with miRNA sets selected by both DEG and AUC-ROC
+approaches, showed better classification performance than each miRNA. To
+avoid unwanted complexity and poor predictive values, a two-layer
+architecture was designed. The first layer was used to discriminate
+between healthy and not-healthy samples, and the second layer was was
+used to separate CAD from MI candidates. As expected, in both
+approaches, a hypertuned SVM model could flawlessly separate healthy and
+not-healthy samples using distinct miRNA sets. ML models are also
 capable of effectively separating CAD from MI patients. Although both
-miRNA sets had nearly the same AUC-ROC with their best model, the
+miRNA sets had nearly the same AUC-ROC using the best model, their
 accuracy, sensitivity, and specificity were different. The model trained
-with AUC-selected miRNAs had better performance in all predictive
-values.
-
-Numerous studies have reported different biological processes can affect
-the expression of miRNAs in PBMCs. However, there are still
-controversies regarding the exact role of miRNAs in the function of
-immune cells and the correlation of specific pathological conditions
-with miRNA profiles. Several studies have proved the activation of
-specific miRNA types in PBMCs under cardiovascular events (S. Li et al.
-2015; Yao et al. 2016; Liu et al. 2017; Horita, Farquharson, and Stephen
-2021; Bhansali et al. 2022). For instance, there is evidence that the
-elevation of miR-186 suppresses the expression of
-Cystathionine-\(\gamma\)-lyase, leading to the subsequent secretion of
-pro-inflammatory cytokines and cellular lipid accumulation. Besides,
-macrophage-derived miR-186 may promote atherosclerotic plaques (Yao et
-al. 2016). In line with this claim, we found that miR-186 is
-up-regulated in both CAD and MI candidates related to control
+with AUC-selected miRNAs showed better performance in all predictive
+values, which is logical because of the higher number of miRNAs in the
+set.
+
+Numerous studies have reported that different biological processes can
+affect the miRNA expression in PBMCs. However, the exact role of miRNAs
+in the function of immune cells and the correlation between specific
+pathological conditions and miRNA profiles remain controversial. Several
+studies have proven the activation of particular miRNA types in PBMCs
+under cardiovascular events H. Li et al. (2018). For instance, there is
+evidence that elevation of miR-186 suppresses the expression of
+cystathionine-\(\gamma\)-lyase, leading to the subsequent secretion of
+pro-inflammatory cytokines and cellular lipid accumulation. In addition,
+macrophage-derived miR-186 may promote atherosclerotic plaque formation
+(Yao et al. 2016). In line with this claim, we found that miR-186 was
+up-regulated in both CAD and MI candidates compared to their control
 counterparts. Surprisingly, the obtained data indicated that the
-expression of miR-186 is higher in CAD patients in comparison to MI
-(Figure @ref(fig:DEMexp)), To be specific, miR-186 is the only
+expression of miR-186 was higher in patients with CAD than in patients
+with MI (Figure @ref(fig:Expall)). Specifically, miR-186 was the only
 differentially expressed miRNA between CAD and MI, with a clear
@@
 As mentioned before, miR-21 was also up-regulated in both MI and CAD
-patients in comparison to healthy controls. Moreover, the expression
-value of miR-21 was significantly higher in MI than that of the CAD
-group (Table @ref(tab:mirExptable)). It is thought that the
-up-regulation of miRNA-21 in PBMCs is a compensatory reaction to reduce
-T\(_{reg}\) lymphocyte number in response to the reduction of
-TGF\(\beta1\) secretion into the plasma through a
-TGF\(\beta1\)/smad-independent pathway. In line with previous and
-present data, miR-21 can modulate the activity of PBMCs following the
-occurrence of cardiovascular diseases (S. Li et al. 2015).
-
-Recent data have supported the elevation of miR-32 in CAD patients with
-the calcification of coronary artery. Of note, miR-32 stimulates the
-calcification of mouse vascular smooth muscle through the regulation of
-bone morphogenetic protein-1, runt-related transcription factor-2
-(RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA
-protein (Liu et al. 2017). Likewise, there are some reports associated
-with the activity of PBMC miR-32 under several pathologies (Zeng et al.
-2021; Wang et al. 2020). The exact role of PBMC miR-32 after
-cardiovascular events remained to be elucidated.
+patients compared to healthy controls. Moreover, the expression value of
+miR-21 was significantly higher in the MI group than in the CAD group
+(Table @ref(tab:mirExptable)). It is thought that the up-regulation of
+miR-21 in PBMCs is a compensatory reaction to reduce the T\(_{reg}\)
+lymphocyte number in response to the reduction in TGF\(\beta1\)
+secretion into the plasma through a TGF\(\beta1\)/smad-independent
+pathway. In line with the previous and present data, miR-21 can modulate
+the activity of PBMCs following the occurrence of cardiovascular
+diseases (S. Li et al. 2015).
+
+Recent data have supported the elevation of miR-32 levels in CAD samples
+with calcification of the coronary artery. Notably, miR-32 promotes
+vascular smooth muscle calcification in mice by controlling the activity
+of several proteins, including bone morphogenetic protein-1,
+runt-related transcription factor-2 (RUNX2), osteopontin, and
+bone-specific phosphoprotein matrix GLA protein. (Liu et al. 2017).
+Likewise, some reports are associated with the activity of miR-32 in
+PBMCs in several pathologies (Zeng et al. 2021; Wang et al. 2020). The
+exact role of miR-32 in PBMCs after cardiovascular events remains
+unclear.
 
 Molecular analyses have indicated the regulatory role of miRNAs selected
-by the AUC-ROC approach in PBMCs after a cardiovascular event. Two
-common miRNAs in DEGs and AUC-ROC approachs, miR-21 and miR-186, were
-covered in DEGs miRNA set. Based on numerous reports mir-29a can be
-activated in different diseases (Horita, Farquharson, and Stephen
-2021).Data analysis indicated that miR-29a is significantly up-regulated
-in CAD patients in comparison to healthy and MI groups (Table
-@ref(tab:mirExptable)). Increased miR-29a is associated with the
-progression of atherosclerosis, and the combination of miR-29a and
-ox-LDL was offered as a valid biomarker set for paraclinical
-classification (Huang et al. 2016). However, the role of miR-29a in the
-function of PBMCs in CAD patients has not been completely examined.
-
-Data indicated that miR-197 is also significantly up-regulated in both
-CAD/healthy and MI/healthy groups. In previous studies, it has been
-shown that miRNA-197 may play a critical role in regulating the
-anti-inflammatory response of IL-35 by affecting pro/ anti-inflammatory
-cytokine secretion, M1/M2 macrophage ratio, T\(_{reg}\) lymphocyte
-proliferation, and T cell suppression suggesting the potential
-diagnostic role of miR-197 in adverse cardiovascular events (Bhansali et
-al. 2022).
+using the AUC-ROC approach in PBMCs after a cardiovascular event. The
+biological importance of two common miRNAs in the DEGs and AUC-ROC
+approaches, miR-21 and miR-186, have already been discussed. Based on
+numerous reports, miR-29a can be activated in different diseases
+(Horita, Farquharson, and Stephen 2021). Data analysis indicated that
+miR-29a was significantly up-regulated in CAD patients compared to the
+healthy and MI groups (Table @ref(tab:mirExptable)). Increased miR-29a
+is associated with the progression of atherosclerosis, and the
+combination of miR-29a and ox-LDL has been suggested as a valid
+biomarker set for paraclinical classification (Huang et al. 2016).
+However, the role of miR-29a in the function of PBMCs from patients with
+CAD has not been thoroughly examined.
+
+The data indicated that miR-197 was significantly up-regulated in both
+the CAD/healthy and MI/healthy groups. Previous studies have
+demonstrated that miR-197 may play a crucial role in controlling the
+anti-inflammatory response of IL-35 by influencing the secretion of
+cytokines that can either promote or suppress inflammation, the ratio of
+M1/M2 macrophages, and the proliferation of T\(_{reg}\) lymphocytes,
+which are responsible for suppressing immune responses (Bhansali et al.
+2022). Alongside our findings, it can be concluded that miR-197 could be
+a useful diagnostic tool for predicting adverse cardiovascular events.
 
 The findings of this study demonstrate the potential of miR-296 as a
-biomarker with high discriminatory power for distinguishing between
-samples from individuals with MI and CAD. MiR-296 has been identified as
-a key regulator in the development and advancement of atherosclerosis by
+biomarker with high discriminatory power to distinguish between samples
+from individuals with MI and CAD. MiR-296 has been identified as a key
+regulator in the development and advancement of atherosclerosis by
 controlling the expression of target genes associated with various
-biological processes including angiogenesis, inflammation, cholesterol
-metabolism, hypertension, cellular proliferation, and apoptosis (H. Li
-et al. 2018). In a study it has been shown that miR-296 expression
-levels are significantly increased in the PBMCs of CAD patients compared
-to healthy controls, suggesting its involvement in regulating
-pro-inflammatory cytokines like IL-6 and TNF-a (Fard et al. 2020). These
-findings suggest that miR-296 may have a significant impact on the
-pathogenesis of atherosclerosis and could potentially serve as a
-diagnostic biomarker for CAD or MI.
+biological processes, including angiogenesis, cholesterol metabolism,
+inflammation, cellular proliferation, hypertension, and apoptosis (H. Li
+et al. 2018). In a previous study, miR-296 expression levels were found
+to be significantly increased in the PBMCs of CAD patients compared to
+healthy controls, suggesting its involvement in regulating
+proinflammatory cytokines such as IL-6 and TNF-\(\alpha\) (Fard et al.
+2020). These findings suggested that miR-296 may have a significant
+impact on the pathogenesis of atherosclerosis and could potentially
+serve as a diagnostic biomarker for CAD or MI.
 
@@
 In summary, we derived a set of miRNA biomarkers by comparing MI samples
-to both healthy and CAD samples. We found that the SVM model performed
-best in both the first layer, which separated healthy and not-healthy
-samples, and the second layer, which classified MI/CAD samples. The set
-of miRNAs selected based on their AUC-ROC values had better performance
-in the second layer. Overall, our two-layer structure achieved a
-accuracy of 0.96. This demonstrates the potential for combining
-bioinformatics and machine learning techniques to identify novel
-biomarkers and gain a deeper understanding of myocardial infarction.
+with both healthy and CAD samples. We found that the SVM model performed
+best in both the first layer, which separated healthy and unhealthy
+samples, and the second layer, which classified the MI/CAD samples. The
+set of miRNAs selected based on their AUC-ROC values performed better in
+the second layer. Overall, our two-layer structure achieved an accuracy
+of 0.96. This demonstrates the potential of combining bioinformatics and
+machine learning techniques to identify novel biomarkers and gain a
+deeper understanding of myocardial infarction.
 
 \hypertarget{references}{%
-\section{References}\label{references}}
-
-\begin{align}
-a^2+b^2=c^2
-\end{align}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
+R Core Team. 2022. \emph{R: A Language and Environment for Statistical
+Computing}. Vienna, Austria: R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-ML}{}}%
@@
 \url{https://doi.org/10.1016/j.biotechadv.2021.107739}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-RStudio}{}}%
+RStudio Team. 2020. \emph{RStudio: Integrated Development Environment
+for r}. Boston, MA: RStudio, PBC. \url{http://www.rstudio.com/}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ggplot}{}}%
+Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
+Analysis}. Springer-Verlag New York.
+\url{https://ggplot2.tidyverse.org}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-186}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cluster}{}}%
+Yu, Guangchuang, Li-Gen Wang, Yanyan Han, and Qing-Yu He. 2012.
+{``clusterProfiler: An r Package for Comparing Biological Themes Among
+Gene Clusters.''} \emph{OMICS: A Journal of Integrative Biology} 16 (5):
+284--87. \url{https://doi.org/10.1089/omi.2011.0118}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-magic}{}}%
@@
 
-\bibliographystyle{spmpsci}
-\bibliography{ForArticleRefrence.bib,discussion.bib,Introduction.bib,mir296.bib}
+\bibliographystyle{spbasic}
+\bibliography{For Article
+Refrence.bib,discussion.bib,Introduction.bib,mir296.bib}
 

### 2023-08-07 Revised Version/Art/sn-article.tex
The version after revision

%Version 2.1 April 2023
% See section 11 of the User Manual for version history
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing

%%=======================================================%%
%% to print line numbers in the margin use lineno option %%
%%=======================================================%%

%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%======================================================%%
%% to compile with pdflatex/xelatex use pdflatex option %%
%%======================================================%%

%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%Note: the following reference styles support Namedate and Numbered referencing. By default the style follows the most common style. To switch between the options you can add or remove ?Numbered? in the optional parenthesis. 
%%The option is available for: sn-basic.bst, sn-vancouver.bst, sn-chicago.bst, sn-mathphys.bst. %

%%\documentclass[sn-nature]{sn-jnl}% Style for submissions to Nature Portfolio journals
%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
\documentclass[sn-mathphys,Numbered]{sn-jnl}% Math and Physical Sciences Reference Style
%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
%%\documentclass[sn-vancouver,Numbered]{sn-jnl}% Vancouver Reference Style
%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style 
%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
%%\documentclass[default]{sn-jnl}% Default
%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
\usepackage{tabularx}%
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

%\jyear{2021}%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

\begin{document}

\title[Article Title]{Unlocking the Potential of microRNAs: Machine Learning Identifies Key Biomarkers for Myocardial Infarction Diagnosis}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author[1]{\fnm{Mehrdad} \sur{Samadishadlou}}

\author[2, 3]{\fnm{Reza} \sur{Rahbarghazi}}

\author[4]{\fnm{Zeynab} \sur{Piryaei}}

\author[5]{\fnm{Mahdad} \sur{Esmaeili}}

\author[6]{\fnm{Çığır} \sur{Biray Avcı}}

\author[1*]{\fnm{Farhad} \sur{Bani}}\email{bainf@tbzmed.ac.ir}

\author[4*]{\fnm{Kaveh} \sur{Kavousi}}\email{kkavousi@ut.ac.ir}

\affil[1]{\orgdiv{Department of Medical Nanotechnology}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}

\affil[2]{\orgdiv{Stem Cell Research Center}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}

\affil[3]{\orgdiv{Department of Applied Cell Sciences}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}

\affil[4]{\orgdiv{Laboratory of Complex Biological Systems and Bioinformatics (CBB), Department of Bioinformatics}, \orgname{Institute of Biochemistry and Biophysics (IBB)}, \orgaddress{\street{University of Tehran}, \city{Tehran}, \country{Iran}}}

\affil[5]{\orgdiv{Medical Bioengineering Department}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}

\affil[6]{\orgdiv{Medical Biology Department}, \orgname{School of Medicine}, \orgaddress{\street{Ege University}, \city{İzmir}, \country{Türkiye}}}

%%================================%%
%% Sample for structured abstract %%
%%================================%%

\abstract{\textbf{Background:} MicroRNAs (miRNAs) play a crucial role in regulating adaptive and
maladaptive responses in cardiovascular diseases, making them attractive
targets for potential biomarkers. However, their potential as novel
biomarkers for diagnosing cardiovascular diseases requires systematic
evaluation.

\textbf{Methods:} In this study, we aimed to identify a key set of miRNA
biomarkers using integrated bioinformatics and machine learning
analysis. We combined and analyzed three gene expression datasets from
the Gene Expression Omnibus (GEO) database, which contains peripheral
blood mononuclear cell (PBMC) samples from individuals with myocardial
infarction (MI), stable coronary artery disease (CAD), and healthy
individuals. Additionally, we selected a set of miRNAs based on their
area under the receiver operating characteristic curve (AUC-ROC) for
separating the CAD and MI samples. We designed a two-layer architecture
for sample classification, in which the first layer isolates healthy
samples from unhealthy samples, and the second layer classifies stable
CAD and MI samples. We trained different machine learning models using
both biomarker sets and evaluated their performance on a test set.

\textbf{Results:} We identified hsa-miR-21-3p, hsa-miR-186-5p, and hsa-miR-32-3p as the
differentially expressed miRNAs, and a set including hsa-miR-186-5p, hsa-miR-21-3p,
hsa-miR-197-5p, hsa-miR-29a-5p, and hsa-miR-296-5p as the optimum set of miRNAs selected by
their AUC-ROC. Both biomarker sets could distinguish healthy from
not-healthy samples with complete accuracy. The best performance for the
classification of CAD and MI was achieved with an SVM model trained
using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.96 and
an accuracy of 0.94 on the test data.

\textbf{Conclusions:} Our study demonstrated that miRNA
signatures derived from PBMCs could serve as valuable novel biomarkers
for cardiovascular diseases.}

\keywords{microRNA, Machine Learning, Myocardial Infarction, Bioinformatics, Biomarker}

%%\pacs[JEL Classification]{D8, H51}

%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

\maketitle

\section{Introduction}\label{introduction}

Cardiovascular diseases (CVDs) are the leading cause of human mortality,
accounting for 32\% of all global deaths. It is estimated that
approximately 85\% of CVD mortality is due to myocardial infarction (MI) \cite{CVD}. MI is an acute coronary
syndrome characterized by sudden blockage and stenosis of the coronary
artery and subsequent myocardial ischemia, leading to extensive
cardiomyocyte damage and necrosis \cite{Macro}.

Over the last 50 years, numerous attempts have been made to use
biomarkers to facilitate diagnosis, assess the risk, follow-up therapy, and determine therapeutic efficacy in CVD candidates. Based on released
guidelines, cardiac troponins (cTns) are used as a highly sensitive and
accurate approach for detecting MI. Despite these inherent advantages,
the high sensitivity of cTn-based assays has also led to more
false-positive results \cite{17}, necessitating the advent
and development of new modalities with pathological value. To improve
the diagnostic value of existing MI biomarkers, a combination of
complementary biological markers, such as microRNAs (miRNAs) and other
genetic factors, has been proposed. Previous research supports the
notion that miRNAs exhibit great potential as alternative biomarkers for
CVD detection and follow-up \cite{miR}. It has been suggested
that miRNAs possess 18-22 nucleotides and play a crucial role in the
regulation of gene expression. Evidence indicates that miRNAs are
involved in the pathogenesis of cardiac tissue injury \cite{m1}. Several biological processes, such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque
formation, and cardiac rhythm, are regulated by miRNAs \cite{m2}. 
% Change S
Circulating and tissue-specific miRNAs have shown promise as diagnostic and prognostic biomarkers across a range of cardiovascular diseases, including MI and other conditions such as CAD, heart failure, atrial fibrillation, cardiac hypertrophy, and fibrosis \cite{11-15, new2}. The use of miRNAs as diagnostic and prognostic biomarkers in CVDs is supported by their stability and rapid release into circulation after myocardial injury \cite{11-15}. In CAD, altered expression of miRNAs like miR-1, miR-133a, miR-208a/b, and miR-499, which are abundantly expressed in the heart, has been reported in patients compared to healthy controls. Additional miRNAs including miR-21, miR-208a/b, miR-133a/b, and the miR-30 family are frequently dysregulated in acute coronary syndrome (ACS) versus stable CAD  \cite{new3}. Furthermore, miRNAs like miR-3113-5p, miR-223-3p, miR-499a-5p, and miR-133a-3p demonstrate potential as biomarkers to identify patients at risk of sudden cardiac death \cite{new4}. Moreover, miRNAs have shown diagnostic potential in other CVDs. For instance, miR-21 has been associated with cardiac injury and has been implicated in the pathology and recurrence of MI. Elevated levels of miR-21 have been observed in ACS patients and have been linked to cardiomyocyte apoptosis and cardiac hypertrophy. Similarly, miR-26 has been implicated in the pathology and recurrence of MI \cite{new5}. In addition to their diagnostic potential, miRNAs have also shown promise as prognostic biomarkers for adverse myocardial effects, sudden death, and risk assessment in MI and other CVDs. For example, miR-101 and miR-150 have been associated with flawed left ventricular contractility after MI, while miR-16 and miR-27a have been linked to an increased risk of adverse left ventricular remodeling \cite{11-15, new3}. These miRNAs may provide valuable prognostic information and aid in risk stratification for post-MI complications.

Numerous studies have investigated the potential of miRNAs as biomarkers for MI, revealing promising findings. For instance, miR-1 has been proposed as a potential biomarker for MI \cite{new3}. This miRNA has shown increased expression levels in patients with MI, suggesting its potential diagnostic value. Additionally, other miRNAs, such as miR-19b-3p, miR-208a, miR-223-3p, miR-483-5p, and miR-499a-5p, have demonstrated promising diagnostic accuracy for MI within a short time window after the onset of symptoms \cite{new4}. A recent systematic review compared the peak time and diagnostic accuracy of miRNAs and conventional biomarkers in MI. The results revealed miR-1-3p, miR-19b-3p, miR-208a, miR-223-3p, miR-483-5p, and miR-499a-5p had superior peak times within 4 hours and better accuracy versus cTn and Creatine kinase-MB, indicating their promise for early diagnosis. The strengths of miRNAs included their early peak expression, satisfactory sensitivity and specificity, and higher accuracy especially within the first few hours of symptom onset compared to conventional biomarkers \cite{sys}.
% Change E

It has been postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in patients with CVD. Specifically, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) \cite{36}. 
PBMCs are a fraction of white blood cells, including monocytes,
lymphocytes, macrophages, and other cells of the immune system \cite{PBMC-miR}. Emerging data indicate that PBMCs can be used as a valid
source of biomarkers for monitoring various pathological conditions. Of
note, the alteration of mRNAs and miRNAs under pathological conditions
provides valuable information about different kinds of disorders. PBMCs
can recapitulate the conditions of target tissues, thus providing a
highly sensitive and specific source of biomarkers \cite{Meder6}. Combined with these conditions, these cells are repositories of
dysregulated genes and miRNA expression profiles in CVDs \cite{PBMC-miR, Meder6}.

In recent years, the advent and application of machine learning (ML) has
been an exciting prospect for advancing scientific research. Although
the concept of ML and its initial algorithms were conceived many years
ago, recent improvements in computing power and access to vast amounts
of data have demonstrated that ML techniques outperform classical
statistical methods in various fields. Furthermore, the progress made in
omics technologies has enabled the analysis of massive and intricate
biological datasets, consisting of hundreds to thousands of samples,
which makes it possible for ML to extract valuable biological insights
and information from such data \cite{ML-Bio}. Consequently, ML
provides innovative methods for merging and interpreting diverse types
of omics data, leading to the identification of new biomarkers. These
biomarkers can aid in precise disease prediction, patient
stratification, and the development of novel therapeutic approaches
\cite{ML}.

In this study, we aimed to identify potential miRNA biomarkers in
patients with MI by combining and analyzing three different microarray
datasets from PBMCs. The integration of omics data with bioinformatics
and ML techniques could be a promising tool in the discovery of new and
more accurate biomarkers for monitoring MI. Additionally, this approach
can deepen our understanding of the underlying mechanisms of MI and aid
in the development of valid diagnostic biomarkers and patient
stratification.

\section{Methods}\label{methods}

\subsection{Microarray data
collection}\label{microarray-data-collection}

Microarray datasets were obtained from the Gene Expression Omnibus (GEO)
database (\url{https://www.ncbi.nlm.nih.gov/geo/}).
% Change S
To obtain robust classification performance between MI, healthy control, and CAD samples, sufficiently large sample sizes for each group are required. For this purpose, the GSE59867 dataset was selected, as it contains sizable numbers of both MI and CAD samples. To provide an equally large set of healthy controls, the GSE56609 and GSE54475 datasets containing healthy samples were also included. Combining these three datasets enabled comparative analysis between MI, CAD, and healthy control groups with adequate statistical power.
% Change E
All samples were produced using Affymetrix Human Gene 1.0 ST Array
platform (GPL6244). 
% Change S
This platform contains 189 miRNA probes based on the annotation data from the GEO database.
% Change E
Only healthy, CAD, and early-stage MI samples were
selected from these datasets for further analysis.
% Change S
Early-stage MI samples were analyzed to enable detection of miRNA biomarkers specific to the initial ischemia and infarction event, before extensive myocardial necrosis and remodeling occurs. Using samples from the early phase enhances identification of miRNA signals related to plaque rupture and MI onset versus stable CAD. Additionally, early-stage samples allow investigation of mechanisms initiating myocardial injury.
%Change E
The basic information
for the three datasets evaluated in this study is provided in Table
~\ref{tab:datasets}. Bioinformatics analyses 
%Change S
including  preprocessing, differential expression analysis, and functional and pathway enrichment
analyses
%Change E
were conducted using R, ver.
4.2.0 \cite{R}, and RStudio \cite{RStudio}. All plots and
graphics of these sections were created using the ggplot2 R package \cite{ggplot}.

\begin{table}[h]
\centering
\caption{Sample information on the GEO microarray dataset.}
\label{tab:datasets}
\begin{tabular}{cccccc}
\toprule
Dataset & Platform & Healthy & CAD & MI & Reference \\
\midrule
GSE59867 & GPL6244 & - & 46 & 111 & \cite{67} \\
GSE56609 & GPL6244 & 46 & - & - & \cite{09} \\
GSE54475 & GPL6244 & 5 & - & - & \cite{75} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Preprocessing}\label{preprocessing}

The raw data in the form of CEL files from all datasets were obtained
from GEO. To prepare the data for analysis, we utilized the fRMA package
\cite{fRMA} to facilitate preprocessing
of individual microarray samples and their consistent combination. For
each dataset, background correction was applied using the RMA algorithm,
followed by quantile normalization based on the reference distribution.
To account for probe-specific effects, batch effects were eliminated
during summarization and gene expression variances were estimated
accordingly. In cases where multiple probe sets matched the same gene,
the mean log-fold change was retained. Consequently, fRMA can serve as a
technique to remove batch effects across diverse datasets generated by
identical microarray platforms \cite{BER}. To ensure the
effectiveness of the batch effect removal, we employed principal
component analysis (PCA) and relative log expression (RLE) plots to
visualize the data before and after applying fRMA.

\subsection{Differential expression
analysis}\label{differential-expression-analysis}

The barcode algorithm was introduced by McCall et al. \cite{Barcode}, aimed to convert actual expression values into binary
barcode values. Extensive sample collections were gathered and
normalization was performed using fRMA across multiple platforms,
including the Affymetrix Human Gene 1.0 ST Array (GPL6244) platform. By
utilizing these normalized datasets, the distribution of the observed
intensities for both the expressed and unexpressed genes was estimated.
The determination of whether a gene was expressed or not was based on
the following equation, where a value of 1 indicates expression and a
value of 0 indicates non-expression:

\begin{equation}
\hat{x}_{ij} =
\begin{cases}
1 & \text{if } x_{ij} \geq \mu^{ne} + C \times \sigma^{ne} \\
0 & \text{otherwise}
\end{cases}
\end{equation}

In the barcode algorithm, the normalized intensity of gene $i$ in
sample $j$ is denoted as $x_{ij}$. A user-defined parameter, $C$,
was introduced along with the standard deviation ($\sigma^{ne}$) and
mean ($\mu^{ne}$) of the non-expressed distribution. Based on these
values, the barcode representation of a sample was generated as a vector
consisting of ones and zeros. 
%Change S
The ones and zeros generated by the barcode algorithm refer to binary calls of whether or not a gene is estimated to be expressed (1) or not expressed (0) in each individual sample. 
%Change E
The barcode function within the R fRMA package was employed to implement the barcode
algorithm, utilizing the default value of $C$.

To assess the differences in expressed ratios between the MI and healthy
control groups, Fisher's exact test was performed on the barcode values
of individual genes. Genes that exhibited a false discovery rate (FDR)
below $0.05$, calculated using the Benjamini-Hochberg procedure to
account for multiple testing issues were identified as differentially
expressed genes (DEGs). The same procedures were applied to the CAD
versus healthy control comparison, as well as to the MI versus CAD
group, to identify DEGs specific to each comparison.

% Change S
\subsubsection{Differentially expressed miRNAs}\label{differentially-expressed-miRNAs}

The differentially expressed miRNAs were defined as those miRNAs within the total DEGs (i.e. they had an FDR $< 0.05$ resulted from the Fisher's exact test comparing the sample groups).

% Change E

\subsection{Functional and pathway enrichment
analyses}\label{functional-and-pathway-enrichment-analyses}

The R clusterProfiler package \cite{cluster} was utilized to perform
Kyoto Encyclopedia of Genes and Genomes (KEGG) pathway enrichment
analysis and Gene Ontology (GO) functional annotation on the set of
DEGs. GO analysis encompassed three categories: biological process (BP),
cellular component (CC), and molecular function (MF). For statistical
significance, an adjusted p-value threshold of less than 0.05 was
employed. Enrichment analyses were conducted separately for DEGs
specific to the MI-healthy and CAD-healthy comparisons. All the default
parameters provided by the package were used in the analyses.

\subsection{ML procedure}\label{ml-procedure}

ML analysis was performed using Python software, ver. 3.9, Numpy \cite{numpy}, Pandas \cite{Pandas}, and Scikit-Learn packages
\cite{scikit-learn}. Whenever hypertuning was needed, the Scikit-opt
package \cite{scikitopt} was used. In all ML analyses, the datasets
were divided into training and test sets at a 0.7:0.3 ratio, and all
reported results are the average of 10-fold cross-validation.

Two different approaches were used to select miRNAs for model training.
The first approach was to use differentially expressed miRNAs.
% Change
To capture additional miRNAs with high discriminatory power for distinguishing MI from CAD despite not reaching differentially expression criteria, a secondary approach was used. miRNAs were selected based on having individual area under the receiver operating characteristic curves (AUC-ROCs) exceeding 0.8 for separating MI and CAD. This AUC-based approach identifies miRNAs with the best classification performance, unconstrained by statistical cutoffs. Using both the differentially expressed and AUC-based selection provides complementary methods to uncover miRNA biomarkers from both a biological and diagnostic perspective.
% Change.

\subsubsection{Differentially expressed miRNAs}\label{mirnas-in-degs}

In this approach, a two-layer architecture is deployed to the data to
maximize the prediction values. The first layer predicted whether a
sample was healthy or not, and the second layer separated MI from CAD in
the samples that were predicted as not healthy in the first layer. To
this end, a distinct ML model was trained for each layer. Because there
were a limited number of miRNAs in the DEGs, both layers were trained
with all of them. For further comparison with the models' performance,
the ROC curve of each miRNA for classifying healthy and not-healthy, as
well as CAD and MI, was generated using a logistic regression model.

\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples}

A support vector machine (SVM) model using RBF kernels was trained and
hypertuned using all miRNAs in the DEGs. 
% Change
To account for the substantial class imbalance between the healthy and not-healthy groups, with 51 samples in the minority healthy class compared to 157 combined CAD and MI samples, adjustments were made to the sample weights used during model training. Without compensating for the imbalance, the machine learning model would be biased towards the majority class and potentially ignore the minority class. To counteract this, the sample weights were empirically tuned, with the weight for healthy samples set to 1 and the weight for not-healthy samples set to 0.5. These values were determined through iterative testing to produce a model with strong performance on both classes despite the imbalance.
% Change
The ROC curve
and confusion matrix for the model are reported.

\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples}

Different models were investigated to achieve the highest classification
performance. To do so, SVM (with linear, polynomial, and RBF kernels),
logistic regression (LR), random forests (RF), k-nearest neighbor (kNN),
gradient boosting (GB), XGBoost (XGB) and decision tree (DT) models were
trained. All models were trained with their preset parameters using
10-fold cross-validation. The criteria for selecting the best model were
the highest accuracy and AUC-ROC for the test set. The best model was
hypertuned using the scikit-opt package \cite{scikitopt} for the best
classification performance. The ROC curve and confusion matrix for the
best model are reported.

\subsubsection{miRNAs with the highest
AUC-ROC}\label{mirnas-with-the-highest-auc-roc}

As in the previous approach, a two-layer strategy was employed. The
first layer classified samples into healthy and not-healthy, and the
second layer separated the MI and CAD samples. However, to keep the
number of miRNAs as low as possible, miRNAs were selected from the
second layer and their performance was evaluated in the first layer. The
AUC-ROC of all miRNAs for classifying MI and CAD samples was calculated,
and miRNAs with AUC-ROC $> 0.8$ were selected. ROC curves for each
selected miRNA for separating healthy samples from not-healthy samples
and MI from CAD samples were also plotted for further comparison.

\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-1}

An SVM model with an RBF kernel is trained using the selected set of
miRNAs. Additionally, the model was hypertuned to find the
hyperparameters for the highest AUC-ROC and accuracy. The same sample
weights as in the previous approach (1 for healthy and 0.5 for
not-healthy samples) were used. The ROC curve and confusion matrix for
the model were reported.

\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-1}

The selected miRNA set was used to train different algorithms to
determine the best model. Similar to the previous approach, the SVM
(with linear, polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT
models were trained. All models were trained with their preset
parameters using 10-fold cross-validation. The models with the highest
AUC-ROC and accuracy on the test set were selected and hypertuned using
the scikit-opt package \cite{scikitopt}. The ROC curve and confusion
matrix for the best model were reported.

\section{Results}\label{results}

\subsection{Preprocessing}\label{preprocessing-1}

The PCA plots of the samples are shown in Figures ~\ref{fig:PCA}A and B.
Healthy samples were separated from the CAD or MI samples in the primary
data and after conducting fRMA. In the RLE plot, there was a distinct
difference between the dataset means for all samples before fRMA was
performed (Figure~\ref{fig:PCA}C). All datasets were rearranged to
approximately 0 in the RLE plot after fRMA was conducted (Figure
~\ref{fig:PCA}D). Moreover, there was an apparent change in the
interquantile distances, but the values were still greater than 0.1.

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{PCA and RLE of trains} \caption{Principal component analysis plots for (A) primary data and (B) the data after fRMA, and the relative log expression plots for (C) primary data and (D) the data after fRMA.}
\label{fig:PCA}
\end{figure}

\subsection{Differential expression
analysis}\label{differential-expression-analysis-1}

\begin{table}
\centering
\caption{Total, up-, and down-regulated DEGs and differentially expressed miRNAs.}
\label{tab:DEGstab}
\begin{tabularx}{\textwidth}{XXXXX}
\hline
& Total DEGs & Up-regulated DEGs & Down-regulated DEGs & miRNAs\\
\hline
MI vs. Healthy & 860 & 323 & 537 & hsa-miR-186-5p, hsa-miR-21-3p, hsa-miR-32-3p\\
CAD vs. Healthy & 670 & 262 & 408 & hsa-miR-186-5p, hsa-miR-21-3p, hsa-miR-32-3p\\
MI vs. CAD & 260 & 144 & 116 & hsa-miR-186-5p\\
\hline
\end{tabularx}
\end{table}

According to the cutoff criterion of $FDR < 0.05$, there were 860 DEGs
between MI and healthy samples. Among them, 323 were up-regulated, and
537 were down-regulated in the MI group compared to the healthy group.
In the CAD and healthy group comparison, we found 670 DEGs, of which 262
and 408 DEGs were up- and down-regulated, respectively, in CAD samples.
In the MI and CAD groups, the number of DEGs was 260, and the numbers of
up- and down-regulated genes in MI samples were 144 and 116,
respectively, compared to CAD samples. The data are summarized in Table
~\ref{tab:DEGstab}.

\begin{figure}
\centering 
\includegraphics[width=0.45\linewidth]{venn-1}
\caption{Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD.}\label{fig:venn}
\end{figure}

The Venn diagram in Figure~\ref{fig:venn} shows that the CAD and MI
samples shared most of their DEGs. From 860 DEGs of MI/healthy and 670
DEGs of CAD/healthy, 531 genes were common, which is 62\% of MI/healthy
DEGs and 79\% of CAD/healthy DEGs.

% Change S
\subsubsection{Differentially expressed miRNAs}\label{mirnas-in-degs}
Among the DEGs for MI/healthy and CAD/healthy comparison, hsa-miR-186-5p, hsa-miR-32-3p, and hsa-miR-21-3p were identified as differentially expressed miRNAs. The only differentially expressed miRNA in MI/CAD comparison was hsa-miR-186-5p (Table~\ref{tab:DEGstab}). The expression profiles of the three
miRNAs are shown in Figure~\ref{fig:Expall}.
% Change E

\subsection{GO and KEGG enrichment analyses of the
DEGs}\label{go-and-kegg-enrichment-analyses-of-the-degs}

To explore the biological classification of the DEGs, we performed GO
and KEGG pathway enrichment analyses on the MI/healthy and CAD/healthy
DEGs. For MI/healthy, GO enrichment analysis in the BP category
suggested that the DEGs were enriched in ``immune response-regulating
signaling pathway,'' ``lymphocyte differentiation,'' ``immune
response-regulating cell surface receptor signaling pathway,'' and
``leukocyte activation involved in immune response'' (Figure
~\ref{fig:MIHEnrich}A). In the CC category, DEGs were enriched in
``secretory granule membrane,'' ``azurophil granule,'' ``ficolin-1-rich
granule,'' ``tertiary granule,'' and ``ficolin-1-rich granule membrane''
(Figure~\ref{fig:MIHEnrich}B). In the MF category, DEGs were involved in
``cadherin binding'' and ``MHC class I protein binding'' (Figure
~\ref{fig:MIHEnrich}C). KEGG pathway analysis indicated that the DEGs
were related to the following pathways: ``Chemokine signaling pathway,''
``Lipid and atherosclerosis,'' and ``Hematopoietic cell lineage''
(Figure~\ref{fig:MIHEnrich}D).

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{Enrichment MI-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}
\label{fig:MIHEnrich}
\end{figure}

The enrichment results for the CAD/healthy DEGs were as follows. In the
BP category, GO enrichment suggested that the DEGs were enriched in
``positive regulation of defense response,'' ``positive regulation of
innate immune response,'' ``mononuclear cell differentiation,'' and
``positive regulation of response to external stimulus'' (Figure
~\ref{fig:CADHEnrich}A). In the CC category, DEGs were enriched in
``azurophil granule,'' ``ficolin-1-rich granule,'' and ``ficolin-1-rich
granule membrane'' (Figure~\ref{fig:CADHEnrich}B). In the MF category,
DEGs were involved in ``lipoprotein particle receptor binding'' and
``NF-$\kappa$B binding'' (Figure~\ref{fig:CADHEnrich}C). KEGG pathway
analysis showed that the DEGs were related to the following pathways:
``Chemokine signaling pathway,'' ``Lipid and atherosclerosis,'' and
``Hematopoietic cell lineage'' (Figure~\ref{fig:CADHEnrich}D).

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{Enrichment CAD-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}
\label{fig:CADHEnrich}
\end{figure}

\subsection{Machine Learning}\label{machine-learning}

\subsubsection{Differentially expressed miRNAs}\label{mirnas-in-degs-1}

The ROC curves of each miRNA in each layer are presented in Figure~\ref{fig:miRROC}. Using the
logistic regression model, the AUC-ROC values of hsa-miR-21-3p, hsa-miR-32-3p, and
hsa-miR-186-5p for separating healthy and not-healthy samples were 0.98, 0.99,
and 0.90, respectively (Figure~\ref{fig:miRROC}A). The accuracy of each
miRNA for classifying the samples into healthy and not-healthy groups on
the test set for hsa-miR-21-3p, hsa-miR-32-3p, and hsa-miR-186-5p was 0.92, 0.98, and 0.89,
respectively. The ROC curve of each miRNA for classifying MI and CAD
samples is presented in Figure~\ref{fig:miRROC}B. The AUC-ROC and
accuracy for hsa-miR-21-3p, hsa-miR-32-3p, and hsa-miR-186-5p in the test set were 0.85;
0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.

\begin{table}
\centering
\caption{Investigated miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.}
\label{tab:mirExptable}
\begin{tabular}{lcccccc}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{CAD/Healthy} & \multicolumn{2}{c}{MI/Healthy} & \multicolumn{2}{c}{MI/CAD} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
& logFC & adj. p-value & logFC & adj. p-value & logFC & adj. p-value\\
\midrule
hsa-miR-186-5p & 1.4 & 3.60e-25 & 0.9 & 6.76e-20 & -0.5 & 1.05e-09\\
hsa-miR-21-3p & 1.4 & 1.31e-17 & 2.3 & 2.07e-47 & 0.8 & 2.96e-11\\
hsa-miR-32-3p & 2.5 & 8.39e-43 & 2.2 & 3.10e-59 & -0.3 & 7.60e-04\\
hsa-miR-197-5p & 0.5 & 2.95e-20 & 0.7 & 1.59e-47 & 0.2 & 8.58e-09\\
hsa-miR-29a-5p & 0.7 & 7.76e-29 & 0.1 & 1.70e-01 & -0.5 & 2.14e-10\\
hsa-miR-296-5p & -0.1 & 5.00e-02 & 0.1 & 2.00e-02 & 0.2 & 6.15e-06\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering 
\includegraphics[width=0.9\linewidth]{Expressionforall}
\caption{Expression profile of all miRNAs in two approaches in different sample classes.}\label{fig:Expall}
\end{figure}

\begin{figure}
\centering 
\includegraphics[width=0.95\linewidth]{miRs ROCs}
\caption{ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples.}\label{fig:miRROC}
\end{figure}

\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-2}

Although single miRNAs had an acceptable performance for this layer,
their predictive value could be further improved by using them as a set.
The ROC curve for the SVM model with an RBF kernel trained with all
three miRNAs is presented in Figure~\ref{fig:DEMROC}A. The model had a
better performance in classification than single miRNAs.The AUC-ROC for
the model was 1, and its accuracy on the test set was also 1. In Figure
~\ref{fig:DEMCF}A, the confusion matrix for the model is presented.

\begin{figure}
\centering 
\includegraphics[width=0.95\linewidth]{DEMs ROCs h not h cad mi}
\caption{ROC curve for the model trained with differentially expressed miRNAs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMROC}
\end{figure}

\begin{figure}
\centering 
\includegraphics[width=0.9\linewidth]{DEMs Conf. matrix}
\caption{Confusion matrix for the model trained with differentially expressed miRNAs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMCF}
\end{figure}

\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-2}

Different models were trained using the expression values of three
differentially expressed miRNAs. The models' AUC-ROC and the accuracy of
the test set are shown in Figure~\ref{fig:DEMmodels}. The best model
from both the AUC-ROC and accuracy points of view was the SVM model with
a linear kernel. The AUC-ROC and accuracy for this model with its preset
values were 0.93 and 0.82, respectively. The model was hypertuned for C
and gamma hyperparameters, and therefore the model showed better
performance. The ROC curve of the hypertuned model is presented in
Figure~\ref{fig:DEMROC}B. For this model, the AUC-ROC reached 0.95, and
the accuracy was improved to 0.85 (Table~\ref{tab:DEGsML}). Moreover,
the sensitivity and specificity for the model on the test set were 0.91
and 0.71, respectively. The confusion matrix for the hypertuned model is
illustrated in Figure~\ref{fig:DEMCF}B.

\begin{figure}
\centering
\includegraphics[width=0.65\linewidth]{Models DEGs mirs}
\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three differentially expressed miRNAs.}\label{fig:DEMmodels}
\end{figure}

\begin{table}
\centering
\caption{AUC-ROC and accuracy for SVM with a linear kernel as the best model trained with differentially expressed miRNAs on the training and test sets before and after hypertuning}
\label{tab:DEGsML}
\begin{tabular}{cccccc}
\toprule
\multicolumn{2}{c}{} & \multicolumn{2}{c}{Preset parameters} & \multicolumn{2}{c}{Hypertuned} \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6}
Model & Metrics & train & test & train & test\\
\midrule
SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.92 & 0.95\\
& Accuracy & 0.83 & 0.82 & 0.84 & 0.85\\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{AUC-ROC approach}\label{auc-roc-approach}

After calculating the AUC-ROC for each miRNA to classify of MI and CAD
samples, the miRNAs with AUC-ROC $> 0.8$ were selected. The miRNAs
selected were hsa-miR-29a-5p, hsa-miR-197-5p, hsa-miR-186-5p, hsa-miR-21-3p, and hsa-miR-296-5p. The
expression levels of these miRNAs in healthy, CAD, and MI samples are
presented in Figure~\ref{fig:Expall}. The ROC curves of the selected
miRNAs in both layers are shown in Figure~\ref{fig:miRROC}.

\paragraph{First layer for the isolation of healthy and not-healthy
samples:}\label{first-layer-for-the-isolation-of-healthy-and-not-healthy-samples-3}

Using the selected set, an SVM model with an RBF kernel was trained to
separate healthy and not-healthy samples. The ROC curve for the model is
presented in Figure~\ref{fig:AUCROC}A, and the confusion matrix is
illustrated in Figure~\ref{fig:AUCCF}A. Both the AUC-ROC and accuracy of
the model on the test set were 1.

\begin{figure}
\centering 
\includegraphics[width=0.95\linewidth]{AUC best model ROCs}
\caption{ROC curve for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI sample classification. (C) SVM with polynomial kernel for CAD and MI sample classification. }\label{fig:AUCROC}
\end{figure}

\begin{figure}
\centering 
\includegraphics[width=0.9\linewidth]{AUC Conf. matrix}
\caption{Confusion matrix for models trained with the set of miRNAs selected by AUC-ROC on test set classification; (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI sample classification. (C) SVM with polynomial kernel for CAD and MI sample classification.}\label{fig:AUCCF}
\end{figure}

\paragraph{Second layer for separating the MI and CAD
samples:}\label{second-layer-for-separating-the-mi-and-cad-samples-3}

To find the best model for this set of miRNAs, different models were
trained using their preset values. The AUC-ROC and accuracy results for
the test set are presented in Figure~\ref{fig:AUCmodels}. The best model
from the AUC-ROC point of view was the SVM with a linear kernel, and
from the accuracy point of view, it was the SVM model with an RBF
kernel. For the SVM-linear model, the AUC-ROC and accuracy were 0.93 and
0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84,
respectively. Both models were hyper-tuned, and the ROC curve for their
best performance is presented in Figure~\ref{fig:AUCROC}B and C. The
AUC-ROC and accuracy for the SVM-linear model were modified to 0.92 and
0.88, respectively. For the SVM-RBF, these values increased to 0.96 and
0.94, respectively (Table~\ref{tab:AUCML}). The sensitivities for the
SVM-linear and SVM-RBF models were 0.91 and 0.97, respectively; and
their specificities were 0.79 and 0.86, respectively. The confusion
matrix for both models is illustrated in Figure~\ref{fig:AUCCF}B and C.

\begin{figure}
\centering 
\includegraphics[width=0.65\linewidth]{Models AUC 5 mirs CV}
\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with AUC-selected miRNAs.}\label{fig:AUCmodels}
\end{figure}

\begin{table}
\centering
\caption{AUC-ROC and accuracy for SVM with the linear and RBF kernels as the best models trained with miRNAs selected based on their AUC-ROC on the train and test sets before and after hypertuning}
\label{tab:AUCML}
\begin{tabular}{cccccc}
\toprule
\multicolumn{2}{c}{} & \multicolumn{2}{c}{Preset parameters} & \multicolumn{2}{c}{Hypertuned} \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6}
Model & Metrics & Train & Test & Train & Test\\
\midrule
SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.93 & 0.92\\
& Accuracy & 0.85 & 0.82 & 0.90 & 0.88\\
SVM-RBF & AUC-ROC & 0.90 & 0.92 & 0.96 & 0.96\\
& Accuracy & 0.86 & 0.84 & 0.96 & 0.94\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}\label{discussion}

The prevalence of MI can lead to high mortality rates in the clinical
setting. However, early diagnosis and the application of suitable
treatment protocols can reduce mortality and improve MI prognosis
(\cite{CVD, 17, 18}). Studies have suggested that changes in miRNA expression
may play a significant role in the progression of MI and the subsequent
remodeling \cite{23}. It is believed that miRNA
expression is altered during the various biological processes correlated
with MI within the myocardium or other related tissues \cite{24}. Although several studies have focused on examining free
circulating miRNAs in serum samples for the detection of cardiac tissue
injuries \cite{11-15}, more information is needed to fully
comprehend the miRNAs found in different blood subcomponents, such as
plasma, platelets, and PBMCs. Based on previous findings, PBMCs play a
crucial role in the destabilization and rupture of plaques as well as in
the initial inflammatory reactions in individuals experiencing
myocardial infarction (MI) \cite{Meder6, Meder23}.
Moreover, PBMCs have specific miRNA profiles that are altered under
certain pathological conditions, making them great candidates as disease
biomarkers \cite{Meder6}.

PBMCs can respond to several insulting conditions, such as MI, in the
shortest possible time with notable changes in their miRNA profile
\cite{Meder6}. Considering their regulatory roles, subtle
changes in the transcription of miRNAs can be monitored even before
alterations in mRNA and protein levels \cite{miR}. These
features make miRNAs a valid early-stage diagnostic tool for the
detection of minor and major cell injuries. To date, few studies have
compared the miRNA profiles in PBMCs from patients with MI and other
CADs and healthy samples to find a robust set of identical miRNAs to
differentiate these pathological conditions.

In this study, we combined three GEO datasets for healthy, CAD, and MI
samples. Having these sample sets alongside bioinformatics analysis and
ML methods enabled us to identify potential biomarker sets and effective
therapeutic targets. The results of the DEG analysis (Table
~\ref{tab:DEGstab} and Figure~\ref{fig:venn}) prove the close
relationship between the MI and CAD samples. Interestingly, functional
enrichment analysis demonstrated that DEGs in both CAD/healthy and
MI/healthy were strongly correlated with the immune cell response, which
is a major part of PBMCs. Two sets of miRNAs were selected as biomarker
sets for sample classification. Hsa-miR-21-3p; hsa-miR-32-3p; and hsa-miR-186-5p were
selected as differentially expressed miRNAs, and hsa-miR-186-5p; hsa-miR-21-3p;
hsa-miR-29a-5p; hsa-miR-197-5p; and hsa-miR-296-5p were selected based on their AUC-ROC
values. As shown in Figure~\ref{fig:miRROC}, all miRNAs selected with
both approaches had AUC-ROCs $> 0.9$ for isolating healthy and
not-healthy samples except for hsa-miR-296-5p and hsa-miR-29a-5p. The data confirmed
that the real challenge was to classify CAD and MI samples because of
the close overlap. Of the six miRNAs under investigation in both
approaches, except for hsa-miR-32-3p, all miRNAs had an AUC-ROC $> 0.8$ for
the discrimination of CAD and MI samples. As expected, the high AUC-ROC
values of the miRNAs confirmed their high potential as biomarkers.

ML models trained with miRNA sets selected by both DEG and AUC-ROC
approaches, showed better classification performance than each miRNA. To
avoid unwanted complexity and poor predictive values, a two-layer
architecture was designed. The first layer was used to discriminate
between healthy and not-healthy samples, and the second layer was was
used to separate CAD from MI candidates. As expected, in both
approaches, a hypertuned SVM model could flawlessly separate healthy and
not-healthy samples using distinct miRNA sets. ML models are also
capable of effectively separating CAD from MI patients. Although both
miRNA sets had nearly the same AUC-ROC using the best model, their
accuracy, sensitivity, and specificity were different. The model trained
with AUC-selected miRNAs showed better performance in all predictive
values, which is logical because of the higher number of miRNAs in the
set.

Numerous studies have reported that different biological processes can
affect the miRNA expression in PBMCs. However, the exact role of miRNAs
in the function of immune cells and the correlation between specific
pathological conditions and miRNA profiles remain controversial. Several
studies have proven the activation of particular miRNA types in PBMCs
under cardiovascular events \cite{296}. For instance, there is
evidence that elevation of hsa-miR-186-5p suppresses the expression of
cystathionine-$\gamma$-lyase, leading to the subsequent secretion of
pro-inflammatory cytokines and cellular lipid accumulation. In addition,
macrophage-derived hsa-miR-186-5p may promote atherosclerotic plaque formation
\cite{186}. In line with this claim, we found that hsa-miR-186-5p was
up-regulated in both CAD and MI candidates compared to their control
counterparts. Surprisingly, the obtained data indicated that the
expression of hsa-miR-186-5p was higher in patients with CAD than in patients
with MI (Figure~\ref{fig:Expall}). Specifically, hsa-miR-186-5p was the only
differentially expressed miRNA between CAD and MI, with a clear
up-regulation in CAD, indicating its main role in the promotion of
atherosclerosis.

As mentioned before, hsa-miR-21-3p was also up-regulated in both MI and CAD
patients compared to healthy controls. Moreover, the expression value of
hsa-miR-21-3p was significantly higher in the MI group than in the CAD group
(Table~\ref{tab:mirExptable}). It is thought that the up-regulation of
hsa-miR-21-3p in PBMCs is a compensatory reaction to reduce the T$_{reg}$
lymphocyte number in response to the reduction in TGF$\beta1$
secretion into the plasma through a TGF$\beta1$/smad-independent
pathway. In line with the previous and present data, hsa-miR-21-3p can modulate
the activity of PBMCs following the occurrence of cardiovascular
diseases \cite{21}.

Recent data have supported the elevation of hsa-miR-32-3p levels in CAD samples
with calcification of the coronary artery. Notably, hsa-miR-32-3p promotes
vascular smooth muscle calcification in mice by controlling the activity
of several proteins, including bone morphogenetic protein-1,
runt-related transcription factor-2 (RUNX2), osteopontin, and
bone-specific phosphoprotein matrix GLA protein \cite{32-2}.
Likewise, some reports are associated with the activity of hsa-miR-32-3p in
PBMCs in several pathologies \cite{magic, 32}. The
exact role of hsa-miR-32-3p in PBMCs after cardiovascular events remains
unclear.

Molecular analyses have indicated the regulatory role of miRNAs selected
using the AUC-ROC approach in PBMCs after a cardiovascular event. The
biological importance of two common miRNAs in the DEGs and AUC-ROC
approaches, hsa-miR-21-3p and hsa-miR-186-5p, have already been discussed. Based on
numerous reports, hsa-miR-29a-5p can be activated in different diseases \cite{29aa}. Data analysis indicated that
hsa-miR-29a-5p was significantly up-regulated in CAD patients compared to the
healthy and MI groups (Table~\ref{tab:mirExptable}). Increased hsa-miR-29a-5p
is associated with the progression of atherosclerosis, and the
combination of hsa-miR-29a-5p and ox-LDL has been suggested as a valid
biomarker set for paraclinical classification \cite{29a}.
However, the role of hsa-miR-29a-5p in the function of PBMCs from patients with
CAD has not been thoroughly examined.

The data indicated that hsa-miR-197-5p was significantly up-regulated in both
the CAD/healthy and MI/healthy groups. Previous studies have
demonstrated that hsa-miR-197-5p may play a crucial role in controlling the
anti-inflammatory response of IL-35 by influencing the secretion of
cytokines that can either promote or suppress inflammation, the ratio of
M1/M2 macrophages, and the proliferation of T$_{reg}$ lymphocytes,
which are responsible for suppressing immune responses \cite{197}. Alongside our findings, it can be concluded that hsa-miR-197-5p could be
a useful diagnostic tool for predicting adverse cardiovascular events.

The findings of this study demonstrate the potential of hsa-miR-296-5p as a
biomarker with high discriminatory power to distinguish between samples
from individuals with MI and CAD. Hsa-miR-296-5p has been identified as a key
regulator in the development and advancement of atherosclerosis by
controlling the expression of target genes associated with various
biological processes, including angiogenesis, cholesterol metabolism,
inflammation, cellular proliferation, hypertension, and apoptosis \cite{296}. In a previous study, hsa-miR-296-5p expression levels were found
to be significantly increased in the PBMCs of CAD patients compared to
healthy controls, suggesting its involvement in regulating
proinflammatory cytokines such as IL-6 and TNF-$\alpha$ \cite{296a}. These findings suggested that hsa-miR-296-5p may have a significant
impact on the pathogenesis of atherosclerosis and could potentially
serve as a diagnostic biomarker for CAD or MI.

\section{Conclusion}\label{conclusion}

In summary, we derived a set of miRNA biomarkers by comparing MI samples
with both healthy and CAD samples. We found that the SVM model performed
best in both the first layer, which separated healthy and unhealthy
samples, and the second layer, which classified the MI/CAD samples. The
set of miRNAs selected based on their AUC-ROC values performed better in
the second layer. Overall, our two-layer structure achieved an accuracy
of 0.96. This demonstrates the potential of combining bioinformatics and
machine learning techniques to identify novel biomarkers and gain a
deeper understanding of myocardial infarction.

\section{Abbreviations}\label{abbreviations}

\begin{description}
  \item \textbf{CVD}: Cardiovascular disease
  \item \textbf{MI}: Myocardial infarction
  \item \textbf{cTn}: Cardiac troponin
  \item \textbf{miRNA}: microRNA
  \item \textbf{ACS}: Acute coronary syndrome
  \item \textbf{PBMC}: Peripheral blood mononuclear cell 
  \item \textbf{ML}: Machine learning 
  \item \textbf{GEO}: Gene Expression Omnibus 
  \item \textbf{PCA}: Principal component analysis 
  \item \textbf{RLE}: Relative log expression 
  \item \textbf{FDR}: False discovery rate 
  \item \textbf{DEG}: Differentially expressed gene
  \item \textbf{KEGG}: Kyoto Encyclopedia of Genes and Genomes 
  \item \textbf{GO}: Gene Ontology  
  \item \textbf{BP}: Biological process 
  \item \textbf{CC}: Cellular component 
  \item \textbf{MF}: Molecular function 
  \item \textbf{SVM}: Support vector machine 
  \item \textbf{LR}: Logistic regression 
  \item \textbf{RF}: Random forests
  \item \textbf{kNN}: K-nearest neighbor 
  \item \textbf{GB}: Gradient boosting 
  \item \textbf{XGB}: XGBoost 
  \item \textbf{DT}: Decision tree 
  \item \textbf{AUC-ROC}: Area under the receiver operating characteristic curve 
\end{description}

\backmatter

%%===================================================%%
%% For presentation purpose, we have included        %%
%% \bigskip command. please ignore this.             %%
%%===================================================%%
\bigskip
\begin{flushleft}%
Editorial Policies for:

\bigskip\noindent
Springer journals and proceedings: \url{https://www.springer.com/gp/editorial-policies}

\bigskip\noindent
Nature Portfolio journals: \url{https://www.nature.com/nature-research/editorial-policies}

\bigskip\noindent
\textit{Scientific Reports}: \url{https://www.nature.com/srep/journal-policies/editorial-policies}

\bigskip\noindent
BMC journals: \url{https://www.biomedcentral.com/getpublished/editorial-policies}
\end{flushleft}

\bibliography{sn-article.bib}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated

\section*{Declarations}

\subsection*{Funding}
This is a report of result from Ph.D. thesis registered in Tabriz University of Medical Sciences with the Number 66372 . This work was extracted from Mehrdad Samadishadlou's thesis titled “Developing and manufacturing of a paper-based Nanobiosensor in order to diagnosing myocardial infarction using a set of blood microRNAs”.

\subsection*{Competing interests}
The authors declare that they have no competing interests.

\subsection*{Ethics approval}
The study was approved by the research ethics committee of Tabriz University of Medical Sciences (approval ID: IR.TBZMED.VCR.REC.1399.388, date of approval: 2021/1/11).

\subsection*{Consent for publication}
All authors gave consent for the publication of the article.

\subsection*{Availability of data and materials}
The datasets generated and/or analysed during the current study are available in the Gene Expression Omnibus (GEO, \url{https://www.ncbi.nlm.nih.gov/geo/}), reference numbers GSE59867, GSE56609 and GSE54475. All data generated or analysed during this study are included in this published article.

\subsection*{Authors' contributions}

F.B. and K.K. conceived the idea and coordinated the project.
M.S. researched, collected the data, performed the analyzes, assembled the results, and drafted the manuscript.
Z.P. and M.E.contributed to data analyze.
Ç.B.A. was foreign supervisor and collaborator.
F.B., R.R. and K.K. edited and revised the manuscript.
F.B. and K.K. are author responsible for contact and ensures communication.
All authors read the content of final manuscript.

\end{document}

### 2023-08-07 Revised Version/Article/sn-article.tex
The version after revision

--- 2023-08-07 Revised Version/Art/sn-article.tex
@@
 \usepackage{listings}%
-\usepackage{tabularx}%
 %%%%
@@
 
-\author[2, 3]{\fnm{Reza} \sur{Rahbarghazi}}
-
-\author[4]{\fnm{Zeynab} \sur{Piryaei}}
-
-\author[5]{\fnm{Mahdad} \sur{Esmaeili}}
-
-\author[6]{\fnm{Çığır} \sur{Biray Avcı}}
+\author[2]{\fnm{Reza} \sur{Rahbarghazi}}
+
+\author[3]{\fnm{Zeynab} \sur{Piryaei}}
+
+\author[4]{\fnm{Mahdad} \sur{Esmaeili}}
+
+\author[5]{\fnm{Çığır} \sur{Biray Avcı}}
 
@@
 
-\author[4*]{\fnm{Kaveh} \sur{Kavousi}}\email{kkavousi@ut.ac.ir}
-
-\affil[1]{\orgdiv{Department of Medical Nanotechnology}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}
-
-\affil[2]{\orgdiv{Stem Cell Research Center}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}
-
-\affil[3]{\orgdiv{Department of Applied Cell Sciences}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}
-
-\affil[4]{\orgdiv{Laboratory of Complex Biological Systems and Bioinformatics (CBB), Department of Bioinformatics}, \orgname{Institute of Biochemistry and Biophysics (IBB)}, \orgaddress{\street{University of Tehran}, \city{Tehran}, \country{Iran}}}
-
-\affil[5]{\orgdiv{Medical Bioengineering Department}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}
-
-\affil[6]{\orgdiv{Medical Biology Department}, \orgname{School of Medicine}, \orgaddress{\street{Ege University}, \city{İzmir}, \country{Türkiye}}}
+\author[2*]{\fnm{Kaveh} \sur{Kavousi}}\email{kkavousi@ut.ac.ir}
+
+\affil[1]{\orgdiv{Department of Medical Nanotechnology,}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}
+
+\affil[2]{\orgdiv{Department of Applied Cell Sciences}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}
+
+\affil[3]{\orgdiv{Laboratory of Complex Biological Systems and Bioinformatics (CBB), Department of Bioinformatics}, \orgname{Institute of Biochemistry and Biophysics (IBB)}, \orgaddress{\street{University of Tehran}, \city{Tehran}, \country{Iran}}}
+
+\affil[4]{\orgdiv{Medical Bioengineering Department}, \orgname{Faculty of Advanced Medical Sciences}, \orgaddress{\street{Tabriz University of Medical Sciences}, \city{Tabriz}, \country{Iran}}}
+
+\affil[5]{\orgdiv{Medical Biology Department}, \orgname{School of Medicine}, \orgaddress{\street{Ege University}, \city{İzmir}, \country{Türkiye}}}
 
@@
 
-\textbf{Results:} We identified hsa-miR-21-3p, hsa-miR-186-5p, and hsa-miR-32-3p as the
-differentially expressed miRNAs, and a set including hsa-miR-186-5p, hsa-miR-21-3p,
-hsa-miR-197-5p, hsa-miR-29a-5p, and hsa-miR-296-5p as the optimum set of miRNAs selected by
+\textbf{Results:} We identified miR-21, miR-186, and miR-32 as the only miRNAs among the
+differentially expressed genes, and a set including miR-186, miR-21,
+miR-197, miR-29A, and miR-296 as the optimum set of miRNAs selected by
 their AUC-ROC. Both biomarker sets could distinguish healthy from
@@
 Over the last 50 years, numerous attempts have been made to use
-biomarkers to facilitate diagnosis, assess the risk, follow-up therapy, and determine therapeutic efficacy in CVD candidates. Based on released
+biomarkers to facilitate diagnosis, assess the risk, follow-up therapy,
+and determine therapeutic efficacy in CVD candidates. Based on released
 guidelines, cardiac troponins (cTns) are used as a highly sensitive and
@@
 regulation of gene expression. Evidence indicates that miRNAs are
-involved in the pathogenesis of cardiac tissue injury \cite{m1}. Several biological processes, such as angiogenesis, cardiomyocyte growth and contractility, lipid metabolism, plaque
-formation, and cardiac rhythm, are regulated by miRNAs \cite{m2}. 
-% Change S
-Circulating and tissue-specific miRNAs have shown promise as diagnostic and prognostic biomarkers across a range of cardiovascular diseases, including MI and other conditions such as CAD, heart failure, atrial fibrillation, cardiac hypertrophy, and fibrosis \cite{11-15, new2}.
-The use of miRNAs as diagnostic and prognostic biomarkers in CVDs is supported by their stability and rapid release into circulation after myocardial injury \cite{11-15}.
-In CAD, altered expression of miRNAs like miR-1, miR-133a, miR-208a/b, and miR-499, which are abundantly expressed in the heart, has been reported in patients compared to healthy controls.
-Additional miRNAs including miR-21, miR-208a/b, miR-133a/b, and the miR-30 family are frequently dysregulated in acute coronary syndrome (ACS) versus stable CAD  \cite{new3}.
-Furthermore, miRNAs like miR-3113-5p, miR-223-3p, miR-499a-5p, and miR-133a-3p demonstrate potential as biomarkers to identify patients at risk of sudden cardiac death \cite{new4}.
-Moreover, miRNAs have shown diagnostic potential in other CVDs.
-For instance, miR-21 has been associated with cardiac injury and has been implicated in the pathology and recurrence of MI.
-Elevated levels of miR-21 have been observed in ACS patients and have been linked to cardiomyocyte apoptosis and cardiac hypertrophy.
-Similarly, miR-26 has been implicated in the pathology and recurrence of MI \cite{new5}.
-In addition to their diagnostic potential, miRNAs have also shown promise as prognostic biomarkers for adverse myocardial effects, sudden death, and risk assessment in MI and other CVDs.
-For example, miR-101 and miR-150 have been associated with flawed left ventricular contractility after MI, while miR-16 and miR-27a have been linked to an increased risk of adverse left ventricular remodeling \cite{11-15, new3}.
-These miRNAs may provide valuable prognostic information and aid in risk stratification for post-MI complications.
-
-Numerous studies have investigated the potential of miRNAs as biomarkers for MI, revealing promising findings.
-For instance, miR-1 has been proposed as a potential biomarker for MI \cite{new3}.
-This miRNA has shown increased expression levels in patients with MI, suggesting its potential diagnostic value.
-Additionally, other miRNAs, such as miR-19b-3p, miR-208a, miR-223-3p, miR-483-5p, and miR-499a-5p, have demonstrated promising diagnostic accuracy for MI within a short time window after the onset of symptoms \cite{new4}.
-A recent systematic review compared the peak time and diagnostic accuracy of miRNAs and conventional biomarkers in MI.
-The results revealed miR-1-3p, miR-19b-3p, miR-208a, miR-223-3p, miR-483-5p, and miR-499a-5p had superior peak times within 4 hours and better accuracy versus cTn and Creatine kinase-MB, indicating their promise for early diagnosis.
-The strengths of miRNAs included their early peak expression, satisfactory sensitivity and specificity, and higher accuracy especially within the first few hours of symptom onset compared to conventional biomarkers \cite{sys}.
-% Change E
-
-It has been postulated that the function and diagnostic properties of miRNAs are beyond the myocardium in patients with CVD.
-Specifically, the expression of miRNAs can vary in different biofluids and cell components such as serum and peripheral blood mononuclear cells (PBMCs) \cite{36}.
+involved in the pathogenesis of cardiac tissue injury \cite{m1}. Several biological processes, such as angiogenesis,
+cardiomyocyte growth and contractility, lipid metabolism, plaque
+formation, and cardiac rhythm, are regulated by miRNAs \cite{m2}. These elements can easily circulate in biofluids and could be
+considered theranostic targets in terms of CVDs \cite{m1}. It has been postulated that the function and diagnostic
+properties of miRNAs are beyond the myocardium in patients with CVD.
+Specifically, the expression of miRNAs can vary in different biofluids
+and cell components such as serum and peripheral blood mononuclear cells
+(PBMCs) \cite{36}.
+
 PBMCs are a fraction of white blood cells, including monocytes,
@@
 Microarray datasets were obtained from the Gene Expression Omnibus (GEO)
-database (\url{https://www.ncbi.nlm.nih.gov/geo/}).
-% Change S
-To obtain robust classification performance between MI, healthy control, and CAD samples, sufficiently large sample sizes for each group are required.
-For this purpose, the GSE59867 dataset was selected, as it contains sizable numbers of both MI and CAD samples.
-To provide an equally large set of healthy controls, the GSE56609 and GSE54475 datasets containing healthy samples were also included.
-Combining these three datasets enabled comparative analysis between MI, CAD, and healthy control groups with adequate statistical power.
-% Change E
+database (\url{https://www.ncbi.nlm.nih.gov/geo/}). To obtain sufficient
+classification power between MI, healthy, and CAD samples, a relatively
+large sample size is needed Therefore, GSE59867 for the MI and CAD
+samples and GSE56609 and GSE54475 for the healthy samples were selected.
 All samples were produced using Affymetrix Human Gene 1.0 ST Array
-platform (GPL6244). 
-% Change S
-This platform contains 189 miRNA probes based on the annotation data from the GEO database.
-% Change E
-Only healthy, CAD, and early-stage MI samples were
-selected from these datasets for further analysis.
-% Change S
-Early-stage MI samples were analyzed to enable detection of miRNA biomarkers specific to the initial ischemia and infarction event, before extensive myocardial necrosis and remodeling occurs.
-Using samples from the early phase enhances identification of miRNA signals related to plaque rupture and MI onset versus stable CAD.
-Additionally, early-stage samples allow investigation of mechanisms initiating myocardial injury.
-%Change E
-The basic information
+platform (GPL6244). Only healthy, CAD, and early-stage MI samples were
+selected from these datasets for further analysis. The basic information
 for the three datasets evaluated in this study is provided in Table
-~\ref{tab:datasets}. Bioinformatics analyses 
-%Change S
-including  preprocessing, differential expression analysis, and functional and pathway enrichment
-analyses
-%Change E
-were conducted using R, ver.
-4.2.0 \cite{R}, and RStudio \cite{RStudio}. All plots and
+~\ref{tab:datasets}. Bioinformatics analyses were conducted using R, ver.
+4.2.0 \cite{R}, RStudio \cite{RStudio}. All plots and
 graphics of these sections were created using the ggplot2 R package \cite{ggplot}.
@@
 values, the barcode representation of a sample was generated as a vector
-consisting of ones and zeros. 
-%Change S
-The ones and zeros generated by the barcode algorithm refer to binary calls of whether or not a gene is estimated to be expressed (1) or not expressed (0) in each individual sample. 
-%Change E
-The barcode function within the R fRMA package was employed to implement the barcode
+consisting of ones and zeros, representing the estimated expression
+(ones) and non-expression (zeros) of each gene. The barcode function
+within the R fRMA package was employed to implement the barcode
 algorithm, utilizing the default value of $C$.
@@
 group, to identify DEGs specific to each comparison.
-
-% Change S
-\subsubsection{Differentially expressed miRNAs}\label{differentially-expressed-miRNAs}
-
-The differentially expressed miRNAs were defined as those miRNAs within the total DEGs (i.e. they had an FDR $< 0.05$ resulted from the Fisher's exact test comparing the sample groups).
-
-% Change E
 
@@
 Two different approaches were used to select miRNAs for model training.
-The first approach was to use differentially expressed miRNAs.
-% Change
-To capture additional miRNAs with high discriminatory power for distinguishing MI from CAD despite not reaching differentially expression criteria, a secondary approach was used.
-miRNAs were selected based on having individual area under the receiver operating characteristic curves (AUC-ROCs) exceeding 0.8 for separating MI and CAD.
-This AUC-based approach identifies miRNAs with the best classification performance, unconstrained by statistical cutoffs.
-Using both the differentially expressed and AUC-based selection provides complementary methods to uncover miRNA biomarkers from both a biological and diagnostic perspective.
-% Change.
-
-\subsubsection{Differentially expressed miRNAs}\label{mirnas-in-degs}
+The first approach was to use differentially expressed miRNAs. In the
+second approach, miRNAs with individual area under the receiver operating characteristic curves (AUC-ROCs) over 0.8 for separating MI from CAD were selected. The results of these two approaches can
+provide an informative comparison between the predictive capabilities of
+sets of miRNAs selected with different logics.
+
+\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs}
 
@@
 the samples that were predicted as not healthy in the first layer. To
-this end, a distinct ML model was trained for each layer. Because there
+this end, a distinct ML model was trained for each layer. Because, there
 were a limited number of miRNAs in the DEGs, both layers were trained
@@
 A support vector machine (SVM) model using RBF kernels was trained and
-hypertuned using all miRNAs in the DEGs. 
-% Change
-To account for the substantial class imbalance between the healthy and not-healthy groups, with 51 samples in the minority healthy class compared to 157 combined CAD and MI samples, adjustments were made to the sample weights used during model training.
-Without compensating for the imbalance, the machine learning model would be biased towards the majority class and potentially ignore the minority class.
-To counteract this, the sample weights were empirically tuned, with the weight for healthy samples set to 1 and the weight for not-healthy samples set to 0.5.
-These values were determined through iterative testing to produce a model with strong performance on both classes despite the imbalance.
-% Change
-The ROC curve
+hypertuned using all miRNAs in the DEGs. To handle the severe imbalance
+in the number of samples (51 for the healthy group and 157 for the
+not-healthy group), the sample weights for the healthy and the
+not-healthy samples were set to 1 and 0.5, respectively. The ROC curve
 and confusion matrix for the model are reported.
@@
 \label{tab:DEGstab}
-\begin{tabularx}{\textwidth}{XXXXX}
-\hline
+\begin{tabular*}{\textwidth}{@{\extracolsep\fill}lcccc}
+\toprule
 & Total DEGs & Up-regulated DEGs & Down-regulated DEGs & miRNAs\\
-\hline
-MI vs. Healthy & 860 & 323 & 537 & hsa-miR-186-5p, hsa-miR-21-3p, hsa-miR-32-3p\\
-CAD vs. Healthy & 670 & 262 & 408 & hsa-miR-186-5p, hsa-miR-21-3p, hsa-miR-32-3p\\
-MI vs. CAD & 260 & 144 & 116 & hsa-miR-186-5p\\
-\hline
-\end{tabularx}
+\midrule
+MI vs. Healthy & 860 & 323 & 537 & miR-186, miR-21, miR-32\\
+CAD vs. Healthy & 670 & 262 & 408 & miR-186, miR-21, miR-32\\
+MI vs. CAD & 260 & 144 & 116 & miR-186\\
+\bottomrule
+\end{tabular*}
 \end{table}
@@
 DEGs and 79\% of CAD/healthy DEGs.
-
-% Change S
-\subsubsection{Differentially expressed miRNAs}\label{mirnas-in-degs}
-Among the DEGs for MI/healthy and CAD/healthy comparison, hsa-miR-186-5p, hsa-miR-32-3p, and hsa-miR-21-3p were identified as differentially expressed miRNAs.
-The only differentially expressed miRNA in MI/CAD comparison was hsa-miR-186-5p (Table~\ref{tab:DEGstab}).
-The expression profiles of the three
-miRNAs are shown in Figure~\ref{fig:Expall}.
-% Change E
 
@@
 
-\subsubsection{Differentially expressed miRNAs}\label{mirnas-in-degs-1}
-
-The ROC curves of each miRNA in each layer are presented in Figure~\ref{fig:miRROC}. Using the
-logistic regression model, the AUC-ROC values of hsa-miR-21-3p, hsa-miR-32-3p, and
-hsa-miR-186-5p for separating healthy and not-healthy samples were 0.98, 0.99,
+\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs-1}
+
+Among the DEGs, miR-186, miR-32, and miR-21 were identified as
+differentially expressed miRNAs. The expression profiles of the three
+miRNAs are shown in Figure~\ref{fig:Expall}. The ROC curves of each
+miRNA in each layer are presented in Figure~\ref{fig:miRROC}. Using the
+logistic regression model, the AUC-ROC values of miR-21, miR-32, and
+miR-186 for separating healthy and not-healthy samples were 0.98, 0.99,
 and 0.90, respectively (Figure~\ref{fig:miRROC}A). The accuracy of each
 miRNA for classifying the samples into healthy and not-healthy groups on
-the test set for hsa-miR-21-3p, hsa-miR-32-3p, and hsa-miR-186-5p was 0.92, 0.98, and 0.89,
+the test set for miR-21, miR-32, and miR-186 was 0.92, 0.98, and 0.89,
 respectively. The ROC curve of each miRNA for classifying MI and CAD
 samples is presented in Figure~\ref{fig:miRROC}B. The AUC-ROC and
-accuracy for hsa-miR-21-3p, hsa-miR-32-3p, and hsa-miR-186-5p in the test set were 0.85;
+accuracy for miR-21, miR-32, and miR-186 in the test set were 0.85;
 0.70; and 0.86, and 0.78; 0.67; and 0.74, respectively.
@@
 \midrule
-hsa-miR-186-5p & 1.4 & 3.60e-25 & 0.9 & 6.76e-20 & -0.5 & 1.05e-09\\
-hsa-miR-21-3p & 1.4 & 1.31e-17 & 2.3 & 2.07e-47 & 0.8 & 2.96e-11\\
-hsa-miR-32-3p & 2.5 & 8.39e-43 & 2.2 & 3.10e-59 & -0.3 & 7.60e-04\\
-hsa-miR-197-5p & 0.5 & 2.95e-20 & 0.7 & 1.59e-47 & 0.2 & 8.58e-09\\
-hsa-miR-29a-5p & 0.7 & 7.76e-29 & 0.1 & 1.70e-01 & -0.5 & 2.14e-10\\
-hsa-miR-296-5p & -0.1 & 5.00e-02 & 0.1 & 2.00e-02 & 0.2 & 6.15e-06\\
+miR-186 & 1.4 & 3.60e-25 & 0.9 & 6.76e-20 & -0.5 & 1.05e-09\\
+miR-21 & 1.4 & 1.31e-17 & 2.3 & 2.07e-47 & 0.8 & 2.96e-11\\
+miR-32 & 2.5 & 8.39e-43 & 2.2 & 3.10e-59 & -0.3 & 7.60e-04\\
+miR-197 & 0.5 & 2.95e-20 & 0.7 & 1.59e-47 & 0.2 & 8.58e-09\\
+miR-29A & 0.7 & 7.76e-29 & 0.1 & 1.70e-01 & -0.5 & 2.14e-10\\
+miR-296 & -0.1 & 5.00e-02 & 0.1 & 2.00e-02 & 0.2 & 6.15e-06\\
 \bottomrule
@@
 \includegraphics[width=0.95\linewidth]{DEMs ROCs h not h cad mi}
-\caption{ROC curve for the model trained with differentially expressed miRNAs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMROC}
+\caption{ROC curve for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMROC}
 \end{figure}
@@
 \includegraphics[width=0.9\linewidth]{DEMs Conf. matrix}
-\caption{Confusion matrix for the model trained with differentially expressed miRNAs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMCF}
+\caption{Confusion matrix for the model trained with miRNAs in DEGs on test set classification; (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI sample classification.}\label{fig:DEMCF}
 \end{figure}
@@
 \includegraphics[width=0.65\linewidth]{Models DEGs mirs}
-\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three differentially expressed miRNAs.}\label{fig:DEMmodels}
+\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs.}\label{fig:DEMmodels}
 \end{figure}
@@
 samples, the miRNAs with AUC-ROC $> 0.8$ were selected. The miRNAs
-selected were hsa-miR-29a-5p, hsa-miR-197-5p, hsa-miR-186-5p, hsa-miR-21-3p, and hsa-miR-296-5p. The
+selected were miR-29a, miR-197, miR-186, miR-21, and miR-296. The
 expression levels of these miRNAs in healthy, CAD, and MI samples are
@@
 is a major part of PBMCs. Two sets of miRNAs were selected as biomarker
-sets for sample classification. Hsa-miR-21-3p; hsa-miR-32-3p; and hsa-miR-186-5p were
-selected as differentially expressed miRNAs, and hsa-miR-186-5p; hsa-miR-21-3p;
-hsa-miR-29a-5p; hsa-miR-197-5p; and hsa-miR-296-5p were selected based on their AUC-ROC
+sets for sample classification. miR-21; miR-32; and miR-186 were
+selected as differentially expressed miRNAs, and miR-186; miR-21;
+miR-29a; miR-197; and miR-296 were selected based on their AUC-ROC
 values. As shown in Figure~\ref{fig:miRROC}, all miRNAs selected with
 both approaches had AUC-ROCs $> 0.9$ for isolating healthy and
-not-healthy samples except for hsa-miR-296-5p and hsa-miR-29a-5p. The data confirmed
+not-healthy samples except for miR-296 and miR-29a. The data confirmed
 that the real challenge was to classify CAD and MI samples because of
 the close overlap. Of the six miRNAs under investigation in both
-approaches, except for hsa-miR-32-3p, all miRNAs had an AUC-ROC $> 0.8$ for
+approaches, except for miR-32, all miRNAs had an AUC-ROC $> 0.8$ for
 the discrimination of CAD and MI samples. As expected, the high AUC-ROC
@@
 under cardiovascular events \cite{296}. For instance, there is
-evidence that elevation of hsa-miR-186-5p suppresses the expression of
+evidence that elevation of miR-186 suppresses the expression of
 cystathionine-$\gamma$-lyase, leading to the subsequent secretion of
 pro-inflammatory cytokines and cellular lipid accumulation. In addition,
-macrophage-derived hsa-miR-186-5p may promote atherosclerotic plaque formation
-\cite{186}. In line with this claim, we found that hsa-miR-186-5p was
+macrophage-derived miR-186 may promote atherosclerotic plaque formation
+\cite{186}. In line with this claim, we found that miR-186 was
 up-regulated in both CAD and MI candidates compared to their control
 counterparts. Surprisingly, the obtained data indicated that the
-expression of hsa-miR-186-5p was higher in patients with CAD than in patients
-with MI (Figure~\ref{fig:Expall}). Specifically, hsa-miR-186-5p was the only
+expression of miR-186 was higher in patients with CAD than in patients
+with MI (Figure~\ref{fig:Expall}). Specifically, miR-186 was the only
 differentially expressed miRNA between CAD and MI, with a clear
@@
 
-As mentioned before, hsa-miR-21-3p was also up-regulated in both MI and CAD
+As mentioned before, miR-21 was also up-regulated in both MI and CAD
 patients compared to healthy controls. Moreover, the expression value of
-hsa-miR-21-3p was significantly higher in the MI group than in the CAD group
+miR-21 was significantly higher in the MI group than in the CAD group
 (Table~\ref{tab:mirExptable}). It is thought that the up-regulation of
-hsa-miR-21-3p in PBMCs is a compensatory reaction to reduce the T$_{reg}$
+miR-21 in PBMCs is a compensatory reaction to reduce the T$_{reg}$
 lymphocyte number in response to the reduction in TGF$\beta1$
 secretion into the plasma through a TGF$\beta1$/smad-independent
-pathway. In line with the previous and present data, hsa-miR-21-3p can modulate
+pathway. In line with the previous and present data, miR-21 can modulate
 the activity of PBMCs following the occurrence of cardiovascular
@@
 
-Recent data have supported the elevation of hsa-miR-32-3p levels in CAD samples
-with calcification of the coronary artery. Notably, hsa-miR-32-3p promotes
+Recent data have supported the elevation of miR-32 levels in CAD samples
+with calcification of the coronary artery. Notably, miR-32 promotes
 vascular smooth muscle calcification in mice by controlling the activity
@@
 bone-specific phosphoprotein matrix GLA protein \cite{32-2}.
-Likewise, some reports are associated with the activity of hsa-miR-32-3p in
+Likewise, some reports are associated with the activity of miR-32 in
 PBMCs in several pathologies \cite{magic, 32}. The
-exact role of hsa-miR-32-3p in PBMCs after cardiovascular events remains
+exact role of miR-32 in PBMCs after cardiovascular events remains
 unclear.
@@
 biological importance of two common miRNAs in the DEGs and AUC-ROC
-approaches, hsa-miR-21-3p and hsa-miR-186-5p, have already been discussed. Based on
-numerous reports, hsa-miR-29a-5p can be activated in different diseases \cite{29aa}. Data analysis indicated that
-hsa-miR-29a-5p was significantly up-regulated in CAD patients compared to the
-healthy and MI groups (Table~\ref{tab:mirExptable}). Increased hsa-miR-29a-5p
+approaches, miR-21 and miR-186, have already been discussed. Based on
+numerous reports, miR-29a can be activated in different diseases \cite{29aa}. Data analysis indicated that
+miR-29a was significantly up-regulated in CAD patients compared to the
+healthy and MI groups (Table~\ref{tab:mirExptable}). Increased miR-29a
 is associated with the progression of atherosclerosis, and the
-combination of hsa-miR-29a-5p and ox-LDL has been suggested as a valid
+combination of miR-29a and ox-LDL has been suggested as a valid
 biomarker set for paraclinical classification \cite{29a}.
-However, the role of hsa-miR-29a-5p in the function of PBMCs from patients with
+However, the role of miR-29a in the function of PBMCs from patients with
 CAD has not been thoroughly examined.
 
-The data indicated that hsa-miR-197-5p was significantly up-regulated in both
+The data indicated that miR-197 was significantly up-regulated in both
 the CAD/healthy and MI/healthy groups. Previous studies have
-demonstrated that hsa-miR-197-5p may play a crucial role in controlling the
+demonstrated that miR-197 may play a crucial role in controlling the
 anti-inflammatory response of IL-35 by influencing the secretion of
@@
 M1/M2 macrophages, and the proliferation of T$_{reg}$ lymphocytes,
-which are responsible for suppressing immune responses \cite{197}. Alongside our findings, it can be concluded that hsa-miR-197-5p could be
+which are responsible for suppressing immune responses \cite{197}. Alongside our findings, it can be concluded that miR-197 could be
 a useful diagnostic tool for predicting adverse cardiovascular events.
 
-The findings of this study demonstrate the potential of hsa-miR-296-5p as a
+The findings of this study demonstrate the potential of miR-296 as a
 biomarker with high discriminatory power to distinguish between samples
-from individuals with MI and CAD. Hsa-miR-296-5p has been identified as a key
+from individuals with MI and CAD. MiR-296 has been identified as a key
 regulator in the development and advancement of atherosclerosis by
@@
 biological processes, including angiogenesis, cholesterol metabolism,
-inflammation, cellular proliferation, hypertension, and apoptosis \cite{296}. In a previous study, hsa-miR-296-5p expression levels were found
+inflammation, cellular proliferation, hypertension, and apoptosis \cite{296}. In a previous study, miR-296 expression levels were found
 to be significantly increased in the PBMCs of CAD patients compared to
 healthy controls, suggesting its involvement in regulating
-proinflammatory cytokines such as IL-6 and TNF-$\alpha$ \cite{296a}. These findings suggested that hsa-miR-296-5p may have a significant
+proinflammatory cytokines such as IL-6 and TNF-$\alpha$ \cite{296a}. These findings suggested that miR-296 may have a significant
 impact on the pathogenesis of atherosclerosis and could potentially
@@
   \item \textbf{miRNA}: microRNA 
-  \item \textbf{ACS}: Acute coronary syndrome
   \item \textbf{PBMC}: Peripheral blood mononuclear cell 
@@
 F.B. and K.K. are author responsible for contact and ensures communication.
-All authors read the content of final manuscript.
 
